---
title: "Subnational Micronutrient Deficiency"
subtitle: "District ranking from public data, for programme targeting"
date: today
params:
  country: "ghana"
format:
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    fig-width: 7
    fig-height: 5.2
    geometry: margin=2.2cm
    colorlinks: true
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    embed-resources: true
    theme: cosmo
keep-md: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
# The dashboard's own loader, so the printed page and the screen read the same
# tables, caveats and thresholds. global.R expects to run from dashboard/.
.here <- here::here("dashboard")
.owd  <- setwd(.here)
suppressPackageStartupMessages(source("global.R"))
setwd(.owd)
suppressPackageStartupMessages({ library(ggplot2); library(sf); library(dplyr); library(knitr); library(kableExtra) })
theme_set(theme_minimal(base_size = 10))

CK <- params$country
CLAB <- unname(meta$countries[[CK]])
OCS <- outcomes_for(CK)
dat <- setNames(lapply(OCS, function(oc) get_country_admin2(CK, oc)), OCS)
NAT <- idx_national[idx_national$country_key == CK, ]
pct  <- function(x, d = 1) ifelse(is.finite(x), sprintf(paste0("%.", d, "f%%"), 100 * x), "--")
cnt  <- function(x) ifelse(is.finite(x), formatC(round(x), big.mark = ",", format = "d"), "--")
num  <- function(x, d = 2) ifelse(is.finite(x), formatC(x, format = "f", digits = d), "--")
lab  <- function(oc) unname(meta$outcome_labels[oc])
```

# What this is

`r SITE_SCOPE_HEADLINE` `r SITE_SCOPE_BODY`

This brief covers **`r CLAB`**. Its survey measured deficiency in
`r NAT$n_surveyed[1]` of the country's `r NAT$n_districts[1]` districts. The
model, built from `r Q$n_predictors` public data layers and scored against the
four national biomarker surveys, ranks every district, including the
`r NAT$n_districts[1] - NAT$n_surveyed[1]` the survey never reached.

The product is the ranking. The percentages are the ranking anchored to the
survey's national figure, so their order is the model's and their level is
the survey's. Read them in bands, not to the decimal. Section
\ref{sec-confidence} says how far to trust each outcome.

Data build: `r data_build_time` (`r PROTOCOL_LABEL`).

# National picture

```{r national}
nat <- do.call(rbind, lapply(OCS, function(oc) {
  d <- sf::st_drop_geometry(dat[[oc]]); n <- NAT[NAT$outcome == oc, ]
  data.frame(Outcome = lab(oc), `National prevalence (survey)` = pct(n$national_prev),
             `Surveyed districts` = sprintf("%d of %d", n$n_surveyed, n$n_districts),
             `People affected` = cnt(sum(d$people_affected, na.rm = TRUE)),
             `Of` = cnt(sum(d$population, na.rm = TRUE)), check.names = FALSE, stringsAsFactors = FALSE)
}))
kable(nat, align = "lrrrr", booktabs = TRUE,
      caption = sprintf("The survey's national prevalence and the planning count, %s", CLAB)) |>
  kable_styling(latex_options = c("hold_position"), font_size = 9)
```

"People affected" applies the planning prevalence to the population of the
group each biomarker describes (children 6 to 59 months, or women 15 to 49).
It is a planning count, not a measurement.

# Where deficiency is concentrated

```{r maps}
#| results: asis
#| fig-width: 6.6
#| fig-height: 5.0
#| dev: png
#| dpi: 200
for (oc in OCS) {
  d <- dat[[oc]]
  cat(sprintf("\n## %s\n\n", lab(oc)))
  cv <- biomarker_caveats[[oc]]
  if (!is.null(cv)) cat(sprintf("*%s*\n\n", cv))
  p <- ggplot(d) +
    geom_sf(aes(fill = priority), colour = "white", linewidth = 0.08) +
    geom_sf(data = d[isTRUE(TRUE) & !is.na(d$surveyed) & d$surveyed, ], fill = NA, colour = "grey15", linewidth = 0.35) +
    scale_fill_gradientn(colours = c("#ffffcc", "#fd8d3c", "#bd0026"), limits = c(0, 100), na.value = "grey88",
                         name = "Priority\nscore") +
    theme_void(base_size = 9) + theme(legend.position = "right")
  print(p)
  cat("\n\nPriority score 100 = ranked worst in the country. Districts outlined in dark grey had their own survey clusters.\n\n")
}
```

# Districts to look at first

```{r top-districts}
#| results: asis
for (oc in OCS) {
  df <- sf::st_drop_geometry(dat[[oc]])
  df <- df[is.finite(df$rank_worst), ]
  df <- df[order(df$rank_worst), ][seq_len(min(10, nrow(df))), ]
  out <- data.frame(Rank = df$rank_worst, District = df$Admin2, Region = df$Admin1,
                    `Chance of worst fifth` = ifelse(is.finite(df$p_worst_fifth), pct(df$p_worst_fifth, 0), "--"),
                    `Planning prevalence` = pct(df$prev_anchored),
                    `Survey estimate` = ifelse(is.finite(df$survey_prev), pct(df$survey_prev), "not surveyed"),
                    `WHO class` = df$who_class, check.names = FALSE, stringsAsFactors = FALSE)
  cat("\n\n")
  print(kable(out, align = "rllrrrl", booktabs = TRUE,
              caption = sprintf("Ranked worst: %s", lab(oc))) |>
          kable_styling(latex_options = c("hold_position"), font_size = 8))
  cat("\n\n")
}
```

"Chance of worst fifth" is how often the district landed in the worst fifth
when the model was refitted 40 times with that district hidden. It exists for
surveyed districts, where there is a survey to check against.

# How well the ranking holds up here {#sec-confidence}

```{r accuracy}
BC <- EV$benchmarks_cells; VC <- EV$ceiling; TCc <- EV$targeting_cells
acc <- do.call(rbind, lapply(OCS, function(oc) {
  g <- function(arm, tg) if (is.null(BC)) NA else g1(BC$spearman[BC$country == CLAB & BC$outcome == oc & BC$estimand == "infill" & BC$arm == arm & BC$target == tg])
  ce <- if (is.null(VC)) NA else g1(VC$ceiling_vc[VC$country == CLAB & VC$outcome == oc & VC$rung == "admin2" & VC$target == "prev"])
  cp <- if (is.null(TCc)) NA else g1(mean(TCc$capture_top20[TCc$country == CLAB & TCc$outcome == oc & TCc$estimand == "infill" & TCc$arm == "domain_index"], na.rm = TRUE))
  data.frame(Outcome = lab(oc), `Model, district hidden` = num(g("domain_index", "prev")),
             `Survey's regional average` = num(g("region_mean_jk", "prev")), `Ceiling` = num(ce),
             `Burden in the worst fifth` = pct(cp, 0), check.names = FALSE, stringsAsFactors = FALSE)
}))
kable(acc, align = "lrrrr", booktabs = TRUE,
      caption = sprintf("Ranking accuracy on prevalence with each district hidden from the fit, against the survey's own regional averages and the ceiling the survey's noise allows; chance is %s. Not computed where the country has too few districts to fold.", num(Q$null_d))) |>
  kable_styling(latex_options = c("hold_position"), font_size = 8)
```

`r GENERAL_CAVEAT`

**Use the ranking more than the number.** Which districts are worst off is the
part that holds up. The planning prevalence borrows its level from the national
survey and should be read in WHO bands.

**The survey's own district values are noisy.** Most districts hold one or two
survey clusters, so even a perfect predictor could not reach 1; the ceiling
column above is what it could reach. Differences in accuracy under 0.03 are ties.

**Some nutrients are measured better than others.** The note under each map says
which.

# Method, in short

Each district is described by `r Q$n_predictors` public data layers in
`r Q$n_domains` groups: satellite imagery, climate, soil, crops, livestock,
malaria and other morbidity, prices and household-survey aggregates. Inside
each country every layer is replaced by its rank across districts; each group
is reduced to a few summary axes; each axis is weighted by how strongly it
tracked deficiency in the surveyed districts; the weighted axes are summed.
Nothing is tuned. For this brief the model was fitted on `r CLAB`'s surveyed
districts and applied to every district. No variable from the biomarker survey
is used as a predictor.

Across the four countries the model ranks hidden districts at
`r num(Q$infill)` against `r num(Q$infill_jk)` for the survey's own regional
averages and `r num(Q$null_d)` for chance, and ranks the districts of a country
it has never seen at `r num(Q$tr)` (`r num(Q$cs)` from climate and soil layers
alone). The interactive version, with the evidence behind every number, is at
<https://amertens.shinyapps.io/micronutrient-burden/>.
