---
title: "Subnational Micronutrient Deficiency"
subtitle: "Working estimates for program targeting"
date: today
params:
  country: "ghana"
format:
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    fig-width: 7
    fig-height: 5.2
    geometry: margin=2.2cm
    colorlinks: true
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    embed-resources: true
    theme: cosmo
keep-md: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
# The dashboard's own loader. Sourcing it rather than re-reading the RDS files
# means this brief shares the app's data, palette, thresholds, caveat text and
# model registry — so the printed page cannot disagree with the screen.
# global.R expects to run from dashboard/, hence the setwd dance.
.here <- here::here("dashboard")
.owd  <- setwd(.here)
suppressPackageStartupMessages(source("global.R"))
setwd(.owd)

suppressPackageStartupMessages({
  library(ggplot2); library(sf); library(dplyr); library(knitr); library(kableExtra)
})
theme_set(theme_minimal(base_size = 10))

CK <- params$country
CLAB <- unname(meta$countries[[CK]])
PD <- pred_model_data(DEFAULT_PRED_MODEL)

# Outcomes this country actually carries, in the registry's order.
OCS <- intersect(names(meta$outcome_labels),
                 unique(PD$outcome[PD$country == CLAB]))

# One joined sf per outcome, reused by the maps and the tables.
dat <- setNames(lapply(OCS, function(oc) {
  d <- get_country_admin2(CK, oc, admin2_bnds, PD, admin2_pop)
  attach_prevalence_range(d, CLAB, oc)
}), OCS)

pct  <- function(x, d = 1) ifelse(is.finite(x), sprintf(paste0("%.", d, "f%%"), 100 * x), "--")
cnt  <- function(x) ifelse(is.finite(x), formatC(round(x), big.mark = ",", format = "d"), "--")

# Share of the deficient population reached by treating the worst districts
# covering `frac` of the population, versus reaching that share blind.
targeting <- function(d, frac = 0.20) {
  df <- as.data.frame(d)
  df <- df[is.finite(df$pred_prev) & is.finite(df$population) & df$population > 0, ]
  if (!nrow(df)) return(NULL)
  w <- df$population; b <- df$pred_prev * w
  o <- order(df$pred_prev, decreasing = TRUE)
  cx <- cumsum(w[o]) / sum(w); cy <- cumsum(b[o]) / sum(b)
  reach <- stats::approx(c(0, cx), c(0, cy), xout = frac, rule = 2)$y
  list(reach = reach, lift = reach / frac, frac = frac,
       curve = data.frame(pop = c(0, cx), burden = c(0, cy)))
}
```

# What this is

`r SITE_SCOPE_HEADLINE` `r SITE_SCOPE_BODY`

This brief covers **`r CLAB`**. It maps the estimated prevalence of
micronutrient deficiency at district level, including districts that were never
surveyed, so that a limited program can be aimed where need is greatest.

Read the percentages in bands, not to the decimal. The part that holds up is
*which districts are worst off*; the exact figure is a planning number, and how
firm it is varies by nutrient. Section \ref{sec-confidence} says where to be
careful.

Estimates come from the `r tolower(PRED_MODEL_INFO[[DEFAULT_PRED_MODEL]]$label)`
model: `r PRED_MODEL_INFO[[DEFAULT_PRED_MODEL]]$blurb`
Uncertainty ranges are drawn from the survey-blended small-area model.

Data build: `r data_build_time`.

# National picture

```{r national}
nat <- do.call(rbind, lapply(OCS, function(oc) {
  a <- national_aggregate(dat[[oc]])
  data.frame(Outcome = unname(meta$outcome_labels[oc]),
             Estimated = pct(a$pred_prev_natl),
             Affected = cnt(a$pop_at_risk_natl),
             Of = cnt(a$pop_total),
             stringsAsFactors = FALSE)
}))
names(nat) <- c("Outcome", "Estimated prevalence", "People affected", "Population")
kable(nat, align = "lrrr", booktabs = TRUE,
      caption = sprintf("Population-weighted national estimates, %s", CLAB)) |>
  kable_styling(latex_options = c("hold_position"), font_size = 9)
```

Counts use the population of the group each biomarker describes (children 6--59
months, or women 15--49), projected to the survey year.

```{r signal-coverage}
#| results: asis
sig <- vapply(OCS, function(oc) has_district_signal(dat[[oc]]$pred_prev), logical(1))
if (any(!sig)) cat(sprintf(paste(
  "**%d of these %d outcomes carry district-level signal.** For the other %d --",
  "%s -- the model returns the same figure for every district, so the national",
  "number above is all there is. Those outcomes are marked in the sections",
  "below and no map is drawn for them.\n\n"),
  sum(sig), length(sig), sum(!sig),
  # Not tolower(): it turns "Vitamin A deficiency" into "vitamin a deficiency".
  paste(unname(meta$outcome_labels[OCS[!sig]]), collapse = "; ")))
```

# Is subnational targeting worth it? {#sec-targeting}

```{r targeting-text}
#| results: asis
lead <- OCS[which.max(vapply(OCS, function(oc) {
  t <- targeting(dat[[oc]]); if (is.null(t) || !is.finite(t$lift)) 0 else t$lift
}, numeric(1)))]
tg <- targeting(dat[[lead]])
if (!is.null(tg)) cat(sprintf(paste(
  "Taking **%s** as the example: reaching the worst-affected **%.0f%%** of",
  "%s's population using these district estimates would cover about **%.0f%%**",
  "of everyone affected. Reaching the same share of the population without",
  "district data would cover roughly %.0f%%. That is a **%.1f-fold** gain for",
  "the same program size.\n\n"),
  tolower(meta$outcome_labels[lead]), 100 * tg$frac, CLAB,
  100 * tg$reach, 100 * tg$frac, tg$lift))
```

```{r targeting-curve}
#| fig-cap: "Burden concentration. The further the curve bows above the diagonal, the more targeting buys you. A straight diagonal means deficiency is spread evenly and a national approach is as good."
curves <- do.call(rbind, lapply(OCS, function(oc) {
  t <- targeting(dat[[oc]]); if (is.null(t)) return(NULL)
  transform(t$curve, Outcome = unname(meta$outcome_labels[oc]))
}))
ggplot(curves, aes(pop, burden, colour = Outcome)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey55") +
  geom_line(linewidth = 0.7) +
  # A hard expand of 0 clips the "100%" tick label at the panel edge.
  scale_x_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.04))) +
  scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.04))) +
  labs(x = "Share of population reached, worst districts first",
       y = "Share of affected people reached", colour = NULL) +
  theme(legend.position = "bottom", legend.text = element_text(size = 7)) +
  guides(colour = guide_legend(nrow = 3))
```

# Where deficiency is concentrated

```{r maps}
#| results: asis
#| fig-width: 6.6
#| fig-height: 5.0
#| dev: png
#| dpi: 200
# Rasterised deliberately. Drawn as vectors, Ghana's 260 district polygons and
# Malawi's 243 embed every vertex in the PDF and push it past 12 MB, which is
# useless for the emailing this brief exists to support. At 200 dpi the maps
# stay sharp in print and the file lands near 1 MB.
for (oc in OCS) {
  d <- dat[[oc]]
  cat(sprintf("\n## %s\n\n", meta$outcome_labels[oc]))
  cv <- biomarker_caveats[[oc]]
  if (!is.null(cv)) cat(sprintf("*%s*\n\n", cv))
  # A flat cell must not be drawn as a choropleth: identical values across every
  # district produce a single-colour map with a one-tick legend, which looks
  # like a rendering fault and reads as "no deficiency variation" rather than
  # "no information". State it instead.
  if (!has_district_signal(d$pred_prev)) {
    v <- d$pred_prev[is.finite(d$pred_prev)]
    cat(sprintf(paste(
      "> **No district-level signal.** The model returns about %s for every",
      "district, so there is no map to draw. The routinely available indicators",
      "carry nothing that separates districts for this nutrient in %s. Use the",
      "national figure above; do not rank districts on this outcome.\n\n"),
      pct(stats::median(v)), CLAB))
    next
  }
  p <- ggplot(d) +
    geom_sf(aes(fill = pred_prev), colour = "white", linewidth = 0.08) +
    geom_sf(data = d[!is.na(d$obs_prev), ], fill = NA,
            colour = "grey15", linewidth = 0.35) +
    scale_fill_gradientn(colours = c("#2c7bb6", "#abd9e9", "#ffffbf", "#fdae61", "#d7191c"),
                         labels = scales::percent, na.value = "grey88",
                         name = "Estimated\nprevalence") +
    theme_void(base_size = 9) +
    theme(legend.position = "right")
  print(p)
  cat("\n\nDistricts outlined in dark grey had their own survey; the rest are model estimates.\n\n")
}
```

# Districts to look at first

```{r top-districts}
#| results: asis
for (oc in OCS) {
  df <- as.data.frame(dat[[oc]])
  df <- df[is.finite(df$pred_prev), ]
  if (!nrow(df)) next
  # Ranking identical values puts whichever district sorted first at the top of
  # a "worst affected" table. That is an artefact of sort order, not a finding.
  if (!has_district_signal(df$pred_prev)) {
    cat(sprintf(paste("\n\n**%s** -- omitted: every district returns the same",
                      "estimate, so a ranking would be arbitrary.\n\n"),
                meta$outcome_labels[oc]))
    next
  }
  df <- df[order(-df$pred_prev), ][seq_len(min(10, nrow(df))), ]
  out <- data.frame(
    District = df$Admin2,
    Estimated = pct(df$pred_prev),
    Range = ifelse(is.finite(df$prev_lo) & is.finite(df$prev_hi),
                   sprintf("%s - %s", pct(df$prev_lo, 0), pct(df$prev_hi, 0)), "--"),
    Affected = cnt(df$pop_at_risk),
    Class = ifelse(is.na(df$who_class), "--", df$who_class),
    Survey = ifelse(is.na(df$obs_prev), "modelled", "surveyed"),
    stringsAsFactors = FALSE)
  names(out) <- c("District", "Estimated", "95% range", "People affected",
                  "WHO class", "Source")
  cat("\n\n")
  print(kable(out, align = "lrrrll", booktabs = TRUE,
              caption = sprintf("Highest estimated prevalence: %s",
                                meta$outcome_labels[oc])) |>
          kable_styling(latex_options = c("hold_position"), font_size = 8))
  cat("\n\n")
}
```

# What to trust {#sec-confidence}

`r GENERAL_CAVEAT`

**Some nutrients are measured better than others.** Vitamin A in women and B12
rest on weaker biomarkers; the note under each map above says which.

**Use the ranking more than the number.** Which districts are worst off is the
part that holds up. The 95% ranges above show how much room there is around any
single figure, and they cover uncertainty in the prevalence estimate only.

**Districts without their own survey carry more risk.** In this country
`r sum(vapply(OCS, function(oc) sum(!is.na(dat[[oc]]$obs_prev)), numeric(1))[1])`
of `r nrow(dat[[OCS[1]]])` districts were surveyed directly; the rest are
estimated from satellite, climate and geospatial indicators.

# Method, in short

Each district's prevalence is modelled directly from routinely available
indicators — satellite imagery, climate reanalysis, modelled disease burden,
food security and price data, and district-level survey indicators. No variable
from the biomarker survey itself is used as a predictor, so the same model can
be applied where no survey exists.

Alternatives available in the interactive dashboard:

```{r model-table}
mt <- data.frame(
  Model = vapply(names(PRED_MODEL_INFO), function(k) PRED_MODEL_INFO[[k]]$label, character(1)),
  What  = vapply(names(PRED_MODEL_INFO), function(k) PRED_MODEL_INFO[[k]]$blurb, character(1)),
  stringsAsFactors = FALSE)
kable(mt, row.names = FALSE, booktabs = TRUE, caption = "Models offered") |>
  column_spec(2, width = "10cm") |>
  kable_styling(latex_options = c("hold_position"), font_size = 8)
```

The interactive version, with the full technical appendix, is at
<https://amertens.shinyapps.io/micronutrient-burden/>.
