---
title: "Subnational Micronutrient Deficiency"
subtitle: "Four-country overview"
date: today
format:
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    fig-width: 7
    fig-height: 4.6
    geometry: margin=2.2cm
    colorlinks: true
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    embed-resources: true
    theme: cosmo
keep-md: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
.owd <- setwd(here::here("dashboard"))
suppressPackageStartupMessages(source("global.R"))
setwd(.owd)
suppressPackageStartupMessages({ library(ggplot2); library(dplyr); library(knitr); library(kableExtra) })
theme_set(theme_minimal(base_size = 10))
pct <- function(x, d = 1) ifelse(is.finite(x), sprintf(paste0("%.", d, "f%%"), 100 * x), "--")
num <- function(x, d = 2) ifelse(is.finite(x), formatC(x, format = "f", digits = d), "--")
BC <- EV$benchmarks_cells; VC <- EV$ceiling; TCc <- EV$targeting_cells
grid <- idx_national |> rowwise() |> mutate(
  label = unname(meta$outcome_labels[outcome]),
  acc = if (is.null(BC)) NA_real_ else g1(BC$spearman[BC$country == country & BC$outcome == outcome & BC$estimand == "infill" & BC$arm == "domain_index" & BC$target == "prev"]),
  acc_jk = if (is.null(BC)) NA_real_ else g1(BC$spearman[BC$country == country & BC$outcome == outcome & BC$estimand == "infill" & BC$arm == "region_mean_jk" & BC$target == "prev"]),
  ceiling = if (is.null(VC)) NA_real_ else g1(VC$ceiling_vc[VC$country == country & VC$outcome == outcome & VC$rung == "admin2" & VC$target == "prev"]),
  capture = if (is.null(TCc)) NA_real_ else g1(mean(TCc$capture_top20[TCc$country == country & TCc$outcome == outcome & TCc$estimand == "infill" & TCc$arm == "domain_index"], na.rm = TRUE))) |>
  ungroup()
```

# What this covers

`r SITE_SCOPE_HEADLINE` `r SITE_SCOPE_BODY`

A ranking of districts by likely micronutrient deficiency for every district of
The Gambia, Ghana, Sierra Leone and Malawi, built from `r Q$n_predictors` public
data layers and scored against the four national biomarker surveys. Country
detail is in the individual briefs; this is the portfolio view.

Data build: `r data_build_time` (`r PROTOCOL_LABEL`). Interactive version:
<https://amertens.shinyapps.io/micronutrient-burden/>

```{r coverage}
cov <- grid |> group_by(Country = country) |>
  summarise(Outcomes = n(), Districts = max(n_districts), Surveyed = max(n_surveyed), .groups = "drop")
kable(cov, align = "lrrr", booktabs = TRUE,
      caption = "Coverage. Surveyed counts districts with their own biomarker clusters; the model ranks all districts.") |>
  kable_styling(latex_options = "hold_position", font_size = 9)
```

# National prevalence, from the surveys

```{r heat}
#| fig-cap: "National prevalence measured by each survey. Blank cells are combinations the survey did not measure."
#| fig-height: 4.2
ggplot(grid, aes(x = country, y = label, fill = national_prev)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = pct(national_prev, 0)), size = 2.9, colour = ifelse(grid$national_prev > 0.55, "white", "grey15")) +
  scale_fill_gradientn(colours = c("#2c7bb6", "#abd9e9", "#ffffbf", "#fdae61", "#d7191c"), labels = scales::percent, name = "Prevalence") +
  labs(x = NULL, y = NULL) + theme(panel.grid = element_blank(), legend.position = "right")
```

# How well the ranking holds up, cell by cell

```{r acc-table}
tab <- grid |> transmute(Country = country, Outcome = label,
                         `Model, district hidden` = num(acc), `Survey's regional average` = num(acc_jk),
                         Ceiling = num(ceiling), `Burden in the worst fifth` = pct(capture, 0)) |>
  arrange(Country, Outcome)
kable(tab, align = "llrrrr", booktabs = TRUE, longtable = TRUE,
      caption = sprintf("Ranking accuracy on prevalence with each district hidden from the fit, the survey's own regional average scored the same way, the ceiling the survey's noise allows, and the share of the country's deficient people in the fifth of districts the model ranks worst. Chance is %s. Sierra Leone's 14 districts cannot be folded.", num(Q$null_d))) |>
  kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 8)
```

```{r acc-plot}
#| fig-cap: "The model against the survey's own regional averages, each district hidden in turn. Points above the diagonal are cells where the model ranks better than the survey's regional average does."
#| fig-height: 4.4
g <- grid[is.finite(grid$acc) & is.finite(grid$acc_jk), ]
ggplot(g, aes(acc_jk, acc, colour = country)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey55") +
  geom_hline(yintercept = Q$null_d, linetype = 3, colour = "grey55") +
  geom_point(size = 2.4) +
  labs(x = "Survey's regional average (ranking accuracy)", y = "Model, district hidden (ranking accuracy)", colour = NULL) +
  theme(legend.position = "bottom")
```

# A country the model has never seen

```{r transport}
#| fig-cap: "Ranking accuracy in each country when the model is fitted on the other three, by outcome, biomarker level. The grey band is chance. This is the case for applying the model where there is no biomarker survey."
#| fig-height: 4.4
tr <- if (is.null(BC)) NULL else BC[BC$estimand == "country" & BC$arm == "domain_index" & BC$target == "level", ]
if (!is.null(tr) && nrow(tr)) {
  tr$label <- unname(meta$outcome_labels[tr$outcome])
  ggplot(tr, aes(spearman, label, colour = country)) +
    annotate("rect", xmin = -Inf, xmax = Q$null_d, ymin = -Inf, ymax = Inf, fill = "grey88", alpha = 0.7) +
    geom_vline(xintercept = 0, colour = "grey60") +
    geom_point(size = 2.4) +
    labs(x = "Ranking accuracy in the held-out country", y = NULL, colour = NULL) +
    theme(legend.position = "bottom")
} else {
  ggplot() + theme_void() + annotate("text", 0, 0, label = "Transport cells not built")
}
```

Across the 22 country-outcome combinations the full index averages
`r num(Q$tr)` (positive in `r Q$tr_pos`), an index from climate and soil layers
alone `r num(Q$cs)`, against `r num(Q$null_d)` for chance. The ranking crosses
borders; the prevalence level does not, so a new country needs one national
survey number to anchor it.

`r GENERAL_CAVEAT`

# What to trust

**Use the ranking more than the number.** The percentages in the briefs are the
ranking anchored to each survey's national figure.

**The ceiling is the survey's.** A perfect predictor could reach about
`r num(Q$ceiling_prev)` on prevalence given the surveys' own district noise;
the model reaches `r num(Q$infill_prev)`, about two-thirds of the way.

**Some nutrients are measured better than others.** Vitamin A in women and B12
rest on weaker markers, folate is measured in three countries, and zinc in
Malawi only.
