---
title: "Subnational Micronutrient Deficiency"
subtitle: "Technical annex"
date: today
format:
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    fig-width: 7
    fig-height: 4.6
    geometry: margin=2.2cm
    colorlinks: true
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    embed-resources: true
    theme: cosmo
keep-md: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
# Static counterpart to the dashboard's Technical appendix: Model diagnostics,
# Benchmarks, Resolution, Transportability and the importance views. Same
# bundles the app reads, so the two cannot disagree.
.owd <- setwd(here::here("dashboard"))
suppressPackageStartupMessages(source("global.R"))
setwd(.owd)
suppressPackageStartupMessages({
  library(ggplot2); library(dplyr); library(knitr); library(kableExtra)
})
theme_set(theme_minimal(base_size = 10))
lab <- function(oc) unname(meta$outcome_labels[oc])
pct <- function(x, d = 1) ifelse(is.finite(x), sprintf(paste0("%.", d, "f%%"), 100 * x), "--")
num <- function(x, d = 2) ifelse(is.finite(x), sprintf(paste0("%.", d, "f"), x), "--")
empty <- function(msg) { cat(sprintf("\n*%s*\n\n", msg)); invisible(NULL) }
```

# Scope

`r SITE_SCOPE_HEADLINE`

This annex holds the modelling diagnostics behind the country briefs: how well
the models discriminate, whether their probabilities are calibrated, how they
compare with simpler alternatives, and how far they transport to a country they
were not trained on. It is the printed form of the dashboard's **Technical
appendix**, for reviewers who want to check the machinery rather than the
findings.

Data build: `r data_build_time`.

# How predictions are validated

Two complementary designs:

**Cluster-blocked cross-validation.** Individuals from the same survey cluster
always land in the same fold, so the model is tested on people spatially
separated from those it trained on. Without this, spatial autocorrelation makes
performance look better than it is.

**Leave-one-country-out (LOCO).** Train on three countries, test on the fourth.
This estimates performance in a country with no biomarker data — the case that
matters for scaling up, and the least flattering.

# Discrimination and calibration

```{r binary}
#| results: asis
b <- model_diag$binary
if (is.null(b) || !nrow(b)) empty("Binary diagnostics not built.") else {
  b$Outcome <- lab(b$outcome)
  t <- b |>
    transmute(Country = country, Outcome, N = n,
              Prevalence = pct(prevalence), `ROC-AUC` = num(roc_auc),
              `PR-AUC` = num(pr_auc), Brier = num(brier, 3),
              `Brier skill` = num(brier_skill, 3)) |>
    arrange(Country, Outcome)
  print(kable(t, align = "llrrrrrr", booktabs = TRUE, longtable = TRUE,
              caption = "Cluster-blocked cross-validated discrimination.") |>
          kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 7.5))
}
```

ROC-AUC runs from 0.5 (no better than chance) to 1.0. **PR-AUC is the more
informative of the two when prevalence is low** — with a rare outcome, ROC-AUC
can look high simply because the model correctly predicts the many non-events.
Brier skill compares squared error against always predicting the overall
prevalence: at or below zero, the model adds nothing beyond knowing the average.

```{r roc}
#| fig-cap: "ROC curves by country. The diagonal is chance."
#| fig-height: 5.6
r <- model_diag$roc
if (!is.null(r) && nrow(r)) {
  r$Outcome <- lab(r$outcome)
  ggplot(r, aes(fpr, tpr, colour = Outcome)) +
    geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey60") +
    geom_line(linewidth = 0.5) +
    facet_wrap(~country) +
    labs(x = "False positive rate", y = "True positive rate", colour = NULL) +
    theme(legend.position = "bottom", legend.text = element_text(size = 6)) +
    guides(colour = guide_legend(nrow = 3))
} else ggplot() + theme_void()
```

```{r calib}
#| fig-cap: "Calibration. Points on the diagonal mean that when the model says 30%, about 30% of people are deficient. Below the line means over-prediction."
#| fig-height: 5.2
cal <- model_diag$calibration
if (!is.null(cal) && nrow(cal)) {
  cal$Outcome <- lab(cal$outcome)
  ggplot(cal, aes(mean_pred, mean_obs, colour = Outcome)) +
    geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey60") +
    geom_point(size = 1.1, alpha = 0.8) +
    geom_line(linewidth = 0.4, alpha = 0.7) +
    facet_wrap(~country, scales = "free") +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = "Mean predicted", y = "Observed", colour = NULL) +
    theme(legend.position = "bottom", legend.text = element_text(size = 6)) +
    guides(colour = guide_legend(nrow = 3))
} else ggplot() + theme_void()
```

# District-level accuracy

```{r admin2err}
#| results: asis
e <- benchmarks_data$admin2_error
if (is.null(e) || !nrow(e)) empty("District error table not built.") else {
  t <- e |>
    transmute(Country = country, Outcome = lab(outcome), Districts = n_admin2,
              `MAE (pp)` = num(mae_pp), `RMSE (pp)` = num(rmse_pp),
              `Correlation r` = num(pearson_r),
              `Bias (pp)` = num(mean_bias_pp)) |>
    arrange(Country, Outcome)
  print(kable(t, align = "llrrrrr", booktabs = TRUE, longtable = TRUE,
              caption = "Predicted vs surveyed district prevalence, within country.") |>
          kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 7.5))
}
```

Correlation is the measure that matters most here: it says whether the model
**ranks** districts correctly, which is what targeting depends on. A negative
value means the ordering is worse than useless for that combination.

::: {.callout-important}
## This table is not the estimator the briefs use

These are the *individual-level SuperLearner* diagnostics, kept because they
are the pipeline's long-running benchmark. The dashboard and the country briefs
show the area-level recipe model instead, and the two can disagree sharply on
the same cell.

Sierra Leone child iron is the clearest case: it appears above with a
correlation of 0.60, while the recipe model returns an identical estimate for
all 14 districts and so has no district-level ranking at all. Both numbers are
correct about their own model. **Where the two conflict, the briefs' figure is
the one in use** — see the final section for the full list of combinations
where the deployed estimator carries no district signal.
:::

# Method comparison

```{r bench}
#| results: asis
ac <- benchmarks_data$area_comparison
if (is.null(ac) || !nrow(ac)) empty("Area comparison not built.") else {
  t <- ac |>
    transmute(Country = country, Outcome = lab(outcome), Approach = approach,
              Districts = n_admin2, `MAE (pp)` = num(mae_pp),
              `Correlation r` = num(pearson_r)) |>
    arrange(Country, Outcome, Approach)
  print(kable(head(t, 60), align = "lllrrr", booktabs = TRUE, longtable = TRUE,
              caption = "Individual-level versus area-level modelling, within country (first 60 rows).") |>
          kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 7))
}
```

The area-level approach is the primary one. Modelling district prevalence
directly, rather than modelling individuals and averaging up, is the single
largest gain in this project: almost every predictor is resolved at district
level, so individual rows are pseudo-replicated and the honest district-ranking
signal looks close to chance.

# Transportability

```{r transport-tab}
#| results: asis
at <- benchmarks_data$area_transport
if (is.null(at) || !nrow(at)) empty("Area-transport metrics not built.") else {
  t <- at |>
    transmute(Outcome = lab(outcome), `Held out` = held_out,
              `Train n` = n_train, `Test n` = n_test,
              `Predictors kept` = n_selected,
              `Correlation r` = num(pearson_r)) |>
    arrange(Outcome, `Held out`)
  print(kable(t, align = "llrrrr", booktabs = TRUE, longtable = TRUE,
              caption = "Leave-one-country-out: train on three, predict the fourth.") |>
          kable_styling(latex_options = c("hold_position", "repeat_header"), font_size = 7.5))
}
```

```{r transport-scatter}
#| fig-cap: "District-level transport. Each point is a district in the held-out country, its survey value against the prediction from a model that never saw that country."
d <- transport_cal$adm2_area
if (!is.null(d) && nrow(d)) {
  d$Outcome <- lab(d$outcome)
  ggplot(d, aes(true_prev, pred_prev, colour = Outcome)) +
    geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey55") +
    geom_point(size = 0.8, alpha = 0.5) +
    facet_wrap(~country, scales = "free") +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = "Survey-measured", y = "Predicted", colour = NULL) +
    theme(legend.position = "bottom", legend.text = element_text(size = 6)) +
    guides(colour = guide_legend(nrow = 3))
} else ggplot() + theme_void()
```

**This is the weakest link and the most important one to read honestly.**
Applying the model to a country with no biomarker survey is the use case that
would let this scale, and it is where performance degrades most. Levels in
particular do not transport — biomarker values are not comparable across
surveys and laboratories, so a transported prediction should be scored on
ranking and anchored to a known national figure.

# What drives the predictions

```{r ablation}
#| fig-cap: "Predictor families by importance. Each family is permuted in turn; the bar is the resulting drop in ROC-AUC. Larger means the model relies on it more."
#| fig-height: 5.4
ab <- importance_data$ablation
if (!is.null(ab) && nrow(ab)) {
  base <- ab[ab$domain_removed == "none", c("country", "outcome", "auc")]
  names(base)[3] <- "auc_base"
  d <- merge(ab[ab$domain_removed != "none", ], base, by = c("country", "outcome"))
  d$drop <- d$auc_base - d$auc
  d$Outcome <- lab(d$outcome)
  ggplot(d, aes(domain_removed, Outcome, fill = drop)) +
    geom_tile(colour = "white", linewidth = 0.4) +
    facet_wrap(~country, scales = "free_y") +
    scale_fill_gradientn(colours = c("#f7f7f7", "#fdae61", "#d7191c"), name = "AUC drop") +
    labs(x = NULL, y = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
          axis.text.y = element_text(size = 6), panel.grid = element_blank())
} else ggplot() + theme_void()
```

Permutation importance understates predictors that overlap: when two families
carry the same information, removing either one alone changes little even
though the information matters. Read near-zero cells as "redundant or
uninformative", not as "irrelevant".

# Resolution: district versus survey cluster

```{r resolution}
#| results: asis
cr <- cluster_res$comparison
if (is.null(cr) || !nrow(cr)) empty("Cluster-resolution comparison not built.") else {
  t <- cr |>
    transmute(Country = country, Outcome = lab(outcome),
              `r (district)` = num(r_admin2), `r (cluster)` = num(r_cluster),
              `MAE district` = num(mae_admin2), `MAE cluster` = num(mae_cluster)) |>
    arrange(Country, Outcome)
  print(kable(t, align = "llrrrr", booktabs = TRUE,
              caption = "Modelling at district resolution versus at survey-cluster GPS buffers.") |>
          kable_styling(latex_options = "hold_position", font_size = 7.5))
}
```

# Known limitations

`r GENERAL_CAVEAT`

```{r nosignal-tech}
#| results: asis
PD <- pred_model_data(DEFAULT_PRED_MODEL)
ns <- do.call(rbind, lapply(names(meta$countries), function(ck) {
  cl <- unname(meta$countries[[ck]])
  ocs <- intersect(names(meta$outcome_labels), unique(PD$outcome[PD$country == cl]))
  do.call(rbind, lapply(ocs, function(oc) {
    v <- PD$pred_prev[PD$country == cl & PD$outcome == oc]
    v <- v[is.finite(v)]
    if (!length(v) || has_district_signal(v)) return(NULL)
    data.frame(Country = cl, Outcome = lab(oc), Estimate = pct(stats::median(v)),
               Spread = sprintf("%.2f pp", 100 * diff(range(v))))
  }))
}))
if (!is.null(ns) && nrow(ns)) {
  cat(sprintf(paste("\n**No district-level signal in %d combinations.** The fit",
    "reduces to an intercept and every district returns the same estimate.",
    "These are legitimate null results, but any subnational figure for them is",
    "the national figure repeated.\n\n"), nrow(ns)))
  print(kable(ns, align = "llrr", booktabs = TRUE,
              caption = "Combinations with no usable district variation") |>
          kable_styling(latex_options = "hold_position", font_size = 8))
}
```

Further caveats: surveys span 2013--2018 and different laboratories, so
biomarker values are not perfectly comparable between countries. Each country
contributes a single timepoint, so predictions assume the proxy--outcome
relationship is stable over time. Population denominators carry their own
modelling uncertainty. And Admin-2 sits below the resolution the surveys were
designed to represent — many districts contain a single survey cluster, so the
direct estimates the models are fitted against are themselves imprecise.

The interactive version, including per-district explanations, is at
<https://amertens.shinyapps.io/micronutrient-burden/>.
