---
title: "National Micronutrient Prevalence Prediction"
subtitle: "BRINDA & VMNIS SuperLearner Analysis"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: cosmo
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(dplyr)
library(ggplot2)
library(knitr)
library(patchwork)

TARGETS_STORE <- here::here("national_prediction", "_targets")
```

## Overview

This report summarizes national-level micronutrient deficiency prevalence
predictions using SuperLearner ensemble models fitted to two data sources:

- **BRINDA**: Biomarkers Reflecting Inflammation and Nutritional Determinants
  of Anemia — prevalence of iron, vitamin A (RBP), and zinc deficiency.
- **VMNIS**: WHO Vitamin and Mineral Nutrition Information System — mean levels
  and prevalence of deficiency for zinc, vitamin B12, retinol, vitamin D,
  folate, and haemoglobin.

National-level proxy predictors (DHS, IHME, and other indicators) are used as
covariates. Models are fitted using `sl3` SuperLearner with leave-country-out
cross-validation (2 folds).

## Data Availability

### VMNIS coverage across LMICs

```{r vmnis-heatmap, fig.width=10, fig.height=8}
tar_read(fig_heatmap_pw, store = TARGETS_STORE)
```

```{r vmnis-heatmap-alt, fig.width=10, fig.height=8}
tar_read(fig_heatmap_pw_alt, store = TARGETS_STORE)
```

### All populations

```{r vmnis-heatmap-all, fig.width=14, fig.height=10}
tar_read(fig_heatmap_all, store = TARGETS_STORE)
```

## Model Performance

### Summary table

```{r perf-table}
all_perf <- tar_read(all_perf, store = TARGETS_STORE)

all_perf %>%
  select(source, outcome, R2_vs_mean, correlation, rmse, mae, mean_bias) %>%
  mutate(across(where(is.numeric), ~round(.x, 3))) %>%
  kable(
    col.names = c("Source", "Outcome", "R² (vs. mean)", "Correlation",
                  "RMSE", "MAE", "Bias"),
    caption = "Cross-validated SuperLearner performance across all outcomes"
  )
```

### Performance comparison

```{r perf-plot, fig.width=12, fig.height=8}
tar_read(fig_performance, store = TARGETS_STORE)
```

### Notes on learner stack

Outcomes using the **simple** stack (mean + LASSO only) due to fitting issues
with the full stack:

- Vitamin B12 (mean and prevalence)
- Vitamin D (mean and prevalence)

All other outcomes use the **full** stack (mean, LASSO, ridge, ranger, XGBoost,
plus screened pipelines with HAL, GAM, Bayesian GLM, and polspline).

## Observed vs. Predicted

### BRINDA outcomes

```{r obs-pred-brinda, fig.width=10, fig.height=8}
tar_read(fig_obs_vs_pred_brinda, store = TARGETS_STORE)
```

### VMNIS outcomes

```{r obs-pred-vmnis, fig.width=12, fig.height=8}
tar_read(fig_obs_vs_pred_vmnis, store = TARGETS_STORE)
```

## Variable Importance

### BRINDA — Iron (Ferritin)

```{r varimp-brinda-ferr, fig.width=8, fig.height=6}
fig <- tar_read(fig_varimp_brinda_ferr, store = TARGETS_STORE)
if (!is.null(fig)) fig else cat("Variable importance not available")
```

### BRINDA — Zinc

```{r varimp-brinda-zinc, fig.width=8, fig.height=6}
fig <- tar_read(fig_varimp_brinda_zinc, store = TARGETS_STORE)
if (!is.null(fig)) fig else cat("Variable importance not available")
```

### VMNIS — Zinc

```{r varimp-vmnis-zinc, fig.width=8, fig.height=6}
fig <- tar_read(fig_varimp_vmnis_zinc, store = TARGETS_STORE)
if (!is.null(fig)) fig else cat("Variable importance not available")
```

### VMNIS — Retinol

```{r varimp-vmnis-retinol, fig.width=8, fig.height=6}
fig <- tar_read(fig_varimp_vmnis_retinol, store = TARGETS_STORE)
if (!is.null(fig)) fig else cat("Variable importance not available")
```

## Session Info

```{r session-info}
sessionInfo()
```
