---
title: "Admin-2 Prevalence Predictions: Dataset Description and Comparative Analysis"
subtitle: "Individual-Level vs. Area-Level SuperLearner Predictions"
author: "Micronutrient Prediction Project"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    theme: cosmo
    self-contained: true
execute:
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(scales)

theme_set(theme_minimal(base_size = 12))

TARGETS_STORE <- here("_targets_full")

load_target <- function(name) {
  tryCatch(targets::tar_read_raw(name, store = TARGETS_STORE),
           error = function(e) NULL)
}

countries <- c("gambia", "ghana", "sierraleone", "malawi")
outcomes  <- c("child_vitA", "women_vitA", "child_iron", "women_iron",
               "women_folate", "women_b12", "child_zinc", "women_zinc")

country_labels <- c(gambia = "Gambia", ghana = "Ghana",
                     sierraleone = "Sierra Leone", malawi = "Malawi")
outcome_labels <- c(
  child_vitA = "Vitamin A (children)", women_vitA = "Vitamin A (women)",
  child_iron = "Iron (children)", women_iron = "Iron (women)",
  women_folate = "Folate (women)", women_b12 = "B12 (women)",
  child_zinc = "Zinc (children)", women_zinc = "Zinc (women)")
```

# Overview

This document describes the Admin-2 (district-level) micronutrient deficiency prevalence prediction datasets produced by the proxy-only SuperLearner pipeline. These datasets are intended for external sharing and programmatic use.

**Two prediction approaches are compared:**

1. **Individual-level SL**: A SuperLearner ensemble trained on individual biomarker data with ~900-1,100 proxy predictors, then aggregated to Admin-2 by taking the survey-weighted mean of individual predictions.

2. **Area-level SL**: A SuperLearner trained directly on Admin-2 survey-weighted prevalence as the outcome, using GEE satellite raster zonal means as predictors. Two loss functions are tested:
   - **MSE (continuous)**: Standard squared-error loss with post-hoc clamping to [0,1]
   - **NLL (binomial)**: Negative log-likelihood / cross-entropy, which naturally handles bounded [0,1] prevalence (inspired by beta regression theory)


# Data Sources

## Survey data

| Country | Survey | Year | Children (n) | Women (n) | Clusters | Admin-2 units |
|---------|--------|------|-------------|----------|---------|--------------|
| Gambia | GMS | 2018 | ~1,020 | ~1,413 | ~100 | ~40 |
| Ghana | DHS+ | 2017 | ~1,165 | ~1,005 | ~400 | ~216 |
| Sierra Leone | MNS | 2013 | ~486 | ~774 | ~60 | ~14 |
| Malawi | MNS | 2015-16 | ~1,100 | ~800 | ~87 | ~87 |

## Predictor data (area-level models)

Area-level models use **GEE satellite raster zonal means** extracted for each Admin-2 polygon:

- NDVI, EVI (vegetation indices)
- Land surface temperature (day/night)
- FLDAS climate variables (soil moisture, temperature, precipitation)
- TRMM/GPM precipitation
- ESA WorldCereal crop classification
- SoilGrids soil properties (pH, carbon, nitrogen, iron, zinc, etc.)
- GHSL settlement classification
- CCNL nighttime lights
- GPW population demographics

Total: ~150-308 GEE variables per country after extraction.


# Admin-2 Coverage

```{r coverage-table}
coverage_rows <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    svy <- load_target(paste0("svy_admin2_", ctry, "_", oc))
    if (is.null(svy) || nrow(svy) == 0) next
    coverage_rows[[paste0(ctry, "_", oc)]] <- data.frame(
      Country = country_labels[ctry],
      Outcome = outcome_labels[oc],
      `Surveyed Admin-2` = nrow(svy),
      `Mean n per Admin-2` = round(mean(svy$n_svy, na.rm = TRUE), 0),
      `Min n` = min(svy$n_svy, na.rm = TRUE),
      `Max n` = max(svy$n_svy, na.rm = TRUE),
      `Mean prev (%)` = round(mean(svy$svy_prev, na.rm = TRUE) * 100, 1),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }
}
if (length(coverage_rows) > 0) {
  cov_df <- bind_rows(coverage_rows)
  kable(cov_df, caption = "Admin-2 survey coverage by country and outcome") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1, valign = "top")
}
```


# Comparative Analysis: Individual-Level vs. Area-Level SL

```{r load-comparison}
area_comp <- load_target("area_comparison_all")
area_loco <- load_target("area_loco_comparison")
```

## Within-Country Comparison

```{r within-country-table}
if (!is.null(area_comp) && nrow(area_comp) > 0) {
  comp_display <- area_comp |>
    mutate(outcome_label = outcome_labels[outcome]) |>
    select(Country = country, Outcome = outcome_label, Approach = approach,
           `N Admin-2` = n_admin2, `MAE (pp)` = mae_pp, `RMSE (pp)` = rmse_pp,
           `Pearson r` = pearson_r, `Bias (pp)` = mean_bias_pp)

  kable(comp_display, caption = "Within-country Admin-2 prediction accuracy: three approaches compared") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10) |>
    collapse_rows(columns = 1:2, valign = "top")
} else {
  cat("Area comparison targets not yet available. Run the pipeline first.\n")
}
```

```{r within-country-heatmap, fig.height=8, fig.width=10}
if (!is.null(area_comp) && nrow(area_comp) > 0) {
  comp_plot <- area_comp |>
    mutate(
      outcome_label = outcome_labels[outcome],
      ctry_outcome = paste0(country, " - ", outcome_label)
    )

  ggplot(comp_plot, aes(x = approach, y = ctry_outcome, fill = mae_pp)) +
    geom_tile(colour = "white", linewidth = 0.8) +
    geom_text(aes(label = sprintf("%.1f", mae_pp)), size = 3.5, fontface = "bold") +
    scale_fill_viridis_c(name = "MAE (pp)", option = "mako", direction = -1) +
    labs(x = NULL, y = NULL,
         title = "Admin-2 Prediction Accuracy: MAE (percentage points)",
         subtitle = "Lower = better | Three approaches compared") +
    theme_minimal(base_size = 11) +
    theme(axis.text.x = element_text(angle = 25, hjust = 1),
          panel.grid = element_blank())
}
```

## Cross-Country (LOCO) Area-Level Comparison

```{r loco-area-table}
if (!is.null(area_loco) && nrow(area_loco) > 0) {
  loco_display <- area_loco |>
    mutate(
      outcome_label = outcome_labels[outcome],
      loss = ifelse(model_type == "continuous", "MSE", "NLL")
    ) |>
    select(`Held out` = held_out, Outcome = outcome_label, Loss = loss,
           `N train` = n_train, `N test` = n_test,
           `Obs prev (%)` = obs_prev, `Pred prev (%)` = pred_prev,
           `MAE (pp)` = mae_pp, `Pearson r` = pearson_r) |>
    mutate(
      `Obs prev (%)` = round(`Obs prev (%)` * 100, 1),
      `Pred prev (%)` = round(`Pred prev (%)` * 100, 1)
    )

  kable(loco_display,
        caption = "Area-level LOCO: trained on 3 countries, tested on held-out country") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10) |>
    collapse_rows(columns = 1:2, valign = "top")
} else {
  cat("Area-level LOCO targets not yet available.\n")
}
```


# Data Dictionary

The shared Admin-2 prediction CSV files contain the following columns:

| Column | Type | Description |
|--------|------|-------------|
| `Admin1` | character | Admin-1 (region/province) name |
| `Admin2` | character | Admin-2 (district) name |
| `country` | character | Country name |
| `outcome` | character | Outcome tag (e.g., child_vitA, women_iron) |
| `svy_prev` | numeric | Survey-weighted prevalence (0-1 scale) |
| `svy_prev_se` | numeric | Standard error of survey prevalence |
| `n_svy` | integer | Number of survey respondents in this Admin-2 |
| `sl_prev` | numeric | Individual-level SL predicted prevalence (aggregated) |
| `area_mse_prev` | numeric | Area-level SL predicted prevalence (MSE loss) |
| `area_nll_prev` | numeric | Area-level SL predicted prevalence (NLL/binomial loss) |
| `has_survey` | logical | Whether this Admin-2 has direct survey data |

**Note:** For Admin-2 units without survey data (`has_survey = FALSE`), `svy_prev` and `sl_prev` are `NA`. Area-level predictions (`area_mse_prev`, `area_nll_prev`) are available for all Admin-2 units with GEE raster coverage.


# Quality Flags and Limitations

**Reliable predictions** (higher confidence):

- Admin-2 units with `n_svy >= 30` and `svy_prev_se < 0.05`
- Individual-level SL predictions for outcomes with within-country AUC > 0.70
- Area-level predictions where LOO MAE < 5 percentage points

**Uncertain predictions** (use with caution):

- Admin-2 units with `n_svy < 15` --- survey estimates are unstable
- Predictions for outcomes with AUC near 0.50 (no individual-level discrimination)
- Area-level extrapolations to unsurveyed Admin-2 units (no direct validation possible)
- Sierra Leone predictions (fewer clusters, older survey)
- Malawi women_folate (continuous model only, 1 binary event)

**Caveats for external users:**

1. Predictions are cross-validated but not externally validated against independent surveys
2. Temporal extrapolation: surveys are from 2013-2018; predictions assume stable proxy-outcome relationships
3. Admin-2 boundaries follow GADM v4.1; name matching may differ from national definitions
4. Survey-weighted prevalences use design-based inference; SL predictions are model-based
5. The pipeline uses proxy-only predictors --- no biomarker data from the same survey
