---
title: "What the Proxy Predictor Set Can Deliver"
subtitle: "Out-of-sample accuracy for micronutrient deficiency prediction across four African countries"
author: "Andrew N. Mertens, University of California, Berkeley"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    fig-width: 9
    fig-align: center
    embed-resources: true
    code-fold: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(here); library(dplyr); library(readr); library(knitr)
source(here("R", "plot_forest.R"))

nat  <- read_csv(here("results","tables","national_estimates_all.csv"), show_col_types = FALSE)
dec  <- read_csv(here("results","tables","corrected","decision_value.csv"), show_col_types = FALSE)
tr   <- read_csv(here("results","tables","transportability_area_loco_metrics.csv"), show_col_types = FALSE)
wi   <- read_csv(here("results","tables","corrected","area_recipe_within.csv"), show_col_types = FALSE)
cmp  <- read_csv(here("results","tables","area_comparison_all.csv"), show_col_types = FALSE)

n_inside <- sum(nat$pred_prev >= nat$obs_lo & nat$pred_prev <= nat$obs_hi)
mae_nat  <- mean(abs(nat$pred_prev - nat$obs_prev)) * 100
lift_max <- max(dec$lift_vs_no_targeting)
n_lift   <- sum(dec$lift_vs_no_targeting > 1)
tr_ok    <- sum(tr$pearson_perm_p < 0.05, na.rm = TRUE)
tr_n     <- sum(!is.na(tr$pearson_r))
```

## Summary

Four national biomarker surveys, **294 harmonised proxy predictors**, **554 districts**,
**24 country–outcome combinations**. Every interval below is a bootstrap or
design-based confidence interval on out-of-sample performance — nothing here is
an in-sample fit.

| | |
|:--|:--|
| **`r n_inside`/`r nrow(nat)`** | predicted national prevalences fall inside the survey's own 95% CI |
| **`r sprintf("%.2f pp", mae_nat)`** | mean absolute error on national prevalence |
| **`r sprintf("%.2f×", lift_max)`** | best targeting lift; above 1.0× in `r n_lift` of `r nrow(dec)` combinations |
| **`r tr_ok`/`r tr_n`** | district rankings transport to a country the model never saw |

## National prevalence recovery

The strongest result in the programme, and the one that has survived every
change to the pipeline.

```{r fig-national}
#| fig-height: 8
forest_national()
```

## Targeting value

If a programme could reach only the top 20% of districts, how many more true
deficiency cases would the model's ranking find than no targeting?

```{r fig-targeting}
#| fig-height: 8
forest_targeting()
```

## Transport to an unseen country

The hardest test: train on three countries, predict district prevalence in a
fourth using no biomarker data from it.

```{r fig-transport}
#| fig-height: 6
forest_transport()
```

## Within-country district ranking

The most demanding resolution, and the results are correspondingly mixed.

```{r fig-within}
#| fig-height: 8
forest_within()
```

## The design choice that pays

The model supplies the *ranking*; the design-based survey supplies the *level*.
Combining them beats either alone across all 24 combinations.

```{r tbl-anchor}
kable(table_anchor_arms(), align = "lrrrr")
```

## What this brief does not claim

These are the strongest results in the programme, selected to show where the
proxy predictor set carries usable information. Four limits belong beside them.

**National prevalence is recovered far better than district prevalence.** The
national result is the robust one. District maps should be read as rankings
rather than as absolute prevalence.

**Absolute levels do not transport across borders even where rankings do**,
with national bias reaching 42 pp in the worst case. This is why the anchoring
step above matters: it is not a refinement, it is what makes a transported
district map interpretable at all.

**A covariate-free spatial smoother matches or beats the covariate models at
district level.** Across 24 combinations the smoother reaches a mean rank
correlation of 0.30 against 0.19 for the best honest covariate model, so
district-scale gains come substantially from geography rather than from the
proxy variables themselves. The covariate models pull clearly ahead only at
admin-1 (0.44 vs 0.21).

**No individual predictor is demonstrably associated with district prevalence
once spatial structure is accounted for.** Testing all 294 predictors in all 24
combinations — 7,104 tests — under a null that shuffles the outcome within
regions, **no predictor survives false-discovery-rate control in any
combination**, and penalised regression at a cross-validated penalty retains a
median of zero predictors (13 of 18 combinations retain none). Between-region
association exists but is confounded with geography. This is the same finding
as the spatial smoother matching the full covariate set, seen from the
association side, and it is why the district-level results in this brief should
be read as *rankings anchored to survey totals* rather than as evidence that
particular proxies drive deficiency.

**Covariate selection inflates apparent accuracy.** Screening predictors on all
data before cross-validating adds a mean of **+0.18** to the rank correlation,
positive in 20 of 20 measurable cells and up to +0.51 — enough to turn cells
with zero real skill into respectable-looking ones. All figures in this brief
screen inside the cross-validation fold.

```{r tbl-approaches}
cmp %>%
  filter(!is.na(pearson_r)) %>%
  group_by(Approach = approach) %>%
  summarise(Cells = n(), `Mean r` = round(mean(pearson_r), 3),
            `MAE (pp)` = round(mean(mae_pp), 2), .groups = "drop") %>%
  arrange(desc(`Mean r`)) %>%
  kable(align = "lrrr",
        caption = paste("All approaches. MRP and the SL rows are scored",
                        "IN-SAMPLE and are not comparable with the",
                        "out-of-fold (OOF) rows; the labels say which."))
```

---

*Proxy modelling for vitamin and mineral deficiencies — Ghana, Sierra Leone,
Malawi, The Gambia. With funding support from the Gates Foundation. All
intervals are out-of-sample: 1,000-replicate bootstrap for correlations,
design-based for survey prevalence.*
