---
title: "Supplementary Appendix: Subnational micronutrient deficiency prediction in sub-Saharan Africa from proxy indicators"
author: "Andrew N. Mertens"
date: today
format:
  docx:
    number-sections: true
    tbl-cap-location: top
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(here); library(dplyr); library(readr); library(knitr)
TBL <- here("results", "tables")
rd <- function(...) { p <- file.path(TBL, ...); if (file.exists(p))
  suppressMessages(read_csv(p, show_col_types = FALSE)) else NULL }

# Counts are READ, not typed. The previous appendix hard-coded "152
# common-vocabulary variables" and went stale without anyone noticing, because
# nothing tied the prose to the artefact it described.
harm <- suppressMessages(tryCatch(
  read_csv(here("data","covariates","harmonized","coverage.csv"), show_col_types = FALSE),
  error = function(e) NULL))
n_shared <- if (!is.null(harm)) sum(harm$in_all, na.rm = TRUE) else NA_integer_
n_canon  <- if (!is.null(harm)) nrow(harm) else NA_integer_
src <- suppressMessages(tryCatch(
  read_csv(here("metadata","covariates","source_registry.csv"), show_col_types = FALSE),
  error = function(e) NULL))
n_prov <- if (!is.null(src)) dplyr::n_distinct(src$provider) else NA_integer_
n_dom  <- if (!is.null(src)) dplyr::n_distinct(src$domain) else NA_integer_
cmp <- rd("area_comparison_all.csv")
n_cells <- if (!is.null(cmp)) nrow(distinct(cmp, country, outcome)) else NA_integer_
```

This appendix documents, in full, the methods summarised at a high level in the
main text. It covers the data sources, the data cleaning and linkage pipeline,
the harmonisation of outcome definitions across surveys, the modelling methods,
the procedures used to assess predictive accuracy, and the sensitivity analyses.
It reflects the analysis as currently implemented in the project's reproducible
`targets` pipeline. The most granular file-by-file data lineage is maintained
alongside this appendix in `docs/data_sources_appendix.qmd`.

# S1. Data sources

## S1.1 Biomarker (outcome) surveys

Outcomes come from four nationally representative biomarker surveys, one per
country, each conducted in a different year and with a partly different
biomarker panel: **The Gambia, 2018** (Gambia National Micronutrient Survey);
**Ghana, 2017** (Ghana Micronutrient Survey); **Sierra Leone, 2013** (Sierra
Leone Micronutrient Survey); and **Malawi, 2015–16** (Malawi Micronutrient
Survey).

A fifth country, Tanzania (2009–10 TDHS, vitamin A only), was added and then
withdrawn: a project collaborator identified its dried-blood-spot RBP
measurements as unreliable, consistent with a unit-label error in the survey's
own biomarker documentation and a 77–92 percentage-point national-level bias
under leave-one-country-out evaluation. See `archive/src/Tanzania/` and the
archive manifest for the preserved scripts and the full rationale.

Within each survey, biomarkers were extracted for women of reproductive age
(15–49 years) and children 6–59 months where measured. Each survey is
nationally representative but modest in size (roughly 1,000–1,400 individuals
per outcome), is georeferenced at the survey-cluster level, and carries survey
design information.

**Populations not analysed.** The Malawi survey is structurally different from
the other three: it sampled four populations rather than two — preschool
children (n = 1,233), **school-age children 5–14 years (n = 800)**, women
(n = 838) and **men (n = 228)**. Only preschool children and women enter the
analysis. Malawi's school-age children already carry derived `vitA_def` and
`iron_def` flags with RBP, ferritin, CRP and AGP on 758 of 800, and are
therefore available at no additional data cost; they are not analysed here
because they cannot enter any cross-country model and because 5–14 years spans
puberty, so they require a separate outcome rather than pooling with 6–59
months.

**Biomarkers not analysed.** Soluble transferrin receptor is present in all
four surveys and is unused. It is substantially less inflammation-sensitive
than ferritin and would support body-iron-stores estimation, sTfR-based iron
deficiency, and iron-deficiency anaemia. Urinary iodine is available in The
Gambia (n = 1,383) and Sierra Leone (n = 1,422); WHO treats urinary iodine as a
population-level rather than an individual-level indicator, which suits an
area-level model but not a person-level one. Malawi additionally measured
red-blood-cell folate (n = 797) alongside serum folate.

## S1.2 Predictor domains

```{r}
#| results: asis
# Stated as a property of the REGISTRY, which is what these counts describe.
# They are not a count of what reaches the harmonised set: the predictor
# inventory is keyed on pre-harmonisation names, so no reliable cross-walk from
# a canonical predictor back to a registry row exists, and quoting a registry
# count as though it described the 294 would be an unsupported claim.
if (!is.na(n_prov)) cat(sprintf(
  paste("The source registry (`metadata/covariates/source_registry.csv`)",
        "documents %d providers across %d thematic domains.\n"),
  n_prov, n_dom))
```

With one exception (the survey biomarker matrix, excluded from the deployable
models) all are publicly available or accessible through common interfaces.
Domains comprise: the survey biomarker matrix (`gw_`, excluded from proxy-only
models); DHS admin-2 indicator aggregates; MICS aggregates; IHME GBD modelled
estimates; LSMS aggregates (Ghana only); Malaria Atlas Project rasters
(epidemiology, interventions, and the inherited blood disorders that confound
ferritin-based iron assessment); Google Earth Engine rasters, extracted both as
cluster-buffer summaries at 10/25/50 km and as admin-2 polygon zonal means; WFP
food prices; FluNet influenza surveillance; food security (HungerMap/HFID and
IPC / Cadre Harmonisé phase classifications); SoilGrids and iSDAsoil soil
chemistry; MapSPAM crop shares; and ESPEN helminth prevalence.

**Harmonisation.** Earlier versions of this analysis used a common-vocabulary
set of 152 area-level variables selected by name intersection. That set has been
replaced by an explicitly harmonised one, built in four documented stages with
every decision recorded in `metadata/covariates/`: raw provider columns are
mapped onto canonical names; near-duplicate variants of one construct (month
climatologies, buffer radii, soil depth summaries, year stamps) are collapsed;
values are converted to canonical units; and variables that cannot be reconciled
across countries are excluded with a stated reason.

The exclusions are substantive rather than cosmetic. Of Ghana's 755 raw admin-2
columns, 301 were dropped because they are not quantities: cross-band summaries
averaged over non-commensurable bands, categorical land-cover class codes whose
zonal mean is meaningless, quality-assurance bands, and population columns in
which the provider replicated five-year epochs across intervening years. A
further 40 were excluded for cross-country irreconcilability, including twelve
whose medians ranged from 742 to 23,784 across countries with maxima to
$2.45\times10^7$ — a defective export rather than a scale factor, and one that
had been inside the previous 152-variable set and was therefore manufacturing a
country effect — and seven outcome-adjacent DHS indicators excluded as leakage
(anaemia is the downstream consequence of iron deficiency; nutrient-specific
supplementation is targeted *by* measured deficiency).

```{r}
#| results: asis
if (!is.na(n_shared)) cat(sprintf(
"The resulting set comprises **%d predictors** shared by all four analysed countries, from %d canonical variables defined overall. The intersection step is not the binding constraint: %d of %d canonical variables are common to all four.\n",
n_shared, n_canon, n_shared, n_canon))
```

District centroid coordinates (`geo_lon`, `geo_lat`) are included as ordinary
covariates. They had previously been excluded by name, which meant no covariate
model could use the one signal that demonstrably carries district-level
information.

# S2. Data cleaning and pipeline architecture

Cleaning and assembly run in two stages. Stage 1 comprises per-country cleaning
scripts that read each raw survey and external source, standardise names and
types, derive inflammation-adjusted biomarkers and binary deficiency flags, and
write a per-country merged analytic table. Stage 2 is the reproducible `targets`
pipeline that reads those tables and runs all modelling, evaluation and
reporting.

Key conventions: all `gw_` survey items are removed from the proxy-only pool by
prefix matching; high-cardinality categoricals are dropped or collapsed;
geometry and raw temporal columns are removed after use in linkage; and missing
data are retained as `NA`, with imputation performed downstream inside model
fitting and, in the corrected methods, strictly within each training fold.

**Untracked-input hazards.** Two inputs are read directly from disk by
functions inside the pipeline and are therefore invisible to the dependency
graph: the DHS clean admin-2 files and the external-predictor cache. Both are
now covered by content stamps (`dhs_clean_stamp()`, `ext_cache_stamp()`) that
force a rebuild when the file changes. The second was added after the external
caches were refreshed without any downstream target rebuilding, which left one
country serving 0 of 57 available `map2_` columns and 0 of 13 `wfp_` columns.

# S3. Geographic and temporal harmonisation, and record linkage

## S3.1 Geographic units and cluster-to-area assignment

Administrative polygons are taken from GADM (admin-1 and admin-2; admin-3 for
Sierra Leone in a resolution sensitivity analysis). Each survey cluster is
assigned to its polygons by point-in-polygon location of the cluster GPS
coordinate.

Cluster coordinates displaced into inland water by the DHS/MICS confidentiality
procedure are reassigned to the nearest land district rather than dropped: every
measured displacement fell inside the documented bounds (up to 2 km urban, 5 km
rural, 10 km for a random 1% of rural clusters), so dropping them would discard
real observations. Water polygons are excluded from the prediction side, since a
lake is not a populated prediction unit.

## S3.2 Linkage of predictors to outcomes

Cluster-level sources (earth-observation buffer summaries, MAP rasters) are
extracted at each cluster's GPS point and joined on cluster identifier.
Area-level sources are joined by administrative unit, with fuzzy string matching
where survey and external names are spelled inconsistently and unmatched units
left missing rather than guessed.

**The join key is the (Admin1, Admin2) pair, not the admin-2 name.** GADM
admin-2 names are not unique within a country — Malawi has 256 polygons under
243 names, with four pairs of genuinely different districts sharing a name. A
name-only join silently averages or discards one member of each pair. Nine
consumers were found still joining on the name after the initial migration; all
now resolve the key per join and fall back to the name only when a side lacks
`Admin1`, warning when they do. The practical consequence had been that Malawi's
area-level rows were scored on 90 districts against 87 surveyed ones, so they
were not comparable with the individual-level rows in the same table.

## S3.3 Temporal alignment

For sources with multiple time points, the version closest to each country's
fieldwork year is selected. Climate variables use the survey year; slowly
changing layers use the most recent global version; multi-band rasters whose
bands are temporal subdivisions are collapsed to the survey-year annual mean.

# S4. Outcome definitions and harmonisation

## S4.1 Deficiency thresholds

Vitamin A (women and children): retinol-binding protein below 0.70 µmol/L.
Iron: serum ferritin below 15 µg/L in women and 12 µg/L in children, with The
Gambia defined on the log-ferritin scale at the equivalent cut-points. Folate
(women): serum folate below 10 nmol/L, with Malawi's native rule harmonised to
the common serum cut-point. Vitamin B12 (women): serum B12 below 148 pmol/L.
Zinc (Malawi only): serum zinc below the IZiNCG cut-point appropriate to time of
day — 66 µg/dL for morning draws and 59 µg/dL for afternoon draws in women. We
verified that the survey's derived flag applies the time-specific rule: it
agrees with it in 100% of cases, and 47% of Malawi's draws were taken in the
afternoon, so applying the morning cut-point uniformly would overstate women's
zinc deficiency by 12.5 percentage points.

Public-health severity tiers (none / mild / moderate / severe) used for the
map's WHO classification are separate from these individual cut-points and
follow the standard prevalence bands.

## S4.2 Inflammation adjustment

Ferritin, and to a lesser extent retinol-binding protein, rises with
inflammation, so unadjusted values understate deficiency in high-infection
settings. Where C-reactive protein and alpha-1-acid glycoprotein were measured,
values are corrected before the cut-point is applied.

**The adjustment is harmonised across countries for vitamin A but not for
iron.** All eight vitamin A country–outcome combinations use a common BRINDA
CRP+AGP adjustment. All eight iron combinations retain each survey's native
adjustment: a log-scale regression correction in The Gambia, Thurnham in Ghana,
a BRINDA regression in Sierra Leone, and a regression adjustment in Malawi. The
raw components required for a uniform adjustment (marker, CRP, AGP) are present
in all four surveys, and harmonising them did not alter cross-country transport
in our testing — the between-country offset lies in raw ferritin rather than in
the adjustment — but the asymmetry is a real limitation and is recorded in
`metadata/covariates/adjustment_inventory.csv`.

# S5. Modelling methods

## S5.1 Estimation strategies compared

Strategies fall into the area-level small-area-estimation family (the primary
analysis) and the individual-level SuperLearner aggregated to admin-2 (a
sensitivity analysis), plus frequentist, Bayesian and transportability-aware
comparators.

## S5.2 Individual-level SuperLearner

Run in one of two modes with a shared implementation, so the learner count is
derived from the library specification at fit time. Fast mode uses 5 base
learners (mean, lasso, elastic net, ranger, XGBoost). Full mode uses 12: the
fast set plus ridge, two further ranger configurations, a deeper XGBoost, and
three BART learners at 50, 100 and 200 trees — BART is the learner that holds up
on the rarest outcomes, where the others collapse toward the mean. A Gaussian
process was evaluated as a 13th learner and dropped: refitting Ghana child iron
with and without it gave an identical binary AUC to 16 decimal places, so it
received effectively zero ensemble weight at $O(n^3)$ cost. For binary outcomes
below 15% prevalence, class-weighted XGBoost and ranger learners are appended at
fit time. Cross-validation folds are formed at the survey-cluster level.

## S5.3 Area-level SuperLearner

The area-level ensemble repeats the fit with the admin-2 polygon as the unit:
one row per district, the target being survey-weighted district prevalence.
Because covariates are resolved at district level, the effective sample size is
the number of polygons (14–90 per country).

Area-level predictions are weighted by inverse sampling variance on the logit
scale, computed by the delta method with Agresti–Coull shrinkage, so that
districts contribute in proportion to their precision rather than equally.

Two uninformed baselines are carried in every comparison — a national-mean null
and a covariate-free spatial smoother on district centroids — so that "better
than the alternatives" can be distinguished from "better than doing nothing".

**Response scale.** The recipe's `scale` field previously had no effect: a
clamped logit transform was applied unconditionally. This matters because the
clamp maps every district observed at exactly 0% prevalence onto a single
extreme value, and eight of the `r ifelse(is.na(n_cells), 24, n_cells)`
country–outcome combinations have more than 40% of districts at exactly zero
(84% for Ghana women's vitamin A, 87% for Malawi's). The transform is now
dispatched on the declared scale. We tested the raw prevalence scale as an
alternative and it did not improve on the logit default for this recipe (S8).

## S5.4 Frequentist and Bayesian comparators

Ordinary least squares on up to eight top-correlated covariates and a
cross-validated elastic net, each evaluated on raw and logit scales, with a
country-mean floor. The Fay–Herriot area-level model uses auxiliary covariates
and known sampling variances; the BYM2 model is a Besag–York–Mollié
disease-mapping model fit by INLA, with adjacency built by Queen's contiguity
and made block-diagonal across training countries for leave-one-country-out
evaluation. Both now reject degenerate design standard errors — any value below
$0.25p(1-p)/n$ is replaced — because most districts contain a single primary
sampling unit, which collapses the design standard error to approximately
$10^{-17}$ and would otherwise be treated as near-perfect precision.

## S5.5 Transportability-aware comparators

A country random-intercept model whose held-out predictions use the marginal
mean; a two-stage model predicting a country mean plus a within-country
deviation; forward selection maximising average leave-one-training-country-out
correlation; a thin-plate spatial GAM on centroids; a group lasso over whole
domains; and a stacked ensemble with country-disjoint folds.

**Transport target.** Because district ranking transports across borders while
the absolute level does not, the transport recipe supports fitting the
within-country z-score of the transformed prevalence, which makes both the
between-country level and spread structurally unlearnable so the model carries
only spatial pattern; both are restored at prediction time from an anchor. See
S8 for the measured effect and for the version of this idea that fails.

## S5.6 Post-hoc calibration, feature augmentation, and intervals

Binary probabilities are recalibrated by Platt scaling, out-of-fold in the
corrected analysis. Outcome-aware feature augmentation adds a small number of
biologically motivated interactions before transport fitting. Uncertainty on
transport correlations is characterised by resampling admin-2 polygons within
each held-out country (1,000 bootstrap replicates).

# S6. The corrected (P1–P8) methods: the analysis of record

Initial benchmarking showed that an analysis at this scale can overstate its own
performance through subtle leakage and optimistic evaluation. The corrected
methods address eight such issues and are the analysis of record.

**P1** Leakage-free fitting: all preprocessing inside training folds only, with
folds formed by survey cluster and, separately, by spatial region. **P2**
Out-of-fold calibration. **P3** Decision-value targeting: reach and lift at the
top 20% of districts. **P4** Split-conformal prediction intervals with a
finite-sample correction. **P5** Sampling-error-aware error decomposition.
**P6** Out-of-support trust flags driving the dashboard's certainty layer.
**P7** Partial-pooling empirical-Bayes estimator. **P8** Join and centring
fixes by construction.

# S7. Assessment of predictive accuracy

Within-country accuracy is estimated by cross-validation on each country's
admin-2 set, with metrics computed on out-of-fold predictions only.
Cross-country transportability is assessed by leave-one-country-out holdout.

Metrics for the area-level task: Pearson and Spearman correlation (spatial
ranking), mean absolute and root-mean-squared error in percentage points, mean
bias, and calibration slope. For binary classifiers: ROC AUC, Brier score and
skill, calibration error, intercept and slope. The corrected analysis adds
sampling-adjusted RMSE, conformal coverage, and decision value.

**Reliability ceiling.** Because a median admin-2 district contributes only
6–36 biomarker measurements, every correlation is reported against an estimated
ceiling $r_{\max}$ derived from a design-based sampling-noise decomposition,
together with the share of that ceiling attained.

# S8. Sensitivity analyses

Individual-level versus area-level fitting; loss function and scale; outcome
harmonisation for folate; admin-3 resolution for Sierra Leone; raw versus
recalibrated classifiers; corrected versus production pipeline; predictor-set
restrictions; and an all-methods comparison on child iron. In addition:

**Selection optimism.** Two otherwise identical pipelines, both cross-validated
by holding out whole regions and differing only in whether the covariate screen
saw the held-out regions, differ by a mean of **+0.182** in rank correlation,
positive in all 20 measurable cells and reaching +0.51. Analyses that screen
predictors on all data before cross-validating should be read with that
magnitude in mind.

**Bivariate association with FDR control.** Every harmonised predictor was
tested against district prevalence in every cell (7,104 tests) under a naive
null and a restricted permutation that shuffles within admin-1 regions. Naive
testing gives 3.4-fold enrichment over chance and 310 survivors of
Benjamini–Hochberg control; the spatially restricted null gives 1.5-fold and
**no predictor survives FDR control in any cell**. Elastic net at a
cross-validated penalty retains a median of zero predictors, with 13 of 18 cells
retaining none.

**Geographic resolution.** The same three arms — null, spatial smoother,
covariate ensemble — were run through identical leave-one-unit-out folds at
admin-2 and admin-1. At admin-2 the covariate ensemble and the spatial smoother
are indistinguishable ($r$ 0.156 vs 0.137) and the smoother has the lower error;
at admin-1 the ensemble roughly triples its correlation (0.437 vs 0.209). A
hybrid selecting the unit a priori from region count reaches $r$ 0.304 across
all four countries. Classifying districts into WHO severity categories for child
vitamin A, the covariate model is correct 33% of the time at admin-2 against 32%
for assigning every district the national prevalence; at admin-1 the comparison
is 56% against 35%.

**Continuous versus binary outcome.** Modelling the district mean of the log
biomarker and converting to prevalence gave 10.80 pp mean absolute error against
9.40 pp for the binary model. An arm converting the *observed* district mean
through the same transformation — performing no prediction — reached 7.49 pp and
$r = 0.672$, beating every other arm including the null, which locates the
deficit in predicting the mean rather than in the conversion. Below 5%
prevalence the continuous route reduced error from 4.61 to 2.99 pp, though that
average is dominated by one cell in which the binary model produced a 7.67 pp
error on a 0.61% outcome.

**Transport target.** Fitting each district's deviation from its own country's
national prevalence does not help (rank correlation 0.201 to 0.140), and an arm
predicting the national figure for every district with no covariates achieved
lower error than any covariate arm. Removing the between-country *spread* as
well — fitting the within-country z-score — raises mean Spearman from 0.146 to
0.199, better in 3 of 4 outcomes. The accompanying error and bias gains derive
mostly from the anchor rather than from the model and should not be quoted
together with the rank gain.

**Changes that did not replicate.** Fitting on the raw rather than clamped logit
scale gave −0.020 mean Spearman, and adding centroid coordinates to the
predictor set gave 0.000, against +0.048 and +0.063 reported in earlier
exploratory work on a smaller covariate set. Both are reported as negative
results and both capabilities are retained in the code.

# S9. Reproducibility

The analysis is a single dependency-aware `targets` pipeline that builds the
data tables, fits every model, runs the corrected methods, computes all metrics,
and writes the result tables consumed by the manuscript, the appendix and the
dashboard. Corrected outputs are written to `results/tables/corrected/`.

Configuration settings that `targets` cannot observe — the covariate
vocabulary, benchmarking mode, and validation switches — are applied through
invalidation traps in `scripts/_pipeline_setup.R`, shared by every runner. A
runner that builds only a subset of the graph passes that subset as a scope, so
traps are applied only to targets the run will rebuild; applying them more
broadly deletes targets that the run then never restores.

The 2026-08 methods-review audit that produced the reliability-ceiling
estimator, the uninformed baselines and the pre-resampling variable-selection
correction is documented in `docs/parsimony_findings.md`, with its scripts and
output archived under `archive/sandbox_parsimony/`.
