---
title: "What geospatial covariates can and cannot do for sub-national micronutrient estimation: a protocol-first re-analysis of four national biomarker surveys"
shorttitle: "Geospatial covariates and sub-national micronutrient estimation"
author:
  - name: Andrew N. Mertens
    affiliation: University of California, Davis, Institute for Global Nutrition
    corresponding: true
    email: anmertens@ucdavis.edu
date: today
format:
  docx:
    number-sections: true
    fig-cap-location: bottom
    tbl-cap-location: top
    fig-width: 6.5
    fig-height: 4.5
    fig-dpi: 300
execute:
  echo: false
  warning: false
  message: false
keywords:
  - micronutrient deficiency
  - small area estimation
  - DHS
  - transportability
  - evaluation protocol
  - vitamin A
  - iron deficiency
bibliography: references.bib
abstract: |
  **Background.** Sub-national maps of micronutrient deficiency guide the
  targeting of fortification, supplementation and food-system programmes, but
  biomarker surveys are rare and their district-level estimates rest on a
  handful of respondents. Published accuracies for models that predict
  district prevalence from geospatial and survey proxies are seldom reported
  with the reliability of their validation target, the information their
  comparison baseline was permitted to see, or the replication of their random
  folds. An audit of our own earlier analysis found that each of these choices
  had decided a headline conclusion, in both directions.

  **Methods.** We harmonised biomarker measurements for iron, vitamin A,
  folate, vitamin B12 and zinc deficiency from national surveys in The Gambia
  (2018), Ghana (2017), Sierra Leone (2013) and Malawi (2015–16): 24
  country–outcome cells on 206 districts. Districts were linked to 473
  predictors in 24 domains from Earth observation, soil, climate, agriculture,
  livestock, infection, food-price and survey-indicator sources. A protocol
  specified before re-analysis names three estimands (filling in unsurveyed
  districts inside a surveyed country, extrapolating to an unseen region, and
  transporting to a country held entirely out of training), recomputes every
  baseline under the model's own folds and jackknifes any baseline derived
  from the survey, replicates every random split ten times, weights districts
  by an effective sample size from measured design effects, rank-normalises
  predictors within country before pooling, and calibrates transport against
  a country-block permutation null. The estimator of record is a zero-tuning
  index over domain principal components; a penalised regression, a
  covariate-free spatial smoother and the survey's own regional averages are
  scored on identical folds.

  **Results.** Within a surveyed country, covariates beat the honest survey
  baseline (rank correlation 0.39 against 0.31 on the biomarker concentration,
  0.29 against 0.19 on prevalence) but a covariate-free spatial smoother
  scores 0.39 where the full covariate model scores 0.39. Rankings transport
  to a country never used in training: mean Spearman 0.27 over 22
  country–outcome combinations at district level (17 positive) and 0.31 at
  the first sub-national tier, against permutation nulls whose 95th
  percentiles are 0.08 and 0.16; an index built only from remotely sensed
  climate and soil layers transports better, 0.37 (22 of 22 positive) and
  0.46 (20 of 22). Absolute prevalence does not transport. Directing effort to
  the fifth of districts the model ranks worst reaches 24% of national
  deficiency burden against 19% for regional survey averages and 20% at
  random. A national biomarker sample of five percent of a survey plus the
  transported ranking gives district levels with a 10 percentage-point error,
  against 20 for a district survey of the same size, which needs about 40% of
  the full sample to do better. A variance-components ceiling puts the
  attainable prevalence correlation at 0.44, one fifth of the split-half
  ceiling being cluster effect; the model is at that ceiling in three of 24
  cells. Four data blocks added in this revision (livestock density, distance
  to water and coast, helminth burden, and IHME surfaces re-extracted as
  zonal means) move the index by less than 0.01 each. Evaluation design moved
  results more than any model did: a published within-country correlation of
  0.058 was a single fold draw (replicated median 0.22), a baseline the models
  lost to had read the scored district's own respondents, a regional headline
  of 0.50–0.56 rested on three countries, and the reliability ceiling used to
  argue that districts were unresolvable was biased low by a factor of about
  4.7.

  **Conclusions.** The defensible product is a ranked district priority list
  for any country, and a level-calibrated map for a country willing to measure
  one national prevalence. Within a surveyed country geography does most of
  the work. We recommend that studies of this kind name their estimand, state
  what their baseline saw, replicate every random division, report the
  reliability of the validation target, and show that each significance
  screen could have detected an effect.
---

```{r setup}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(knitr)
})
theme_set(theme_minimal(base_size = 11))
# Every number in the body is read from the protocol tables at render time
# (results/tables/protocol_v2, re-run 2026-09-07 on the 473-column set).
root <- here::here()
P2   <- file.path(root, "results", "tables", "protocol_v2")
CL   <- file.path(root, "results", "tables", "cluster_level")
HD   <- file.path(root, "data", "covariates", "harmonized")
rd <- function(...) { f <- file.path(...); if (file.exists(f)) read.csv(f, stringsAsFactors = FALSE, check.names = FALSE) else NULL }
kb <- function(d, digits = 2, ...) knitr::kable(d, digits = digits, row.names = FALSE, booktabs = TRUE, ...)
df <- function(...) data.frame(..., check.names = FALSE, stringsAsFactors = FALSE)
get <- function(x) if (length(x) && is.finite(x[1])) x[1] else NA_real_
MD  <- rd(HD, "predictors_admin2_shared_metadata.csv")
BM  <- rd(P2, "benchmarks_v2_summary.csv")
TG  <- rd(P2, "targets_v2.csv")
arm_label <- c(domain_index = "Zero-tuning domain index", spatial_plus_domain = "Spatial smoother + covariates",
               spatial = "Spatial smoother only", domain_enet = "Elastic net on domain PCs",
               raw_enet = "Elastic net on all columns", region_mean_jk = "Survey: jackknifed regional mean",
               null_train_mean = "Null: training mean")
outcome_label <- c(child_vitA = "Child vitamin A", women_vitA = "Women's vitamin A", child_iron = "Child iron",
                   women_iron = "Women's iron", women_folate = "Women's folate", women_b12 = "Women's B12",
                   child_zinc = "Child zinc", women_zinc = "Women's zinc")
```

::: {.callout-note}
## About this version

This is the second version of the manuscript. It adopts the paper outline of
the project's findings notes (`docs/findings/TWO_READINGS_2026-09f.md`) for
the Results and Discussion, writes the two readings there as one voice, and
moves the individual-level SuperLearner and corrected-methods analyses to a
supplement. The first version (`docs/manuscript_mcn.qmd`) is kept unchanged
for comparison. Every table here is generated from the protocol result tables
at render time, so a re-run of the pipeline re-renders the paper.
:::

# Introduction

Micronutrient deficiencies remain a leading driver of preventable morbidity
and mortality among children and women in low- and middle-income countries,
contributing to anaemia, impaired growth, immune dysfunction and adverse
pregnancy outcomes [@stevens2022micronutrient; @who2020anaemia]. National
biomarker surveys collated under the WHO Vitamin and Mineral Nutrition
Information System [@who_vmnis] and the BRINDA pooled analyses
[@brinda2017overview] document broad regional contrasts, but their spatial
resolution is rarely finer than the province, and many countries have no
survey at all. Sub-national targeting of fortification, supplementation and
food-system programmes therefore rests on indirect inference from sparse
data.

Two traditions have tried to fill the gap. Small-area estimation, from the
Fay–Herriot model [@fay1979estimates] to the Bayesian disease-mapping
framework [@besag1991bayesian; @mercer2015small; @wakefield2019small;
@riebler2016intuitive], borrows strength across neighbouring areas with known
sampling variances. Ensemble machine learning on satellite-derived covariates
[@vanderlaan2007super; @osgoodzimmerman2018mapping; @local2021maps] has
produced high-resolution maps of indirect nutritional indicators. Both
traditions report accuracies for district-level prediction; what they rarely
report is the reliability of the district target those accuracies are scored
against, the information their comparison baseline was permitted to see, or
the spread across the random divisions of the data that produced the number.

That observation motivates this paper. An internal audit of our own earlier
analysis of the same four surveys found that each of those omissions had
decided a headline conclusion: a published within-country correlation of
0.058 was one draw from a distribution whose median was 0.22; a
covariate-free baseline the models appeared to lose to had been computed from
the scored district's own respondents; a significance screen reported as
finding nothing could not have found anything at any effect size; the
reliability ceiling used to argue that districts were unresolvable was biased
low by a factor of about 4.7; and a regional transport headline of 0.50–0.56
rested on three of four countries because of a spelling mismatch in a
population file. Corrections ran in both directions.

We therefore re-analysed the four surveys under a protocol specified before
the re-analysis, and we report the protocol as the paper's main contribution.
Three questions are answered separately because they have different answers:
whether covariates can fill in unsurveyed districts inside a surveyed country,
whether they can extrapolate to a region none of whose districts was seen,
and whether they transport to a country held entirely out of training. Each is
scored against the cheapest alternative a programme actually has, against a
permutation null, and against the reliability of the target itself. The
result is a deliverable stated at the resolution the evidence supports: a
ranked district priority list that transports across borders, and a level that
does not.

# Data

## Surveys and outcomes

We assembled individual biomarker measurements from four national surveys:
The Gambia Micronutrient Survey 2018 (fieldwork January–April 2018), the Ghana
Micronutrient Survey 2017 (April–June 2017), the Sierra Leone Micronutrient
Survey 2013 (November–December 2013) and the Malawi Micronutrient Survey
2015–16 (December 2015–February 2016). Fieldwork dates were recovered for every
survey cluster so that time-varying predictors could be matched to the
fieldwork window. Outcomes are deficiency indicators for children aged 6–59
months (vitamin A from retinol-binding protein or retinol, iron from ferritin,
zinc where measured) and women aged 15–49 (vitamin A, iron, folate, vitamin
B12, zinc where measured), each defined uniformly across surveys at the WHO
cut-off after BRINDA inflammation adjustment where acute-phase proteins were
available [@brinda2017adjustment; @who_vmnis_thresholds]. Two targets are
carried for every cell: the survey-weighted district prevalence, and the
survey-weighted district mean of the log biomarker concentration, which
retains the information a dichotomised indicator discards.

A fifth survey, Tanzania 2010, was withdrawn before analysis after its
dried-blood-spot vitamin A measurements were judged unreliable by a
collaborator and failed independent consistency checks; Côte d'Ivoire, with
covariates but no biomarker survey, is used only as an out-of-sample
prediction target and not for evaluation.

```{r tbl-cells}
#| tbl-cap: "The 24 country–outcome cells: districts with a biomarker estimate, respondents, and survey-weighted national prevalence. Districts are the second sub-national tier (Malawi's Traditional Authorities for in-fill; its 27 districts for region extrapolation and the regional transport tier)."
if (!is.null(TG)) {
  d <- TG |> group_by(Country = country, Outcome = outcome_label[outcome]) |>
    summarise(Districts = n(), Respondents = sum(n_raw, na.rm = TRUE),
              `Prevalence (%)` = round(100 * weighted.mean(y_prev, n_raw, na.rm = TRUE), 1),
              `Median n per district` = round(median(n_raw, na.rm = TRUE)), .groups = "drop")
  kb(d, digits = 1)
}
```

Most districts contribute between six and forty biomarker measurements, and
in three of the four surveys most districts contain a single primary sampling
unit. The precision of the district target, not the number of districts, is
the binding constraint on everything that follows, and Section 3 and Section
4.5 treat it explicitly.

## Predictors

Every predictor available at district resolution was assembled, not Earth
observation alone. The vocabulary in this version holds
`r if (!is.null(MD)) nrow(MD) else 473` columns in
`r if (!is.null(MD)) length(unique(MD$domain)) else 24` domains
(@tbl-domains): Google Earth Engine zonal aggregates [@gorelick2017google]
of climate, land cover, greenness, terrain, built environment and night-time
lights; the 64-dimension AlphaEarth satellite embedding; SoilGrids and
iSDAsoil soil chemistry; MapSPAM crop composition; Gridded Livestock of the
World densities; Malaria Atlas Project surfaces; IHME modelled surfaces of
child growth failure, anaemia, HIV, vaccination and water and sanitation,
extracted as population-weighted zonal means of the published 5 km rasters;
WHO ESPEN helminth and schistosomiasis programme data; WFP market prices,
FAOSTAT food supply and food-security indicators; distance to surface water
and to the coast; and survey-indicator aggregates from the DHS and MICS rounds
nearest each biomarker survey, with every biomarker-adjacent indicator removed
as leakage.

```{r tbl-domains}
#| tbl-cap: "Predictor domains in the shared vocabulary, with the number of columns and the sources each draws on. A domain enters the models as its leading principal components to 80% of variance."
if (!is.null(MD)) {
  src <- function(s) { s <- as.character(s); dplyr::case_when(grepl("^DHS", s) ~ "DHS / MICS", grepl("AlphaEarth", s) ~ "AlphaEarth", grepl("SoilGrids", s) ~ "SoilGrids / iSDA",
    grepl("^GEE", s) ~ "Earth Engine", grepl("IHME", s) ~ "IHME", grepl("Malaria Atlas", s) ~ "Malaria Atlas", grepl("MapSPAM", s) ~ "MapSPAM", grepl("Koppen", s) ~ "Koppen / AEZ",
    grepl("WFP", s) ~ "WFP", grepl("FAOSTAT", s) ~ "FAOSTAT", grepl("Livestock", s) ~ "GLW4", grepl("ESPEN", s) ~ "ESPEN", grepl("Surface Water", s) ~ "Earth Engine", grepl("GFDx", s) ~ "GFDx", grepl("HFID", s) ~ "HFID", TRUE ~ "WorldPop / GHS / RWI") }
  d <- MD |> mutate(Source = src(source)) |> group_by(Domain = domain) |>
    summarise(Columns = n(), Sources = paste(sort(unique(Source)), collapse = ", "), .groups = "drop") |> arrange(desc(Columns))
  kb(d, digits = 0)
}
```

Which administrative level a source natively lives at matters as much as what
it measures. The IHME surfaces are a worked example: their published district
rollups are keyed by district name, and joining them by name left a quarter
of Ghana's surveyed districts without a value after the 2019 district splits
and could only reach Malawi at its districts, which are this project's
regions. Re-extracting the same surfaces as zonal means over the spine
polygons removed both problems, and the two versions agree at Spearman
0.6–0.98 where both exist. Every source's native level, join method and known
gaps are recorded in the variable annotation sheet (Supplement S2).

Predictors that could not be reconciled across surveys were excluded with a
stated reason; national values (FAOSTAT supply, fortification standards) are
broadcast to every district and enter the pooled models only. Nothing
survey-derived from the same respondents as an outcome is used to predict it.

# The evaluation protocol {#sec-protocol}

The protocol was specified before the re-analysis and applied to every
comparison in this paper. It rests on five rules and three measurement
choices; the effect of each in isolation is reported in Section 5.

**One. Name the estimand.** "Can covariates predict district prevalence" is
three questions with three answers. *In-fill* predicts unsurveyed districts
inside a surveyed country (5-fold cross-validation over districts); *region
extrapolation* predicts a region none of whose districts was seen (exhaustive
leave-one-region-out); *country transport* predicts a country held entirely
out of training (leave-one-country-out, scored at the district and at the
first sub-national tier). Every result is reported against a named estimand.

**Two. Match the baseline's information to the model's.** A comparison is
interpretable only if the benchmark has seen what the model has seen. The
regional-mean baseline is computed under the same folds as the model and
jackknifed, so the scored district never contributes to the regional average
it is compared with. A covariate-free spatial smoother on district centroids
is scored on the identical folds, so "covariates versus geography" is not a
comparison across fold schemes.

**Three. Replicate every random split.** A single random division of 14–87
districts is a high-variance draw: across ten draws of the same scheme, the
median range of a cell's rank correlation is 0.17 and the maximum 0.71. Every
random scheme is repeated ten times and reported as its median; the region
scheme is exhaustive and carries no draw.

**Four. Calibrate transport against a null that respects the design.**
Because outcomes are correlated within a country, counts of positive cells
have a fat null: with no transport at all, 15–16 of 22 cells come out positive
five percent of the time. The primary transport criterion is therefore the
mean rank correlation over cells against the 95th percentile of a
country-block permutation null; cell counts are secondary.

**Five. Report the reliability of the target.** Every within-country
correlation is read against a ceiling for the cell: the correlation a perfect
predictor of the true district value could achieve given the sampling noise in
the survey estimate. Section 4.5 estimates it from a
respondent-in-cluster-in-district-in-region variance model, which separates
the cluster effect a split-half estimator counts as geography.

Three measurement choices follow. Districts are weighted by an effective
sample size, Kish's $n_\text{eff}$ divided by a design effect measured at the
national level for each country and outcome (median 2.4; the conventional 1.5
understates it). Predictors are rank-normalised within country before pooling,
because the same indicator sits on different scales in different surveys and
z-scoring does not remove that. And predictors enter in three competing
representations so that model capacity is a measured factor: all columns
under elastic net; the leading principal components of each domain to 80% of
variance, oriented from the training rows only; and a zero-tuning **domain
index**, a marginal-correlation-weighted composite over those components with
no cross-validated hyperparameter, which is the estimator of record.

The protocol is pre-registered for the next countries added to the panel
(`docs/findings/PREREGISTRATION_NEW_COUNTRIES_2026-09.md`): seven
predictions with thresholds fixed in advance, of which the central one is that
the two-domain climate-and-soil index transports at least as well as the full
vocabulary.

# Results

## Filling in unsurveyed districts inside a surveyed country

Within a surveyed country the covariates beat the honest survey baseline and
tie the covariate-free smoother (@tbl-infill). Against a jackknifed regional
mean, the only sub-national alternative a programme actually has, the domain
index reaches a mean rank correlation of
`r sprintf("%.2f", get(BM$mean_spearman[BM$estimand == "infill" & BM$arm == "domain_index" & BM$target == "level"]))`
against
`r sprintf("%.2f", get(BM$mean_spearman[BM$estimand == "infill" & BM$arm == "region_mean_jk" & BM$target == "level"]))`
on the biomarker concentration and
`r sprintf("%.2f", get(BM$mean_spearman[BM$estimand == "infill" & BM$arm == "domain_index" & BM$target == "prev"]))`
against
`r sprintf("%.2f", get(BM$mean_spearman[BM$estimand == "infill" & BM$arm == "region_mean_jk" & BM$target == "prev"]))`
on prevalence. A spatial smoother with no covariates at all scores
`r sprintf("%.2f", get(BM$mean_spearman[BM$estimand == "infill" & BM$arm == "spatial" & BM$target == "level"]))`
on the concentration, within 0.01 of the full covariate model, and covariates
fitted to the smoother's residuals add a median of zero. Inside a surveyed
country, most of what the covariates supply is a smooth surface that
neighbouring districts already provide.

```{r tbl-infill}
#| tbl-cap: "In-fill: mean Spearman correlation between predicted and survey-estimated district values over 24 country–outcome cells, 5-fold cross-validation over districts, median of ten fold draws. The regional-mean baseline is jackknifed so the scored district never contributes to it."
if (!is.null(BM)) {
  d <- BM |> filter(estimand == "infill", arm %in% names(arm_label)) |> select(target, arm, mean_spearman, cells_positive) |>
    pivot_wider(names_from = target, values_from = c(mean_spearman, cells_positive)) |>
    mutate(Model = arm_label[arm]) |> arrange(desc(mean_spearman_level)) |>
    select(Model, `Biomarker level` = mean_spearman_level, `Positive cells` = cells_positive_level, `Prevalence` = mean_spearman_prev, `Positive cells ` = cells_positive_prev)
  kb(d)
}
```

Capacity is a liability at these sample sizes. The zero-tuning index over
domain components outperforms a penalised regression over every column
(`r sprintf("%.2f", get(BM$mean_spearman[BM$estimand == "infill" & BM$arm == "domain_index" & BM$target == "prev"]))`
against
`r sprintf("%.2f", get(BM$mean_spearman[BM$estimand == "infill" & BM$arm == "raw_enet" & BM$target == "prev"]))`
on prevalence), and the penalised fit turns negative under region
extrapolation on prevalence (@tbl-region). With 14–87 units, a regression over
hundreds of correlated columns fits the country it was trained in. A
SuperLearner ensemble that contains the index among its candidates does not
beat it under any of four meta-learner losses (Supplement S1).

## Extrapolating to a region none of whose districts was seen

Leaving out a whole region is where a smoother must extrapolate past the edge
of its data. The index and the smoother stay close; the survey has nothing to
say about a region it has not seen, and the training-mean null collapses
(@tbl-region). Ghana (16 regions) and Malawi (27 districts) are the countries
in which this estimand is well determined; The Gambia (6) and Sierra Leone (4)
contribute one number per region.

```{r tbl-region}
#| tbl-cap: "Region extrapolation: mean Spearman over 24 cells under exhaustive leave-one-region-out. The jackknifed regional mean is not an arm here because the held-out region has no survey estimate by construction."
if (!is.null(BM)) {
  d <- BM |> filter(estimand == "region", arm %in% names(arm_label)) |> select(target, arm, mean_spearman, cells_positive) |>
    pivot_wider(names_from = target, values_from = c(mean_spearman, cells_positive)) |>
    mutate(Model = arm_label[arm]) |> arrange(desc(mean_spearman_level)) |>
    select(Model, `Biomarker level` = mean_spearman_level, `Positive cells` = cells_positive_level, `Prevalence` = mean_spearman_prev, `Positive cells ` = cells_positive_prev)
  kb(d)
}
```

## Transporting to a country never used in training

Rankings transport across borders; levels do not (@tbl-transport). Under
leave-one-country-out the domain index ranks a held-out country's districts
with a mean Spearman of
`r sprintf("%.2f", get(BM$mean_spearman[BM$estimand == "country" & BM$arm == "domain_index" & BM$target == "level"]))`
on the biomarker concentration over the 22 evaluable cells, positive in
`r get(BM$cells_positive[BM$estimand == "country" & BM$arm == "domain_index" & BM$target == "level"])`,
and its regions with
`r { a1 <- rd(P2, "admin1_transport.csv"); sprintf("%.2f", get(mean(a1$spearman[a1$arm == "domain_index" & a1$target == "level"], na.rm = TRUE))) }`;
the country-block permutation null's 95th percentiles are
`r { nc <- rd(P2, "transport_null_calibration.csv"); sprintf("%.2f and %.2f", get(nc$null_mean_q95[nc$tier == "admin2"]), get(nc$null_mean_q95[nc$tier == "admin1"])) }`.
An index built only from the climate and soil domains transports better at
both tiers, 0.37 at district level (positive in all 22) and 0.46 at the
regional tier (20 of 22). That two-domain index was chosen on the same four
countries it is scored on, which is why it enters the next country as a
pre-registered prediction rather than a result; honest nested selection of
domains inside the training fold recovers about 0.04 of its margin.

```{r tbl-transport}
#| tbl-cap: "Country transport: mean Spearman in a country held entirely out of training, over 22 cells, at the district and at the first sub-national tier, for the full-vocabulary index and the pre-registered climate + soil index, with the 95th percentile of the country-block permutation null."
a1 <- rd(P2, "admin1_transport.csv"); cs1 <- rd(P2, "climate_soil_admin1.csv"); nd <- rd(P2, "nested_domain_selection.csv"); nc <- rd(P2, "transport_null_calibration.csv")
mrow <- function(x, ...) if (!is.null(x)) get(mean(x$spearman[...], na.rm = TRUE)) else NA_real_
if (!is.null(BM)) kb(df(Tier = c("District (Admin-2)", "Regional (Admin-1)"),
  `Full index, level` = c(get(BM$mean_spearman[BM$estimand == "country" & BM$arm == "domain_index" & BM$target == "level"]), mrow(a1, a1$arm == "domain_index" & a1$target == "level")),
  `Full index, prevalence` = c(get(BM$mean_spearman[BM$estimand == "country" & BM$arm == "domain_index" & BM$target == "prev"]), mrow(a1, a1$arm == "domain_index" & a1$target == "prev")),
  `Climate + soil, level` = c(mrow(nd, nd$arm == "fixed_cs" & nd$target == "level"), mrow(cs1, cs1$set == "climate_soil" & cs1$target == "level")),
  `Climate + soil, prevalence` = c(mrow(nd, nd$arm == "fixed_cs" & nd$target == "prev"), mrow(cs1, cs1$set == "climate_soil" & cs1$target == "prev")),
  `Null 95th percentile` = c(if (!is.null(nc)) get(nc$null_mean_q95[nc$tier == "admin2"]) else NA, if (!is.null(nc)) get(nc$null_mean_q95[nc$tier == "admin1"]) else NA)))
```

Absolute prevalence does not transport: biomarker levels differ between
surveys by more than any covariate can explain (raw ferritin differs six-fold
across the four), so a transported map is a ranking. Which layers carry the
ranking is shown by ablation (@tbl-ablation): dropping each domain in turn and
rebuilding the components, soil, climate, agriculture and livestock density
are the only domains whose removal costs transport; removing the 138 survey
indicator columns improves it. Adding training countries helps at about 0.05
per country on the concentration target, so the panel is not saturated in
countries even if it is saturated in layers.

```{r tbl-ablation}
#| tbl-cap: "Which domains carry transport: change in mean district-level Spearman (biomarker level, 22 cells) when each domain is dropped from the vocabulary and the components rebuilt, and the transport of the domain on its own. Positive 'cost of dropping' means the domain is load-bearing."
da <- rd(P2, "domain_ablation_loco_summary.csv")
if (!is.null(da)) { d <- da |> filter(target == "level") |> arrange(desc(delta_drop)) |> transmute(Domain = domain, `Cost of dropping` = delta_drop, `Alone` = only) |> head(10); kb(d, digits = 3) }
```

The transported ranking is not an urban–rural map. Partialling a composite of
night lights, population density, built surface and travel time out of the
climate-and-soil index moves its transport by less than 0.02, and urbanicity
alone does not transport.

## What the ranking buys a programme

Used for targeting, the ranking reaches more deficiency burden than the
survey's own averages (@tbl-burden). Directing effort to the fifth of
districts the model ranks worst reaches
`r { nt <- rd(P2, "nce_targeting_summary.csv"); sprintf("%.0f%%", 100 * get(nt$mean_capture[nt$estimand == "infill" & nt$arm == "domain_index"])) }`
of national deficiency burden in-fill, against
`r { nt <- rd(P2, "nce_targeting_summary.csv"); sprintf("%.0f%%", 100 * get(nt$mean_capture[nt$estimand == "infill" & nt$arm == "region_mean_jk"])) }`
for jackknifed regional survey averages, which are themselves no better than
random selection, and 47% for an oracle that knows the true worst fifth. The
margin is modest per cell and concentrates where regions hold several
districts; in Malawi the selected districts carry 43% child iron deficiency
against 26% nationally and hold half the national burden. Under transport the
burden margin disappears (lift 1.0), so the transport result is about ordering
and should not be restated as a targeting gain.

```{r tbl-burden}
#| tbl-cap: "Burden captured by the worst-ranked fifth of districts, in-fill, mean over 24 cells: share of national deficiency burden reached, lift over random selection, and cells with lift above one."
nt <- rd(P2, "nce_targeting_summary.csv")
if (!is.null(nt)) { d <- nt |> filter(estimand == "infill") |> mutate(Model = c(oracle_ceiling = "Oracle: true worst fifth", arm_label)[arm]) |>
  select(Model, `Burden reached` = mean_capture, `Lift vs random` = mean_lift, `Cells with lift > 1` = cells_lift_gt1); kb(d) }
```

Classified into the WHO public-health bands for vitamin A (none, mild,
moderate, severe), the in-fill predictions place a district in the correct
band 52% of the time and within one band 90% of the time; a transported
ranking anchored to the country's national prevalence does as well (50% and
87%).

A national estimate is enough to put levels on the ranking (@tbl-anchor).
A survey design that measures only a national prevalence from five percent of
a full survey's sample (fifty to seventy respondents) and lets the transported
covariates order the districts gives a district-level error of about 10
percentage points, against 20 for a district survey of the same size and 12
for a regional one; the district survey needs about 40% of the full sample,
roughly 400 respondents, before it does better on level error. It loses on
burden captured at every sample size, because burden concentrates in populous
districts that a survey measures precisely, so the design is a level
calibration for a ranking, not a substitute for a survey.

```{r tbl-anchor}
#| tbl-cap: "Survey design: median district error in percentage points over 22 cells, by share of the full survey sample. National anchor + ranking measures only a national prevalence and orders districts by the transported climate + soil index; the district and regional surveys spend the same sample on direct estimates."
ar <- rd(P2, "anchor_and_rank_summary.csv")
if (!is.null(ar)) { d <- ar |> filter(rank_from == "prev", set == "climate_soil", fraction %in% c(0.05, 0.15, 0.25, 0.4, 0.6)) |>
  mutate(design = c(A1_anchor_rank = "National anchor + ranking", A2_region_anchor_rank = "Regional anchors + ranking", B_district_survey = "District survey", C_regional_survey = "Regional survey")[design],
         fraction = sprintf("%.0f%%", 100 * fraction)) |> select(fraction, design, mae) |> pivot_wider(names_from = design, values_from = mae) |> rename(`Share of full sample` = fraction); kb(d, digits = 1) }
```

## How much is attainable: the reliability ceiling {#sec-ceiling}

A district estimate built from a dozen respondents is noisy, and no model can
correlate with the true district value better than the survey estimate does.
A split-half ceiling, which correlates two random halves of each district's
respondents, counts the survey cluster as geography: respondents in one
cluster resemble each other whether or not their district differs from the
next. A respondent-in-cluster-in-district-in-region variance model reproduces
the split-half ceiling when the cluster is included and removes it when it is
not (@tbl-ceiling). One fifth of the split-half ceiling is cluster effect; the
honest ceiling averages 0.44 on prevalence, six of 24 cells have a ceiling
below 0.30, and the model already reaches the ceiling in three cells. Child
zinc has no district geography at all: its between-district variance is zero,
and what a split-half estimator called geography was the cluster. The earlier
analytic ceiling, which assumed a design effect of 1.5, was biased low by
about 4.7 and is withdrawn (Section 5).

```{r tbl-ceiling}
#| tbl-cap: "Reliability ceiling on prevalence by country, district tier: share of districts with a single survey cluster, the split-half ceiling, the variance-components ceiling with the cluster effect removed, and the in-fill correlation achieved by the domain index."
vc <- rd(P2, "variance_components_ceiling.csv")
if (!is.null(vc)) { d <- vc |> filter(rung == "admin2", target == "prev") |> group_by(Country = country) |>
  summarise(`Single-cluster districts` = sprintf("%.0f%%", 100 * mean(share_single_cluster)), `Split-half ceiling` = round(mean(ceiling_within_vc), 2), `Honest ceiling` = round(mean(ceiling_vc), 2), `Achieved, in-fill` = { a <- mean(achieved_spearman, na.rm = TRUE); ifelse(is.finite(a), sprintf("%.2f", a), "not folded (14 districts)") }, .groups = "drop"); kb(d) }
```

## What does not help, which is itself a result

Nothing added to the vocabulary in this revision moves the index. Food prices
and night-time temperature matched to each cluster's fieldwork month, the
satellite embedding as a domain of its own, and four blocks added for this
version, livestock density, distance to surface water and to the coast,
helminth burden and control, and the IHME surfaces re-extracted as zonal
means, each change in-fill and transport by less than 0.01 when added to the
base vocabulary (@tbl-addons). The fieldwork-month temperature columns are
among the most consistently associated with deficiency in the marginal scan
(four of four countries for child vitamin A) yet improve no fold, because a
covariate defined by when the survey happened cannot rank districts that have
not been surveyed. Livestock density alone matches the full vocabulary at
district transport on the concentration target, and enters the ablation as
the second load-bearing domain; the vocabulary is saturated, and the two
remotely sensed domains carry it.

```{r tbl-addons}
#| tbl-cap: "Add-on blocks scored under one harness: median change in Spearman when the block is added to the base vocabulary (in-fill, 18 cells; district transport, 22 cells) and the transport of the block alone."
tags <- c(food_tm = "Food prices, fieldwork-matched", climate_tm = "Night temperature and rainfall, fieldwork-matched", alphaearth = "Satellite embedding as own domain",
          livestock = "Livestock density (GLW4)", water_distance = "Distance to water and coast", espen = "Helminth burden and control (ESPEN)", ihme_raster = "IHME surfaces as zonal means")
rows <- lapply(names(tags), function(t) { i <- rd(P2, paste0("addon_", t, "_infill.csv")); l <- rd(P2, paste0("addon_", t, "_loco.csv")); if (is.null(i) || is.null(l)) return(NULL)
  dd_i <- function(x) { w <- x |> filter(target == "level") |> group_by(country, outcome) |> summarise(d = mean(spearman[set == "plus"], na.rm = TRUE) - mean(spearman[set == "base"], na.rm = TRUE), .groups = "drop"); median(w$d, na.rm = TRUE) }
  dd_l <- function(x, s1) { w <- x |> filter(tier == "admin2", target == "level") |> group_by(heldout, outcome) |> summarise(d = mean(spearman[set == s1], na.rm = TRUE) - mean(spearman[set == "base"], na.rm = TRUE), .groups = "drop"); median(w$d, na.rm = TRUE) }
  df(Block = tags[[t]], `In-fill, level` = dd_i(i), `Transport, level` = dd_l(l, "plus"), `Alone, transport` = dd_l(l, "addon_only")) })
rows <- rows[!sapply(rows, is.null)]; if (length(rows)) kb(bind_rows(rows), digits = 3)
```

Two further sensitivity analyses are reported in the supplement. Fitting the
same arms at the survey cluster instead of the district (323 clusters with
undisplaced GPS, covariates as 2 km and 5 km buffer means) does not beat the
district fit on any estimand and is worse for transport; a matched-vocabulary
comparison isolates the fitting unit from the layers. And a person-level
SuperLearner on the survey questionnaire, the district proxies, or both, has
discriminative skill only for iron (Brier skill 0.06–0.14) and none for the
vitamins; the questionnaire and the proxies carry the same weak signal at the
individual level (Supplement S1).

# What the protocol changes {#sec-changes}

Evaluation design moved the results of this study more than any modelling
choice did, and the same errors are available to any group doing this work.
@tbl-changes sets each headline under the earlier evaluation next to its value
under the protocol. Three corrections favoured the covariates, three ran
against them, and two were data-handling defects found only because the
protocol made a dropped country or a mis-filed domain visible.

```{r tbl-changes}
#| tbl-cap: "Each headline result under the earlier evaluation and under the protocol, with the change that moved it."
kb(df(
  Result = c("Within-country rank correlation", "Covariate-free survey baseline", "Reliability ceiling (prevalence)", "Regional transport",
             "District transport", "Regional anchoring gain", "Predictor significance screen", "Covariates vs geography, cells won",
             "Individual-level arm", "Satellite embedding", "IHME block for The Gambia", "Population join", "DHS anaemia prevalence as a predictor", "Vitamin A level target"),
  `Earlier evaluation` = c("0.058 (one fold draw)", "r 0.516, beat every covariate arm", "0.098 analytic, then 0.47–0.61 split-half",
             "0.50–0.56, 12 of 12", "0.28 penalised fit, 21 of 22", "0.164 to 0.413, better in 20 of 24", "0 of 294 survive FDR",
             "6 of 24", "0.516, headline", "64 columns filed under agriculture", "25 of 34 columns empty", "Sierra Leone dropped from every population-joined run", "In the vocabulary, all four countries", "Each survey's own adjusted RBP (raw in Malawi)"),
  `Under the protocol` = c("0.22 replicated; 0.39 (level) and 0.29 (prevalence) for the index", "0.076 jackknifed; loses to the index",
             "0.44 variance-components; 21% of the split-half ceiling was cluster effect", "0.31 over 22 cells, 17 positive; three countries only before",
             "0.27 zero-tuning index, 17 of 22; 0.29 penalised, 20 of 22", "0.147 vs 0.156 unanchored; withdrawn", "Screen had power 0.005; sign replication reported instead",
             "10 of 24 (miscount); tied on matched folds", "Trained on the scored district's respondents; supplement only", "Own domain; +0.01 on transport",
             "One year per country dropped every series ending earlier; fixed per indicator", "Spelling normalised; all 22 cells restored", "Removed; iron transport moved by at most 0.04 in one cell, the index by 0.00", "Same BRINDA-adjusted RBP as the prevalence; Malawi child vitamin A level +0.05 to +0.07"),
  `What moved it` = c("Replicated folds", "Jackknife", "Variance model with cluster term", "BUG-01 found by the per-country cell count", "Estimator of record stated",
             "Jackknifed control arm", "Attained power reported", "Recount; matched folds", "Estimand named", "Metadata audit", "Per-country coverage audit", "Per-country cell count", "Per-column leakage audit (AU-01)", "Audit of target construction (AU-01)")))
```

Two lessons generalise. First, a baseline or benchmark computed from data that
includes the unit being scored reports a gain that is partly the unit's own
answer; the jackknifed arm cost one additional run and reversed a finding.
Second, a dropped country, a mis-filed domain or a blanked block does not
announce itself in a column count: the vocabulary had 451 columns before and
after the IHME defect, and the regional headline had 12 cells before and after
Sierra Leone vanished. Per-country coverage and per-country cell counts, not
totals, are the checks that found them.

# Limitations

Four countries, three of them West African, bound the transport result; the
pre-registration for the next countries is the test that can move it. The
deliverable is a ranking, not a level: absolute prevalence differs between
surveys by more than covariates explain, and the level must come from a
national measurement. Within a surveyed country geography supplies most of the
attainable accuracy, so the covariates earn their place mainly where no survey
exists to smooth from. Sierra Leone's 14 districts cannot be folded for
in-fill and contribute one number per region. Several targets are degenerate at
district resolution, most clearly child zinc, whose between-district variance
is zero, and the rare outcomes (women's vitamin A below 3% in every survey)
have ceilings below 0.30. The two-domain index was selected on the countries
it is scored on. Iron is inflammation-adjusted by each survey's native method,
vitamin A uniformly; soluble transferrin receptor, available in all four
surveys and less inflammation-sensitive than ferritin, is unused. Iodine, with
urinary iodine in two surveys, cannot be validated across borders and is
reported within country only (0.35–0.48 in The Gambia).

# Recommendations for the field

The transferable result is procedural. Studies that predict sub-national
prevalence from covariates should:

1. **Name the estimand.** In-fill, region extrapolation and country transport
   have materially different answers and are routinely reported
   interchangeably.
2. **State what the baseline saw**, and jackknife any baseline derived from the
   same survey so the scored unit never contributes to it.
3. **Replicate every random division** of the data and report the spread; a
   single draw of 14–87 districts is good to about ±0.09 and can be off by
   ±0.35.
4. **Report the reliability of the validation target** from a variance model
   that separates the survey cluster from geography, and read every
   correlation against it.
5. **Show that each significance screen could have detected an effect** at
   the permutation resolution used; where it cannot, replace it with an
   evidence standard that has power, such as replication of sign across
   surveys.
6. **Audit coverage per country, not in total**, because the defects that
   moved this paper's headlines were invisible in column and cell counts.

Applied to our own work, these conditions reversed conclusions in both
directions, and we would not have identified any of them from the modelling
results alone.

# Conclusions

Proxy indicators carry replicated, nutrient-specific signal for sub-national
micronutrient deficiency, and the honest product is a ranked district
priority list: mean rank correlation about 0.3 at either tier for a country
with no survey, higher for an index built from remotely sensed climate and
soil alone, plus a level calibration that needs one national estimate. Within
a surveyed country a spatial smoother delivers most of what the covariates do.
The vocabulary is saturated, the headroom against the honest ceiling is small,
and claims about levels, burden or survey savings should be read as upper
bounds. The protocol that produced these statements, and the side-by-side
record of what it changed, are the parts of this work most worth carrying to
the next survey.

# Acknowledgements {.unnumbered}

Funded by the Bill & Melinda Gates Foundation (grant number INV-XXXXXX). The
funder had no role in study design, analysis, decision to publish, or
manuscript preparation.

# Author contributions {.unnumbered}

ANM conceived the study, conducted the analysis, and drafted the manuscript.
Co-authors to be added on revision.

# Data availability {.unnumbered}

Survey biomarker data are available from the originating survey programmes
following registration. Pipeline code, the protocol scripts
(`scripts/protocol_v2/`), the pre-registration and the findings notes are in
the project repository.

# Supplement S1: the individual-level SuperLearner and the corrected-methods analyses {.unnumbered}

The analyses in this supplement were the main results of the first version of
this manuscript. They are retained because they document the estimators the
protocol was compared against and because two of their findings stand; they
are not the analysis of record, for the reasons stated in Section 5.

## S1.1 Individual-level SuperLearner aggregated to districts {.unnumbered}

The first version fitted a person-level ensemble (twelve learners: penalised
regression, random forests, gradient boosting and BART; a Gaussian process was
dropped after receiving zero ensemble weight while dominating runtime) on the
harmonised survey items and averaged its out-of-fold predictions to the
district. It reported the highest raw within-country score of any arm (median
Pearson r 0.516 over 24 cells; median absolute error 8.55 percentage points).
The folds split respondents rather than districts, so a district's own
respondents appeared in training; the number is not an out-of-district
comparison and was withdrawn as a headline. The aggregated predictions also
showed the compression the small-area literature predicts [@mercer2015small;
@wakefield2019small; @niculescu2005predicting]: a probabilistic classifier
under uncertainty shrinks toward the marginal class probability, and naive
aggregation pulls every district mean toward the national prevalence, giving
large negative biases for high-prevalence outcomes. Fitting at the unit of
deliverable is the standing recommendation.

Under the protocol the person-level model was re-fitted with survey weights,
outer folds cut by district so no district's respondents are split across
training and test, and three predictor sets: the survey questionnaire alone,
the district proxies alone, and both (@fig-individual). Discrimination is
near chance for the vitamins and folate in every set and present only for
iron (AUC 0.75–0.78, Brier skill 0.06–0.14 in Ghana and Malawi); the
questionnaire and the proxies carry the same weak signal at the individual
level, and adding the proxies to the questionnaire changes the Brier skill by
less than 0.01 in most cells.

```{r fig-individual}
#| fig-cap: "Person-level SuperLearner, outer 5-fold cross-validation by district: Brier skill over the national prevalence for the survey questionnaire alone, the district proxies alone, and both, by country and outcome."
#| fig-width: 8
#| fig-height: 4.2
il <- rd(P2, "individual_level_models.csv")
if (!is.null(il) && nrow(il)) {
  d <- il |> mutate(Outcome = outcome_label[outcome], set = factor(c(survey_only = "Survey variables only", proxies_only = "District proxies only", both = "Both")[set], levels = c("Survey variables only", "District proxies only", "Both")))
  ggplot(d, aes(x = 100 * brier_skill, y = Outcome, colour = set)) + geom_vline(xintercept = 0, linetype = 2, colour = "grey50") +
    geom_point(size = 2.5, position = position_dodge(width = 0.5)) + facet_wrap(~ country, nrow = 1) +
    labs(x = "Brier skill over the national prevalence (%)", y = NULL, colour = NULL) + theme(legend.position = "bottom")
}
```

## S1.2 The corrected-methods (P1–P8) layer {.unnumbered}

Before the protocol, a set of corrected methods was built to remove the
leakages found in the production pipeline: preprocessing inside each fold and
folds blocked by survey cluster and by region (P1); out-of-fold recalibration
of binary predictions (P2); decomposition of district error into a survey
sampling component and a model component (P3); split-conformal and
design-based intervals per district (P4); a decision-value score for ranking
the highest-burden districts (P5); out-of-support flags for districts whose
covariates fall outside the surveyed range (P6); empirical-Bayes partial
pooling (P7); and per-district trust labels (P8). Their outputs are in
`results/tables/corrected/`. Two of their findings stand and are absorbed by
the protocol: honest cluster-blocked folds lower within-country discrimination
by a few points of AUC relative to random folds, and a substantial part of the
apparent district error is survey sampling noise. The protocol supersedes the
layer by naming the estimand, matching baselines, replicating folds and
weighting by measured effective sample size, none of which the P1–P8 layer
did.

```{r tbl-s1-honesty}
#| tbl-cap: "Corrected methods, P1: within-country AUC under random cross-validation with full-data preprocessing against cluster-blocked and region-blocked folds with in-fold preprocessing."
cc <- rd(root, "results", "tables", "corrected", "cv_honesty_compare.csv")
if (!is.null(cc)) {
  lab <- c("random CV + full-data preprocessing (optimistic)" = "Optimistic", "cluster-block CV" = "Cluster-block", "region/spatial-block CV" = "Region-block")
  d <- cc |> mutate(scheme = lab[scheme]) |> filter(!is.na(scheme)) |> select(country, outcome, scheme, auc) |> pivot_wider(names_from = scheme, values_from = auc) |> arrange(country, outcome)
  kb(d)
}
```

```{r tbl-s1-error}
#| tbl-cap: "Corrected methods, P3: district prediction error decomposed into a survey-sampling component and a model component."
ce <- rd(root, "results", "tables", "corrected", "admin2_error_sampling_adjusted.csv")
if (!is.null(ce)) kb(ce |> select(country, outcome, n_admin2, naive_rmse_pp, sampling_rmse_pp, adjusted_rmse_pp, pearson_r), col.names = c("Country", "Outcome", "n areas", "Naive RMSE (pp)", "Sampling RMSE (pp)", "Adjusted RMSE (pp)", "r"))
```

## S1.3 Other analyses of the first version {.unnumbered}

The first version also reported a benchmark of 21 comparator methods under
leave-one-country-out (Bayesian small-area models, penalised regressions,
invariance-based transport methods and two spatial smoothers, of which the
smoothers led on rank correlation), an area-level prevalence recipe whose
comparative claims did not survive a leakage-corrected re-ranking, a test of
modelling the continuous biomarker instead of the binary, a test of what
target a model should carry across a border (within-country z-scores
transport; deviations from the national level do not), a national-level model
on the WHO VMNIS panel that cannot supply a district map's level, and a
comparison of admin-1 against admin-2 as the unit of estimation. Their text and
tables are in `docs/manuscript_mcn.qmd`; each conclusion that survives is
stated in the body above, and each that does not is listed in @tbl-changes.

# Supplement S2: data provenance and the variable annotation sheet {.unnumbered}

Every predictor's source, native administrative level, join method, coverage
by country and known gaps are recorded in the variable annotation sheet
(`results/tables/protocol_v2/variable_sheet.csv`) and the source-block
metadata (`data/covariates/harmonized/*_metadata.csv`). Four data-handling
defects found during this work are recorded there with their fix: the
population file spelling that dropped Sierra Leone from every population-joined
run; the satellite embedding filed under agriculture; the one-year-per-country
selection that blanked 25 of The Gambia's 34 IHME and GFDx columns; and the
survey year of The Gambia entered as 2021 in two scripts (fieldwork was
January–April 2018). Remaining gaps are source gaps: the fortification
database has no rows for The Gambia, IHME does not model circumcision or four
sanitation indicators there, and ESPEN exposes programme endemicity classes
rather than survey prevalence.

# Supplement S3: per-cell transport {.unnumbered}

```{r tbl-s3-cells}
#| tbl-cap: "Country transport by cell: Spearman correlation in the held-out country for the zero-tuning domain index, biomarker level and prevalence, district tier."
cells <- rd(P2, "benchmarks_v2_cells.csv")
if (!is.null(cells)) {
  d <- cells |> filter(estimand == "country", arm == "domain_index") |> group_by(country, outcome, target) |> summarise(rho = mean(spearman, na.rm = TRUE), .groups = "drop") |>
    pivot_wider(names_from = target, values_from = rho) |> transmute(Country = country, Outcome = outcome_label[outcome], `Level` = level, `Prevalence` = prev) |> arrange(Country, Outcome)
  kb(d)
}
```

# References {.unnumbered}

::: {#refs}
:::
