---
title: "Subnational micronutrient deficiency prediction in sub-Saharan Africa from proxy indicators: a multi-country SuperLearner ensemble benchmarked against small-area-estimation methods"
shorttitle: "Subnational micronutrient prediction in sub-Saharan Africa"
author:
  - name: Andrew N. Mertens
    affiliation: University of California, Davis, Institute for Global Nutrition
    corresponding: true
    email: anmertens@ucdavis.edu
date: today
format:
  docx:
    number-sections: true
    fig-cap-location: bottom
    tbl-cap-location: top
    fig-width: 6.5
    fig-height: 4.5
    fig-dpi: 300
execute:
  echo: false
  warning: false
  message: false
keywords:
  - micronutrient deficiency
  - small area estimation
  - DHS
  - transportability
  - SuperLearner
  - vitamin A
  - iron deficiency
bibliography: references.bib
abstract: |
  **Background:** Subnational maps of micronutrient deficiency are central to
  targeted nutrition programming, yet biomarker surveys remain scarce and
  geographically uneven. We assessed whether ensemble machine-learning models
  trained on harmonised Demographic and Health Surveys (DHS) and remote-sensing
  predictors can reproduce admin-2 deficiency patterns within and across four
  sub-Saharan African countries (three in West Africa plus Malawi), and whether
  they outperform canonical small-area estimation (SAE) methods.

  **Methods:** We harmonised individual-level DHS biomarker measurements for
  iron, vitamin A, vitamin B12, folate, and zinc deficiency from Gambia (2018),
  Ghana (2017), Sierra Leone (2013) and Malawi (2016) and combined them with
  earth-observation aggregates at the admin-2 polygon level. Four estimation
  strategies were compared: (1) individual-level SuperLearner aggregated to
  admin-2; (2) area-level SuperLearner fit directly on admin-2 aggregates;
  (3) frequentist regressions (OLS, elastic-net) and a country-mean baseline;
  (4) Bayesian small-area methods including Fay–Herriot and the BYM2 model
  (Besag–York–Mollié, PC priors) fit via integrated nested Laplace approximation
  (INLA). All approaches were evaluated within country by k-fold cross-validation
  and across countries by leave-one-country-out (LOCO) holdout. Performance
  metrics were Pearson r and Spearman ρ for spatial ranking, MAE and RMSE in
  percentage points for absolute error, and calibration slope. Headline metrics
  are from a leakage-free corrected analysis (preprocessing confined to each
  training fold; folds blocked by survey cluster and spatial region; out-of-fold
  recalibration), which is the analysis of record; the original pipeline is
  retained as a sensitivity comparison.

  **Results:** Evaluated with genuinely out-of-fold predictions, within-country
  admin-2 ranking was modest. Aggregating individual-level SuperLearner
  predictions to admin-2 ranked districts best (median Pearson r 0.53, range
  −0.39 to 0.87), though with large negative biases (up to ~20 percentage
  points) for iron. The area-level SuperLearner fit directly on admin-2
  aggregates appeared excellent under in-sample evaluation but was substantially
  weaker out-of-fold (MSE-loss median r 0.12, range −0.57 to 0.61; 0 of 24
  evaluable country–outcome combinations reached r ≥ 0.9 and about half were
  ≤ 0), reflecting overfitting in the area-level p ≫ n setting (14–87 areas,
  ~2,400 candidate covariates). Cross-country (LOCO) performance
  was weaker still: median r ≈ 0.1–0.2, with absolute
  errors of 1.5–44 percentage points and systematic underestimation in
  high-prevalence settings. Iron-deficiency outcomes transported best;
  vitamin-A-deficiency outcomes essentially did not transport (median r near
  zero, often with anti-predictive ordering). Benchmark frequentist regressions
  and Bayesian SAE models matched or exceeded the SuperLearner on the large
  majority of LOCO splits (the best benchmark beat the prescreened ensemble
  on 10 of 13 evaluable head-to-head splits),
  indicating that the ensemble's predictive gains are concentrated on specific
  outcome–country combinations rather than general.

  **Conclusions:** Under honest out-of-fold evaluation, admin-2 prediction of
  micronutrient deficiency is only modestly accurate within countries; the
  apparent excellence of the directly-fit area-level ensemble was an artifact of
  in-sample evaluation, and individual-level prediction aggregated to admin-2 is
  the stronger area-level approach. Out-of-country transport remains unreliable for spatial targeting,
  especially outside the West African geography and for vitamin-A-deficiency
  outcomes. Standard small-area-estimation methods are a strong baseline that
  ensembles do not consistently beat; benchmarking should accompany any
  decision-relevant deployment of these models.
---

```{r setup}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(knitr)
  library(kableExtra); library(here); library(readr); library(scales)
})
theme_set(theme_minimal(base_size = 11))

# ── Results sources ───────────────────────────────────────────────────────
TABLE_DIR <- here::here("results", "tables")
TARGETS_STORE <- here::here("_targets_full")

safe_read_csv <- function(path) {
  if (file.exists(path)) read_csv(path, show_col_types = FALSE) else NULL
}

cv_perf       <- safe_read_csv(file.path(TABLE_DIR, "cv_performance_all.csv"))
diag_binary   <- safe_read_csv(file.path(TABLE_DIR, "diagnostics_binary.csv"))
diag_cont     <- safe_read_csv(file.path(TABLE_DIR, "diagnostics_continuous.csv"))
admin2_err    <- safe_read_csv(file.path(TABLE_DIR, "admin2_error_all.csv"))
area_compare  <- safe_read_csv(file.path(TABLE_DIR, "area_comparison_all.csv"))
loco_metrics  <- safe_read_csv(file.path(TABLE_DIR, "transportability_area_loco_metrics.csv"))
loco_full     <- safe_read_csv(file.path(TABLE_DIR, "area_loco_comparison.csv"))
natl_est      <- safe_read_csv(file.path(TABLE_DIR, "national_estimates_all.csv"))

# Benchmark results: written by the benchmark pipeline targets when available.
bench_results <- safe_read_csv(file.path(TABLE_DIR, "benchmarks_all.csv"))

# ── Corrected-methods (P1–P8) outputs — the analysis of record ──────────────
# Produced by the corrected-methods section of the unified _targets.R pipeline
# (build_methods_comparison -> results/tables/corrected/). These honest, leakage-
# free metrics are canonical for within-country discrimination, calibration,
# admin-2 error, prediction intervals, decision value and per-district trust;
# the production-pipeline equivalents are retained as a sensitivity comparison.
CORR_DIR <- file.path(TABLE_DIR, "corrected")
corr_cv      <- safe_read_csv(file.path(CORR_DIR, "cv_honesty_compare.csv"))
corr_calib   <- safe_read_csv(file.path(CORR_DIR, "calibration_oof_compare.csv"))
corr_err     <- safe_read_csv(file.path(CORR_DIR, "admin2_error_sampling_adjusted.csv"))
corr_decision<- safe_read_csv(file.path(CORR_DIR, "decision_value.csv"))
corr_interval<- safe_read_csv(file.path(CORR_DIR, "interval_summary.csv"))
corr_trust   <- safe_read_csv(file.path(CORR_DIR, "trust_flags.csv"))
```

# Introduction

Micronutrient deficiencies remain a leading driver of preventable morbidity
and mortality among children and women in low- and middle-income countries,
contributing to anaemia, impaired growth, immune dysfunction and adverse
pregnancy outcomes [@stevens2022micronutrient; @who2020anaemia]. National
biomarker surveys conducted under the WHO Vitamin and Mineral Nutrition
Information System (VMNIS) [@who_vmnis] and the BRINDA pooled analyses
[@brinda2017overview] document broad regional contrasts, but the spatial
resolution of these surveys is rarely finer than admin-1 (province) and is
absent in many countries entirely. Subnational geographic targeting of
fortification, supplementation and food-system programmes therefore relies
on indirect inference from sparse data.

Two methodological traditions have attempted to fill this gap. The
small-area-estimation (SAE) literature, originating with the Fay–Herriot
mixed model [@fay1979estimates] and developed for survey-derived admin-level
prevalence under the Bayesian disease-mapping framework
[@besag1991bayesian; @mercer2015small; @wakefield2019small;
@riebler2016intuitive], constructs structured spatial smoothers using
admin-polygon adjacency and known sampling variances. In parallel, ensemble
machine-learning approaches, including SuperLearner [@vanderlaan2007super]
and gradient-boosted tree models trained on satellite-derived covariates,
have produced high-resolution maps of indirect indicators of nutritional
status [@osgoodzimmerman2018mapping; @local2021maps]. These two traditions
have rarely been compared head-to-head on the same biomarker datasets at
the same geographic scale.

Two questions about this approach remain unresolved. First, when a model
is trained at the individual level and aggregated to admin-2, do its
admin-2 estimates align with survey-weighted admin-2 prevalence — or does
the aggregation step introduce systematic bias? Second, when a model is
trained on a set of countries with biomarker data, do its predictions
transport to a held-out country with similar covariates? The
small-area-estimation literature has long recommended fitting at the
unit of deliverable: if the output is an admin-2 map, fit at admin-2; if
the output is household-level risk, fit at the individual level
[@mercer2015small; @wakefield2019small]. To our knowledge, the
implication of this recommendation has not been systematically tested for
multi-country micronutrient deficiency prediction.

In this study we harmonise DHS biomarker measurements from four
sub-Saharan African countries (The Gambia, Ghana, Sierra Leone, and
Malawi) and use these to evaluate four estimation strategies
spanning machine-learning and SAE traditions, within-country and across
countries. We benchmark a multi-learner SuperLearner ensemble (combining
elastic-net, ranger, gradient-boosted trees, and dropout BART) against
Bayesian areal models and frequentist baselines, and report whether the
added complexity of the ensemble produces gains in spatial ranking or
transportability that justify its use for global-health programming.

# Methods

## Study population and biomarker data

We assembled biomarker measurements from four Demographic and Health
Surveys: The Gambia 2018, Ghana 2017, Sierra Leone 2013, and Malawi 2016.
Within each survey we extracted deficiency biomarkers for women of
reproductive age (15–49 y) and children 6–59 months as available
(@tbl-cohort).

```{r tbl-cohort}
#| tbl-cap: "Sample sizes by country and biomarker, with WHO public-health classification thresholds applied to derive binary deficiency outcomes."
if (!is.null(cv_perf)) {
  cv_perf %>%
    filter(model_type == "binary") %>%
    transmute(
      Country = country,
      Outcome = outcome,
      n = n,
      `Observed prevalence` = ifelse(is.finite(brier) & !is.na(auc),
        sprintf("%.1f%%", 100 *
          (diag_binary$prevalence[match(paste(country, outcome),
                                          paste(diag_binary$country, diag_binary$outcome))])),
        "—")
    ) %>%
    kable(booktabs = TRUE)
}
```

Each biomarker was converted to a binary deficiency indicator using WHO
or BRINDA-recommended thresholds, with adjustment for inflammation where
acute-phase proteins were available [@brinda2017adjustment;
@who_vmnis_thresholds]. Survey weights from DHS were retained throughout
and used to compute admin-2 (district-equivalent) prevalence with
appropriate clustered variance estimation. Where available, sample
weights, primary sampling units and stratification were retained for
admin-2 prevalence estimation.

## Predictors

Two predictor sets were assembled. The **individual-level** set included
all proxy survey items recommended by Mertens et al. (under review) — a
harmonised harmonisable set of approximately 200 socio-demographic and
household indicators across the four DHS surveys (mother education,
household assets, water and sanitation, livelihoods, exposure to malaria
interventions). This set excludes biomarker survey items (the `gw_`
prefix in the analytic data dictionary) so that predictions can be made
in target populations that lack biomarker measurement.

The **area-level** set comprised over 2,500 earth-observation aggregates
extracted from Google Earth Engine [@gorelick2017google] for each
admin-2 polygon using exact-extract weighted by polygon overlap. Layers
included Worldpop population density, ESA WorldCover land use,
accessibility surfaces, NDVI/EVI temporal aggregates, atmospheric
parameters, elevation, and night-time lights. Variables present in fewer
than three countries' GEE extracts were excluded from cross-country
analyses, producing a common-vocabulary set of 152 variables.

## Estimation strategies

We compared estimation strategies spanning the machine-learning and
small-area-estimation traditions, organised into two families. The
**area-level small-area estimation (SAE)** family fits one row per admin-2
polygon, with survey-weighted prevalence as the target: an area-level
SuperLearner on the polygon-aggregated covariates; frequentist baselines (a
country-mean floor, OLS, and cross-validated elastic-net); Bayesian areal
models (the Fay–Herriot model [@fay1979estimates; @mercer2015small] and the
BYM2 disease-mapping model fit by integrated nested Laplace approximation
[@riebler2016intuitive; @rue2009approximate]); and a set of
transportability-aware comparators (a country random-intercept model, a
two-stage country-mean-plus-within-country-deviation model, a forward
selection that optimises a leave-one-training-country-out criterion, a
spatial-spline GAM, a data-source group-lasso, and a stacked ensemble of
these). The **individual-level SuperLearner aggregated to admin-2** fits a
person-level ensemble and averages its predictions up to the district; it
shares the same inputs but is reported as a sensitivity analysis, because
aggregating shrunken person-level probabilities biases area means (Results).

A subsequent methods-development analysis established that the area-level
SuperLearner's weak honest performance reflects *unregularised* fitting in the
p ≫ n regime rather than an inherent limitation of area-level modelling, and
that modelling the area prevalence directly is in fact the strongest approach
once the estimator is matched to the regime (Results). We therefore add an
**area-level prevalence recipe** as the primary district estimator. It models
the admin-2 survey-weighted prevalence directly under a **binomial (count) loss
weighted by the survey sample size** (which leaves district ranking unchanged
relative to a Gaussian logit fit but markedly improves calibration); reduces the
covariate space by a mild **global near-null pre-filter** (covariates whose
marginal association with the outcome falls in the lowest 70% are dropped — a
step validated as leakage-free below) followed by an in-fold supervised screen;
and fits a penalised regression (elastic net). The covariate set is
**regime-specific**: a malaria- and modelled-health-burden–enriched set (adding
Malaria Atlas Project falciparum/vivax layers and IHME anaemia/stunting surfaces
to the earth-observation covariates) for within-country gap-filling, and an
earth-observation-only ("universal") set for cross-country transport, because
the malaria and modelled-burden relationships are country-specific and degrade
transport. Two opt-in variants are provided: a domain-wise **late-fusion stack**
(one regularised learner per data source, stacked) for outcomes whose signal is
distributed across domains, and, for vitamin A, a **distributional** estimator
that models the continuous biomarker and integrates over the deficiency
threshold. The recipe and the evidence for each choice are documented in the
project's area-level recipe specification.

The full list of strategies with their loss functions and implementations is
given in Supplementary Appendix Table S1, and the complete modelling detail
(learner libraries and hyper-parameters, adjacency construction, post-hoc
calibration, outcome-aware feature augmentation, and the bootstrap interval
procedure) is in Supplementary Appendix Section S5.

## Honest evaluation and the corrected-methods analysis of record

Initial benchmarking surfaced several ways an analysis at this scale can
flatter itself, so the headline results below come from a set of corrected,
leakage-free methods (referred to as the P1–P8 methods) that constitute the
analysis of record; the original production-pipeline equivalents are retained
as a sensitivity comparison. In brief, the corrected methods (i) perform all
preprocessing inside each cross-validation fold and form folds by survey
cluster and by spatial region, so that no information leaks from held-out
areas; (ii) recalibrate binary predictions out-of-fold rather than in-sample;
(iii) separate the component of admin-2 prediction error that is attributable
to survey sampling noise from genuine model error; (iv) attach
split-conformal prediction intervals and design-based survey intervals to each
district; (v) score the decision-relevant task of correctly ranking the
highest-burden districts; (vi) flag districts whose covariates fall outside the
surveyed training range as out-of-support; and (vii) produce partial-pooling
(empirical-Bayes) area estimates. The full specification is in Supplementary
Appendix Section S6.

Two further safeguards apply to the area-level recipe. First, every reported
correlation carries an area-bootstrap 95% confidence interval and a
non-parametric permutation-null *p*-value (the permutation re-labels districts
within country), so signal can be distinguished from sampling noise at the small
effective sample sizes involved. Second, the global near-null pre-filter is
shown to be leakage-free by an honest permutation test that **re-derives the
filter on each permuted outcome**: the resulting null distribution of the
out-of-sample correlation remains centred at zero (95th percentile 0.17–0.35
across cells) while the observed correlations are 0.29–0.65 (permutation
*p* ≤ 0.02), confirming that dropping obvious near-null covariates on the full
data introduces negligible optimism. The empirical-Bayes / Fay–Herriot layer is
used for honest interval width only and not to shrink the point level, because
the covariates carry district ranking but not absolute level; the level is set
by anchoring to a known national prevalence.

## Evaluation

We evaluated two questions corresponding to two deployment regimes.
**Within-country** spatial accuracy (gap-filling unsurveyed districts in a
country with some survey data) was estimated by cross-validation on each
country's admin-2 set with metrics on the union of out-of-fold predictions; for
the area-level recipe we report both **district-block** cross-validation (whole
districts held out) and the more demanding **region-block** scheme (whole admin-1
regions held out), using the enriched covariate set. **Cross-country
transportability** (predicting a country with no biomarker survey) was assessed
by leave-one-country-out (LOCO) holdout on the pooled admin-2 dataset using the
universal covariate set and scored by rank correlation, which is invariant to
the untransportable absolute level: the model is trained on three countries'
admin-2 areas (n = 119–192) and evaluated on the fourth's (n = 14–87). Metrics reported are Pearson and Spearman
correlation, mean absolute and root-mean-squared error in percentage
points, mean bias in percentage points, and calibration slope from a
linear regression of observed on predicted prevalence.

For binary classifiers we additionally report area under the ROC curve
(AUC), Brier score and Brier skill score, expected and maximum
calibration error, and the calibration intercept and slope from a
Cox-style logistic recalibration. For continuous biomarker regressions
we report R² and Pearson r.

All analyses were executed via the `targets` R package
[@landau2021targets] for reproducible dependency-aware execution. Code
and pipeline definitions are available at
\textless<repository-URL placeholder>\textgreater.

# Results

## Sample characteristics

The pooled admin-2 dataset comprised 206 administrative units across the
four countries (Gambia 30, Ghana 75, Sierra Leone 14, Malawi 87). Sample
sizes per country and biomarker are presented in @tbl-cohort. Observed
admin-2 deficiency prevalence varied widely across outcomes and
countries, with iron deficiency in children ranging from 5% (Sierra
Leone) to 64% (Gambia), and vitamin-A deficiency in women remaining
below 5% in most countries.

## Within-country admin-2 estimation

Under genuinely out-of-fold evaluation, within-country admin-2 ranking
was modest, and the apparent superiority of the directly-fit area-level
SuperLearner did not survive honest cross-validation (@tbl-area-comparison).
The area-level SuperLearner reached a median out-of-fold Pearson r of only
0.12 (MSE loss; 0.14 for NLL) between predicted and survey-weighted admin-2
prevalence, with 0 of 24 evaluable country–outcome combinations exceeding
r ≥ 0.9 and roughly half at or below zero (range −0.57 to 0.61). In
contrast, aggregating the individual-level SuperLearner's out-of-fold
predictions to admin-2 ranked districts substantially better (median r
0.53, range −0.39 to 0.87), albeit with larger negative biases for iron.

This reversal reflects the area-level p ≫ n setting. Because the proxy
and earth-observation predictors are resolved at the admin-2 level, the
effective sample size for area-level fitting is the number of polygons —
14 in Sierra Leone, 30 in Gambia, 75 in Ghana, and 87 in Malawi — against
roughly 150 screened candidate covariates, leaving few residual degrees of
freedom. An earlier in-sample evaluation — predicting the training areas
from the fully-fitted ensemble — produced near-perfect correlations (r up
to 1.00), but these were overfitting artifacts rather than transportable
signal; the present numbers use genuine out-of-fold prediction. The
individual-level model, fit on hundreds to thousands of individuals,
generalises better once aggregated than the *unregularised* area-level
ensemble. We report the number of areas alongside individual sample sizes
throughout. Crucially, however, the area-level deficit is a property of the
*estimator*, not of area-level modelling: once the estimator is matched to the
p ≫ n regime — the area-level recipe described next — modelling the area
prevalence directly becomes the strongest within-country approach.

```{r tbl-area-comparison}
#| tbl-cap: "Within-country admin-2 prediction performance for three approaches: individual SuperLearner aggregated to admin-2, area-level SuperLearner with MSE loss, and area-level SuperLearner with NLL loss. Pearson r between predicted and observed admin-2 prevalence; MAE in percentage points. From results/tables/area_comparison_all.csv."
if (!is.null(area_compare)) {
  area_compare %>%
    select(country, outcome, approach, pearson_r, mae_pp, mean_bias_pp) %>%
    arrange(country, outcome, approach) %>%
    kable(booktabs = TRUE,
          col.names = c("Country","Outcome","Approach","r","MAE (pp)","Bias (pp)"))
}
```

In contrast, the aggregated individual-level SuperLearner exhibited a
characteristic compression pattern: predictions correlated moderately
with admin-2 prevalence (median r ≈ 0.50) but were systematically pulled
toward the country mean. For high-prevalence outcomes this produced
large negative biases of 8 to 20 percentage points (e.g., Gambia
child-iron MAE 20.8 pp, bias −20.4 pp; women-iron MAE 18.0 pp, bias
−18.0 pp). The mechanism is straightforward and consistent with the
small-area-estimation literature [@mercer2015small;
@wakefield2019small]: a probabilistic individual-level classifier under
uncertainty shrinks toward the marginal class probability, and naive
aggregation propagates this shrinkage to area-level means. The recipe
to avoid it — *fit at the unit of deliverable* — is the
small-area-estimation lesson.

## The area-level prevalence recipe

Fitting at the unit of deliverable, done correctly, is the strongest
within-country estimator. The area-level recipe — admin-2 prevalence under a
binomial count loss, a leakage-free near-null pre-filter followed by an in-fold
supervised screen, and a penalised regression on the malaria/IHME-enriched
covariate set — substantially exceeds both the unregularised area-level
SuperLearner (median r 0.12) and the individual-level aggregate for the
outcomes that carry signal (@tbl-area-recipe). Under district-block
cross-validation the recipe reached Pearson r = 0.64 for child iron and 0.56 for
women's vitamin B12 in Ghana, 0.60 for women's iron in Gambia, with
permutation-null *p* < 0.05 for iron and several women's micronutrients in
Gambia and Ghana; under the more demanding region-block scheme the same cells
retained r = 0.20–0.55. The binomial count loss left district ranking
essentially unchanged relative to a Gaussian logit fit but markedly improved
calibration (slopes 0.5–1.1, versus systematic over-shrinkage under the Gaussian
loss). Enriching the earth-observation covariates with Malaria Atlas Project and
IHME anaemia/stunting layers recovered signal in the highest-malaria setting:
Malawi child-iron rose from null with soil/climate features alone to r = 0.20
(district, *p* = 0.03). The two opt-in variants behaved as designed — the
distributional estimator improved vitamin A (Gambia child vitamin A r 0.18 →
0.46; Gambia women vitamin A r 0.42), while domain-wise late-fusion stacking
helped distributed-signal outcomes (Gambia women iron r 0.65) but collapsed the
sparse Malawi child-iron signal (0.20 → 0.03), confirming that dimension
reduction must not be a default. Across all 18 evaluable country–outcome cells,
genuine permutation-significant within-country signal was concentrated in iron
(and several women's micronutrients) in Gambia and Ghana, was weaker but present
for iron and folate in Malawi, and was absent for zinc.

```{r tbl-area-recipe}
#| tbl-cap: "Area-level recipe within-country performance (binomial loss, enriched features), district- and region-block cross-validation. Pearson r with area-bootstrap 95% CI and permutation-null p; calibration slope. From results/tables/corrected/area_recipe_within.csv."
.arp <- "results/tables/corrected/area_recipe_within.csv"
if (file.exists(.arp)) {
  read.csv(.arp) %>%
    transmute(Country = country, Outcome = outcome, Scheme = scheme, n = n_area,
              r = round(pearson_r, 2),
              `95% CI` = sprintf("[%.2f, %.2f]", pearson_ci_lo, pearson_ci_hi),
              `perm p` = round(pearson_perm_p, 3), Slope = round(calib_slope, 2)) %>%
    arrange(Country, Outcome, desc(Scheme)) %>%
    kable(booktabs = TRUE)
}
```

## Cross-country transportability

LOCO holdout produced markedly weaker performance (@tbl-loco;
@fig-loco-scatter). Pearson r across the 16 outcome–holdout
combinations had a median of 0.21 and ranged from −0.14 to 0.53 (one
holdout — women vitamin A in Gambia — was unestimable because no
covariate survived selection). Iron deficiency outcomes transported best
(median r 0.28), while vitamin-A deficiency in women essentially did not
transport across countries (median r ≈ −0.12, with anti-predictive
ordering in two of three estimable holdouts).

The area-level recipe, applied with the universal (earth-observation-only)
feature set and scored by rank correlation, modestly improved transport but did
not overturn this picture. Iron deficiency transported to the West-African
held-out countries (Spearman 0.42 and 0.49 for child iron, 0.75 and 0.35 for
women's iron, in Gambia and Ghana respectively; all permutation *p* ≤ 0.01) but
not to Malawi (Spearman ≈ 0.03–0.05, non-significant), and vitamin A did not
transport in any country. The malaria/IHME enrichment that aided within-country
prediction *degraded* transport (mean child-iron Spearman 0.30 with universal
features versus 0.20 enriched), because the malaria–iron and modelled-burden
relationships are country-specific rather than universal. Cross-country
prediction is therefore a ranking-only product, valid within an
epidemiologically similar neighbourhood and requiring national-anchor
calibration to recover the absolute level.

```{r tbl-loco}
#| tbl-cap: "Leave-one-country-out (LOCO) transportability of the area-level SuperLearner. From results/tables/transportability_area_loco_metrics.csv."
if (!is.null(loco_metrics)) {
  loco_metrics %>%
    select(outcome, held_out, n_train, n_test, pearson_r, spearman_r,
           mae_pp, nat_bias_pp, calib_slope) %>%
    kable(booktabs = TRUE,
          col.names = c("Outcome","Held out","n train","n test",
                          "r","ρ","MAE (pp)","Bias (pp)","Calib slope"))
}
```

```{r fig-loco-scatter}
#| fig-cap: "Observed vs predicted admin-2 deficiency prevalence under leave-one-country-out cross-validation, by outcome and held-out country. The reference line is y = x. Predicted prevalences are from the area-level SuperLearner trained on the three non-held-out countries."
# Generated by scripts/fig_loco_scatter.R from
# results/tables/transportability_area_loco_predictions.csv. Panel Pearson r
# matches @tbl-loco (same unweighted definition).
knitr::include_graphics(
  here::here("results", "figures", "loco_scatter.png"),
  error = FALSE)
```

Spatial patterns transport more readily within the West African
geography. Holding out Malawi — the only training country outside West
Africa — produced anti-correlated predictions for three of four
outcomes, suggesting a regional generalisation gap. Held-out Ghana
provided the most favourable LOCO performance (median r 0.36),
plausibly because its admin-2 covariate distribution is well-represented
in the remaining three countries' training data.

National-level aggregation is consistently more accurate than admin-2.
National prevalence estimates were within ±5 percentage points of
observed for all 24 country–outcome combinations
(@tbl-national-estimates), with a median absolute difference of 0.4
percentage points (median relative difference 3.8%). The
within-country individual-level SL therefore remains useful at the
national level even where it fails at admin-2 resolution.

However, when the model is *transported* across countries the held-out
national estimate is systematically biased toward the pooled training
mean (@fig-transport-calibration). The bias is outcome-specific and
concerns the national *level*, not the within-country admin-2 ranking
of @tbl-loco. Two outcomes transport with broadly correct calibration:
vitamin-A deficiency, whose national prevalence sits close to the
cross-country mean in every survey (per-outcome calibration slope 0.93
in children, 0.65 in women), and — once its cutoff is harmonised to the
common serum definition (folate < 10 nmol/L; Malawi's native < 3 ng/mL
rule is otherwise non-comparable) — folate deficiency, whose held-out
prevalence preserves the correct ordering though compressed in magnitude
(slope 0.37; Ghana 55%→18%, Sierra Leone 79%→30%, Malawi 2%→1%). The
remaining outcomes fail. Iron deficiency, whose national prevalence
ranges from 6% (Sierra Leone children) to 64% (Gambia children), is
pulled to the training mean so strongly that the ordering inverts
(per-outcome slope −0.72 in children, −0.18 in women), over-predicting
low-prevalence countries (Sierra Leone children 6%→40%, Malawi 10%→51%)
and under-predicting high-prevalence ones (Gambia children 64%→22%).
Vitamin B12 — despite being the strongest *within-country* signal
(continuous cross-validated R² up to 0.23 in Malawi) — does not
transport at all: held-out discrimination falls to near chance
(AUC 0.53–0.59) and prediction becomes unstable under the rare-outcome
class-weighting (held-out Malawi predicted at 80% against a true 2.8%).
Cross-country transport failure is therefore a level-calibration error
driven by between-country prevalence heterogeneity — most severe for the
outcomes and countries that differ most from the training distribution —
and within-country accuracy is a poor guide to which outcomes transport.

```{r fig-transport-calibration}
#| fig-cap: "Cross-country transport calibration. Predicted versus true national deficiency prevalence for the pooled individual-level SuperLearner under leave-one-country-out (LOCO) holdout — one panel per outcome, one point per held-out country (trained on the remaining countries), panels ordered worst to best calibration. The dashed line is perfect calibration (y = x); each panel header reports the fitted calibration slope (1.0 = perfectly calibrated; <1 = predictions compressed toward the cross-country mean; <0 = inverted ordering). Vertical segments mark each fold's calibration gap. Vitamin-A national prevalence, close to the cross-country mean, is well recovered and folate (harmonised to a common serum < 10 nmol/L cutoff) preserves the correct ordering; iron is pulled to the training mean and inverts, and vitamin B12 does not transport (near-chance discrimination, with unstable prediction under rare-outcome class-weighting). Folate and B12 are women-only and available in three countries; zinc is Malawi-only and cannot be assessed for transport. From results/tables/transportability_loco_results.csv."
knitr::include_graphics(
  here::here("results", "figures", "transport_calibration_loco_faceted.png"),
  error = FALSE)
```

**Parsimonious model with soil micronutrient predictors.** A series of
post-hoc analyses confirmed that the strongest transportable signal in
the existing 152-variable GEE feature pool is concentrated in the
SoilGrids covariates. Univariate leave-one-country-out scoring of each
of the 152 GEE features identified eight SoilGrids variables (soil
aluminium and calcium standard deviation at two depth profiles, soil
magnesium standard deviation, soil zinc mean and standard deviation,
soil total carbon mean) as the most consistently transportable single
predictors; no NDVI, accessibility, night-lights, or population-density
variable cracked the top eight in the per-variable LOCO ranking. A
weighted generalised additive model with a thin-plate spline on
admin-2 polygon centroid coordinates plus these eight soil features as
linear terms achieved mean LOCO Pearson r = 0.330 across 16
outcome–holdout combinations on the four main outcomes — substantially
above any other parsimonious comparator we tested and roughly
comparable to the area-level SuperLearner ensemble at a fraction of
the compute cost. Bootstrap 95% confidence intervals on the
combination's Pearson r were strictly above zero for seven of the
sixteen holdouts, with the wins concentrated on iron and vitamin A
in Gambia and Ghana. We caveat that the eight-feature recipe and the
spatial gradient information may not generalise outside the West
African training footprint — extending the training set to East and
Southern Africa is the most important external-validity step.

**Sierra Leone testing power at admin-3 resolution.** The admin-2
Sierra Leone LOCO sample (n = 14 districts) is small enough that the
holdout estimates are dominated by sampling noise; we documented this
by re-running the analysis at admin-3 (chiefdom) resolution after
spatial-joining DHS cluster coordinates to GADM admin-3 polygons. At
admin-3 resolution, Sierra Leone has 44 chiefdoms with biomarker
measurements (a 3.1-fold increase over admin-2). With Sierra Leone
held out as the test country and the other three countries as training
at admin-2 resolution, LOCO Spearman correlations on the four
continuous biomarker outcomes (child / women vitamin A, women B12,
women folate) moved from a range of −0.07 to −0.35 at admin-2 toward
or above zero at admin-3, including a meaningfully positive +0.30 for
women's vitamin B12. We interpret the headline admin-2 anti-predictive
estimates for Sierra Leone as primarily a small-sample artefact and
not as evidence of fundamental non-transportability.

```{r tbl-national-estimates}
#| tbl-cap: "National prevalence: model-derived vs survey-derived. From results/tables/national_estimates_all.csv."
if (!is.null(natl_est)) {
  natl_est %>%
    transmute(Country = country, Outcome = outcome,
              `Observed (95% CI)` = sprintf("%.1f (%.1f–%.1f)%%",
                obs_prev*100, obs_lo*100, obs_hi*100),
              `Predicted` = sprintf("%.1f%%", pred_prev*100),
              `Abs diff (pp)` = round(abs_diff*100, 1)) %>%
    kable(booktabs = TRUE)
}
```

## Benchmark comparison against canonical methods

Five comparator methods were evaluated alongside the area-level
SuperLearner: country-mean baseline, OLS with up to 8 covariates,
elastic-net penalised regression, Fay–Herriot, and the BYM2 disease
mapping model fit in INLA (@tbl-benchmarks). Results across the four
primary LOCO outcomes indicate that the prescreened SuperLearner
ensemble does not consistently outperform simpler methods, and in fact
loses to them on most splits. Among the 13 outcome–holdout splits for
which both the prescreened ensemble and at least one canonical benchmark
(country-mean baseline, OLS, elastic-net, Fay–Herriot, or BYM2) were
estimable, the best benchmark method beat the ensemble on 10 and the
ensemble won 3 by Pearson r. The most dramatic disagreement was for
child iron deficiency with Sierra Leone held out, where the ensemble
produced anti-predictive ordering (r = −0.19) while a full-covariate
OLS recovered the signal (r = 0.51).

The ensemble's few wins were concentrated on outcomes with non-linear
spatial signal — most clearly women-iron deficiency with Gambia held out
(ensemble r = 0.59, versus 0.51 for elastic-net and −0.03 for OLS) and
with Ghana held out (ensemble r = 0.40 versus 0.27 for elastic-net) — and
in these cases ranger was the winning base learner in the ensemble.
Bayesian SAE models (Fay–Herriot, BYM2)
provided small additional gains in calibration but did not
substantially shift the Pearson r ranking, consistent with the
observation that these methods derive most of their advantage from
within-country spatial smoothing, which is partially lost in LOCO
holdout when the held-out country contributes no observed data to the
graph.

```{r tbl-benchmarks}
#| tbl-cap: "LOCO transportability: prescreened SuperLearner vs benchmark methods. Pearson r between predicted and observed admin-2 prevalence. From results/tables/benchmarks_all.csv."
if (!is.null(bench_results)) {
  bench_results %>%
    filter(eval_type == "loco", model_type == "continuous") %>%
    pivot_wider(id_cols = c(outcome, held_out),
                names_from = method, values_from = pearson_r) %>%
    kable(booktabs = TRUE, digits = 3)
} else {
  cat("Benchmark results not yet available — placeholder. See preliminary",
      "smoke-test results in the discussion.")
}
```

## Supplementary all-methods comparison on a representative outcome

To provide a comprehensive supplementary comparison, we ran every method
and SuperLearner library variant tested in this work on a single
representative outcome — child iron deficiency, chosen because it has
the strongest transportable signal in our data and biomarker
measurements in all four countries. Two evaluation tracks were used:
within-country five-fold cross-validation on each country's admin-2
polygons (separately per country) and leave-one-country-out
transportability across the four-country pool (@fig-all-methods-iron).

```{r fig-all-methods-iron}
#| fig-cap: "Supplementary all-methods comparison for child iron deficiency. (a) Within-country 5-fold cross-validation Pearson r per method, with one dot per country and X-marks denoting the mean across the four countries. (b) Leave-one-country-out transportability Pearson r per method, with one dot per held-out country. Methods are ordered by mean LOCO r. The supplementary figure source data are in results/tables/all_methods_child_iron.csv."
#| fig-width: 11
#| fig-height: 6
fig_path <- here::here("results","figures","all_methods_child_iron.png")
if (file.exists(fig_path)) knitr::include_graphics(fig_path)
```

For LOCO, spatial smoothing on admin-2 polygon centroids (mean Pearson r
= 0.48) and spatial + top-8 SoilGrids covariates (0.46) were the two
strongest methods, both substantially outperforming the prescreened
SuperLearner ensemble (0.22) and the simpler invariance-filter and OLS
recipes (0.39 each). Within-country, the picture flips: nonparametric
flexibility helps when training data is in-distribution, with random
forest (mean Pearson r = 0.23 across four countries) and the
prescreened SuperLearner (0.17) outperforming the linear methods, while
spatial-only and spatial + soil predict best on the two countries with
enough admin-2 polygons (Ghana and Malawi) but are unstable on the
small (n = 14) Sierra Leone CV split. The conclusion is consistent
across outcomes (see Methods §Estimation strategies and the exploratory
findings logs, `docs/exploratory_loco_findings.md`): non-linear ensembles help
within-country, parsimonious linear models with soil micronutrient features
help out-of-country.

## Calibration and reliability of binary classifiers

For binary deficiency classifiers, several diagnostic flags emerged
(@tbl-calibration). The Brier skill score was negative — indicating
predictions worse than the no-skill baseline — for 13 of 23
country–outcome combinations, though in most the deficit was small
(Brier skill > −0.04 for 10 of the 13). Post-hoc Platt recalibration
(see Methods) eliminated the negative values entirely: all 23 calibrated
classifiers achieved non-negative Brier skill. Calibration slopes were
strongly inverted
(< 0) for several models in Malawi, Sierra Leone, and Ghana,
particularly at very low prevalence (vitamin A in women, vitamin B12 in
women) where AUC remained high (> 0.7) but Brier skill was near zero.
This pattern — high AUC concealing poor calibration — is a known
artefact of imbalanced binary classification and we therefore
recommend that decision thresholds derived from these models be
calibrated against observed prevalence before deployment
[@niculescu2005predicting; @vansmedenetal2019sample].

```{r tbl-calibration}
#| tbl-cap: "Binary classifier calibration. Brier skill > 0 indicates information gain over a no-skill baseline. From results/tables/diagnostics_binary.csv."
if (!is.null(diag_binary)) {
  diag_binary %>%
    transmute(Country = country, Outcome = outcome,
              `Prev` = sprintf("%.1f%%", prevalence*100),
              AUC = sprintf("%.2f", roc_auc),
              `Brier skill` = sprintf("%+.3f", brier_skill),
              `Calib slope` = sprintf("%+.2f", calib_slope),
              ECE = sprintf("%.3f", ece)) %>%
    kable(booktabs = TRUE)
}
```

## Honest evaluation: the corrected analysis of record

The metrics in this section are the analysis of record, produced by the
leakage-free corrected (P1–P8) methods described in Methods; the
production-pipeline equivalents elsewhere are retained for comparison.

**Within-country discrimination shrinks under honest cross-validation.**
Replacing random cross-validation and full-data preprocessing with
cluster-blocked and spatial-block folds and strictly in-fold preprocessing
lowers within-country discrimination, confirming that the optimistic design
overstated performance (@tbl-cv-honesty). Across the four primary outcomes the
honest (cluster-block) AUC is typically a few points below the optimistic AUC
and frequently near chance, consistent with the modest area-level signal seen
throughout.

```{r tbl-cv-honesty}
#| tbl-cap: "Within-country binary discrimination (AUC) under optimistic versus honest cross-validation, from the corrected methods (results/tables/corrected/cv_honesty_compare.csv). Honest schemes block folds by survey cluster and by spatial region and confine all preprocessing to the training fold."
if (!is.null(corr_cv)) {
  lab <- c("random CV + full-data preprocessing (optimistic)" = "Optimistic",
           "cluster-block CV"                                 = "Cluster-block (honest)",
           "region/spatial-block CV"                          = "Spatial-block (honest)")
  corr_cv %>%
    mutate(scheme = lab[scheme]) %>%
    filter(!is.na(scheme)) %>%
    select(country, outcome, scheme, auc) %>%
    tidyr::pivot_wider(names_from = scheme, values_from = auc) %>%
    arrange(country, outcome) %>%
    kable(booktabs = TRUE, digits = 2,
          col.names = c("Country", "Outcome", "Optimistic",
                        "Cluster-block (honest)", "Spatial-block (honest)"))
}
```

**Out-of-fold calibration.** Recalibrating binary predictions out-of-fold
(rather than in-sample) gives an honest picture of calibration; the in-sample
Platt fit is optimistic by construction (slope forced toward 1). Raw and
out-of-fold-calibrated slopes are reported per outcome in
`results/tables/corrected/calibration_oof_compare.csv`.

**Admin-2 error, separating sampling noise from model error.** A substantial
part of the apparent district-level error is survey sampling noise rather than
model error. After removing the sampling-variance component, the
sampling-adjusted RMSE is the honest measure of how far model predictions sit
from the true district prevalence (@tbl-corr-error).

```{r tbl-corr-error}
#| tbl-cap: "Admin-2 prediction error decomposed into a survey-sampling component and a model component (results/tables/corrected/admin2_error_sampling_adjusted.csv). Naive RMSE is the raw predicted-vs-survey error; the sampling component is what survey noise alone would produce; the adjusted RMSE is the honest model error."
if (!is.null(corr_err)) {
  corr_err %>%
    select(country, outcome, n_admin2, naive_rmse_pp, sampling_rmse_pp,
           adjusted_rmse_pp, pearson_r) %>%
    kable(booktabs = TRUE, digits = 2,
          col.names = c("Country", "Outcome", "n areas", "Naive RMSE (pp)",
                        "Sampling RMSE (pp)", "Adjusted RMSE (pp)", "r"))
}
```

**Decision value for targeting.** Because the operational task is to find the
highest-burden districts, we score how well the predictions concentrate true
burden in the top-ranked districts. Targeting the top 20% of districts by
predicted prevalence reaches more of the true burden than untargeted allocation
(lift > 1) for most outcomes, with the largest lift on the outcomes that have
the most spatial signal (@tbl-decision).

```{r tbl-decision}
#| tbl-cap: "Decision value of district ranking (results/tables/corrected/decision_value.csv). reach@20% is the share of true burden captured by the top 20% of districts ranked by predicted prevalence; lift is that reach divided by the 20% an untargeted allocation would reach; hit-rate@top20 is the share of the predicted top-20% districts that are truly in the top 20%."
if (!is.null(corr_decision)) {
  corr_decision %>%
    select(country, outcome, reach_at_20pct, lift_vs_no_targeting, hit_rate_top20) %>%
    kable(booktabs = TRUE, digits = 2,
          col.names = c("Country", "Outcome", "Reach@20%", "Lift", "Hit-rate@top20"))
}
```

**Prediction intervals are reasonably calibrated.** Split-conformal intervals
attached to each surveyed district achieve empirical held-out coverage close to
the 90% target (range across outcomes typically 0.85–1.00;
`results/tables/corrected/interval_summary.csv`), at the cost of wide intervals
for the noisiest outcomes.

**Per-district trust flags.** Each district is flagged for reliability by three
criteria: whether its covariates fall outside the surveyed training range
(out-of-support), small survey sample, and a wide prediction interval. The
resulting OK-to-use versus do-not-rely labels are in
`results/tables/corrected/trust_flags.csv` and drive the dashboard's certainty
layer; out-of-support is the dominant driver of do-not-rely flags in
unsurveyed districts.

# Discussion

## Summary

We harmonised individual-level biomarker measurements from four
sub-Saharan African DHS (three West African plus Malawi) to evaluate
whether ensemble machine-learning models can
reproduce admin-2 micronutrient deficiency prevalence. Three findings
have programmatic significance.

First, within-country admin-2 accuracy is modest once evaluated
out-of-fold, and—contrary to in-sample appearances—aggregating
individual-level predictions to admin-2 (median Pearson r 0.53)
outperforms fitting an ensemble *directly at the admin-2 level* (median
r 0.12, MSE-loss variant), which overfits badly in the area-level p ≫ n
setting (14–87 areal units against ~150 screened covariates). The
individual-level aggregate still shows compression toward the country mean
and negative biases of up to 20 percentage points for iron, so neither
approach delivers the near-perfect ranking that in-sample evaluation had
suggested. This qualifies the usual small-area-estimation expectation that
the unit of estimation should match the unit of deliverable
[@mercer2015small; @wakefield2019small]: with so few areal units, the
data-richer individual-level fit generalises more reliably than a model
trained directly on the areas.

Second, cross-country transportability is limited. Trained on three West
African countries and tested on a fourth, the SuperLearner achieved
median Pearson r of 0.21 across 16 outcome–holdout combinations, with
absolute errors of 1.5–44 percentage points and frequent anti-predictive
ordering. Iron-deficiency outcomes were the most transportable; vitamin-A
deficiency was not transportable. Malawi, the only training country
outside West Africa, produced anti-correlated predictions when held out
across three of four outcomes — a regional generalisation gap that should
caution against using these models for spatial targeting in unseen
geographies without local validation.

Third, the SuperLearner does not reliably outperform standard SAE methods.
A plain 8-covariate OLS or an elastic-net penalised regression matched
or beat the ensemble on a majority of LOCO splits. The ensemble's
advantages, when present, concentrated in specific outcome–country
combinations where ranger captured genuine non-linear spatial signal.
Bayesian SAE comparators provided modest additional calibration gains,
particularly for low-prevalence outcomes where shrinkage to area-level
random effects stabilised estimates.

## The aggregation problem and "fit at the unit of deliverable"

The mechanism by which individual-level predictions degrade upon
aggregation merits explicit treatment. A well-calibrated probabilistic
classifier under high noise shrinks individual-level predictions toward
the marginal class probability — this is the correct loss-minimising
behaviour [@niculescu2005predicting]. When such predictions are
aggregated across all individuals in an admin-2 unit, the shrinkage
does not cancel; it pulls every area-level mean toward the country
prevalence. In high-prevalence settings this produces large negative
biases; in low-prevalence settings, positive biases. Calibration of
admin-2 outputs against survey-weighted admin-2 prevalence is therefore
necessary if individual-level models are repurposed as area-level
maps. We follow the small-area-estimation tradition in recommending the
opposite path — fitting at the admin-2 unit directly — when the
deliverable is spatial.

This conclusion is consistent with @mercer2015small and
@wakefield2019small, who emphasise that for DHS-derived admin-level
prevalence the natural model is one fit at the admin level with known
sampling variances, structured (e.g., ICAR or BYM2) random effects,
and auxiliary covariates. Our finding that BYM2 and Fay–Herriot are
competitive with — and on some outcomes superior to — a multi-learner
ensemble is consistent with this tradition.

## Why does the SuperLearner sometimes underperform?

The most striking single result is the child-iron Sierra Leone LOCO
case, where the SuperLearner produces anti-predictive ranking
(r = −0.19) while a plain OLS achieves r = 0.51. We interpret this as
a model-selection failure of the ensemble: with only 14 admin-2 areas
in the held-out set and approximately 152 covariates, the cross-validated
risk estimator within the ensemble has high variance, and the cv-selected
learner (ranger in this case) appears to have over-fit to a
non-transportable feature pattern. Simpler regularised models with
implicit variable shrinkage are more robust under these conditions.

This suggests that ensemble methods should be benchmarked against
parsimonious alternatives whenever the test set is small or the target
population is meaningfully outside the training distribution. The
17.7-hour computational cost of our SuperLearner pipeline relative to a
~1-minute OLS baseline is justified only on a subset of country–outcome
combinations.

## Implications for programming

Admin-2 deficiency maps derived from this approach are interpretable as
*relative ranking* tools within a country with biomarker data — useful
for prioritising districts within a national-level intervention. They
are not suitable for absolute-level inference (calibration slopes drift
substantially) without additional anchoring against direct national
estimates [@stevens2022micronutrient]. For an unsurveyed country, our
results suggest that LOCO-based predictions should be treated as coarse
risk triage at most, and only within the same broad agroecological
region as the training set.

The VMNIS and BRINDA national-level databases [@who_vmnis;
@brinda2017overview] offer one path to external validation. A natural
next step is to apply the best-LOCO parsimonious model — a penalised
regression on area-level GEE aggregates — to additional African
countries with biomarker data already collated in BRINDA/VMNIS, and
to estimate calibration and bias of admin-2 maps in those populations.
A companion analysis using earth-observation rasters for Côte d'Ivoire
(available in the project repository) is in progress.

## Limitations

Our sample comprises four sub-Saharan African countries — three in West
Africa plus Malawi — so the rest of East, Central, and Southern Africa is
unrepresented. Because the predictors are resolved at admin-2, the
effective sample size for area-level models is the number of polygons
(14–87 per country) rather than the underlying individuals, which limits
the degrees of freedom for both within- and cross-country fitting and
makes the within-country cross-validated correlations an optimistic bound
(most acutely for Sierra Leone, with only 14 polygons). We did not formally
quantify the contribution of survey design weights to admin-2 prevalence
estimates; assuming a fixed design-effect of 1.5 (DHS-typical) is an
approximation. We restricted the cross-country common-vocabulary GEE
covariate set to 152 variables, and other satellite-derived predictors
not present in all four countries (notably finer-grained land-use
classes) were excluded. Finally, our raw binary classifiers' calibration
diagnostics flagged 13 of 23 outcomes with negative Brier skill; post-hoc
recalibration restored all of them to non-negative skill, but the
underlying discrimination at very low prevalence remains weak (several
AUCs near 0.5) and these outcomes should be treated as exploratory.

## External validation pathway

The findings presented here are based on a largely West African LOCO
panel (three West African countries plus Malawi) and a single external
held-out country (Côte d'Ivoire). We have
prepared a multi-source national-level benchmark to evaluate future
expansions of the model to East and Southern Africa, comprising: (i) the
Kenya National Micronutrient Survey 2011 [@knms2011], which provides
nationally-representative biomarker prevalence for vitamin A, iron,
zinc, folate, B12 and iodine across pre-school children, school-age
children, pregnant women, non-pregnant women, and men; (ii) modelled
WHO/UNICEF joint anaemia prevalence estimates for all eight focus
countries, retrieved via the WHO Global Health Observatory API; and
(iii) WHO Vitamin and Mineral Nutrition Information System (VMNIS)
historical survey-level prevalence for the same eight countries, from
the latest publicly-released long-format dump. The Stevens et al. 2022
*Lancet Global Health* pooled posterior estimates [@stevens2022micronutrient] are the
natural comparator for our national-level predictions; however, the
country-level Bayesian posteriors are not publicly released, only the
22 input surveys used to fit the model, and none of our cohort countries
are among them. The full assembled benchmark is in
`results/external/{vmnis_national,who_anaemia_country,knms2011_national_summary,kenya_national_benchmark}.csv`.

Three concrete next-step datasets, each requiring a Data Use Agreement
with the originating ministry, would substantially expand the LOCO panel
beyond West Africa: the Kenya National Micronutrient Survey (KNMS 2011),
the Tanzania National Nutrition Survey 2018, and the Ethiopia National
Micronutrient Survey 2016. All three include the full vitamin A / iron /
zinc / folate / B12 / iodine biomarker panel collected under the BRINDA
inflammation-adjustment framework, making harmonisation with our four
training countries straightforward. As a lighter-weight alternative,
the most recent DHS rounds with haemoglobin biomarker (Kenya DHS 2022,
Tanzania DHS 2022, Burkina Faso DHS 2021) can be added with no further
paperwork beyond the user's existing DHS account, providing iron-deficiency
anaemia coverage for three additional East/Sahelian countries. We
provide country-agnostic ingestion stubs (`R/ingest_new_country.R`,
function `stub_new_country_config()`) for each of these candidates;
once a harmonised survey file lands in `data/IPD/<country>/`, the entire
LOCO, calibration, SHAP, and dashboard pipeline regenerates automatically.

## Future methodological directions

The results above locate the limits of the current approach precisely — modest
within-country ranking, transport failure that is predominantly a national-*level*
calibration error rather than a ranking error, and an ensemble that does not beat
parsimonious SAE baselines — and these limits motivate several planned
extensions. We describe them here as directions rather than results.

First, **decision-focused estimation.** For programmatic use the relevant target
is rarely the exact prevalence of a district but whether it lies above a
public-health action threshold, and whether the highest-risk districts are
correctly prioritised. Because the SuperLearner's base learners optimise
likelihood or squared-error loss while the ensembling (or discrete-selection)
step may target any criterion [@vanderlaan2007super], the ensemble can be tuned
directly for an aggregate, asymmetric objective — for example maximising recall of
the truly high-risk areas (top-quantile or above-threshold) subject to a bounded
false-positive rate. This reframes the deliverable toward the ranking signal that,
as shown above, transports more reliably than absolute level, and is
statistically less demanding than level estimation.

Second, **anchoring to known national prevalence.** Because transport failure is
concentrated in the national level rather than the within-country ordering
(@fig-transport-calibration), a one-parameter level recalibration that constrains
a transported prediction's population-weighted mean to a known or historical
national value (e.g. from VMNIS or the BRINDA pooled analyses
[@who_vmnis; @brinda2017overview]) should remove much of the LOCO level bias by
construction, leaving the ranking unchanged. Such anchors are obtainable for many
countries that lack the individual-level data needed to fit a model.

Third, **borrowing strength across co-occurring deficiencies.** The micronutrient
outcomes are biologically correlated but were modelled independently here.
Modelling them jointly through a forward factorisation of their conditional
density — predicting the most-informative outcome first (iron, measured in all
four countries) and supplying predicted upstream deficiencies, with
missing-indicator encoding and imputation, as covariates for subsequent outcomes
— is a standard route to exploiting that covariance and may stabilise the
data-poorest outcomes (zinc, B12). We expect within-country gains to exceed
transport gains, since transported upstream predictions themselves carry error.

Fourth, **inference and value-of-information.** Valid uncertainty on the aggregate
deliverable (national or admin-level prevalence) — via cross-validated
loss-based inference, conformal prediction intervals, or the area-resampling
bootstrap already reported — enables a complementary product that is robust to the
data being imperfect: a value-of-information map that ranks candidate survey
locations by the product of predicted risk and predicted uncertainty, identifying
where a new biomarker survey would be most informative. For a field constrained by
the cost and rarity of biomarker measurement, guiding *where to measure next* may
be as valuable as the prevalence estimates themselves.

Finally, two lower-level refinements follow from the analysis: fitting
area-prevalence learners under a binomial (rather than Gaussian) loss to respect
the bounded, mean-variance-linked response, and adding a principal-components
learner suited to the many correlated predictors spanning distinct domains
(climate, soil, poverty). The most consequential validation step, however, is to
re-run the entire methodology on anaemia (haemoglobin), which is measured in
nearly every sub-Saharan African DHS at large sample size: a trusted,
abundant outcome would separate methodological limitations from the disputed
comparability of the micronutrient biomarkers themselves, and would expand the
cross-country panel far beyond the four countries available here.

## Conclusions

Within countries with biomarker data, admin-2 spatial ranking is only
modestly accurate when evaluated out-of-fold; aggregating individual-level
SuperLearner predictions outperforms fitting an ensemble directly on
admin-2 aggregates, which overfits given the small number of areal units.
Across countries, transportability is
limited and ensemble methods do not consistently outperform standard
small-area-estimation approaches. We recommend that within-country
performance be reported only from out-of-fold (not in-sample) predictions,
that future deployments of such models be benchmarked against Bayesian and
frequentist SAE methods, and that absolute-level predictions in unseen
countries be anchored against direct national estimates before use in
resource-allocation decisions.

# Acknowledgements

Funded by the Bill & Melinda Gates Foundation (grant number INV-XXXXXX).
The funder had no role in study design, analysis, decision to publish,
or manuscript preparation.

# Author contributions

ANM conceived the study, conducted the analysis, and drafted the
manuscript. Co-authors to be added on revision.

# Data availability

DHS biomarker data are available from the DHS Program
(<https://dhsprogram.com/>) following registration. Pipeline code,
including the area-level SuperLearner, frequentist baselines, and
INLA-based BYM2 implementation, are available at
\textless<repository-URL placeholder>\textgreater.

# References

::: {#refs}
:::
