---
title: "Supplementary Appendix: Subnational micronutrient deficiency prediction in sub-Saharan Africa from proxy indicators"
shorttitle: "Supplementary Appendix"
author:
  - name: Andrew N. Mertens
    affiliation: University of California, Davis, Institute for Global Nutrition
date: today
format:
  docx:
    number-sections: true
    toc: true
    toc-depth: 3
    fig-cap-location: bottom
    tbl-cap-location: top
execute:
  echo: false
  warning: false
  message: false
bibliography: references.bib
---

```{r setup}
suppressPackageStartupMessages({
  library(dplyr); library(knitr); library(here); library(readr)
})
```

This appendix documents, in full, the methods summarised at a high level in the
main text. It covers the data sources, the data cleaning and linkage pipeline,
the harmonisation of outcome definitions across surveys, the modelling methods,
the procedures used to assess predictive accuracy, and the sensitivity analyses.
It reflects the analysis as currently implemented in the project's reproducible
`targets` pipeline. The most granular file-by-file data lineage (exact source
filenames, fuzzy-match thresholds, and per-domain missingness rates) is
maintained alongside this appendix in `docs/data_sources_appendix.qmd`, which
this document summarises and extends.

# S1. Data sources

## S1.1 Biomarker (outcome) surveys

Outcomes come from four nationally representative micronutrient or
micronutrient-containing biomarker surveys, one per country, each conducted in a
different year and with a partly different biomarker panel:

- The Gambia, 2018 (Gambia National Micronutrient Survey).
- Ghana, 2017 (Ghana Micronutrient Survey).
- Sierra Leone, 2013 (Sierra Leone Micronutrient Survey).
- Malawi, 2015 to 2016 (Malawi Micronutrient Survey).

Within each survey, biomarkers were extracted for women of reproductive age (15
to 49 years) and for children 6 to 59 months, where measured. One to two surveys
additionally measured adult men. Each survey is nationally representative but
modest in size (roughly 1,000 to 1,400 individuals per outcome), is georeferenced
at the survey-cluster (primary sampling unit) level, and carries survey design
information (sampling weights, primary sampling units, and strata).

Iron is measured by serum ferritin in all four countries. Vitamin A is measured
by retinol-binding protein (a proxy for serum retinol). Vitamin B12, folate, and
zinc are measured by their serum concentrations and are available in a subset of
countries. The cross-country comparability of these biomarkers is itself debated
in the nutrition field, which is one motivation for treating absolute prevalence
levels with caution and emphasising within-country spatial ranking.

## S1.2 Predictor domains

Predictors are drawn from eleven data-source domains. With one exception (the
survey biomarker matrix, which is excluded from the deployable models) all are
publicly available or accessible through common interfaces, so the resulting
models can be applied where no biomarker survey exists.

1. **Survey biomarker matrix (`gw_` prefix).** The original survey items,
   including the biomarkers themselves and survey-derived dietary and household
   items. These are **excluded** from the primary (proxy-only) models so that
   predictions do not depend on having run a biomarker survey. They are retained
   only for the benchmark models that deliberately use full survey information.
2. **DHS (`dhs<year>_`).** Demographic and Health Survey indicators aggregated to
   the administrative unit (mother's education, household assets, water and
   sanitation, child health and nutrition, malaria interventions). Each indicator
   is stamped with the DHS round year, so the column names differ by country and
   the DHS indicators do not share names across countries.
3. **MICS (`mics_`).** Multiple Indicator Cluster Survey aggregates, where
   available, providing additional household and child indicators.
4. **IHME GBD geospatial estimates (`ihme_`).** Modelled estimates of related
   conditions (for example stunting, wasting, diarrhoeal disease, anaemia,
   under-five mortality), aggregated to admin-1 and admin-2.
5. **LSMS (`lsms_`).** World Bank Living Standards Measurement Study aggregates;
   wired for Ghana only.
6. **Malaria Atlas Project (`MAP_`).** Rasters of malaria epidemiology
   (*P. falciparum* and *P. vivax* incidence, parasite rate, mortality, and the
   *Pf* reproductive number), malaria interventions (insecticide-treated-net
   access and use, indoor residual spraying, effective antimalarial treatment),
   and inherited blood disorders (HbS, HbC, G6PD deficiency, Duffy negativity)
   that confound ferritin-based iron assessment.
7. **Google Earth Engine rasters (`gee_`).** Two complementary extractions:
   (a) cluster-buffer summaries, in which each survey cluster's GPS point is
   buffered at 10, 25, and 50 km and raster families are summarised within each
   buffer (monthly precipitation, vegetation index, and land-surface temperature;
   plus single-value population density, built-up surface, settlement model,
   global human modification, and grassland fraction); and (b) admin-2 polygon
   zonal means for the full set of rasters, including soil chemistry, climate,
   vegetation and agriculture, settlement and human modification, accessibility,
   demography, land cover, and livestock layers.
8. **WFP food prices (`wfp_`).** Market food-price series.
9. **FluNet (`flunet_`).** Influenza surveillance indicators.
10. **Food security (`fsec_`).** Household food-insecurity indicators (HungerMap /
    HFID) and IPC / Cadre Harmonisé food-security phase classifications.
11. **SoilGrids / ISRIC (`soil_`, and `gee_a2_Soil*`).** Remotely sensed soil
    micronutrient and chemistry layers (zinc, iron, calcium, magnesium,
    phosphorus, potassium, sulphur, nitrogen, total carbon, cation exchange
    capacity, aluminium).

Two predictor sets are assembled from these domains. The **individual-level set**
is roughly 200 harmonised socio-demographic and household proxy indicators shared
across the four DHS surveys. The **area-level set** is over 2,500
earth-observation aggregates at the admin-2 polygon level; for cross-country
analyses this is restricted to the 152 variables present in at least three
countries' extracts (the common-vocabulary set). Exact per-domain and per-country
counts are tabulated in `docs/data_sources_appendix.qmd`.

# S2. Data cleaning and pipeline architecture

Cleaning and assembly run in two stages. **Stage 1** is a set of per-country
cleaning scripts that read each raw survey and each external source, standardise
variable names and types, derive the inflammation-adjusted biomarkers and binary
deficiency flags, and write a per-country merged analytic table. **Stage 2** is
the reproducible `targets` pipeline that reads those merged tables and runs all
modelling, evaluation, and reporting.

Key cleaning conventions:

- **Outcome-leakage exclusion.** All `gw_` survey items are removed from the
  proxy-only predictor pool by prefix matching, so no biomarker-survey
  information enters the deployable models.
- **High-cardinality categoricals** are dropped or collapsed to avoid
  overfitting on identifiers.
- **Geometry and raw temporal columns** are removed from the analytic feature
  matrix after they have been used for linkage and temporal alignment.
- **Missing data** are retained as `NA` in the merged table; imputation is done
  downstream inside model fitting (and, in the corrected methods, strictly within
  each training fold; see S6).

The final analytic object for each country is a single individual-level table
(`data/IPD/<Country>/<Country>_merged_dataset.rds`) carrying the outcomes, the
survey design columns, and all predictor domains.

# S3. Geographic and temporal harmonisation, and record linkage

## S3.1 Geographic units and cluster-to-area assignment

Administrative polygons are taken from GADM (admin-1 region/province and admin-2
district; admin-3 chiefdom for Sierra Leone, used in a resolution sensitivity
analysis). Each survey cluster is assigned to its admin-1 and admin-2 polygon by
point-in-polygon location of the cluster GPS coordinate.

## S3.2 Linkage of predictors to outcomes

Predictors are linked at the finest feasible resolution:

- **Cluster-level linkage.** Earth-observation cluster-buffer summaries and
  Malaria Atlas rasters are extracted at each cluster's GPS point (buffer means
  for GEE; bilinear interpolation for MAP) and joined on cluster identifier.
  These predictors are therefore constant within a cluster.
- **Area-level linkage.** Sources that are only available as area aggregates
  (DHS, MICS, IHME, LSMS, food security, and the GEE admin-2 zonal means) are
  joined to the survey by administrative unit. Because survey and external admin
  names are spelled inconsistently, names are matched with fuzzy string matching
  (Jaro-Winkler at admin-1, string-distance methods at admin-2, with conservative
  thresholds); unmatched units are left missing rather than guessed. Every
  survey-to-area join is keyed on the pair (country, admin-2), never on admin-2
  name alone, to prevent cross-country mismatches.

## S3.3 Temporal alignment

For sources with multiple time points, the version closest to each country's
survey fieldwork year is selected. Climate variables use the survey year
directly; slowly changing layers (for example soil properties) use the most
recent global version. Multi-band rasters whose bands are temporal subdivisions
of a year are collapsed to the survey-year annual mean.

# S4. Outcome definitions and harmonisation

Each continuous biomarker is converted to a binary deficiency indicator using a
WHO or BRINDA-recommended threshold. Definitions, and the inflammation
adjustment that precedes them, vary across surveys; harmonising them is a
central data-preparation step.

## S4.1 Deficiency thresholds (binary outcomes)

- **Vitamin A** (women and children): retinol-binding protein below 0.70 µmol/L.
- **Iron** (ferritin): serum ferritin below 15 µg/L in women and below 12 µg/L
  in children. The Gambia outcome is defined on the log-ferritin scale at the
  equivalent cut-points (log 15 and log 12); Ghana uses Thurnham-adjusted
  ferritin on the linear scale; Sierra Leone and Malawi use the standard
  ferritin cut-points.
- **Folate** (women): serum folate below 10 nmol/L. Malawi's native rule (below
  3 ng/mL) is otherwise non-comparable, so folate is harmonised to the common
  serum cut-point for cross-country analysis.
- **Vitamin B12** (women): serum B12 below 148 pmol/L.
- **Zinc** (women and children, Malawi only): serum zinc below the WHO
  age/sex/fasting-specific cut-point.

Public-health severity tiers (low / mild / moderate / severe), used for the
map's WHO classification, are separate from these individual cut-points and
follow the standard prevalence bands (for example, for iron, prevalence
thresholds at 5%, 20%, and 40%; for vitamin A, at 2%, 10%, and 20%).

## S4.2 Inflammation adjustment

Ferritin (and, to a lesser extent, retinol-binding protein) rises with acute and
chronic inflammation, so unadjusted values understate iron deficiency in
high-infection settings. Where the acute-phase proteins C-reactive protein and
alpha-1-acid glycoprotein were measured, ferritin is corrected using the BRINDA
regression approach (or the Thurnham correction-factor approach, depending on
the country's original processing) before the deficiency cut-point is applied.
Country-specific details of which adjustment was applied to which biomarker are
documented in `docs/data_sources_appendix.qmd`. A practical consequence is that
the iron-deficiency definition is not perfectly identical across countries, which
contributes to the between-country differences in absolute prevalence and is one
reason transport of absolute levels is unreliable.

# S5. Modelling methods

## S5.1 Estimation strategies compared

Table S1 lists every estimation strategy, its loss function, and its
implementation. The strategies fall into the area-level small-area-estimation
family (the primary analysis) and the individual-level SuperLearner aggregated to
admin-2 (a sensitivity analysis), plus frequentist, Bayesian, and
transportability-aware comparators.

```{r tbl-s1}
#| tbl-cap: "Table S1. Estimation strategies compared, with loss functions and implementations."
data.frame(
  Strategy = c("Individual SuperLearner (aggregated to admin-2)",
               "Area-level SuperLearner", "OLS / elastic-net",
               "Fay–Herriot", "BYM2 (INLA)",
               "Mixed model (random country intercept)",
               "Two-stage (country mean + within-country deviation)",
               "Forward selection with LOCO-CV criterion",
               "GAM (thin-plate spatial spline)",
               "Group-lasso by data source",
               "Stacked LOCO ensemble",
               "Quasi-binomial logit GLM",
               "Quantile (median) regression",
               "Beta regression",
               "Causal DAG-based variable selection"),
  Loss = c("NLL (binary) / MSE (continuous)",
           "MSE on prevalence (and NLL variant)",
           "MSE (weighted by survey n)",
           "Gaussian likelihood + sampling variance",
           "Binomial / Gaussian (survey-n weighted)",
           "REML, weights = survey n",
           "MSE stage 1; weighted MSE stage 2",
           "MSE, weights = survey n",
           "REML, weights = survey n",
           "Grouped L2, weights = survey n",
           "Inherits from base methods",
           "Quasi-binomial deviance, weights = survey n",
           "L1 (median), weights = survey n",
           "Beta likelihood (mean + precision)",
           "MSE on DAG adjustment set"),
  Implementation = c("mlr3superlearner (6 learners), survey-weighted aggregation",
                     "mlr3superlearner on admin-2 aggregates",
                     "lm / cv.glmnet",
                     "sae::eblupFH (REML)",
                     "R-INLA, PC priors (Riebler 2016)",
                     "lme4::lmer",
                     "OLS country means; cv.glmnet on residuals",
                     "stepwise add, LOCO-r screen, max 6 vars",
                     "mgcv::gam, thin-plate spline on centroids, k = 4",
                     "gglasso, groups by data-source prefix",
                     "NNLS stack over base methods, country-disjoint CV",
                     "glm(quasibinomial, logit)",
                     "quantreg::rq at tau = 0.5",
                     "betareg (logit link)",
                     "dagitty adjustment sets, then OLS")
) %>% kable(booktabs = TRUE)
```

## S5.2 Individual-level SuperLearner

The individual-level SuperLearner combines six base learners: a mean (intercept)
learner, penalised generalised linear models (lasso, ridge, and elastic-net via
glmnet), a ranger random forest, gradient-boosted trees (XGBoost, depth 4,
learning rate 0.05, subsample 0.8), and dropout BART. For binary outcomes with
prevalence below 15%, class-weighted XGBoost and ranger learners are added to
counter class imbalance. Cross-validation folds are formed at the survey-cluster
(primary sampling unit) level so that individuals in the same cluster are never
split across folds. To produce an admin-2 map, person-level predictions are
averaged within each district using survey weights.

## S5.3 Area-level SuperLearner

The area-level SuperLearner repeats the ensemble fit with the admin-2 polygon as
the unit of analysis: one row per district, the target being survey-weighted
district prevalence and the predictors being the polygon-aggregated covariates.
Both a mean-squared-error loss and an approximate negative-log-likelihood loss
were evaluated. Because the covariates are resolved at the district level, the
effective sample size for this fit is the number of polygons (14 to 87 per
country), which is the source of the area-level overfitting reported in the main
text.

## S5.4 Frequentist baselines

Ordinary least squares using up to eight top-correlated covariates (capped to
avoid rank deficiency on small training sets) and a 10-fold cross-validated
elastic-net (alpha 0.5, with a ridge fallback for very small samples). Each was
evaluated on both the raw prevalence scale and the logit scale with
back-transformation. A country-mean predictor is included as a floor.

## S5.5 Bayesian small-area models

The Fay–Herriot area-level model uses auxiliary covariates and known
sampling variances derived from survey-weighted prevalence and effective sample
size. The BYM2 model is a Besag–York–Mollié disease-mapping model with
penalised-complexity priors, fit by integrated nested Laplace approximation.
Admin-2 adjacency is built from GADM polygons by Queen's contiguity; for
leave-one-country-out evaluation the adjacency graph is block-diagonal across
training countries (no edges to the held-out country), and held-out districts
receive predictions from the covariate (fixed-effect) component only.

## S5.6 Transportability-aware comparators

Several comparators target cross-country transport directly: a country
random-intercept mixed model whose held-out predictions use the marginal mean
(random effect set to zero, so transport depends only on covariate slopes); a
two-stage model that predicts a country-level mean from country-aggregated
covariates and adds a within-country deviation from centred area covariates; a
forward variable-selection model that chooses variables to maximise the average
correlation across leave-one-training-country-out splits; a thin-plate spatial
GAM on polygon centroids; a group-lasso that selects whole data-source domains;
and a stacked ensemble whose non-negative weights are estimated on
country-disjoint folds so the stack does not overfit within-training-country
structure.

## S5.7 Post-hoc calibration, feature augmentation, and intervals

Binary classifier probabilities were recalibrated with Platt logistic
recalibration; in the corrected analysis of record this is done out-of-fold
(S6). Outcome-aware feature augmentation added a small number of
biologically motivated interaction terms before transport fitting (for example
bed-net coverage by cereal share for iron, vegetation by livelihood density for
vitamin A, soil zinc by staple-crop yield for zinc, livestock density by
accessibility for B12 and folate); each interaction passed through the same
screening as the base covariates. Uncertainty on the transport correlation point
estimates was characterised by resampling admin-2 polygons within each held-out
country with replacement (500 bootstrap replicates).

# S6. The corrected (P1–P8) methods: the analysis of record

Initial benchmarking showed that an analysis at this scale can overstate its own
performance through subtle leakage and optimistic evaluation. The corrected
methods below address eight such issues and are the analysis of record; the
original pipeline outputs are retained as a sensitivity comparison. They are
implemented in `R/corrected/` within the single `targets` pipeline and consume
the same per-slice data the main pipeline builds.

- **P1. Leakage-free fitting and honest cross-validation.** All preprocessing
  (screening, correlation filtering, standardisation, imputation) is performed
  inside each training fold only. Folds are formed by survey cluster and,
  separately, by spatial region, so spatially autocorrelated information cannot
  leak from held-out areas. An optimistic scheme (random folds with full-data
  preprocessing) is fit alongside purely for comparison, which quantifies the
  optimism of the original design.
- **P2. Out-of-fold calibration.** Platt recalibration parameters are estimated
  on training folds and applied to held-out folds, rather than fit in-sample. The
  in-sample calibration (slope forced toward 1) is reported only as the
  optimistic reference.
- **P3. Decision-value targeting.** Predictions are scored on the operational
  task of ranking districts: the share of true burden captured by the top 20% of
  districts ranked by predicted prevalence (reach), the lift of that reach over
  untargeted allocation, and the hit-rate of the predicted top-20% against the
  true top-20%.
- **P4. Honest prediction intervals.** Each district receives a split-conformal
  prediction interval (calibration residuals from a held-out half), with the
  finite-sample correction clamped so the conformal level is valid for small
  calibration sets, plus a design-based survey interval where available.
  Empirical held-out coverage is reported against the nominal level.
- **P5. Sampling-error-aware error.** District-level prediction error is
  decomposed into a survey-sampling component (what survey noise alone would
  produce) and a model component, so the honest model error (sampling-adjusted
  RMSE) is separated from irreducible survey noise.
- **P6. Out-of-support trust flags.** Each district is flagged on three criteria:
  whether its covariates fall outside the range spanned by the surveyed
  (training) districts (out-of-support distance), small survey sample, and a wide
  prediction interval. The combined OK-to-use versus do-not-rely label drives the
  dashboard's certainty layer.
- **P7. Partial-pooling small-area estimator.** An empirical-Bayes estimator
  shrinks each district's direct survey estimate toward the covariate-model
  prediction in proportion to its sampling uncertainty, giving stable estimates
  for sparsely sampled districts.
- **P8. Join and centring bug fixes (by construction).** Every survey-to-area
  join is keyed on (country, admin-2); all standardisation uses training
  statistics only; and out-of-sample imputation uses training medians, never the
  target rows' own values.

# S7. Assessment of predictive accuracy

Two questions are evaluated. **Within-country** accuracy is estimated by 5-fold
cross-validation on each country's admin-2 set (or by the honest cluster- and
spatial-block schemes of S6), with all metrics computed on the union of
out-of-fold predictions, never in-sample. **Cross-country transportability** is
assessed by leave-one-country-out holdout: the model is trained on three
countries' districts and evaluated on the fourth, which contributed no data.

Metrics. For the area-level prevalence task: Pearson and Spearman correlation
between predicted and observed district prevalence (spatial ranking), mean
absolute error and root-mean-squared error in percentage points (absolute
accuracy), mean bias in percentage points, and the calibration slope from a
regression of observed on predicted prevalence. For binary classifiers: area
under the ROC curve, Brier score and Brier skill score, expected and maximum
calibration error, and the calibration intercept and slope. The corrected
analysis adds the sampling-adjusted RMSE (S6 P5), split-conformal interval
coverage (P4), and the decision-value measures (P3). National prevalence
estimates are compared against survey-weighted national values.

A central methodological point is that honest, out-of-fold evaluation is
materially less flattering than in-sample or random-fold evaluation, and that
within-country accuracy is a poor guide to cross-country transportability.

# S8. Sensitivity analyses

The following analyses test the robustness of the main findings and are reported
as sensitivity analyses rather than headline results:

- **Individual-level versus area-level fitting.** The individual-level
  SuperLearner aggregated to admin-2 is compared against the area-level fit; the
  comparison motivates treating the area-level SAE as primary.
- **Loss function and scale.** Area models were evaluated under both
  mean-squared-error and negative-log-likelihood losses, and on both the raw and
  logit prevalence scales.
- **Outcome-definition harmonisation.** Folate transport was assessed before and
  after harmonising Malawi's native cut-point to the common serum definition.
- **Finer spatial resolution.** Sierra Leone was re-analysed at admin-3
  (chiefdom) resolution to show that its anti-predictive admin-2 transport
  estimates are largely a small-sample artefact (14 districts).
- **Raw versus recalibrated classifiers.** Binary calibration diagnostics are
  reported both raw and after out-of-fold Platt recalibration.
- **Corrected versus production pipeline.** The full corrected-versus-original
  comparison (CV honesty, calibration, admin-2 error, intervals, decision value,
  trust, partial pooling) is the basis for adopting the corrected methods as the
  analysis of record.
- **Predictor-set restrictions.** A remote-sensing-only (GEE-only) transport
  analysis, and a parsimonious model using a small set of soil-micronutrient
  features plus a spatial term, test which signal transports.
- **All-methods comparison on a representative outcome.** Every method and
  SuperLearner library variant was run on child iron deficiency, the
  best-transporting outcome, under both within-country and transport evaluation.

# S9. Reproducibility

The entire analysis is a single dependency-aware `targets` pipeline
(`_targets.R`) that builds the data tables, fits every model, runs the corrected
(P1–P8) methods, computes all metrics, and writes the result tables consumed by
the manuscript and the dashboard. Stage 1 cleaning scripts regenerate the merged
per-country analytic tables from the raw sources. The corrected methods write
their canonical outputs to `results/tables/corrected/`. Pipeline definitions and
code are available in the project repository; the deepest data lineage is in
`docs/data_sources_appendix.qmd`.

# References

::: {#refs}
:::
