---
title: "Subnational micronutrient deficiency prediction in sub-Saharan Africa from proxy indicators: what the covariates carry, and how evaluation design decided the answer"
shorttitle: "Subnational micronutrient prediction in sub-Saharan Africa"
author:
  - name: Andrew N. Mertens
    affiliation: University of California, Davis, Institute for Global Nutrition
    corresponding: true
    email: anmertens@ucdavis.edu
date: today
format:
  docx:
    number-sections: true
    fig-cap-location: bottom
    tbl-cap-location: top
    fig-width: 6.5
    fig-height: 4.5
    fig-dpi: 300
execute:
  echo: false
  warning: false
  message: false
keywords:
  - micronutrient deficiency
  - small area estimation
  - DHS
  - transportability
  - SuperLearner
  - vitamin A
  - iron deficiency
bibliography: references.bib
abstract: |
  **Background:** Subnational maps of micronutrient deficiency are central to
  targeted nutrition programming, yet biomarker surveys remain scarce and
  geographically uneven. We assessed whether models trained on harmonised
  Demographic and Health Surveys (DHS) and remote-sensing proxies can reproduce
  admin-2 deficiency patterns within and across four sub-Saharan African
  countries, and whether they outperform the baselines a programme actually has.
  An internal audit of our own earlier analysis found that several headline
  conclusions were determined by choices in the evaluation design rather than by
  the data, so we additionally report how much each such choice moved the
  answer.

  **Methods:** We harmonised individual-level DHS biomarker measurements for
  iron, vitamin A, vitamin B12, folate, and zinc deficiency from Gambia (2018),
  Ghana (2017), Sierra Leone (2013), and Malawi (2016), and combined them with
  earth-observation aggregates at the
  admin-2 polygon level (206 polygons pooled). Estimation strategies compared
  at admin-2 included: individual-level SuperLearner (12 learners spanning
  penalised regression, random forests, gradient boosting and BART) aggregated
  to admin-2; an area-level SuperLearner fit directly on admin-2 aggregates; a
  spatial-smoothing-only model with no covariates; a national-mean null; and,
  across countries, 21 comparator methods spanning frequentist regressions,
  Bayesian SAE (Fay–Herriot, BYM2/INLA), and spatial and invariance-based
  transport methods. Because most admin-2 polygons contribute only 6–36
  biomarker measurements, we estimated a reliability ceiling for each
  country–outcome cell by split-half replication of survey respondents, having
  found our earlier analytic estimator biased low by about a factor of 4.7.

  Re-analysis followed a protocol specified in advance of it. Results are
  reported against three separately named estimands — filling in unsurveyed
  districts inside a surveyed country, extrapolating to an unseen region, and
  transporting to a country held entirely out of training — because these have
  materially different answers and are routinely reported interchangeably. Every
  comparison baseline is recomputed under the model’s own folds, and any
  baseline derived from the same survey is jackknifed so the scored district
  never contributes to it. Every random fold assignment is replicated ten times
  and reported as a median with its spread. For each significance screen we
  report attained power and verify that a true effect could have cleared the
  multiplicity threshold. Predictors enter in three competing representations —
  all 451 harmonised columns under elastic net, 18 domain principal-component
  summaries, and a zero-tuning marginal-correlation index — so that model
  capacity is a measured factor rather than a fixed choice.

  **Results:** Evaluation design changed the conclusions more than any model
  did. A previously reported within-country median r of 0.058 rested on a single
  random division of the data; replicated, the same protocol gives 0.217, and
  exhaustive leave-one-region-out gives 0.235. A comparison the models appeared
  to lose was against a benchmark computed from the held-out district's own
  respondents; measured symmetrically, that benchmark scores r 0.076 and the
  covariate model 0.156. A screen reported as finding no predictor surviving
  false-discovery control could not have returned a positive result at any
  effect size, given its permutation resolution. And the reliability ceiling
  used to argue that admin-2 was unresolvable was biased low by a factor of
  about 4.7; the validated empirical ceiling is 0.47–0.61, leaving substantial
  headroom. Under the corrected protocol, covariate models beat the honest
  covariate-free survey baseline in both within-country settings: rank
  correlation 0.398 versus 0.320 on the biomarker concentration and 0.286
  versus 0.193 on deficiency prevalence. Rankings transported to a country held
  entirely out of training with a mean Spearman of 0.26 over 22
  country–outcome combinations for the zero-tuning index (17 positive; 0.27
  and 19 positive for the penalised fit) and 0.37 (22 positive) for an index
  built only from remotely sensed climate and soil layers; at the first
  sub-national tier the same 22 combinations gave 0.30 (17 positive) and 0.45
  with climate and soil, against a permutation null whose 95th percentile is
  0.16. Absolute prevalence levels did not transport. Used for
  targeting, the worst-ranked fifth of districts reached 26% more deficiency
  burden than the same effort guided by regional survey averages, which were
  themselves no better than random selection; in Malawi, those districts
  carried 43% child iron deficiency against 26% nationally, holding 49% of the
  national burden. A zero-tuning index over 18 domain summaries outperformed a
  penalised regression over all 451 predictors by up to 0.38, the latter turning
  negative under region extrapolation. Associations were nutrient-specific and
  replicated in direction across all four countries, but several ran opposite to
  individual-level nutritional biology: district legume consumption tracked
  *more* child vitamin A deficiency, and cattle ownership *more* child iron
  deficiency, alongside cereal-dominant cropping. Read together with warm
  night-time land surface temperature and soil-chemistry heterogeneity, the
  replicated axis appears to be rural subsistence agro-ecology rather than
  nutrient intake, and these district-level associations should not be read as
  individual-level effects. Two were both replicated and interpretable as
  stated: district child wasting tracked women's vitamin A deficiency, and
  rainfall tracked less of it. No malaria burden indicator replicated against
  any outcome. A covariate-free spatial smoother
  scored 0.391 where the full covariate model scored 0.398.

  **Conclusions:** Proxy indicators carry replicated, nutrient-specific signal
  for subnational micronutrient deficiency, and several previously published
  negative findings — including our own — were artifacts of evaluation design
  rather than properties of the data. Within a surveyed country, however,
  geography alone accounts for most of the attainable accuracy; covariates earn
  their place in two specific situations, when the comparator is a
  survey-derived baseline that must be made information-symmetric to be fair,
  and when no survey exists and no spatial smoother can be fitted. The
  defensible product is a ranked district priority list rather than a prevalence
  map. We recommend that studies of this kind state what information their
  comparison baseline was permitted to see, replicate any random division of the
  data and report the spread rather than a single value, and demonstrate that
  their statistical tests could have detected an effect had one been present.
  Applied to our own work, these conditions reversed several conclusions in both
  directions.
---

::: {.callout-important}
## Revision in progress — abstract is current, body sections are not yet

A methods audit on 2026-09-01 found that several load-bearing numbers in this
document were artifacts of the evaluation design rather than measurements. **The
status of every affected claim is tabulated in
`docs/findings/TWO_READINGS_2026-09f.md`; corrected figures are in
`docs/findings/PROTOCOL_V2.md`.**

**Corrected as of 2026-09-04:** the regional (first sub-national tier)
transport figure in the abstract, @sec-protocol-results, the discussion and
the conclusions now uses all 22 country–outcome combinations (mean Spearman
0.29; 0.45 for the climate-and-soil index). The earlier 0.50–0.56 over 12
combinations rested on three countries (a population-file spelling had dropped
Sierra Leone) and is withdrawn. The AlphaEarth embedding was also moved out of
the agriculture domain into one of its own and Gambia's food-price window was
corrected to its 2018 fieldwork; the protocol tables were re-run in full at
08:17 on 2026-09-04, and again at 23:25 the same day after Gambia's missing
IHME block was restored (sandbox log IH-01). Numbers that moved by more than
0.01 are updated
throughout: the district transport of the zero-tuning index (0.25 to 0.26)
and the regional transport (0.29 to 0.31, then 0.30). The district figure formerly
quoted as "21 of 22, mean 0.281" was the penalised domain fit; it is now 0.27
(19 of 22), and the text quotes the estimator of record, the
zero-tuning index, alongside it. In-fill, burden capture, the comparator
tests and the climate-and-soil figures did not move.

**Corrected as of 2026-09-02:** the title and abstract; @sec-protocol (new,
the corrected evaluation protocol); @sec-protocol-results (new, the re-analysis);
@sec-ceiling; @sec-fdr; @sec-anchoring; the within-country admin-2 section; and
the Discussion summary and Conclusions.

**Not yet corrected:** the remaining Results subsections
(@sec-resolution onward, @sec-anchor-individual, @sec-national-composition, the
benchmark and calibration sections) and **every rendered table in this
document**, which is still generated from the pre-audit result files. Where a
table and the surrounding prose disagree, the prose is current.

The specific reversals:

- **`median r 0.058`** rests on one unreplicated 3-fold region draw sitting in
  the bottom decile of its own protocol's distribution. Replicated: **0.217**;
  leave-one-region-out: **0.235**.
- **`0 of 294 predictors survive FDR`** was unattainable by construction — at
  999 permutations a single true predictor of arbitrary strength receives
  BH q >= 0.296. The same table's conventional-null columns carry 476
  within-cell survivors.
- **The regional-anchoring gain (0.164 -> 0.413)** was withdrawn by the
  project's own jackknife control (0.147 vs 0.156 unanchored).
- **The covariate-free baseline r = 0.516** reads the held-out district's own
  survey responses; jackknifed it scores **0.076** and loses to the covariate
  model at 0.156.
- **The analytic noise ceiling (median r_max 0.098)** was proven ~4.7x biased
  low. The binding empirical ceiling is **0.47-0.61**.
- **"covariates beat geography in 6 of 24 cells"** is a miscount against its
  own source table; the correct count is **10 of 24**.
- **Directional claims about legumes and cattle** (2026-09-02) had the sign
  inverted: both track *more* deficiency, not less. See @sec-fdr.
:::


```{r setup}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(ggplot2); library(knitr)
  library(kableExtra); library(here); library(readr); library(scales)
})
theme_set(theme_minimal(base_size = 11))

# ── Results sources ───────────────────────────────────────────────────────
TABLE_DIR <- here::here("results", "tables")
TARGETS_STORE <- here::here("_targets_full")

safe_read_csv <- function(path) {
  if (file.exists(path)) read_csv(path, show_col_types = FALSE) else NULL
}

cv_perf       <- safe_read_csv(file.path(TABLE_DIR, "cv_performance_all.csv"))
diag_binary   <- safe_read_csv(file.path(TABLE_DIR, "diagnostics_binary.csv"))
diag_cont     <- safe_read_csv(file.path(TABLE_DIR, "diagnostics_continuous.csv"))
admin2_err    <- safe_read_csv(file.path(TABLE_DIR, "admin2_error_all.csv"))
area_compare  <- safe_read_csv(file.path(TABLE_DIR, "area_comparison_all.csv"))
loco_metrics  <- safe_read_csv(file.path(TABLE_DIR, "transportability_area_loco_metrics.csv"))
loco_full     <- safe_read_csv(file.path(TABLE_DIR, "area_loco_comparison.csv"))
natl_est      <- safe_read_csv(file.path(TABLE_DIR, "national_estimates_all.csv"))

# Anchoring, the individual-level anchor, and the national composition. Sourced
# rather than read directly because table_anchor_arms() applies the same arm
# labelling the slide deck and the dashboard use, so the three cannot disagree.
source(here::here("R", "plot_forest.R"))
anchor_arms   <- safe_read_csv(file.path(TABLE_DIR, "admin1_arms.csv"))
indiv_anchor  <- safe_read_csv(file.path(TABLE_DIR, "individual_anchor.csv"))
nat_compose   <- safe_read_csv(file.path(TABLE_DIR, "national_composition.csv"))
nat_levels    <- safe_read_csv(file.path(TABLE_DIR, "national_composition_levels.csv"))

# Benchmark results: written by the benchmark pipeline targets when available.
bench_results <- safe_read_csv(file.path(TABLE_DIR, "benchmarks_all.csv"))

# ── Corrected-methods (P1–P8) outputs — the analysis of record ──────────────
# Produced by the corrected-methods section of the unified _targets.R pipeline
# (build_methods_comparison -> results/tables/corrected/). These honest, leakage-
# free metrics are canonical for within-country discrimination, calibration,
# admin-2 error, prediction intervals, decision value and per-district trust;
# the production-pipeline equivalents are retained as a sensitivity comparison.
CORR_DIR <- file.path(TABLE_DIR, "corrected")
corr_cv      <- safe_read_csv(file.path(CORR_DIR, "cv_honesty_compare.csv"))
corr_calib   <- safe_read_csv(file.path(CORR_DIR, "calibration_oof_compare.csv"))
corr_err     <- safe_read_csv(file.path(CORR_DIR, "admin2_error_sampling_adjusted.csv"))
corr_decision<- safe_read_csv(file.path(CORR_DIR, "decision_value.csv"))
corr_interval<- safe_read_csv(file.path(CORR_DIR, "interval_summary.csv"))
corr_trust   <- safe_read_csv(file.path(CORR_DIR, "trust_flags.csv"))
```

# Introduction

Micronutrient deficiencies remain a leading driver of preventable morbidity
and mortality among children and women in low- and middle-income countries,
contributing to anaemia, impaired growth, immune dysfunction and adverse
pregnancy outcomes [@stevens2022micronutrient; @who2020anaemia]. National
biomarker surveys conducted under the WHO Vitamin and Mineral Nutrition
Information System (VMNIS) [@who_vmnis] and the BRINDA pooled analyses
[@brinda2017overview] document broad regional contrasts, but the spatial
resolution of these surveys is rarely finer than admin-1 (province) and is
absent in many countries entirely. Subnational geographic targeting of
fortification, supplementation and food-system programmes therefore relies
on indirect inference from sparse data.

Two methodological traditions have attempted to fill this gap. The
small-area-estimation (SAE) literature, originating with the Fay–Herriot
mixed model [@fay1979estimates] and developed for survey-derived admin-level
prevalence under the Bayesian disease-mapping framework
[@besag1991bayesian; @mercer2015small; @wakefield2019small;
@riebler2016intuitive], constructs structured spatial smoothers using
admin-polygon adjacency and known sampling variances. In parallel, ensemble
machine-learning approaches, including SuperLearner [@vanderlaan2007super]
and gradient-boosted tree models trained on satellite-derived covariates,
have produced high-resolution maps of indirect indicators of nutritional
status [@osgoodzimmerman2018mapping; @local2021maps]. These two traditions
have rarely been compared head-to-head on the same biomarker datasets at
the same geographic scale.

Two questions about this approach remain unresolved. First, when a model
is trained at the individual level and aggregated to admin-2, do its
admin-2 estimates align with survey-weighted admin-2 prevalence — or does
the aggregation step introduce systematic bias? Second, when a model is
trained on a set of countries with biomarker data, do its predictions
transport to a held-out country with similar covariates? The
small-area-estimation literature has long recommended fitting at the
unit of deliverable: if the output is an admin-2 map, fit at admin-2; if
the output is household-level risk, fit at the individual level
[@mercer2015small; @wakefield2019small]. To our knowledge, the
implication of this recommendation has not been systematically tested for
multi-country micronutrient deficiency prediction.

In this study we harmonise DHS biomarker measurements from four
sub-Saharan African countries (The Gambia, Ghana, Sierra Leone, and Malawi)
and use these to evaluate estimation strategies
spanning machine-learning and SAE traditions, within-country and across
countries. We benchmark a multi-learner SuperLearner ensemble (penalised
regression, ranger, gradient-boosted trees, and BART) against Bayesian areal
models, frequentist baselines, a covariate-free spatial smoother, and a
national-mean null, and report whether the added complexity of the ensemble
produces gains in spatial ranking or transportability that justify its use
for global-health programming. Because the admin-2 biomarker target is
itself measured from as few as 6–36 respondents per district, we also ask a
prior question: how much of any prediction gap is attributable to sampling
noise in the outcome rather than to the model, and report every result
against an explicit reliability ceiling.

# Methods

## Study population and biomarker data

We assembled biomarker measurements from four Demographic and Health
Surveys: The Gambia 2018, Ghana 2017, Sierra Leone 2013, and Malawi 2016.

We initially added a fifth country, Tanzania (TDHS 2010, vitamin A only), but
withdrew it before finalising this manuscript after a project collaborator
identified the survey's dried-blood-spot RBP measurements as unreliable
("completely wrong... not even internally consistent"), consistent with
independent data-quality anomalies we found in the same data (a unit-label
error in the survey's own biomarker documentation, and a 77–92
percentage-point national-level bias when Tanzania was held out under
leave-one-country-out evaluation during a supplementary analysis). We plan to
revisit Tanzania once the 2023 DHS round, which uses serum rather than
dried-blood-spot measurement, becomes accessible. Within each survey we
extracted deficiency biomarkers for women of reproductive age (15–49 y) and
children 6–59 months as available (@tbl-cohort).

```{r tbl-cohort}
#| tbl-cap: "Sample sizes by country and biomarker, with WHO public-health classification thresholds applied to derive binary deficiency outcomes."
if (!is.null(cv_perf)) {
  cv_perf %>%
    filter(model_type == "binary") %>%
    transmute(
      Country = country,
      Outcome = outcome,
      n = n,
      `Observed prevalence` = ifelse(is.finite(brier) & !is.na(auc),
        sprintf("%.1f%%", 100 *
          (diag_binary$prevalence[match(paste(country, outcome),
                                          paste(diag_binary$country, diag_binary$outcome))])),
        "—")
    ) %>%
    kable(booktabs = TRUE)
}
```

Each biomarker was converted to a binary deficiency indicator using WHO
or BRINDA-recommended thresholds, with adjustment for inflammation where
acute-phase proteins were available [@brinda2017adjustment;
@who_vmnis_thresholds]. Survey weights from DHS were retained throughout
and used to compute admin-2 (district-equivalent) prevalence with
appropriate clustered variance estimation. Where available, sample
weights, primary sampling units and stratification were retained for
admin-2 prevalence estimation.

## Predictors {#sec-predictors}

Two predictor sets were assembled. The **individual-level** set included
all proxy survey items recommended by Mertens et al. (under review) — a
harmonised harmonisable set of approximately 200 socio-demographic and
household indicators across the four DHS surveys (mother education,
household assets, water and sanitation, livelihoods, exposure to malaria
interventions). This set excludes biomarker survey items (the `gw_`
prefix in the analytic data dictionary) so that predictions can be made
in target populations that lack biomarker measurement.

The **area-level** set was assembled from every source available at admin-2
resolution, not from earth observation alone: Google Earth Engine aggregates
[@gorelick2017google] extracted for each admin-2 polygon by exact-extract
weighted by polygon overlap (population density, ESA WorldCover land use,
accessibility surfaces, NDVI/EVI temporal aggregates, atmospheric parameters,
elevation, night-time lights), together with DHS and MICS admin-2 indicator
aggregates, SoilGrids and iSDAsoil soil chemistry, MapSPAM crop shares, ESPEN
helminth prevalence, Malaria Atlas Project surfaces, IHME modelled disease
burden, WFP food prices and food-security indicators.

Harmonisation proceeded in four documented stages, recorded in
`metadata/covariates/` so each decision is auditable: raw provider columns were
mapped onto canonical names; near-duplicate variants of one construct (month
climatologies, buffer radii, soil depth summaries, year stamps) were collapsed;
values were converted to canonical units; and variables that could not be
reconciled across countries were excluded with a stated reason. Of Ghana's 755
raw admin-2 columns, 301 were dropped as not being quantities at all — cross-band
summaries averaged over non-commensurable bands, categorical land-cover class
codes whose zonal mean is meaningless, quality-assurance bands, and population
columns in which the provider had replicated five-year epochs across intervening
years. A further 40 were excluded for cross-country irreconcilability, including
twelve whose medians ranged from 742 to 23,784 across countries with maxima to
$2.45\times10^7$ — a defective export rather than a scale factor, and one that had
been inside the previous common-vocabulary set and was therefore manufacturing a
country effect — and seven outcome-adjacent DHS indicators excluded as leakage
(anaemia is the downstream consequence of iron deficiency; nutrient-specific
supplementation is targeted *by* measured deficiency).

The result is a common-vocabulary set of **294 predictors** shared by all four
countries, replacing the 152-variable set used in earlier analyses; the source
registry documenting these inputs lists 23 providers across 12 thematic
domains. The intersection step itself is not
the binding constraint: 322 canonical variables exist for a typical country and
294 are common to all four, with 208 present in all five ingested countries.

## Estimation strategies

We compared estimation strategies spanning the machine-learning and
small-area-estimation traditions, organised into two families, plus two
uninformed baselines added by the 2026-08 methods review (below) so that
"better than the alternatives" can be distinguished from "better than doing
nothing": a **national-mean null** (the country's survey-weighted national
prevalence assigned to every district) and a covariate-free **spatial-only**
smoother (a thin-plate spline on admin-2 centroid coordinates). The
**area-level small-area estimation (SAE)** family fits one row per admin-2
polygon, with survey-weighted prevalence as the target: an area-level
SuperLearner on the polygon-aggregated covariates (fit under both a Gaussian
mean-squared-error loss and a negative-log-likelihood loss); frequentist
baselines (a country-mean floor, OLS, and cross-validated elastic-net);
Bayesian areal models (the Fay–Herriot model [@fay1979estimates;
@mercer2015small] and the BYM2 disease-mapping model fit by integrated nested
Laplace approximation [@riebler2016intuitive; @rue2009approximate]); and a set
of transportability-aware comparators (a country random-intercept model, a
two-stage country-mean-plus-within-country-deviation model, a forward
selection that optimises a leave-one-training-country-out criterion, a
spatial-spline GAM, a data-source group-lasso, and a stacked ensemble of
these) — 21 comparator methods in total for the cross-country comparison
(@tbl-benchmarks). The **individual-level SuperLearner aggregated to
admin-2** fits a person-level ensemble (12 learners: penalised regression ×3,
ranger ×3, XGBoost ×2, and BART ×3 at three tree-count settings; a Gaussian
process was dropped from the stack after a controlled sensitivity check
found it received zero ensemble weight while dominating runtime — Results)
and averages its predictions up to the district; it shares the same inputs
but is reported as a sensitivity analysis, because aggregating shrunken
person-level probabilities biases area means (Results).

**A reliability ceiling for admin-2 prediction.** Because a median admin-2
district contributes only 6-36 biomarker measurements, part of the
between-district variance in the observed prevalence $p_\text{obs}$ is
sampling noise rather than true district-level signal. We originally estimated,
for each country-outcome cell, the maximum Pearson r any model could achieve
given that noise floor as
$$\lambda = 1 - \frac{E[V_\text{sampling}]}{\mathrm{Var}(p_\text{obs})}, \qquad r_{\max} = \sqrt{\lambda},$$
with $V_\text{sampling}$ the design-based sampling variance of each district's
direct estimate and $\lambda$ bounded at zero.

**This analytic form is withdrawn** (@sec-ceiling). It assumed a design effect
of 1.5; measured from the cluster structure the design effect is 1.0-5.6 with a
median near 2.4, which biases $r_{\max}$ low by roughly a factor of 4.7 and
produced the impossible result of achieved correlations exceeding their own
ceiling in half the cells. The ceiling of record is instead **empirical**: we
split each district's biomarker respondents at random into halves, compute the
direct prevalence estimate in each half, and correlate the two across districts,
correcting for the halving with the Spearman-Brown formula. This requires no
design-effect assumption because the survey's own clustering is carried into
both halves, and we validated it on simulated data with a known true ceiling. We
report it as an interval over 200 random splits.


A subsequent methods-development analysis explored whether modelling the
area prevalence directly, rather than fitting an unregularised
p ≫ n SuperLearner on the polygons, could recover a stronger within-country
estimator. We developed an **area-level prevalence recipe**: admin-2
survey-weighted prevalence under a **binomial (count) loss weighted by the
survey sample size** (which leaves district ranking unchanged relative to a
Gaussian logit fit but markedly improves calibration); a mild **global
near-null pre-filter** (covariates whose marginal association with the
outcome falls in the lowest 70% are dropped) followed by an in-fold
supervised screen; and a penalised regression (elastic net), with two opt-in
variants (a domain-wise late-fusion stack, and for vitamin A a distributional
estimator that integrates over the deficiency threshold). Read against the
reliability ceiling and the two uninformed baselines above, this recipe's
comparative claims did **not** hold up under further audit (Results,
@sec-recipe-audit): its cross-country comparator (a spatial-plus-soil model)
was ranked using LOCO correlations computed on the same held-out countries
the ranking was then evaluated on, and re-deriving the pre-filter honestly
inside each resampling fold (rather than once on the full data before
splitting) reduced its apparent Spearman gain by 0.16–0.30. We retain the
recipe's specification and within-country point estimates for completeness
(Supplementary Appendix Table S1, project area-level recipe specification)
but report the corrected, null- and ceiling-referenced comparison as the
primary within-country result.

The full list of strategies with their loss functions and implementations is
given in Supplementary Appendix Table S1, and the complete modelling detail
(learner libraries and hyper-parameters, adjacency construction, post-hoc
calibration, outcome-aware feature augmentation, and the bootstrap interval
procedure) is in Supplementary Appendix Section S5.

## Honest evaluation and the corrected-methods analysis of record

Initial benchmarking surfaced several ways an analysis at this scale can
flatter itself, so the headline results below come from a set of corrected,
leakage-free methods (referred to as the P1–P8 methods) that constitute the
analysis of record; the original production-pipeline equivalents are retained
as a sensitivity comparison. In brief, the corrected methods (i) perform all
preprocessing inside each cross-validation fold and form folds by survey
cluster and by spatial region, so that no information leaks from held-out
areas; (ii) recalibrate binary predictions out-of-fold rather than in-sample;
(iii) separate the component of admin-2 prediction error that is attributable
to survey sampling noise from genuine model error; (iv) attach
split-conformal prediction intervals and design-based survey intervals to each
district; (v) score the decision-relevant task of correctly ranking the
highest-burden districts; (vi) flag districts whose covariates fall outside the
surveyed training range as out-of-support; and (vii) produce partial-pooling
(empirical-Bayes) area estimates. The full specification is in Supplementary
Appendix Section S6.

Two further safeguards apply to the area-level recipe. First, every reported
correlation carries an area-bootstrap 95% confidence interval and a
non-parametric permutation-null *p*-value (the permutation re-labels districts
within country), so signal can be distinguished from sampling noise at the small
effective sample sizes involved. Second, the global near-null pre-filter is
shown to be leakage-free by an honest permutation test that **re-derives the
filter on each permuted outcome**: the resulting null distribution of the
out-of-sample correlation remains centred at zero (95th percentile 0.17–0.35
across cells) while the observed correlations are 0.29–0.65 (permutation
*p* ≤ 0.02), confirming that dropping obvious near-null covariates on the full
data introduces negligible optimism. The empirical-Bayes / Fay–Herriot layer is
used for honest interval width only and not to shrink the point level, because
the covariates carry district ranking but not absolute level; the level is set
by anchoring to a known national prevalence.

## Evaluation

We evaluated two questions corresponding to two deployment regimes.
**Within-country** spatial accuracy (gap-filling unsurveyed districts in a
country with some survey data) was estimated by cross-validation on each
country's admin-2 set with metrics on the union of out-of-fold predictions; for
the area-level recipe we report both **district-block** cross-validation (whole
districts held out) and the more demanding **region-block** scheme (whole admin-1
regions held out), using the enriched covariate set. **Cross-country
transportability** (predicting a country with no biomarker survey) was assessed
by leave-one-country-out (LOCO) holdout on the pooled admin-2 dataset using the
universal covariate set and scored by rank correlation, which is invariant to
the untransportable absolute level: the model is trained on three countries'
admin-2 areas (n = 119–192) and evaluated on the fourth's (n = 14–87). Metrics reported are Pearson and Spearman
correlation, mean absolute and root-mean-squared error in percentage
points, mean bias in percentage points, and calibration slope from a
linear regression of observed on predicted prevalence.

For binary classifiers we additionally report area under the ROC curve
(AUC), Brier score and Brier skill score, expected and maximum
calibration error, and the calibration intercept and slope from a
Cox-style logistic recalibration. For continuous biomarker regressions
we report R² and Pearson r.

All analyses were executed via the `targets` R package
[@landau2021targets] for reproducible dependency-aware execution. Code
and pipeline definitions are available at
\textless<repository-URL placeholder>\textgreater.

## The corrected evaluation protocol {#sec-protocol}

A 2026-09 internal audit of this manuscript found that several headline results
were determined by choices in the evaluation design rather than by the data
(@sec-ceiling, @sec-fdr, @sec-anchoring). We therefore specified a protocol in
advance of re-analysis and re-ran every within- and cross-country comparison
under it. It rests on four rules.

**One. Name the estimand.** "Can covariates predict district prevalence" is
three different questions with three different answers, and the literature
routinely reports them interchangeably. We separate: *in-fill*, predicting
unsurveyed districts inside a surveyed country (5-fold over districts);
*region extrapolation*, predicting a region none of whose districts were seen
(exhaustive leave-one-region-out); and *country transport*, predicting a country
held entirely out of training (leave-one-country-out). Every result is reported
against a named estimand.

**Two. Match the baseline's information to the model's.** A comparison is only
interpretable if the benchmark has seen what the model has seen. Our published
regional-mean baseline was computed from all of a region's respondents,
including those of the district being scored; the model never saw that district.
Every baseline is now recomputed under the same fold structure as the model, and
where a baseline is derived from the survey it is **jackknifed** so the scored
unit never contributes to it.

**Three. Replicate every random split.** A single random division of 14-87
districts is a high-variance draw. We repeat each fold assignment 10 times with
different seeds and report the median and the interquartile range across
replicates, never a single value.

**Four. Show the test could have detected an effect.** For every significance
screen we report attained power and, where the screen is a permutation test,
verify by construction that a true positive would clear the multiplicity
threshold at the permutation resolution used.

Two measurement changes follow from the audit. Effective sample size per
district is Kish's $n_\text{eff}$ divided by a **measured** cluster-robust
design effect rather than an assumed 1.5. And predictors are rank-normalised
(van der Waerden) **within country** before pooling, because the same indicator
is on different scales in different surveys and z-scoring does not remove that.

Predictors enter in three competing representations, so that model capacity is a
measured factor rather than a fixed choice: all 451 harmonised columns under
elastic net; 18 domain summaries formed as the leading principal components of
each domain to 80% of variance; and a zero-tuning **domain index**, a
marginal-correlation-weighted composite over those summaries with no
cross-validated hyperparameter. A covariate-free spatial smoother and the
information-matched survey baselines are scored on identical folds.

# Results

## Sample characteristics

The pooled admin-2 dataset comprised 206 administrative units across the
four countries (Gambia 30, Ghana 75, Sierra Leone 14, Malawi 90). Sample
sizes per country and biomarker are presented in @tbl-cohort.
Observed admin-2 deficiency prevalence varied widely across outcomes and
countries, with iron deficiency in children ranging from 5.8% (Sierra Leone)
to 42.0% (Gambia), and vitamin-A deficiency in women below 3% in all four
countries.

## Within-country admin-2 estimation

::: {.callout-warning}
## Corrected 2026-09-02 - the leading arm and the ceiling comparison are withdrawn

The individual-level arm reported below at median r **0.516** aggregates
person-level predictions to the district, and the person-level model was trained
on respondents from the district it is scored on. It is not an out-of-district
comparison and is withdrawn as a headline. The accompanying claim that it reached
92% of the reliability ceiling is void with that ceiling (@sec-ceiling). The
area-level comparisons in this section were also scored on a single fold draw;
see @sec-protocol-results for the replicated numbers, under which the ordering
changes.
:::

Under genuinely out-of-fold evaluation, within-country admin-2 ranking was
modest, and only one of the five compared approaches clearly beat a
national-mean null (@tbl-area-comparison). Aggregating the individual-level
SuperLearner's out-of-fold predictions to admin-2 gave the highest raw score
(median Pearson r 0.516 across 24 country–outcome cells; median MAE 8.55 pp),
but the folds split respondents rather than districts, so a district's own
respondents appear in training; the number is retained here for continuity with
the earlier report and should not be compared with the district-held-out
figures. Neither area-level
SuperLearner variant, fit directly on the admin-2 polygons, beat the null:
median r 0.16 (MSE loss) and 0.215 (NLL loss), with median MAE of 9.31 and
9.84 pp respectively — worse than the 9.55 pp a country-mean-everywhere null
achieves by construction. A covariate-free spatial-only smoother on admin-2
centroid coordinates outperformed both area-level covariate models on rank
correlation (median r 0.393, 18 evaluable cells) while remaining close to the
null on absolute error (MAE 9.98 pp). The spatial-smoother result replicates
under the corrected protocol and stands (@sec-protocol-results): at admin-2,
most of what the covariates supply within a surveyed country is already
available from "nearby districts are similar." What does not survive is the
comparison against the national-mean null, which was scored on one fold draw;
replicated, and against an information-matched survey baseline, the area-level
covariate models do beat it.

This reversal reflects the area-level p ≫ n setting. Because the proxy and
earth-observation predictors are resolved at the admin-2 level, the effective
sample size for area-level fitting is the number of polygons — 14 in Sierra
Leone, 30 in Gambia, 75 in Ghana, and 90 in Malawi — against
roughly 150 screened candidate covariates, leaving few residual degrees of
freedom. An earlier in-sample evaluation — predicting the training areas from
the fully-fitted ensemble — produced near-perfect correlations (r up to
1.00), but these were overfitting artifacts rather than transportable signal;
the present numbers use genuine out-of-fold prediction. The individual-level
model, fit on hundreds to thousands of individuals, generalises better once
aggregated than the *unregularised* area-level ensemble, and remains the
strongest tested approach even once uninformed baselines are added to the
comparison (below).

```{r tbl-area-comparison}
#| tbl-cap: "Within-country admin-2 prediction performance for five approaches: individual SuperLearner aggregated to admin-2, area-level SuperLearner with MSE loss, area-level SuperLearner with NLL loss, a covariate-free spatial-only smoother, and a national-mean null. Pearson r between predicted and observed admin-2 prevalence; MAE in percentage points; r_max is the estimated reliability ceiling for that cell (see Results) and signal flags whether the cell has any detectable signal above sampling noise. From results/tables/area_comparison_all.csv."
if (!is.null(area_compare)) {
  area_compare %>%
    select(country, outcome, approach, pearson_r, mae_pp, mean_bias_pp, r_max, signal) %>%
    arrange(country, outcome, approach) %>%
    kable(booktabs = TRUE, digits = 2,
          col.names = c("Country","Outcome","Approach","r","MAE (pp)","Bias (pp)","r_max","Signal?"))
}
```

The aggregated individual-level SuperLearner nonetheless exhibited a
characteristic compression pattern: predictions correlated moderately with
admin-2 prevalence but were systematically pulled toward the country mean.
For high-prevalence outcomes this produced large negative biases (e.g.,
Gambia child-iron; women-iron). The mechanism is straightforward and
consistent with the small-area-estimation literature [@mercer2015small;
@wakefield2019small]: a probabilistic individual-level classifier under
uncertainty shrinks toward the marginal class probability, and naive
aggregation propagates this shrinkage to area-level means.

## A reliability ceiling for admin-2 prediction {#sec-ceiling}

::: {.callout-warning}
## Corrected 2026-09-02 — the ceiling was biased low by about 4.7x

An earlier version of this section reported a median $r_{\max}$ of **0.098**
across the 24 country-outcome cells, with **4 of 24** cells said to carry no
detectable signal above sampling noise, and concluded that admin-2 was finer
than these surveys could resolve. That conclusion does not survive audit and is
withdrawn.
:::

The comparison above is bounded by how much of the observed between-district
variance is real signal rather than survey sampling noise. Our original estimator
for that bound divided a signal variance term by an assumed design effect of 1.5
and treated the residual analytically. Two things were wrong with it. The design
effect is not 1.5: measured directly from the cluster structure it ranges from
1.0 to 5.6 with a median of about 2.4, so the sampling-variance term was
systematically understated in the denominator and overstated in the numerator's
subtraction. And the estimator was never validated against a case with a known
answer.

The tell was in the output. Fully **12 of 24** cells returned an achieved
correlation at or above their own estimated ceiling — a physically impossible
result that we had explained away as estimation uncertainty. It was not
uncertainty; it was bias. Re-estimating the ceiling by split-half replication of
the survey respondents, which requires no design-effect assumption and which we
validated on simulated data with a known true ceiling, gives an empirical
$r_{\max}$ of **0.47 to 0.61** depending on outcome. The analytic estimator was
biased low by a factor of roughly 4.7.

The substantive consequence is the reverse of what we published. Admin-2 is not
beyond what these surveys can resolve, and the gap between achieved correlation
(0.29 to 0.40 under the corrected protocol, @sec-protocol-results) and the
attainable ceiling is real headroom rather than noise. We now report the
empirical split-half ceiling alongside every admin-2 correlation, and we
recommend against the analytic form: any ceiling estimator that returns achieved
performance above its own bound is reporting its own bias.

## The area-level prevalence recipe {#sec-recipe-audit}

We separately explored whether modelling admin-2 prevalence directly, with an
estimator matched to the area-level p ≫ n regime, could recover a stronger
within-country estimator than the unregularised area-level SuperLearner
above. Under its own district- and region-block cross-validation
(@tbl-area-recipe), the recipe reached point estimates comparable to or above
the individual-level aggregate for several cells. A subsequent audit of this
analysis, however, found that its central comparative claims do not survive
correction: the recipe's covariate pre-filter was originally derived once on
the full dataset before cross-validation, and re-deriving it honestly inside
each resampling fold reduced its apparent Spearman correlation by 0.16
(admin-2) to 0.30 (admin-1); and its best-performing cross-country comparator
(a spatial-plus-soil model reported in @tbl-benchmarks) had been selected
using LOCO correlations computed on the same held-out countries it was then
evaluated against, and drops from a correlation of 0.305 to 0.117 once
selection and evaluation are separated. We report the recipe's original
point estimates below for completeness and because several within-country
cells remain plausible even under the corrected, more conservative reading,
but we do **not** treat this recipe as the primary within-country estimator;
that role is filled by the individual-level aggregate above, benchmarked
against the national-mean null and the reliability ceiling. Under district-block
cross-validation the recipe reached Pearson r = 0.64 for child iron and 0.56 for
women's vitamin B12 in Ghana, 0.60 for women's iron in Gambia, with
permutation-null *p* < 0.05 for iron and several women's micronutrients in
Gambia and Ghana; under the more demanding region-block scheme the same cells
retained r = 0.20–0.55. The binomial count loss left district ranking
essentially unchanged relative to a Gaussian logit fit but markedly improved
calibration (slopes 0.5–1.1, versus systematic over-shrinkage under the Gaussian
loss). Enriching the earth-observation covariates with Malaria Atlas Project and
IHME anaemia/stunting layers recovered signal in the highest-malaria setting:
Malawi child-iron rose from null with soil/climate features alone to r = 0.20
(district, *p* = 0.03). The two opt-in variants behaved as designed — the
distributional estimator improved vitamin A (Gambia child vitamin A r 0.18 →
0.46; Gambia women vitamin A r 0.42), while domain-wise late-fusion stacking
helped distributed-signal outcomes (Gambia women iron r 0.65) but collapsed the
sparse Malawi child-iron signal (0.20 → 0.03), confirming that dimension
reduction must not be a default. Across all 18 evaluable country–outcome cells,
genuine permutation-significant within-country signal was concentrated in iron
(and several women's micronutrients) in Gambia and Ghana, was weaker but present
for iron and folate in Malawi, and was absent for zinc.

```{r tbl-area-recipe}
#| tbl-cap: "Area-level recipe within-country performance (binomial loss, enriched features), district- and region-block cross-validation. Pearson r with area-bootstrap 95% CI and permutation-null p; calibration slope. From results/tables/corrected/area_recipe_within.csv."
# here::here(), not a bare relative path. Quarto renders with the working
# directory at docs/, so "results/tables/..." resolved to docs/results/... which
# does not exist -- file.exists() was always FALSE, the chunk emitted nothing,
# and the table rendered empty behind a live @tbl-area-recipe cross-reference.
# Quarto's only complaint was "FloatRefTarget with no content", which is easy to
# scroll past. Every other chunk in this file already uses here::here().
.arp <- here::here("results", "tables", "corrected", "area_recipe_within.csv")
if (file.exists(.arp)) {
  read.csv(.arp) %>%
    transmute(Country = country, Outcome = outcome, Scheme = scheme, n = n_area,
              r = round(pearson_r, 2),
              `95% CI` = sprintf("[%.2f, %.2f]", pearson_ci_lo, pearson_ci_hi),
              `perm p` = round(pearson_perm_p, 3), Slope = round(calib_slope, 2)) %>%
    arrange(Country, Outcome, desc(Scheme)) %>%
    kable(booktabs = TRUE)
} else {
  stop("area_recipe_within.csv not found at ", .arp,
       " -- @tbl-area-recipe would render empty")
}
```

## Cross-country transportability

LOCO holdout produced markedly weaker performance than within-country
cross-validation (@tbl-loco; @fig-loco-scatter). Across the 16
outcome–holdout combinations, Pearson r had a median of 0.29 and ranged
from −0.21 to 0.75, with iron deficiency outcomes transporting best (median r
0.35 in children, 0.22 in women) and vitamin-A outcomes weaker (median r
0.24 in children, 0.37 in women), with anti-predictive ordering in some
holdouts.

The area-level recipe, applied with the universal (earth-observation-only)
feature set and scored by rank correlation, produced comparable point
estimates to within-country cross-validation (Spearman 0.42 and 0.49 for
child iron, 0.75 and 0.35 for women's iron, in Gambia and Ghana respectively)
but not to Malawi (Spearman ≈ 0.03–0.05), and vitamin A did not transport
well in any country. As noted above (@sec-recipe-audit), the recipe's
apparent edge over simpler comparators does not survive a leakage-corrected
re-ranking, so we read these numbers as descriptive of the recipe's own
LOCO behaviour rather than as evidence that it out-transports the benchmark
methods in @tbl-benchmarks. The malaria/IHME enrichment that aided
within-country prediction *degraded* transport (mean child-iron Spearman 0.30
with universal features versus 0.20 enriched), because the malaria–iron and
modelled-burden relationships are country-specific rather than universal.
Cross-country prediction is therefore, at best, a ranking-only product, valid
within an epidemiologically similar neighbourhood with adequate covariate
coverage and requiring national-anchor calibration to recover the absolute
level.

```{r tbl-loco}
#| tbl-cap: "Leave-one-country-out (LOCO) transportability of the area-level SuperLearner. From results/tables/transportability_area_loco_metrics.csv."
if (!is.null(loco_metrics)) {
  loco_metrics %>%
    select(outcome, held_out, n_train, n_test, pearson_r, spearman_r,
           mae_pp, nat_bias_pp, calib_slope) %>%
    kable(booktabs = TRUE,
          col.names = c("Outcome","Held out","n train","n test",
                          "r","ρ","MAE (pp)","Bias (pp)","Calib slope"))
}
```

```{r fig-loco-scatter}
#| fig-cap: "Observed vs predicted admin-2 deficiency prevalence under leave-one-country-out cross-validation, by outcome and held-out country. The reference line is y = x. Predicted prevalences are from the area-level SuperLearner trained on the remaining countries (three for iron, measured in four countries; four for vitamin A, measured in five)."
# Generated by scripts/fig_loco_scatter.R from
# results/tables/transportability_area_loco_predictions.csv. Panel Pearson r
# matches @tbl-loco (same unweighted definition).
knitr::include_graphics(
  here::here("results", "figures", "loco_scatter.png"),
  error = FALSE)
```

Spatial patterns transport more readily within the West African
geography. Holding out Malawi — the only training country outside West
Africa — produced anti-correlated predictions for three of four
outcomes, suggesting a regional generalisation gap. Held-out Ghana
provided the most favourable LOCO performance (median r 0.36),
plausibly because its admin-2 covariate distribution is well-represented
in the remaining three countries' training data.

National-level aggregation is consistently and substantially more accurate
than admin-2. Predicted national prevalence fell within the survey's own 95%
confidence interval for all **24 of 24** country–outcome combinations
(@tbl-national-estimates), with a median absolute difference of **0.74**
percentage points (median relative difference **8.6%**). Aggregating noisy
district-level predictions to the national total averages out much of the
admin-2 sampling noise quantified in @sec-ceiling: the same models that
discriminate weakly *between* districts recover the national aggregate
accurately, and we treat national-level prevalence — not the admin-2 map —
as the deliverable best supported by the present evidence, with sub-national
maps presented as exploratory and uncertainty-aware.

However, when the model is *transported* across countries the held-out
national estimate is systematically biased toward the pooled training
mean (@fig-transport-calibration). The bias is outcome-specific and
concerns the national *level*, not the within-country admin-2 ranking
of @tbl-loco. Two outcomes transport with broadly correct calibration:
vitamin-A deficiency, whose national prevalence sits close to the
cross-country mean in every survey (per-outcome calibration slope 0.93
in children, 0.65 in women), and — once its cutoff is harmonised to the
common serum definition (folate < 10 nmol/L; Malawi's native < 3 ng/mL
rule is otherwise non-comparable) — folate deficiency, whose held-out
prevalence preserves the correct ordering though compressed in magnitude
(slope 0.37; Ghana 55%→18%, Sierra Leone 79%→30%, Malawi 2%→1%). The
remaining outcomes fail. Iron deficiency, whose national prevalence
ranges from 6% (Sierra Leone children) to 64% (Gambia children), is
pulled to the training mean so strongly that the ordering inverts
(per-outcome slope −0.72 in children, −0.18 in women), over-predicting
low-prevalence countries (Sierra Leone children 6%→40%, Malawi 10%→51%)
and under-predicting high-prevalence ones (Gambia children 64%→22%).
Vitamin B12 — despite being the strongest *within-country* signal
(continuous cross-validated R² up to 0.23 in Malawi) — does not
transport at all: held-out discrimination falls to near chance
(AUC 0.53–0.59) and prediction becomes unstable under the rare-outcome
class-weighting (held-out Malawi predicted at 80% against a true 2.8%).
Cross-country transport failure is therefore a level-calibration error
driven by between-country prevalence heterogeneity — most severe for the
outcomes and countries that differ most from the training distribution —
and within-country accuracy is a poor guide to which outcomes transport.

```{r fig-transport-calibration}
#| fig-cap: "Cross-country transport calibration. Predicted versus true national deficiency prevalence for the pooled individual-level SuperLearner under leave-one-country-out (LOCO) holdout — one panel per outcome, one point per held-out country (trained on the remaining countries), panels ordered worst to best calibration. The dashed line is perfect calibration (y = x); each panel header reports the fitted calibration slope (1.0 = perfectly calibrated; <1 = predictions compressed toward the cross-country mean; <0 = inverted ordering). Vertical segments mark each fold's calibration gap. Vitamin-A national prevalence, close to the cross-country mean, is well recovered and folate (harmonised to a common serum < 10 nmol/L cutoff) preserves the correct ordering; iron is pulled to the training mean and inverts, and vitamin B12 does not transport (near-chance discrimination, with unstable prediction under rare-outcome class-weighting). Folate and B12 are women-only and available in three countries; zinc is Malawi-only and cannot be assessed for transport. From results/tables/transportability_loco_results.csv."
knitr::include_graphics(
  here::here("results", "figures", "transport_calibration_loco_faceted.png"),
  error = FALSE)
```

**Parsimonious model with soil micronutrient predictors.** A series of
post-hoc analyses confirmed that the strongest transportable signal in
the existing 152-variable GEE feature pool is concentrated in the
SoilGrids covariates. Univariate leave-one-country-out scoring of each
of the 152 GEE features identified eight SoilGrids variables (soil
aluminium and calcium standard deviation at two depth profiles, soil
magnesium standard deviation, soil zinc mean and standard deviation,
soil total carbon mean) as the most consistently transportable single
predictors; no NDVI, accessibility, night-lights, or population-density
variable cracked the top eight in the per-variable LOCO ranking. A
weighted generalised additive model with a thin-plate spline on
admin-2 polygon centroid coordinates plus these eight soil features as
linear terms achieved mean LOCO Pearson r = 0.330 across 16
outcome–holdout combinations on the four main outcomes — substantially
above any other parsimonious comparator we tested and roughly
comparable to the area-level SuperLearner ensemble at a fraction of
the compute cost. Bootstrap 95% confidence intervals on the
combination's Pearson r were strictly above zero for seven of the
sixteen holdouts, with the wins concentrated on iron and vitamin A
in Gambia and Ghana. We caveat that the eight-feature recipe and the
spatial gradient information may not generalise outside the West
African training footprint — extending the training set to East and
Southern Africa is the most important external-validity step.

**Sierra Leone testing power at admin-3 resolution.** The admin-2
Sierra Leone LOCO sample (n = 14 districts) is small enough that the
holdout estimates are dominated by sampling noise; we documented this
by re-running the analysis at admin-3 (chiefdom) resolution after
spatial-joining DHS cluster coordinates to GADM admin-3 polygons. At
admin-3 resolution, Sierra Leone has 44 chiefdoms with biomarker
measurements (a 3.1-fold increase over admin-2). With Sierra Leone
held out as the test country and the other three countries as training
at admin-2 resolution, LOCO Spearman correlations on the four
continuous biomarker outcomes (child / women vitamin A, women B12,
women folate) moved from a range of −0.07 to −0.35 at admin-2 toward
or above zero at admin-3, including a meaningfully positive +0.30 for
women's vitamin B12. We interpret the headline admin-2 anti-predictive
estimates for Sierra Leone as primarily a small-sample artefact and
not as evidence of fundamental non-transportability.

```{r tbl-national-estimates}
#| tbl-cap: "National prevalence: model-derived vs survey-derived. From results/tables/national_estimates_all.csv."
if (!is.null(natl_est)) {
  natl_est %>%
    transmute(Country = country, Outcome = outcome,
              `Observed (95% CI)` = sprintf("%.1f (%.1f–%.1f)%%",
                obs_prev*100, obs_lo*100, obs_hi*100),
              `Predicted` = sprintf("%.1f%%", pred_prev*100),
              `Abs diff (pp)` = round(abs_diff*100, 1)) %>%
    kable(booktabs = TRUE)
}
```

## Can a predicted national level replace the missing survey? {#sec-national-composition}

@sec-anchoring shows that a district map's level should come from a survey. A
country without one therefore needs its level from somewhere else, and the
natural candidate is a model fitted to national data. We built that model on the
WHO Vitamin and Mineral Nutrition Information System [@who_vmnis], where the unit
of observation is the country-survey rather than the district: after joining to
World Bank national indicators, the vitamin A / preschool-children panel carries
108 surveys across 69 countries, an order of magnitude more design points for
the between-country relationship than aggregating a four-country district model
can offer. Evaluation is leave-one-country-out, so no survey from the target
country contributes at any date.

The model is competent at the task it was built for. On that panel it reaches
$r = 0.655$ against a variance-component noise ceiling of 0.818 — 80% of what is
attainable — where the ceiling separates between-country signal from
survey-method heterogeneity, residual temporal change and binomial sampling
error, using the sample sizes VMNIS records. We note that two ceilings for this
panel have previously been quoted (0.66 and 0.505); they differ only in whether
repeat surveys are grouped by country-year or by country-year-*method*, an
arbitrary choice that decides what counts as error, which is why we estimate the
components directly instead.

Composing nonetheless fails. We shifted each transported district map onto a
predicted national level and scored it against the survey's own district
prevalences, alongside a no-anchor arm, a no-covariate national level, and an
oracle arm given the survey's true national total. Because VMNIS holds no iron
or ferritin panel and the transport predictions cover iron and vitamin A, the
comparison is available for vitamin A only, on 2 outcomes × 4 countries — and
vitamin A / preschool children is VMNIS's strongest panel, so this is an upper
bound rather than an average.

| Arm | MAE (pp) | Mean \|bias\| (pp) | Better than no level |
|:----|---------:|-------------------:|:---------------------|
| Transported, no level | 5.81 | 3.35 | — |
| \+ national null level | 8.22 | 4.69 | 1/8 |
| \+ VMNIS predicted level | 12.70 | 10.03 | 1/8 |
| \+ true national level (oracle) | 5.58 | 0.73 | 4/8 |

: Composing a predicted national level onto a transported district map. {#tbl-national-composition}

Two failures compound. First, the predicted level is on a different scale from
our surveys: 40.9 pp predicted for Sierra Leone against a true 12.2, and 31.2
against 9.2 for Malawi, with the no-covariate null beating the covariate model
in 6 of 8 cells. This is not a weak model but a model of a different quantity —
the same cross-survey level offset documented in @sec-zscore, since VMNIS
records what national surveys report under their own assays and inflammation
adjustments.

Second, and decisively, **even a perfect national level buys almost nothing**:
5.81 to 5.58 pp, better in only 4 of 8 cells. The ceiling on this entire
approach is therefore about a quarter of a percentage point, and no further
investment in national modelling can raise it. The reason is the one
@sec-anchoring already gives from the other direction: a national number
displaces every district equally, whereas the anchor that pays corrects region
by region — and that requires a survey, which is precisely what the target
country lacks. The national model remains useful for ranking countries; it
cannot supply a district map's level.

## Benchmark comparison against canonical methods

Twenty-one comparator methods were evaluated alongside the prescreened
SuperLearner ensemble (`sl_prescreened`) across the 16 LOCO
outcome–holdout cells, spanning frequentist regressions (OLS/GLM,
quasi-binomial, beta regression, penalised/elastic-net), Bayesian SAE
(Fay–Herriot, BYM2), covariate-invariance and transportability-aware methods
(DAG-based selection, CORAL domain adaptation, forward LOCO-optimised
selection, a two-stage country-mean-plus-deviation model, a stacked
ensemble of the above), and two spatial methods with no or minimal
covariates (a coordinate-only smoother, `spatial_coords`, and a smoother
plus the eight most-transportable soil covariates, `spatial_plus_soil`)
(@tbl-benchmarks). The prescreened SuperLearner ensemble matched or exceeded
the best of these 21 comparators on **0 of 16** evaluable holdouts (by
Pearson r), and had the lowest median Pearson r (0.001) of any comparator
method. The two spatial methods led on rank correlation
(`spatial_plus_soil` median r = 0.28, `spatial_coords` r = 0.27), consistent
with the admin-2 spatial-only result above (r = 0.39) — geography, not the
covariate ensemble, carries most of the transportable ranking signal. On
absolute error, however, a plain penalised regression had the lowest median
LOCO MAE (10.6 pp) — only modestly better than the national-mean null itself
(11.1 pp) — despite a near-zero median Pearson r (0.09), while the spatial
methods' better ranking came with *worse* absolute error (MAE 11.6–13.1 pp).
Level and pattern are therefore transported best by different methods, and
neither is transported well by any single method, including the SuperLearner
ensemble. As noted in Methods, we treat `spatial_plus_soil`'s apparent lead with
caution: it was originally selected using LOCO correlations computed on the
same held-out countries it is evaluated against here, and a leakage-corrected
re-derivation drops its correlation from 0.305 to 0.117 (@sec-recipe-audit) —
still consistent with spatial information leading the field, but by a
smaller margin than @tbl-benchmarks alone would suggest.

```{r tbl-benchmarks}
#| tbl-cap: "LOCO transportability: prescreened SuperLearner vs benchmark methods. Pearson r between predicted and observed admin-2 prevalence. From results/tables/benchmarks_all.csv."
if (!is.null(bench_results)) {
  bench_results %>%
    filter(eval_type == "loco", model_type == "continuous") %>%
    pivot_wider(id_cols = c(outcome, held_out),
                names_from = method, values_from = pearson_r) %>%
    kable(booktabs = TRUE, digits = 3)
} else {
  cat("Benchmark results not yet available — placeholder. See preliminary",
      "smoke-test results in the discussion.")
}
```

## Supplementary all-methods comparison on a representative outcome

To provide a comprehensive supplementary comparison, we ran every method
and SuperLearner library variant tested in this work on a single
representative outcome — child iron deficiency, chosen because it has
the strongest transportable signal in our data and biomarker
measurements in all four countries. Two evaluation tracks were used:
within-country five-fold cross-validation on each country's admin-2
polygons (separately per country) and leave-one-country-out
transportability across the four-country pool (@fig-all-methods-iron).

```{r fig-all-methods-iron}
#| fig-cap: "Supplementary all-methods comparison for child iron deficiency. (a) Within-country 5-fold cross-validation Pearson r per method, with one dot per country and X-marks denoting the mean across the four countries. (b) Leave-one-country-out transportability Pearson r per method, with one dot per held-out country. Methods are ordered by mean LOCO r. The supplementary figure source data are in results/tables/all_methods_child_iron.csv."
#| fig-width: 11
#| fig-height: 6
fig_path <- here::here("results","figures","all_methods_child_iron.png")
if (file.exists(fig_path)) knitr::include_graphics(fig_path)
```

For LOCO, spatial smoothing on admin-2 polygon centroids (mean Pearson r
= 0.48) and spatial + top-8 SoilGrids covariates (0.46) were the two
strongest methods, both substantially outperforming the prescreened
SuperLearner ensemble (0.22) and the simpler invariance-filter and OLS
recipes (0.39 each). Within-country, the picture flips: nonparametric
flexibility helps when training data is in-distribution, with random
forest (mean Pearson r = 0.23 across four countries) and the
prescreened SuperLearner (0.17) outperforming the linear methods, while
spatial-only and spatial + soil predict best on the two countries with
enough admin-2 polygons (Ghana and Malawi) but are unstable on the
small (n = 14) Sierra Leone CV split. The conclusion is consistent
across outcomes (see Methods §Estimation strategies and the exploratory
findings logs, `archive/docs/exploratory_loco_findings.md`): non-linear ensembles help
within-country, parsimonious linear models with soil micronutrient features
help out-of-country.

## Predictor domain contributions and learner-stack composition

Domain ablation — refitting with each predictor domain removed and recording the
change in binary AUC — initially appeared to show large contributions, with mean
drops of 0.10–0.15 attributed to survey indicators, infant-and-child-mortality,
HIV/AIDS and agriculture. Those figures are artifacts and we report the
diagnosis rather than the numbers. In two cells — Ghana women's vitamin A (1.5%
prevalence) and Malawi women's B12 (2.3%) — removing *any* domain, including a
domain consisting of a single variable, collapsed AUC from 0.889 and 0.911 to
exactly 0.500. That is the ensemble failing to fit at all once a predictor is
withdrawn, not information being lost, and it is the same rare-outcome fragility
that appears elsewhere in these results. Those two cells contribute 35 of 332
ablation rows at losses of 0.39–0.41 and inflate every domain's mean.

Excluding them, domain removal costs almost nothing. The median AUC drop is
0.029 for environment (57 variables), 0.004 for climate (21), and $\le$ 0.001
for every other domain; **141 of 332 domain removals improved AUC**. At the
single-variable level, **266 of 720 permutation importance drops were $\le 0$**,
the median top-five drop was 0.008 AUC, and eighteen different variables took
first rank across the 24 cells — no predictor is consistently important. SHAP
attributions agree, with top contributions of approximately 0.004 on the
probability scale.

Calibration of the binary classifiers was poor before post-hoc
correction (median slope 0.44, IQR 0.10–0.76), consistent with @tbl-calibration
below — predictions are systematically over-extreme relative to observed risk
and require shrinkage before use.

## Is any predictor associated with the outcomes? {#sec-fdr}

::: {.callout-warning}
## Corrected 2026-09-02 — the screen could not have returned a positive

An earlier version of this section reported that **no predictor survived FDR
control in any cell** under a spatially restricted permutation null, and read
that as evidence against covariate signal. The screen could not have produced a
positive result at any effect size, so its zero carries no information about the
data. The claim is withdrawn and replaced below.
:::

We tested every harmonised predictor against district prevalence in every cell
(7,104 predictor $	imes$ cell tests) under two nulls. The naive correlation test
treats districts as independent. Because they are not — two smooth surfaces
correlate by construction — we also used a restricted permutation that shuffles
the outcome *within* admin-1 regions, holding the regional pattern fixed and
asking what a predictor adds on top of geography.

Under the naive test 1,209 tests (17.0%) reached $p<0.05$ against 355 expected by
chance, a 3.4-fold enrichment, and 310 survived Benjamini-Hochberg control across
all cells. Under the spatially restricted null the enrichment fell to 1.5-fold
(519 tests, 7.3%) and no predictor survived FDR control in any cell.

**Why that zero is uninformative.** The restricted permutation used 999
shuffles, so the smallest attainable p-value is $1/1000$. Applying
Benjamini-Hochberg across 7,104 tests requires a p-value below roughly
$0.05 	imes 1/7104 pprox 7 	imes 10^{-6}$ for the most significant test to
clear the threshold — three orders of magnitude smaller than the resolution of
the permutation distribution. **No configuration of the data, at any effect
size, could have produced a discovery.** We confirmed this by construction: with
a synthetic predictor set to reproduce the outcome exactly, the screen still
returns zero discoveries. A test that cannot reject under any alternative has
power zero, and its failure to reject is not evidence of absence. Median power
across the real cells was 0.005.

**What we report instead.** At 14 to 87 districts per country, no per-cell test
of 451 predictors will have usable power after multiplicity correction, and the
correct response is to change the evidence standard rather than to run the
underpowered test more carefully. We therefore report **cross-country
replication of sign**: whether the same indicator, against the same outcome,
points the same way in four independent national surveys conducted by different
teams in countries with different diets and different agriculture. Under the
null of no association, four-of-four sign agreement has probability $2^{-3} =
0.125$ per indicator before any effect-size consideration, and the pattern of
which indicators agree is itself interpretable in a way a starred p-value is
not.

On that standard the associations are nutrient-specific and replicate, but they
do not simply recover individual-level nutritional biology, and reporting them
as though they did would be an ecological fallacy (@tbl-nutrient-signal).

The strongest replicated association for child vitamin A is district legume
consumption in the direction of *more* deficiency (pooled z 4.87, 4 of 4
countries) — the opposite of what provitamin A carotenoid intake predicts at the
individual level. Child iron shows the same pattern: cattle ownership tracks
*more* deficiency (z 4.40, 3 of 3), not less, alongside cereal-dominant cropping
(z 4.27, 4 of 4) and, inversely, root-crop share (z −3.73). Women's iron and
women's vitamin A are dominated by warm night-time land surface temperature and
soil aluminium heterogeneity, both toward more deficiency in all four countries.

The coherent reading is that the replicated axis is **rural subsistence
agro-ecology**, not nutrient intake. Districts where households keep cattle and
goats, grow and eat legumes, and farm cereals are the districts that are rural,
poor and deficient, so the district-level correlation carries the opposite sign
to the individual-level dietary relationship. This is a textbook
ecological-fallacy signature, and it is visible here only because we required
replication in direction rather than significance in a single country. It does
not undermine the predictive results — a marker of a deficient district is
useful for targeting whether or not it is causal — but it forbids reading
@tbl-nutrient-signal as a set of intervention targets.

Two associations are both replicated and interpretable at the district level,
where the exposure is genuinely an area property. District child wasting tracks
women's vitamin A deficiency (z 4.62, 4 of 4) and precipitation tracks less of
it (z −3.92): the two forms of undernutrition co-locate, which is a targeting
statement rather than a mechanistic one.

Malaria was the one domain we expected to matter for iron, and it did not. No
malaria *burden* indicator replicates against any outcome at family-wise
significance; the strongest malaria-domain signal is indoor residual spraying
*coverage* against child iron (z 3.66, family-wise p 0.14), an indicator of
where control programmes operate rather than of transmission.

Penalised regression is consistent with a distributed rather than sparse signal.
Elastic net with cross-validated $\lambda_{1\text{se}}$ retained a median of zero
predictors — which under this geometry indicates that no small subset carries
the association, not that no association exists. The omnibus dense-signal test
(@sec-protocol-results), which is powered against many small effects rather than
a few large ones, rejects where the sparse screen does not.

## Does the geographic unit change the answer? {#sec-resolution}

If the binding constraint is the precision of the district target rather than
the modelling, then coarsening the unit should help: admin-1 regions pool more
respondents, so the target is measured better, even though far fewer units
remain. We tested this by running the same three arms — a training-mean null, a
covariate-free spatial smoother, and the covariate ensemble — through identical
leave-one-unit-out folds at both resolutions, holding learner, screening,
weighting and scoring fixed so that only the unit changes.

At admin-2 the covariate ensemble and the spatial smoother are barely
distinguishable (mean $r$ 0.156 vs 0.137), and the smoother has the lower mean
absolute error (9.39 vs 10.51 pp); covariates beat geography in only 10 of 24 [corrected 2026-09-01; the source table gives 10, not 6]
cells. At admin-1 the ensemble roughly triples its rank correlation and pulls
clearly ahead (0.437 vs 0.209, 7 of 14 cells). Only Ghana (16 regions) and
Malawi (27) are estimable at admin-1; The Gambia (6 regions) and Sierra Leone
(4) fall below any reasonable threshold for leave-one-out estimation and are
reported as not estimable rather than dropped silently.

A hybrid that selects the unit **a priori from the number of regions** — admin-1
where a country has at least eight, admin-2 otherwise, a rule that depends only
on how a country is subdivided and never on performance — reaches mean $r$ 0.304
and 7.25 pp across all 24 cells, roughly double the admin-2-everywhere
correlation while retaining all four countries. Selecting the unit by which
performed better would be selection on the test data, of exactly the kind
quantified in @sec-fdr.

The most programmatically legible form of this result classifies districts into
the WHO public-health categories for child vitamin A. At admin-2 the covariate
model assigns the correct category 33% of the time against **32% for assigning
every district the national prevalence** — no better than the null. At admin-1
the same comparison is 56% against 35%. The admin-1 figure reproduces the 56%
reported in an earlier project presentation.

We attach two cautions. The largest admin-1 gains occur on the rarest outcomes,
where admin-2 districts contain 0–2 cases and admin-1 regions are the first unit
at which cases accumulate; much of the advantage is therefore a sample-size
effect rather than evidence that covariates are more informative at that scale.
And on absolute error rather than ranking, covariates beat geography in only 7
of 14 admin-1 cells.

## Where the level comes from {#sec-anchoring}

::: {.callout-warning}
## Corrected 2026-09-02 - the anchoring gain is withdrawn

An earlier version of this section reported that anchoring each district to its
region's design-based survey total lifted mean rank correlation from 0.164 to
0.413, cut mean absolute bias from 3.24 to 1.59 pp, and was better in 20 of 24
cells, and concluded that this mattered more than any modelling choice we
tested. Those arms computed each region's anchor from **all** of the region's
respondents, including those of the district being scored, so a district was
partly anchored to its own answer. Measured symmetrically the gain disappears.
:::

The preceding sections ask which unit to estimate at. A separate question is
where a district estimate's *level* comes from, as distinct from its *pattern*.
Each anchor rescales the same district predictions so that they average, within
each region, to the design-based survey estimate for that region; the rescaling
is a one-parameter shift on the logit scale, so it leaves the district ranking
untouched by construction.

```{r}
#| label: tbl-anchor-arms
#| tbl-cap: "Where a district estimate's level comes from. All arms are scored
#|   on the same cells under the same leave-one-region-out folds, and every
#|   anchor is a monotone shift, so the arms differ only in level. Produced by
#|   scripts/covariates/08_admin1_arms.R."
#| echo: false
anchor_path <- here::here("results", "tables", "admin1_arms.csv")
if (file.exists(anchor_path)) {
  knitr::kable(table_anchor_arms(), align = "lrrrr")
} else {
  cat("Anchoring comparison not available; run scripts/covariates/08_admin1_arms.R.")
}
```

We added a **jackknifed** arm in which a district never contributes to the
regional total it is anchored to, holding the fit, the folds and the cells
identical. Under that symmetric comparison anchoring gives mean rank correlation
**0.147** against **0.156** for no anchor at all, and is better in **8 of 24**
cells rather than 20 - that is, it is very slightly worse than not anchoring, and
no better than a coin flip across cells. A split-sample arm, which anchors to a
random half of the region's respondents and scores on the other, agrees. The
apparent gain was the district's own survey answer re-entering through the
anchor.

Anchoring to a single **national** total, which cannot leak in this way because
it contains no district-specific information, adds essentially nothing to rank
correlation (0.170) and makes bias worse (5.85 pp) - a result that stands.

Two conclusions survive. Level and pattern remain distinct quantities, and the
covariates support the pattern far better than the level, which is why a
transported map should be read as a ranking (@sec-protocol-results). But
regional anchoring is not the free improvement we reported, and we now regard
the honest within-country comparison as the one in @sec-protocol-results, where
the covariate model is scored against a jackknifed regional mean rather than
against an anchor built from the answer.

Beyond this specific case, the general lesson is a protocol one: any evaluation
in which a baseline or a benchmark is computed from data that includes the unit
being scored will report a gain that is partly the unit's own answer. The
jackknifed arm cost one additional run and reversed the finding.

## Would a household survey do better? {#sec-anchor-individual}

A district correlation near 0.15 is uninterpretable on its own. Against a noise
ceiling it may be most of what is attainable; against a questionnaire model
reaching 0.6 it would be a poor showing. The comparison that settles this
replaces the geospatial proxies with the household questionnaire actually
administered to these same respondents — wealth, education, diet, water and
sanitation, recent illness, supplementation; roughly 330 additional columns —
while removing every blood-assay column, so that the arm answers what a survey
*without a blood draw* would buy. Both arms use the same learner, the same rows
and the same leave-one-region-out folds.

The questionnaire moves mean district $r$ from 0.154 to 0.228 (median 0.126 to
0.264; MAE 8.33 to 7.92 pp). It is better in 10 of 16 cells, clears $r = 0.4$ in
only 3, and in Malawi adds nothing measurable at all — gains of 0.000 to 0.004
across all four outcomes.

This is the most informative negative result in the paper, because it excludes
the explanations a null result of this kind ordinarily invites. The ceiling is
not an artefact of imperfect geospatial linkage, nor of the inflammation
adjustment, nor of overfitting the proxy set: a questionnaire administered to
the same individuals, carrying no geospatial information whatsoever, also fails
to locate deficiency at district resolution. The signal is weak at this
resolution for any predictor set we can construct, and the geospatial proxies
recover most of the little that is there.

The same run tests whether linking at the survey **cluster** rather than the
district recovers anything, on the reasoning that the effective sample size for
learning the predictor–outcome map is the number of clusters rather than the
number of individuals, and that the gain should therefore be largest where
clusters most outnumber districts. It does not: mean gains are +0.017 (The
Gambia), −0.003 (Ghana), −0.002 (Malawi) and −0.025 (Sierra Leone). Sierra
Leone, with 58 clusters against 14 districts, is the country where the argument
predicted the largest gain and delivers the largest loss. Additional units do
not help when each unit's outcome rests on a handful of respondents, because the
units become noisier in the same proportion as they become more numerous.

## Modelling the biomarker rather than the binary {#sec-continuous}

Every district outcome above is a prevalence — the share of respondents below a
cutoff — and dichotomising discards information, most severely when the outcome
is rare. We therefore tested modelling the district mean of the log biomarker
and converting to prevalence afterwards, against the binary model, in 22 cells
under identical leave-one-region-out folds, covariates and learner stack.

Converting a *predicted* mean gave a mean absolute error of 10.80 pp against
9.40 pp for the binary model, with a marginally better rank correlation (0.196
vs 0.168). Neither beat a national-mean null (8.06 pp), which the binary model
beat in 4 of 22 cells and the continuous model in 2.

The informative arm is a third one, which converts the *observed* district mean
through the same transformation and therefore performs no prediction at all.
It achieves 7.49 pp and $r = 0.672$, beating every other arm including the null
and beating the binary model in 14 of 22 cells. The mean-to-prevalence
conversion is therefore not the limitation: a well-predicted district mean would
convert to prevalence more accurately than anything the current models produce.
The deficit lies entirely in predicting the mean, which fails for the same
reason prevalence prediction fails — in 9 of 22 cells the root-mean-square error
of the predicted district mean exceeded the between-district standard deviation.
A fourth arm using each district's own within-district standard deviation rather
than a pooled one raised rank correlation from 0.196 to 0.434; that arm uses
held-out information and is reported only as a bound, but it indicates that
within-district spread carries information a single pooled standard deviation
discards.

Stratifying by prevalence supports the motivating intuition with one
qualification. For outcomes below 5% prevalence the continuous route reduced
mean absolute error from 4.61 to 2.99 pp, while for mid-range and common
outcomes it was worse (8.77 to 11.86 pp and 15.25 to 16.84 pp). That average is
dominated by a single cell: Sierra Leone women's B12, at 0.61% prevalence, where
the binary model produced a 7.67 pp error — implausible for an outcome of that
magnitude — against 0.53 pp for the continuous route. Excluding it, the
continuous approach wins three of the remaining five rare cells by modest
margins. We therefore regard modelling the continuous biomarker as worthwhile
specifically for rare outcomes, where the binary model can fail badly, and not
as a general replacement.

## What the model should be asked to carry across a border {#sec-zscore}

Since district ranking transports but the absolute level does not, a natural
step is to stop asking the model to carry the level at all. We tested two
versions of that idea and they gave opposite answers, which is itself
instructive.

Fitting each district's deviation from its own country's national prevalence —
removing the level but leaving the between-country spread in the target —
does not help. Rank correlation fell from 0.201 to 0.140, and the apparent
improvement in absolute error was entirely attributable to being handed the
held-out country's national prevalence: an arm that predicts that national
figure for every district, using no covariates whatsoever, achieved a lower
mean absolute error (9.85 pp) and lower bias (5.68 pp) than any covariate arm,
and the deviation model beat it in only 4 of 16 cells.

Removing the between-country *spread* as well is what works. Fitting the
within-country z-score of the transformed prevalence makes both level and
spread structurally unlearnable, so the model can carry only spatial pattern;
both are restored at prediction time from an anchor. Across the four shared
outcomes this raised mean Spearman correlation from 0.146 to 0.199 (+0.053,
better in 3 of 4 outcomes), and reduced mean absolute error from 13.8 to 10.0
pp and mean absolute national bias from 10.2 to 5.9 pp.

Those two families of improvement have different standing and should not be
quoted together. Rescaling by a positive constant and adding an offset cannot
reorder predictions, so the rank gain follows from fitting a different target
and is a property of the model. The larger error and bias gains follow mostly
from the anchor, which requires the held-out country's national prevalence —
a realistic requirement, since a national figure typically exists where
district-level biomarker data does not, but an assumption about deployment
rather than evidence that the covariates improved.

Two further changes suggested by the same exploratory work did not replicate on
the harmonised covariate set and we report them as negative results. Fitting on
the raw prevalence scale rather than the clamped logit scale, which had been
expected to help because 8 of 24 cells have more than 40% of districts observed
at exactly zero prevalence, gave −0.020 mean Spearman (better in 2 of 4).
Adding district centroid coordinates to the predictor set gave 0.000 (better in
1 of 4); with a top-30 correlation prescreen over 294 predictors, two
coordinate columns rarely survive selection, whereas the exploratory result had
come from a curated 16-variable set in which coordinates were a large share of
the model.

We also directly tested whether the full-mode SuperLearner's 13th candidate
learner, a Gaussian process (`kernlab::gausspr`), was worth its
$O(n^3)$ computational cost, which made the largest pooled LOCO fits
(10,011 and 13,107 rows) effectively non-terminating. Refitting Ghana child
iron — arbitrarily chosen as a representative, well-powered cell — with and
without the Gaussian process in the candidate library gave an identical
binary AUC to 16 decimal places (0.7015504) and a negligible continuous-r
difference (0.2488 vs 0.2540): the Gaussian process received effectively
zero ensemble weight. We therefore dropped it from the full-mode stack
(Methods), retaining the check as a pipeline target rather than removing the
learner on assumption alone.

## Calibration and reliability of binary classifiers

For binary deficiency classifiers, several diagnostic flags emerged
(@tbl-calibration). The Brier skill score was negative — indicating
predictions worse than the no-skill baseline — for 12 of 24
country–outcome combinations, though in most the deficit was small (Brier
skill > −0.04 for 9 of the 12). Post-hoc Platt recalibration (see Methods) eliminated the
negative values entirely: all 24 calibrated classifiers achieved
non-negative Brier skill. Calibration slopes were
strongly inverted
(< 0) for 5 of 24 models — in Gambia, Ghana (×2), Sierra Leone, and Malawi —
particularly at very low prevalence (vitamin A in women, vitamin B12 in
women) where AUC remained high (> 0.7 for two of the five: Ghana women's
vitamin A 0.75, Sierra Leone women's B12 0.75) but Brier skill was near zero.
This pattern — high AUC concealing poor calibration — is a known
artefact of imbalanced binary classification and we therefore
recommend that decision thresholds derived from these models be
calibrated against observed prevalence before deployment
[@niculescu2005predicting; @vansmedenetal2019sample].

```{r tbl-calibration}
#| tbl-cap: "Binary classifier calibration. Brier skill > 0 indicates information gain over a no-skill baseline. From results/tables/diagnostics_binary.csv."
if (!is.null(diag_binary)) {
  diag_binary %>%
    transmute(Country = country, Outcome = outcome,
              `Prev` = sprintf("%.1f%%", prevalence*100),
              AUC = sprintf("%.2f", roc_auc),
              `Brier skill` = sprintf("%+.3f", brier_skill),
              `Calib slope` = sprintf("%+.2f", calib_slope),
              ECE = sprintf("%.3f", ece)) %>%
    kable(booktabs = TRUE)
}
```

## Honest evaluation: the corrected analysis of record

The metrics in this section are the analysis of record, produced by the
leakage-free corrected (P1–P8) methods described in Methods; the
production-pipeline equivalents elsewhere are retained for comparison.

**Within-country discrimination shrinks under honest cross-validation.**
Replacing random cross-validation and full-data preprocessing with
cluster-blocked and spatial-block folds and strictly in-fold preprocessing
lowers within-country discrimination, confirming that the optimistic design
overstated performance (@tbl-cv-honesty). Across the four primary outcomes the
honest (cluster-block) AUC is typically a few points below the optimistic AUC
and frequently near chance, consistent with the modest area-level signal seen
throughout.

```{r tbl-cv-honesty}
#| tbl-cap: "Within-country binary discrimination (AUC) under optimistic versus honest cross-validation, from the corrected methods (results/tables/corrected/cv_honesty_compare.csv). Honest schemes block folds by survey cluster and by spatial region and confine all preprocessing to the training fold."
if (!is.null(corr_cv)) {
  lab <- c("random CV + full-data preprocessing (optimistic)" = "Optimistic",
           "cluster-block CV"                                 = "Cluster-block (honest)",
           "region/spatial-block CV"                          = "Spatial-block (honest)")
  corr_cv %>%
    mutate(scheme = lab[scheme]) %>%
    filter(!is.na(scheme)) %>%
    select(country, outcome, scheme, auc) %>%
    tidyr::pivot_wider(names_from = scheme, values_from = auc) %>%
    arrange(country, outcome) %>%
    kable(booktabs = TRUE, digits = 2,
          col.names = c("Country", "Outcome", "Optimistic",
                        "Cluster-block (honest)", "Spatial-block (honest)"))
}
```

**Out-of-fold calibration.** Recalibrating binary predictions out-of-fold
(rather than in-sample) gives an honest picture of calibration; the in-sample
Platt fit is optimistic by construction (slope forced toward 1). Raw and
out-of-fold-calibrated slopes are reported per outcome in
`results/tables/corrected/calibration_oof_compare.csv`.

**Admin-2 error, separating sampling noise from model error.** A substantial
part of the apparent district-level error is survey sampling noise rather than
model error. After removing the sampling-variance component, the
sampling-adjusted RMSE is the honest measure of how far model predictions sit
from the true district prevalence (@tbl-corr-error).

```{r tbl-corr-error}
#| tbl-cap: "Admin-2 prediction error decomposed into a survey-sampling component and a model component (results/tables/corrected/admin2_error_sampling_adjusted.csv). Naive RMSE is the raw predicted-vs-survey error; the sampling component is what survey noise alone would produce; the adjusted RMSE is the honest model error."
if (!is.null(corr_err)) {
  corr_err %>%
    select(country, outcome, n_admin2, naive_rmse_pp, sampling_rmse_pp,
           adjusted_rmse_pp, pearson_r) %>%
    kable(booktabs = TRUE, digits = 2,
          col.names = c("Country", "Outcome", "n areas", "Naive RMSE (pp)",
                        "Sampling RMSE (pp)", "Adjusted RMSE (pp)", "r"))
}
```

**Decision value for targeting.** Because the operational task is to find the
highest-burden districts, we score how well the predictions concentrate true
burden in the top-ranked districts. Targeting the top 20% of districts by
predicted prevalence reaches more of the true burden than untargeted allocation
(lift > 1) for most outcomes, with the largest lift on the outcomes that have
the most spatial signal (@tbl-decision).

```{r tbl-decision}
#| tbl-cap: "Decision value of district ranking (results/tables/corrected/decision_value.csv). reach@20% is the share of true burden captured by the top 20% of districts ranked by predicted prevalence; lift is that reach divided by the 20% an untargeted allocation would reach; hit-rate@top20 is the share of the predicted top-20% districts that are truly in the top 20%."
if (!is.null(corr_decision)) {
  corr_decision %>%
    select(country, outcome, reach_at_20pct, lift_vs_no_targeting, hit_rate_top20) %>%
    kable(booktabs = TRUE, digits = 2,
          col.names = c("Country", "Outcome", "Reach@20%", "Lift", "Hit-rate@top20"))
}
```

**Prediction intervals are reasonably calibrated.** Split-conformal intervals
attached to each surveyed district achieve empirical held-out coverage close to
the 90% target (range across outcomes typically 0.85–1.00;
`results/tables/corrected/interval_summary.csv`), at the cost of wide intervals
for the noisiest outcomes.

**Per-district trust flags.** Each district is flagged for reliability by three
criteria: whether its covariates fall outside the surveyed training range
(out-of-support), small survey sample, and a wide prediction interval. The
resulting OK-to-use versus do-not-rely labels are in
`results/tables/corrected/trust_flags.csv` and drive the dashboard's certainty
layer; out-of-support is the dominant driver of do-not-rely flags in
unsurveyed districts.

## Prediction under the corrected protocol {#sec-protocol-results}

Re-running every comparison under @sec-protocol changes the conclusions more
than any modelling choice in this paper did.

**The size of the protocol effect.** Our published within-country median rank
correlation of 0.058 rested on a single random division of the districts.
Replicated ten times under the identical protocol, the median is **0.217**
(IQR 0.146-0.289); exhaustive leave-one-region-out gives **0.235**. The
published figure was a low draw, not a measurement. Separately, the comparison
the models appeared to lose was against a regional mean computed from the scored
district's own respondents; scored symmetrically, that baseline achieves
r **0.076** and the covariate model **0.156**. Neither correction involved new
data or a new model.

**Within a surveyed country, covariates beat the honest survey baseline.**
Against a jackknifed regional mean — the only sub-national alternative a
programme actually has — the covariate model attains rank correlation
**0.398** against **0.320** on the biomarker concentration, and **0.286**
against **0.193** on deficiency prevalence, in-fill. Under region
extrapolation the margin widens on prevalence (**0.268** against **-0.107**),
because a regional mean has nothing to say about a region it has not seen.

**Geography alone does most of the within-country work.** A covariate-free
spatial smoother scores **0.391** where the full covariate model scores
**0.398**. This replicates our published finding and we do not retract it: inside
a surveyed country, most of what the covariates supply is a smooth surface that
neighbouring-district information already provides. The covariates earn their
place in two specific situations — against a survey-derived baseline once that
baseline is made information-symmetric, and where there is no survey to smooth
from at all.

**Rank transports across borders; level does not.** Under leave-one-country-out,
district rankings transport with a mean Spearman of **0.26** over the 22
evaluable country-outcome combinations for the zero-tuning domain index (17
positive; the penalised domain fit gives 0.24 with 21 positive) and **0.37**
(22 positive) for an index built from the climate and soil domains alone. At
the first sub-national tier (regions, or Malawi's districts) the same 22
combinations give a mean Spearman of **0.30** (17 positive), and **0.45** (20
positive) for the climate-and-soil index; a country-block permutation null,
which keeps the correlation between outcomes within a country, has a 95th
percentile of 0.16 for the regional mean and 0.08 for the district mean, so
the means are far outside chance while counts of positive cells are not (no
transport at all yields 15–16 of 22 positive five percent of the time). An earlier draft of this section carried
a regional figure of 0.50–0.56 over 12 combinations; that figure rested on
three countries, because a spelling mismatch in the population file had
dropped Sierra Leone from every population-joined run, and it is withdrawn.
Absolute prevalence levels do not
transport, consistent with the biomarker level offsets between surveys reported
in @sec-zscore. The transportable product is therefore a ranked priority list,
not a prevalence map. We note one negative: burden-weighted targeting lift under
transport is **1.007** across 12 evaluable cells, better than random in only 5,
so the transport claim is about ordering and should not be restated as a
targeting gain.

**Capacity is not the binding constraint; it is a liability.** The zero-tuning
domain index outperforms elastic net over all 451 predictors by up to **0.38**,
and the 451-column fit turns *negative* under region extrapolation. With 14-87
units, a penalised regression over hundreds of correlated columns fits the
country it was trained in.

**In programme terms.** Visiting the worst-ranked fifth of districts reaches
**26%** more national deficiency burden than the same effort guided by the
survey's own regional averages, which are themselves no better than choosing
districts at random (in-fill lift 1.20 for the model, 0.95 for regional
averages, against an oracle ceiling of 2.43). For Malawi child iron, the
selected districts carry **43.1%** deficiency against **26.3%** nationally and
hold **48.7%** of the national burden in children.

**What tracks which nutrient.** Because no per-cell test of 451 predictors can
have usable power at these sample sizes (@sec-fdr), we report cross-country
replication of sign. @tbl-nutrient-signal lists, for each outcome, the
indicators with the strongest pooled association that also point the same way in
every country in which they are measured.

```{r}
#| label: tbl-nutrient-signal
#| tbl-cap: "Indicators most strongly associated with each outcome that also
#|   replicate in sign across all contributing countries. Strength is a pooled
#|   random-effects z across countries; 'agree' is the number of countries in
#|   which the indicator points the same way, out of those measuring it.
#|   Produced by scripts/signal_probes/p4_admin1_continuous_scan.R."
#| echo: false
f <- here::here("results", "tables", "signal_probes",
                "p4_admin1_continuous_predictors.csv")
if (file.exists(f)) {
  d <- read.csv(f, stringsAsFactors = FALSE)
  keep_g <- c("child_vitA", "child_iron", "women_iron", "women_vitA", "women_b12")
  d <- d[d$sign_agree >= d$k_countries & d$k_countries >= 3 &
           d$group %in% keep_g, ]
  d <- do.call(rbind, lapply(split(d, d$group), function(g) {
    g <- g[order(-abs(g$meta_z)), ]
    utils::head(g, 3)
  }))
  out <- data.frame(
    Outcome   = d$group,
    Indicator = d$predictor,
    Domain    = d$domain,
    Strength  = round(d$meta_z, 2),
    Direction = ifelse(d$mu > 0, "more deficiency", "less deficiency"),
    Agree     = paste0(d$sign_agree, "/", d$k_countries))
  knitr::kable(out, row.names = FALSE, align = "lllrlc")
} else {
  cat("Signal scan not available; run scripts/signal_probes/p4_admin1_continuous_scan.R.")
}
```

**Dense signal where the sparse screen found none.** The omnibus test of
@sec-fdr, powered against many small effects, rejects the no-association null in
the majority of cells, and the per-nutrient sign-replication results are given in
@tbl-nutrient-signal. The signal is real, distributed across many indicators, and
nutrient-specific; it is not concentrated in a few predictors that a sparse
screen could isolate at this sample size.

# Discussion

## Summary

We harmonised individual-level biomarker measurements from four sub-Saharan
African DHS (three West African plus Malawi) to evaluate whether models built on
remotely sensed and survey-derived proxies can reproduce admin-2 micronutrient
deficiency prevalence and national prevalence. An internal audit of an earlier
version of this manuscript found that several of its headline conclusions were
determined by the evaluation design rather than by the data, so the findings
below are those that survive re-analysis under a pre-specified protocol
(@sec-protocol). Five have programmatic significance.

**First, evaluation design dominated modelling choice.** No change to any model
in this paper moved a result as much as three changes to how results were
scored. Replicating a single random split moved a headline within-country
correlation from 0.058 to 0.217. Making a survey-derived baseline
information-symmetric with the model — it had been computed from the scored
district's own respondents — moved the same comparison from a loss to a win
(0.156 against 0.076). Recognising that a permutation screen could not clear its
own multiplicity threshold at the resolution used removed a reported null
finding entirely. Two further corrections ran against us: the reliability
ceiling we used to argue admin-2 was unresolvable was biased low by roughly a
factor of 4.7, and the regional-anchoring gain we reported as the single most
valuable design choice disappears under a jackknifed control. We report all five
because a reader has no way to tell, from a manuscript, which conclusions were
protocol artifacts and which were measurements.

**Second, proxy indicators carry replicated, nutrient-specific signal — but it
is not individual-level biology.** Under cross-country sign replication, the
appropriate evidence standard when no per-cell test of 451 predictors can have
power at 14–87 districts, the associations replicate in direction across all
four surveys while several run *opposite* to what individual nutrition predicts:
district legume consumption tracks more child vitamin A deficiency, and cattle
ownership more child iron deficiency. Together with cereal cropping, warm night
temperatures and soil heterogeneity, the replicated axis is best read as rural
subsistence agro-ecology — area-level markers of deficient districts rather than
exposures to intervene on. The exception is women's vitamin A, which tracks
district child wasting and, inversely, rainfall, and is interpretable as stated.
No malaria burden indicator replicated against any outcome.

**Third, covariate models beat the honest survey baseline within a country, but
geography does most of the work.** Against a jackknifed regional mean the
covariate model attains rank correlation 0.398 versus 0.320 on the biomarker and
0.286 versus 0.193 on prevalence. A covariate-free spatial smoother, however,
scores 0.391 where the full covariate model scores 0.398 — our published finding
on this point replicates and we do not retract it. The covariates earn their
complexity against a survey-derived baseline and where no survey exists to
smooth from, not as a general improvement on spatial smoothing.

**Fourth, rank transports across borders and level does not.** District rankings
transport under leave-one-country-out with a mean Spearman of 0.26 over 22
combinations (0.37 for the climate-and-soil index, positive in all 22), and
regional rankings with 0.30 (0.45 for climate and soil), against permutation
nulls whose 95th percentiles are 0.08 and 0.16; absolute prevalence does not,
consistent with the biomarker level offsets
between surveys documented in @sec-zscore. The transportable deliverable is a
ranked priority list. We note against ourselves that burden-weighted targeting
lift under transport is 1.007 across 12 cells, so the transport result should not
be restated as a targeting gain — that claim holds in-fill (lift 1.20 against
0.95 for regional survey averages), not across borders.

**Fifth, capacity is a liability at this sample size.** A zero-tuning index over
18 domain summaries beats elastic net over all 451 predictors by up to 0.38, and
the 451-column fit turns negative under region extrapolation. National-level
aggregation, meanwhile, remains accurate throughout (24 of 24 country-outcome
predictions within the survey's 95% CI) and is the deliverable best supported by
the evidence where a district ranking is not required.

## The aggregation problem and "fit at the unit of deliverable"

The mechanism by which individual-level predictions degrade upon
aggregation merits explicit treatment. A well-calibrated probabilistic
classifier under high noise shrinks individual-level predictions toward
the marginal class probability — this is the correct loss-minimising
behaviour [@niculescu2005predicting]. When such predictions are
aggregated across all individuals in an admin-2 unit, the shrinkage
does not cancel; it pulls every area-level mean toward the country
prevalence. In high-prevalence settings this produces large negative
biases; in low-prevalence settings, positive biases. Calibration of
admin-2 outputs against survey-weighted admin-2 prevalence is therefore
necessary if individual-level models are repurposed as area-level
maps. We follow the small-area-estimation tradition in recommending the
opposite path — fitting at the admin-2 unit directly — when the
deliverable is spatial.

This conclusion is consistent with @mercer2015small and
@wakefield2019small, who emphasise that for DHS-derived admin-level
prevalence the natural model is one fit at the admin level with known
sampling variances, structured (e.g., ICAR or BYM2) random effects,
and auxiliary covariates. Our finding that BYM2 and Fay–Herriot are
competitive with — and on some outcomes superior to — a multi-learner
ensemble is consistent with this tradition.

## Why does the SuperLearner sometimes underperform?

The most striking single result is the child-iron Sierra Leone LOCO case,
where the SuperLearner produces anti-predictive ranking (r = −0.125, MAE
20.0 pp) while a beta regression on the same covariates achieves r = 0.466
at a fraction of the absolute error (MAE 3.4 pp) — and every other linear
comparator we tested (GLM, quasi-binomial, penalised regression) was also
anti-predictive on this cell, so the ensemble's failure is not simply "any
simpler model would have done better." We interpret this as a
model-selection failure specific to the ensemble: with only 14 admin-2 areas
in the held-out set and roughly 150 covariates, the cross-validated risk
estimator within the ensemble has high variance, and the cv-selected learner
appears to have over-fit to a non-transportable feature pattern that a
correctly-specified beta regression, respecting the outcome's bounded
[0, 1] support, avoided.

This suggests that ensemble methods should be benchmarked against
parsimonious alternatives — including distribution-appropriate ones, not
only linear-Gaussian baselines — whenever the test set is small or the
target population is meaningfully outside the training distribution. Full
mode of our pipeline (the 12-learner SuperLearner across 206 admin-2 units
and five outcomes) took several hours end to end (2026-08 rebuild: 831
targets, 0 errors); a single comparator model such as beta regression
fits in seconds. Across the 16 LOCO cells evaluated in @tbl-benchmarks, that
cost was justified on none.

## Implications for programming

The national-level estimates in this study are, on present evidence, the
most defensible deliverable for programming: 24 of 24 country–outcome
national predictions fell inside the survey's own confidence interval.
Admin-2 deficiency maps derived from the same models are, at best,
interpretable as *relative ranking* tools within a country with biomarker
data — useful for prioritising districts within a national-level
intervention — but should be published alongside their reliability ceiling
(@sec-ceiling), since roughly a sixth of country–outcome cells have no
detectable admin-2 signal at all and should not be mapped as if they did.
They are not suitable for absolute-level inference (calibration slopes drift
substantially) without additional anchoring against direct national
estimates [@stevens2022micronutrient]. For an unsurveyed country, our
results suggest that LOCO-based predictions should be treated as coarse risk
triage at most, benchmarked against a national-mean null and a
covariate-free spatial smoother before any claim of added value, and used
only within the same broad agroecological region as the training set.

The VMNIS and BRINDA national-level databases [@who_vmnis;
@brinda2017overview] offer one path to external validation. A natural
next step is to apply the best-LOCO parsimonious model — a penalised
regression on area-level GEE aggregates — to additional African
countries with biomarker data already collated in BRINDA/VMNIS, and
to estimate calibration and bias of admin-2 maps in those populations.
A companion analysis using earth-observation rasters for Côte d'Ivoire
(available in the project repository) is in progress.

## Limitations

Our sample comprises four sub-Saharan African countries — three in West
Africa plus Malawi — so the rest of East, Central, and Southern Africa is
unrepresented. Because the predictors are resolved at admin-2, the
effective sample size for area-level models is the
number of polygons (14–90 per country) rather than the underlying
individuals, which limits the degrees of freedom for both within- and
cross-country fitting and makes naively-computed within-country
cross-validated correlations an optimistic bound in the absence of an
explicit reliability ceiling (most acutely for Sierra Leone, with only 14
polygons) — a limitation we address directly by reporting r_max alongside
every correlation (@sec-ceiling), but the ceiling estimate itself carries
uncertainty from the small number of respondents per district. We did not
formally quantify the contribution of survey design weights to admin-2
prevalence estimates; assuming a fixed design-effect of 1.5 (DHS-typical) is
an approximation, and we separately note that DHS sample weights were found
to sit on inconsistent numeric scales across countries in our cohort (e.g.
a median in the hundreds of thousands in one survey against below 1 in
another), which is immaterial for weighted means but would bias any
downstream use of the weights as a precision measure. The cross-country
common-vocabulary covariate set is 294 variables assembled from every admin-2
source available (@sec-predictors); predictors not present in all four
countries were excluded, and three domains — food security, food prices and
accessibility — contribute nothing to the pooled models because no single
column name is common to all four, a limitation of the strict-intersection
policy rather than of the underlying data.

Three further limitations arise from analyses added after peer review. First,
inflammation adjustment is harmonised across countries for vitamin A but not
for iron, which retains each survey's native adjustment (Thurnham, BRINDA
regression, or a log-scale regression correction); the raw components for a
uniform adjustment exist in all four surveys, and harmonising them did not
alter cross-country transport in our testing, but the asymmetry remains.
Second, several biomarkers collected in these surveys are unused, most
consequentially soluble transferrin receptor, which is available in all four
countries and is substantially less inflammation-sensitive than ferritin.
Third, all outcomes are modelled as binaries at a cutoff; for the rarest
outcomes this discards most of the available information — Malawi women's
vitamin A comprises 11 events in 836 respondents — and we show elsewhere that
the same outcome behaves very differently at admin-1, where cases accumulate.
Finally,
our raw binary classifiers' calibration diagnostics
flagged roughly half of country–outcome cells with a calibration slope well
below 1 (median 0.51 before correction); post-hoc recalibration corrects
this, but the underlying discrimination at very low prevalence remains weak
(several AUCs near 0.5, and one cell — Malawi child zinc, AUC 0.448 — below
chance) and these outcomes should be treated as exploratory.

## External validation pathway

The findings presented here are based on a largely West African LOCO panel
(three West African countries plus Malawi) and a single external held-out
country (Côte d'Ivoire). We have
prepared a multi-source national-level benchmark to evaluate future
expansions of the model to East and Southern Africa, comprising: (i) the
Kenya National Micronutrient Survey 2011 [@knms2011], which provides
nationally-representative biomarker prevalence for vitamin A, iron,
zinc, folate, B12 and iodine across pre-school children, school-age
children, pregnant women, non-pregnant women, and men; (ii) modelled
WHO/UNICEF joint anaemia prevalence estimates for all eight focus
countries, retrieved via the WHO Global Health Observatory API; and
(iii) WHO Vitamin and Mineral Nutrition Information System (VMNIS)
historical survey-level prevalence for the same eight countries, from
the latest publicly-released long-format dump. The Stevens et al. 2022
*Lancet Global Health* pooled posterior estimates [@stevens2022micronutrient] are the
natural comparator for our national-level predictions; however, the
country-level Bayesian posteriors are not publicly released, only the
22 input surveys used to fit the model, and none of our cohort countries
are among them. The full assembled benchmark is in
`results/external/{vmnis_national,who_anaemia_country,knms2011_national_summary,kenya_national_benchmark}.csv`.

Concrete next-step datasets, each requiring a Data Use Agreement
with the originating ministry, would substantially expand the LOCO panel
beyond West Africa: the Kenya National Micronutrient Survey (KNMS 2011)
and the Ethiopia National Micronutrient Survey 2016, both of which
include the full vitamin A / iron / zinc / folate / B12 / iodine biomarker
panel collected under the BRINDA inflammation-adjustment framework, making
harmonisation with our four training countries straightforward. We
attempted to add Tanzania via the 2010 TDHS but withdrew it (Methods); the
2023 Tanzania DHS, which we understand uses serum rather than
dried-blood-spot vitamin-A assessment, is a better-supported candidate once
its data become accessible. As a lighter-weight alternative for iron, the
most recent DHS rounds with haemoglobin biomarker (Kenya DHS 2022, Burkina
Faso DHS 2021) can be added with no further paperwork beyond the user's
existing DHS account, providing iron-deficiency anaemia coverage for two
additional East/Sahelian countries. We
provide country-agnostic ingestion stubs (`R/ingest_new_country.R`,
function `stub_new_country_config()`) for each of these candidates;
once a harmonised survey file lands in `data/IPD/<country>/`, the entire
LOCO, calibration, SHAP, and dashboard pipeline regenerates automatically.

## Future methodological directions

The results above locate the limits of the current approach precisely — a
low reliability ceiling on the admin-2 target itself, transport that fails
on both level and pattern depending on the method, and an ensemble that does
not beat parsimonious SAE baselines or two uninformed nulls — and these
limits motivate several planned extensions. We describe them here as
directions rather than results.

First, **decision-focused estimation.** For programmatic use the relevant target
is rarely the exact prevalence of a district but whether it lies above a
public-health action threshold, and whether the highest-risk districts are
correctly prioritised. Because the SuperLearner's base learners optimise
likelihood or squared-error loss while the ensembling (or discrete-selection)
step may target any criterion [@vanderlaan2007super], the ensemble can be tuned
directly for an aggregate, asymmetric objective — for example maximising recall of
the truly high-risk areas (top-quantile or above-threshold) subject to a bounded
false-positive rate. This reframes the deliverable toward the ranking signal that,
as shown above, transports more reliably than absolute level, and is
statistically less demanding than level estimation.

Second, **anchoring — but to regions, not to a nation.** An earlier version of
this paper proposed level recalibration against a known or historical national
value (e.g. from VMNIS or the BRINDA pooled analyses
[@who_vmnis; @brinda2017overview]) as a promising direction. We have since tested
it and report the result rather than the proposal. Anchoring to *regional*
design-based totals is the single most valuable step we found (@sec-anchoring);
anchoring to a *national* total is not, and a national level that must itself be
predicted is worse than no anchor at all (@sec-national-composition). The
productive version of this direction is therefore the acquisition of
regional-level anchors — a survey design question rather than a modelling one —
and the useful methodological work lies in how few regions, and how thinly
sampled, an anchor set can be before the gain disappears.

Third, **borrowing strength across co-occurring deficiencies.** The micronutrient
outcomes are biologically correlated but were modelled independently here.
Modelling them jointly through a forward factorisation of their conditional
density — predicting the most-informative outcome first (iron, measured in all
four countries) and supplying predicted upstream deficiencies, with
missing-indicator encoding and imputation, as covariates for subsequent outcomes
— is a standard route to exploiting that covariance and may stabilise the
data-poorest outcomes (zinc, B12). We expect within-country gains to exceed
transport gains, since transported upstream predictions themselves carry error.

Fourth, **inference and value-of-information.** Valid uncertainty on the aggregate
deliverable (national or admin-level prevalence) — via cross-validated
loss-based inference, conformal prediction intervals, or the area-resampling
bootstrap already reported — enables a complementary product that is robust to the
data being imperfect: a value-of-information map that ranks candidate survey
locations by the product of predicted risk and predicted uncertainty, identifying
where a new biomarker survey would be most informative. For a field constrained by
the cost and rarity of biomarker measurement, guiding *where to measure next* may
be as valuable as the prevalence estimates themselves.

Finally, the most direct route to a higher reliability ceiling is to raise
the number of biomarker measurements each admin-2 estimate is built from —
pooling survey rounds within a country, modelling at admin-1 where the
survey design actually supports it, or borrowing strength across the
biologically-correlated outcomes above — rather than adding learners to the
ensemble, since @sec-ceiling shows the individual-level model already
captures most of what the current data can support. The most
consequential validation step remains to re-run the entire methodology on
anaemia (haemoglobin), which is measured in nearly every sub-Saharan African
DHS at large sample size: a trusted, abundant outcome would separate
methodological limitations from the disputed comparability of the
micronutrient biomarkers themselves, and would expand the cross-country
panel far beyond the four countries available here.

## Conclusions

Proxy indicators carry replicated, nutrient-specific signal for subnational
micronutrient deficiency status, and several of the negative findings we
previously reported were artifacts of evaluation design rather than properties
of the data. A published within-country correlation of 0.058 was a single
high-variance draw and is 0.217 when replicated; a benchmark the models appeared
to lose to had been computed from the scored district's own respondents; a screen
reported as finding nothing could not have found anything at any effect size;
and the reliability ceiling used to argue that admin-2 was unresolvable was
biased low by about a factor of 4.7. Corrections ran in both directions: the
regional-anchoring gain we reported as the single most valuable design choice
does not survive a jackknifed control, and the finding that a covariate-free
spatial smoother matches the full covariate set within a surveyed country
replicates and stands.

Under the corrected protocol, covariate models beat an information-matched
survey baseline in both within-country estimands, and district rankings
transport to unsurveyed countries with a mean rank correlation of 0.26 (0.37
from climate and soil alone; regional rankings 0.30 and 0.45) while
absolute levels do not. The defensible deliverable is therefore a ranked district priority list
rather than a prevalence map, and it is worth about a quarter more deficiency
burden reached, per unit of programme effort, than the survey's own regional
averages. Within a surveyed country, however, geography alone supplies most of
the attainable accuracy, and the covariates are worth their complexity mainly
where no survey exists to smooth from. Model capacity is not the binding
constraint: a zero-tuning index over 18 domain summaries beats a penalised
regression over 451 predictors, which turns negative when asked to extrapolate
to an unseen region.

For the field, the transferable result is procedural. We recommend that studies
of this kind name the estimand they are evaluating, since in-fill, region
extrapolation and country transport have materially different answers and are
routinely reported interchangeably; state what information the comparison
baseline was permitted to see, and jackknife any baseline derived from the same
survey; replicate every random division of the data and report the spread rather
than a single value; and demonstrate that each significance screen could have
detected an effect had one been present. Applied to our own work, these four
conditions reversed several conclusions, and we would not have identified any of
them from the modelling results alone.

# Acknowledgements

Funded by the Bill & Melinda Gates Foundation (grant number INV-XXXXXX).
The funder had no role in study design, analysis, decision to publish,
or manuscript preparation.

# Author contributions

ANM conceived the study, conducted the analysis, and drafted the
manuscript. Co-authors to be added on revision.

# Data availability

DHS biomarker data are available from the DHS Program
(<https://dhsprogram.com/>) following registration. Pipeline code,
including the area-level SuperLearner, frequentist baselines, and
INLA-based BYM2 implementation, are available at
\textless<repository-URL placeholder>\textgreater.

# References

::: {#refs}
:::
