---
title: "Estimating Subnational Micronutrient Deficiency Prevalence"
subtitle: "A SuperLearner Ensemble Approach — Multi-Country Analysis"
author: "Andrew N. Mertens, UC Davis Institute for Global Nutrition"
date: "2026"
format:
  pptx:
    reference-doc: "C:/Users/andre/Dropbox/MN prediction/BMGF_mn_project_slides.pptx"
  revealjs:
    theme: default
    slide-number: true
    width: 1280
    height: 720
    transition: fade
    incremental: false
    fig-align: center
    toc: false
    footer: "Proxy Micronutrient Deficiency Prediction — Multi-Country"
    highlight-style: github
    self-contained: true
    smaller: true
    scrollable: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(here)
library(scales)
library(patchwork)
library(viridis)

theme_set(theme_minimal(base_size = 13))

TARGETS_STORE <- here("_targets_full")

load_target <- function(name) {
  tryCatch(targets::tar_read_raw(name, store = TARGETS_STORE),
           error = function(e) NULL)
}

country_labels <- c(gambia = "Gambia", ghana = "Ghana",
                     sierraleone = "Sierra Leone", malawi = "Malawi")
outcome_labels <- c(
  child_vitA = "Vitamin A (children)", women_vitA = "Vitamin A (women)",
  child_iron = "Iron deficiency (children)", women_iron = "Iron deficiency (women)",
  women_folate = "Folate deficiency (women)", women_b12 = "Vitamin B12 deficiency (women)",
  child_zinc = "Zinc deficiency (children)", women_zinc = "Zinc deficiency (women)")

countries <- c("gambia", "ghana", "sierraleone", "malawi")
outcomes  <- c("child_vitA", "women_vitA", "child_iron", "women_iron",
               "women_folate", "women_b12", "child_zinc", "women_zinc")

# Load CV performance
cv_list <- list()
for (ctry in countries) for (oc in outcomes) {
  obj <- load_target(paste0("cv_perf_", ctry, "_", oc))
  if (!is.null(obj) && is.data.frame(obj) && nrow(obj) > 0) {
    if (!"country" %in% names(obj)) obj$country <- country_labels[ctry]
    cv_list[[paste0(ctry, "_", oc)]] <- obj
  }
}
cv_perf <- if (length(cv_list) > 0) bind_rows(cv_list) else NULL

# Load diagnostics
diagnostics_all <- load_target("diagnostics_all")
```


# Motivation {background-color="#1a3a5c"}

---

## The Biomarker Data Gap

:::: {.columns}
::: {.column width="60%"}
- Micronutrient deficiency (iron, vitamin A, zinc, folate, B12) drives child and maternal
  morbidity across sub-Saharan Africa
- Nationally representative **biochemical surveys are rare** --- conducted at decadal intervals
- Between survey cycles, **sub-national burden is essentially unknown**
- Program allocation relies on outdated or modelled estimates

**Can we estimate micronutrient deficiency from routinely available proxy data?**
:::
::: {.column width="40%"}
**Current focus micronutrients:**

- Iron deficiency (ferritin)
- Vitamin A deficiency (RBP)
- Zinc deficiency
- Folate deficiency
- Vitamin B12 deficiency

**Countries:** Gambia, Ghana, Sierra Leone, Malawi + Cote d'Ivoire (OOS)
:::
::::

---

## Objective

::: {.callout-note appearance="minimal"}
**Primary objective:** Estimate Admin-1 and national prevalence of micronutrient deficiencies
across four countries using a SuperLearner ensemble with multi-source proxy predictors ---
external data available without a concurrent biomarker survey.
:::

**Secondary objectives:**

1. Quantify prediction uncertainty via split-conformal prediction intervals
2. Assess predictor domain contributions via permutation-based domain ablation
3. Evaluate cross-country transportability via leave-one-country-out (LOCO) validation
4. Generate out-of-sample predictions for Cote d'Ivoire with covariate-shift adjusted CIs
5. Identify key individual predictors via variable importance analysis

---

# Data Architecture {background-color="#1a3a5c"}

---

## Predictor Strategy: 13+ Proxy Data Sources

**Proxy-only models** --- survey-derived (`gw_`) variables excluded to prevent data leakage.

| Domain | Source | Approx. N cols |
|---|---|---|
| DHS cluster aggregates | DHS API | ~400 |
| MICS indicators | MICS compilations | ~150 |
| IHME disease estimates | GBD geospatial | ~60 |
| MAP malaria surfaces | Malaria Atlas Project | ~26 |
| WFP food prices | WFP VAM | ~46 |
| GEE satellite imagery | Google Earth Engine Admin-2 zonal means | ~308 |
| CHIRPS rainfall | UCSB CHIRPS | ~12 |
| WorldPop density | WorldPop | ~5 |
| VIIRS nighttime lights | NOAA/CCNL | ~3 |
| SoilGrids | ISRIC | ~15 |
| GDL development indices | Global Data Lab | ~5 |
| HarvestChoice/SPAM crops | IFPRI | ~20 |
| **Total proxy** | | **~900--1,100** |

---

## Outcome and Population Structure

**Four core outcomes** (fitted independently per country):

| Population | Micronutrient | Binary threshold | Countries |
|---|---|---|---|
| Children | Vitamin A | RBP < 0.70 umol/L | All 4 |
| Women | Vitamin A | RBP < 0.70 umol/L | All 4 |
| Children | Iron | Ferritin < 12 ug/L | All 4 |
| Women | Iron | Ferritin < 15 ug/L | All 4 |
| Women | Folate | Serum folate < 10 nmol/L | GH, SL, MW |
| Women | Vitamin B12 | Serum B12 < 148 pmol/L | GH, SL, MW |
| Children/Women | Zinc | Serum zinc < threshold | MW only |

**24 country x outcome models** fitted in the full pipeline.

---

# Modeling Strategy {background-color="#1a3a5c"}

---

## SuperLearner Ensemble (Full Stack)

**16 base learners**, NNLS metalearner, binomial log-likelihood loss:

| Class | Learners | Notes |
|---|---|---|
| Regularized linear | Lasso, Ridge, Elastic net | L1/L2/mixed penalty |
| Random forest | Ranger x3 | Varying depth and mtry |
| Gradient boosting | XGBoost x2 | Conservative + deep |
| Bayesian trees | BART x3 | 50/100/200 trees |
| Other | Gaussian process | Smooth spatial patterns |

**Additional for rare outcomes (prevalence < 15%):**

- Class-weighted XGBoost (minority upweighting)
- Class-weighted Ranger

**Cross-validation:** 5-fold, cluster-blocked (no cluster in both train and validation)

---

## Uncertainty: Split-Conformal Prediction Intervals

| Step | Action |
|---|---|
| 1 | Split data into training + calibration sets (cluster-blocked) |
| 2 | Fit full SL on training set |
| 3 | Compute nonconformity scores on calibration set |
| 4 | Construct adaptive-width intervals using calibration quantiles |
| 5 | Aggregate to Admin-1/national with intervals |

**For out-of-sample (CIV):** Weighted conformal with covariate-shift density ratios

---

# Results: Within-Country Performance {background-color="#1a3a5c"}

---

## Binary Model AUC Across Countries

```{r auc-heatmap, fig.height=4.5, fig.width=9}
if (!is.null(cv_perf) && "auc" %in% names(cv_perf)) {
  cv_binary <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome])

  if (nrow(cv_binary) > 0) {
    ggplot(cv_binary, aes(x = outcome_label, y = country, fill = auc)) +
      geom_tile(colour = "white", linewidth = 1) +
      geom_text(aes(label = sprintf("%.3f", auc)), size = 4.5, fontface = "bold") +
      scale_fill_viridis_c(name = "AUC", limits = c(0.45, 0.95), option = "plasma",
                            oob = squish) +
      labs(x = NULL, y = NULL,
           title = "Cross-Validated ROC-AUC by Country and Outcome") +
      theme_minimal(base_size = 14) +
      theme(axis.text.x = element_text(angle = 25, hjust = 1),
            panel.grid = element_blank())
  }
}
```

---

## AUC Forest Plot with DeLong CIs

```{r auc-forest, fig.height=5.5, fig.width=10}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$binary_metrics)) {
  bm <- diagnostics_all$binary_metrics |>
    filter(!is.na(roc_auc)) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      label = paste0(country, " (n=", n, ", prev=", round(prevalence*100,1), "%)")
    )

  if (nrow(bm) > 0) {
    # Compute DeLong CIs
    auc_ci_rows <- list()
    for (ctry in countries) {
      for (oc in outcomes) {
        ci <- tryCatch({
          sl <- load_target(paste0("sl_fit_", ctry, "_", oc))
          if (!is.null(sl) && !is.null(sl$bin_fit)) {
            bf <- sl$bin_fit
            roc_obj <- pROC::roc(bf$res$Y, bf$res$yhat_full, quiet = TRUE)
            ci_obj <- pROC::ci.auc(roc_obj, method = "delong")
            data.frame(country = country_labels[ctry], outcome = oc,
                       auc_lo = ci_obj[1], auc_hi = ci_obj[3])
          }
        }, error = function(e) NULL)
        if (!is.null(ci)) auc_ci_rows[[paste0(ctry, "_", oc)]] <- ci
      }
    }
    if (length(auc_ci_rows) > 0) {
      auc_ci <- bind_rows(auc_ci_rows)
      bm <- bm |> left_join(auc_ci, by = c("country", "outcome"))
    }

    p <- ggplot(bm, aes(x = roc_auc, y = reorder(label, roc_auc)))
    if ("auc_lo" %in% names(bm)) {
      p <- p + geom_errorbarh(aes(xmin = auc_lo, xmax = auc_hi),
                               height = 0.3, colour = "grey40", linewidth = 0.5)
    }
    p + geom_vline(xintercept = 0.5, linetype = "dashed", colour = "grey60") +
      geom_point(aes(colour = country), size = 3) +
      scale_x_continuous(limits = c(0.3, 1), breaks = seq(0.3, 1, 0.1)) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      facet_wrap(~ outcome_label, scales = "free_y", ncol = 2) +
      labs(x = "ROC-AUC (95% DeLong CI)", y = NULL) +
      theme_minimal(base_size = 11) +
      theme(strip.text = element_text(face = "bold"))
  }
}
```

---

## Key Performance Findings

::: {.callout-important appearance="minimal"}
**Rebuilt 2026-08** (full mode, 12-learner stack, 4 countries, 831 targets,
0 errors). A fifth country, Tanzania, was added and then withdrawn (see
Full Rebuild slide); the numbers below reflect the four-country pipeline.
:::

:::: {.columns}
::: {.column width="50%"}
**Across all 24 binary country x outcome cells**

- Median AUC **0.594** (range 0.448--0.767)
- Only **6 of 24** reach AUC $\ge$ 0.70; **2 of 24** reach 0.75
- One cell falls **below chance** (Malawi child zinc, 0.448)
- Median continuous $r$ = **0.163** (23 cells)

**Highest AUCs (all rare outcomes)**

| Cell | AUC | events |
|---|---|---|
| Malawi women vitA | 0.767 | 11 |
| Sierra Leone women B12 | 0.752 | 4 |
| Malawi women B12 | 0.747 | 23 |
| Ghana women vitA | 0.745 | 15 |
| Gambia women vitA | 0.737 | 38 |
:::
::: {.column width="50%"}
**Performance is highest where events are fewest**

$r$(AUC, log events) = **-0.521**
(95% CI -0.764 to -0.149, $p$ = 0.009, n = 24)

| Event tertile | median events | median AUC |
|---|---|---|
| Few | 19 | 0.741 |
| Middle | 93 | 0.592 |
| Many | 431 | **0.562** |

This is the signature of small-sample optimism, not signal. Every cell with
enough events to estimate AUC precisely sits near chance.

**Read the many-events tertile as the honest estimate of discrimination.**
:::
::::

---


## Evaluating Models for Rare Outcomes

| Dimension | Metric | What it tells you |
|---|---|---|
| **Discrimination** | ROC-AUC, **PR-AUC** | Can the model rank individuals by risk? |
| **Calibration** | Cal. slope, Cal. intercept | Are predicted probabilities accurate? |
| **Prevalence recovery** | Predicted vs. observed | Does the model get the right burden? |

::: {.callout-important appearance="minimal"}
For **subnational mapping**, prevalence recovery matters more than individual-level discrimination. A model with AUC = 0.65 can still accurately estimate Admin-1 prevalence if well calibrated.
:::

---

# Domain Importance {background-color="#1a3a5c"}

---

## Cross-Outcome Domain Importance

```{r domain-heatmap, fig.height=5.5, fig.width=11}
ablation <- read.csv(here("results", "tables", "domain_ablation_all.csv"))
if (nrow(ablation) > 0) {
  base <- ablation |> filter(domain_removed == "none") |>
    select(outcome, country, base_auc = auc)
  abl <- ablation |>
    left_join(base, by = c("outcome", "country")) |>
    mutate(delta_auc = base_auc - auc) |>
    filter(domain_removed != "none", !is.na(delta_auc), base_auc > 0.51) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      ctry_outcome = paste0(country, " - ", outcome_label)
    )

  if (nrow(abl) > 0) {
    dom_order <- abl |> group_by(domain_removed) |>
      summarise(mean_d = mean(delta_auc, na.rm = TRUE), .groups = "drop") |>
      arrange(desc(mean_d))
    abl$domain_removed <- factor(abl$domain_removed, levels = rev(dom_order$domain_removed))

    lim <- max(abs(abl$delta_auc), na.rm = TRUE) * 1.05

    ggplot(abl, aes(x = domain_removed, y = ctry_outcome, fill = delta_auc)) +
      geom_tile(colour = "white", linewidth = 0.5) +
      geom_text(aes(label = sprintf("%.3f", delta_auc)), size = 2) +
      scale_fill_gradient2(name = expression(Delta~AUC),
        low = "#c0392b", mid = "white", high = "#27ae60",
        midpoint = 0, limits = c(-lim, lim), oob = squish) +
      labs(x = NULL, y = NULL, title = "Domain Importance: AUC Drop When Domain Permuted") +
      theme_minimal(base_size = 10) +
      theme(axis.text.x = element_text(angle = 50, hjust = 1, size = 8),
            panel.grid = element_blank())
  }
}
```

---

## Domain Importance: Key Findings

- **No single domain dominates** --- predictive signal is distributed across themes
- **High redundancy** among domains --- removing one allows others to compensate
- **Top domains by mean importance:** Climate, Infectious Disease, Environment, Water & Sanitation
- **Country-specific patterns:** Malaria indicators more important in high-endemicity countries (Ghana, Sierra Leone)

::: {.callout-note appearance="minimal"}
**Policy implication:** Maintaining diverse data inputs across multiple thematic areas is more important than investing heavily in any single data source.
:::

---

## Individual Variable Importance (Top-2 Models)

```{r varimp, fig.height=5, fig.width=10}
vi_files <- list.files(here("results", "tables"),
                        pattern = "^variable_importance_.*\\.csv$", full.names = TRUE)
if (length(vi_files) > 0) {
  vi_all <- bind_rows(lapply(vi_files, read.csv))
  vi_top <- vi_all |>
    group_by(country, outcome) |> slice_max(importance, n = 15) |> ungroup() |>
    mutate(
      outcome_label = outcome_labels[outcome],
      model_label = paste0(country, " - ", outcome_label),
      var_short = gsub("^(ihme_|MAP_|gee_|dhs[0-9]*_|chirps_|wp_|ntl_|sg_|gdl_|wfp_|hc_|mics_)", "", variable),
      var_short = substr(var_short, 1, 40)
    )

  plots <- list()
  for (ml in unique(vi_top$model_label)) {
    vi_sub <- vi_top |> filter(model_label == ml) |>
      arrange(importance_norm) |>
      mutate(var_short = factor(var_short, levels = var_short))
    plots[[ml]] <- ggplot(vi_sub, aes(x = importance_norm, y = var_short)) +
      geom_col(fill = "#2c7bb6", width = 0.7) +
      labs(x = "Relative importance", y = NULL, title = ml) +
      scale_x_continuous(labels = percent_format(accuracy = 1)) +
      theme_minimal(base_size = 9) +
      theme(axis.text.y = element_text(size = 7))
  }
  if (length(plots) == 2) {
    plots[[1]] + plots[[2]]
  } else if (length(plots) > 0) {
    plots[[1]]
  }
}
```

*Ranger impurity-based importance from 1000-tree RF on same preprocessed data*

---

## Variable Importance: Key Patterns

**Ghana child iron** (AUC = 0.702, rebuilt 2026-08):

- Rainfall (TRMM April, CHIRPS April) --- seasonal precipitation patterns
- Land cover (ESA WorldCereal) --- agricultural land use
- Soil nitrogen, soil iron --- geochemical correlates
- MICS indicators, GDL subnational HDI

**Ghana women B12** (AUC = 0.698, rebuilt 2026-08):

- Rainfall (TRMM July, CHIRPS July) --- wet season patterns
- MAP insecticide-treated net access
- IHME HIV prevalence --- disease burden correlate
- WFP rice prices --- food access indicator

**Common theme:** Climate/rainfall and environmental variables dominate, suggesting strong agro-ecological gradients drive micronutrient status.

---

# Transportability {background-color="#1a3a5c"}

---

## Leave-One-Country-Out (LOCO) Cross-Validation

**Design:** Train on 3 countries, predict on held-out 4th. Repeat for all 4.

```{r loco-table}
ts <- load_target("transportability_summary")
if (!is.null(ts) && nrow(ts) > 0) {
  kable(ts |> mutate(across(where(is.numeric), ~round(., 3))),
        caption = "LOCO transportability results") |>
    kableExtra::kable_styling(font_size = 12, full_width = FALSE)
} else {
  cat("LOCO results: women_iron AUC range 0.51-0.63 across held-out countries.\n")
  cat("Full LOCO for all 4 outcomes currently running.\n")
}
```

**Key findings:**

- LOCO AUC substantially lower than within-country CV AUC (gap ~0.15--0.25)
- Cross-country heterogeneity in biomarker-predictor relationships limits transportability
- **Gambia** best predicted when held out (most similar to pooled training data)
- **Ghana** hardest to predict from other countries (most country-specific signal)

---

## Admin-2 estimation: fit at the unit of deliverable

**Within-country area-level SuperLearner (fit directly on admin-2 aggregates)
dramatically outperforms aggregating an individual-level SL to admin-2:**

| Approach | Pearson r (admin-2, median across 30 country–outcomes) | MAE (pp) | Bias |
|---|---|---|---|
| Individual SL → aggregate to admin-2 | 0.55 (range −0.39 to 0.87) | 4--21 | Often large negative |
| **Area SL (MSE)** — fit at admin-2 | **0.97** (r > 0.9 in 27 of 30) | **1--11** | Near zero |
| Area SL (NLL)                       | 0.94                          | 1--14 | Near zero |

Aggregating individual-level predictions to admin-2 introduces characteristic
*shrinkage compression*: high-prevalence districts are systematically
under-predicted (e.g., Gambia child-iron MAE 20.8 pp, bias −20.4 pp).

> This is the classic small-area-estimation lesson [Mercer 2015, Wakefield
> 2019]: **fit at the unit of deliverable**. If the output is an admin-2 map,
> fit at admin-2; if the output is household-level risk, fit at the individual
> level.

---

## Cross-country (LOCO) admin-2 transportability is limited

LOCO Pearson r between predicted and observed admin-2 prevalence:

| Outcome | Held-out best | Held-out worst | Median |
|---|---|---|---|
| child_iron  | Ghana (r=0.53) | Malawi (r=−0.02) | 0.31 |
| women_iron  | Gambia (r=0.40) | Sierra Leone (r=−0.08) | 0.15 |
| child_vitA  | Ghana (r=0.38) | Malawi (r=0.18) | 0.27 |
| women_vitA  | Ghana (r=0.17) | Sierra Leone (r=−0.15) | 0.03 |

- Iron-deficiency outcomes transport best.
- Vitamin-A in women essentially does not transport.
- **Malawi held out → anti-predictive in 3 of 4 outcomes** — a regional
  generalisation gap (only training country outside West Africa).
- MAE 5--46 pp; mean bias often > 10 pp.

---

## Benchmark comparison: does the SuperLearner beat simpler methods?

We compared the area-level SL against four canonical baselines:

::: {.columns}
::: {.column width="55%"}
| Method | Why include |
|---|---|
| Country-mean baseline | Is the model doing anything? |
| OLS (top-8 covariates) | Is anything more than linear needed? |
| Penalised GLM (elastic-net) | Best frequentist comparator |
| Fay--Herriot (`sae`) | Industry standard SAE |
| BYM2 (INLA, PC priors) | Bayesian areal disease-mapping |
:::
::: {.column width="45%"}
**Preliminary head-to-head, 16 LOCO splits:**

> **SuperLearner wins:** 7
> **Benchmark wins:** 9

The 17.7-h SL pipeline does **not consistently beat** an 8-variable OLS or
penalised regression. Where SL wins, it does so via genuine non-linearity
(ranger consistently the winning learner).
:::
:::

**Stark example:** child-iron Sierra Leone --- SL r = **−0.19** (anti-predictive),
plain OLS r = **0.37**. Ensemble model selection failed; a simpler regularised
model recovered the signal.

---

# Out-of-Sample: Cote d'Ivoire {background-color="#1a3a5c"}

---

## CIV Predictions

**Approach:** Pooled elastic net trained on all 4 countries applied to CIV Admin-2 polygons using GEE satellite predictors only.

**Weighted conformal CIs** account for covariate shift between training countries and CIV.

```{r civ-table}
oos_list <- list()
for (oc in outcomes) {
  obj <- load_target(paste0("oos_civ_", oc))
  if (!is.null(obj) && !is.null(obj$admin1_preds)) {
    a1 <- obj$admin1_preds
    oos_list[[oc]] <- data.frame(
      Outcome = outcome_labels[oc],
      `Admin-1 units` = nrow(a1),
      `Mean prev (%)` = round(mean(a1$pred_prev, na.rm = TRUE) * 100, 1),
      `Range (%)` = sprintf("%.1f--%.1f", min(a1$pred_prev)*100, max(a1$pred_prev)*100),
      check.names = FALSE)
  }
}
if (length(oos_list) > 0) {
  kable(bind_rows(oos_list)) |>
    kableExtra::kable_styling(font_size = 12, full_width = FALSE)
} else {
  cat("CIV OOS predictions available in the results report.\n")
}
```

---

# Discussion {background-color="#1a3a5c"}

---

## Strengths

1. **Multi-country framework** --- same pipeline, predictors, and evaluation across 4 countries
2. **Proxy-only design** --- answers the policy question: can we estimate burden without new surveys?
3. **Diverse predictor ensemble** --- 13+ data sources spanning climate, population, agriculture, health, socioeconomic domains
4. **Data-adaptive ensemble** --- 12-learner SL with NNLS weights; no learner hard-coded
5. **Conformal prediction intervals** --- distribution-free coverage guarantees
6. **Transportability testing** --- LOCO provides honest cross-country assessment
7. **Reproducible `{targets}` pipeline** --- cached, dependency-tracked; ~540 targets

---

## Limitations and Cautions

1. **Rare outcome instability** --- women_vitA has < 20 events in several countries; binary AUC unreliable; Brier skill is negative for ~25% of binary models.
2. **Cross-country heterogeneity** --- LOCO shows 0.15--0.25 AUC drop vs. within-country; admin-2 Pearson r often 0.0--0.3 cross-country; vitamin-A in women essentially does not transport.
3. **Aggregation bias** --- individual-level SL aggregated to admin-2 introduces shrinkage compression and large negative biases in high-prevalence districts. Fit at the unit of deliverable.
4. **SuperLearner does not consistently beat simpler methods** under LOCO --- preliminary head-to-head shows 9 of 16 splits won by benchmark methods (OLS, penalised GLM, Fay--Herriot, BYM2). The ensemble's advantage concentrates on outcomes with genuine non-linear spatial signal.
5. **Calibration drift** --- several models have negative calibration slopes (predictions move opposite to truth), especially in Sierra Leone and Malawi at low prevalence. Continuous biomarker R² is near zero across the board.
6. **Static predictor vintage** --- DHS/MICS data ages; predictive performance may degrade.
7. **Survey support dependence** --- Admin-1/2 units with no nearby clusters are extrapolations.
8. **Limited geographic coverage** --- 4 West African countries; Malawi LOCO failure suggests model does not transport to East / Southern Africa without local validation.

---

## Milestones

| Deliverable | Status |
|---|---|
| Scientific Advisory Board | Established |
| Multi-country pipeline (4 countries) | Complete |
| Full learner stack (12 learners incl. 3 BART sizes) | Complete |
| Conformal prediction intervals | Complete |
| 13+ external predictor sources | Complete |
| LOCO transportability (4 outcomes) | Complete |
| CIV out-of-sample predictions | Complete |
| Domain ablation (permutation-based) | Complete |
| Variable importance (top-2 models) | Complete |
| Folate, B12, zinc outcomes | Complete (where data available) |
| Sierra Leone mean-learner fix | Complete |
| Gridded pixel-level maps | Planned |
| Interactive dashboard (Shiny) | Planned |

---

## Next Steps

:::: {.columns}
::: {.column width="50%"}
**Immediate:**

- Re-run the pipeline: the review fixes change model inputs
- Residualise the VMNIS outcome out-of-fold and re-score
- Prepare Phase 2 report to Gates Foundation
:::
::: {.column width="50%"}
**Longer-term:**

- Gridded (pixel-level) prevalence maps
- Additional countries beyond current 4 + CIV
- Integrate into DInA prioritization framework
- Interactive Shiny dashboard with leaflet maps
- Harmonise survey weight scales across countries
- Expand the VMNIS national panel beyond vitamin A
:::
::::

---

# Methods Review 2026-08 {background-color="#7a2020"}

## What the review changed

::: {.callout-important appearance="minimal"}
A systematic re-audit of the area-level and LOCO models (`sandbox_parsimony/FINDINGS.md`,
24 sections, 34 result tables) found **four data bugs, three inflated performance numbers,
and one clear modelling win**. Several previously reported results do not survive; they are
corrected here rather than quietly dropped.
:::

**Data integrity**

- GADM **water bodies used as prediction units** — `Lake Malawi` (n=21–33) and `Lake Victoria`
  (n=60–110) appear in the *survey* tables. Respondents geocoded onto lakes.
- **Duplicate Admin-2 names** cause silent cartesian joins (inflated one table 256 → 29,183 rows).
- **Fay–Herriot ignored its own sampling variance** — single-PSU districts return SE ≈ 1e-17,
  which passed a `D <= 0` guard. The EBLUP kept ~71% of the raw direct estimate, not ~26%.
- **Survey weights sit on inconsistent scales** — Malawi median 790,000, Ghana 0.98. Harmless for
  weighted means, fatal anywhere weights act as precision.

---

## The outcome is noise-dominated, and nothing said so

A ceiling is now reported next to every correlation:

$$\lambda = 1 - \frac{E[V_\text{sampling}]}{\mathrm{Var}(p_\text{obs})}, \qquad r_{\max} = \sqrt{\lambda}$$

A median Admin-2 district contributes **6–36 biomarker measurements**.

| cell | published CV r | r_max | % of ceiling |
|---|---|---|---|
| Ghana child iron | 0.609 | 0.672 | **77–91%** — saturated |
| Gambia women iron | 0.538 | 0.706 | 68–76% |
| Malawi child iron | 0.038 | 0.199 | 7–19% — real headroom |
| Ghana / Malawi women vit A | −0.039 / 0.176 | **0.00** | nothing to predict |

::: {.callout-note appearance="minimal"}
Two very different situations were being presented identically. Cells whose ceiling is
indistinguishable from zero are now blanked in the results tables, not reported as model failures.
:::

---

## Three published numbers do not survive

| claim | published | honest | why |
|---|---|---|---|
| `spatial_plus_soil` leads LOCO | rho = 0.305 | **rho = 0.117** | its 8 soil features were ranked by LOCO r on the *same* outcomes and held-out countries |
| adaptive variable selection helps | — | no reliable gain | selecting once on all data inflates rho by **+0.30** (Admin-1), **+0.16** (Admin-2) |
| an Admin-1 layer would roughly double accuracy | proposed | **rejected** | ceiling doubles but achieved rho *halves* (0.064 vs 0.137) — 16–32 regions is too few to fit on |

::: {.callout-warning appearance="minimal"}
Any variable-selection step must sit **inside** the resampling loop. Selection done first is the
single largest source of optimism found in this review — larger than every genuine modelling
difference combined.
:::

---

## The clear win: parsimony plus geography

**16 a-priori variables + longitude/latitude in a random forest** beat the 237–487-predictor
production recipe in **14 of 16** country × outcome cells (paired delta-Spearman **+0.063**, t = 4.9).

:::: {.columns}
::: {.column width="55%"}
| model | rho | r / r_max |
|---|---|---|
| **curated 16 + lon/lat, RF** | **0.374** | 0.58 |
| 20 cluster reps + lon/lat, RF | 0.362 | 0.60 |
| *spatial only, no covariates* | 0.347 | 0.53 |
| production (top-30 screen, enet) | 0.310 | 0.49 |
| ridge on all 237–487 | 0.230 | 0.32 |
:::
::: {.column width="45%"}
**18 predictors, not 237.**

A covariate-free lon/lat smoother already out-scores the production recipe — geography must be
**in** the model, not treated as an alternative to it.

Ridge on the full pool is the **worst** covariate model: pool redundancy is actively harmful when
nothing selects against it.
:::
::::

---

## Level and pattern are different problems

LOCO transports the **spatial pattern** modestly and the **level** not at all.

- Best pattern transfer: train on the within-country z-score, rho 0.154 → **0.261**
- Level: the best LOCO model misses national prevalence by **5.4 pp** on average and by a factor
  of **2–3** on iron (Malawi child iron 10.6% actual → 28.4% predicted)

**Fix — national benchmarking, now default-on.** A logit-scale shift solved so the
population-weighted aggregate equals the design-based national estimate:

| | before | after |
|---|---|---|
| national error | 4.33 pp | **0.00 pp** |
| district Spearman change | — | **0.0000** (strictly monotone) |

::: {.callout-note appearance="minimal"}
The survey's *own* district estimates aggregate **4.50 pp** away from the national figure the same
survey produces. Most of the model drift is inherited from the direct estimates, not created by the
model — these surveys are powered for national, and Admin-2 is an unplanned domain.
:::

---

## Absolute accuracy: honest framing

Against the comparator a programme already has — paint the national mean on every district:

| | model | national-mean map |
|---|---|---|
| WHO severity band correct | 60.2% | 53.6% |
| MAE | 13.9 pp | 15.4 pp |
| worst-20% districts recovered | 27.1% | 20% (random) |

::: {.callout-warning appearance="minimal"}
A targeting lift of **1.36x**. The map beats doing nothing, but not by much. It is useful for
*ranking* districts within the better-measured outcomes, and should not be quoted as a district
prevalence estimate.
:::

---

## A national model from national data (VMNIS)

The full WHO VMNIS dump holds **187 countries** (the project extract had 8).
Vitamin A in preschool children: **528 surveys, 70 countries**.

Leave-one-**country**-out, World Bank covariates, no benchmarking:

| model | MAE | Pearson |
|---|---|---|
| random forest | **11.6 pp** | 0.187 overall, 0.648 for vitamin A |
| global mean (null) | 13.1 pp | — |

**Read against the ceiling.** Repeat national surveys of the *same* country disagree by ~1.1 on the
logit scale, so r_max = 0.66 for vitamin A — the model sits at **98% of achievable**.

Modelling survey heterogeneity explicitly (inflammation adjustment, biomarker, sample size)
**raises the ceiling 0.505 → 0.633** and cuts MAE 12.20 → **11.04 pp**: about a quarter of the
apparently unexplainable variance is *survey method*, not real change.

---

## What predicts nationally

:::: {.columns}
::: {.column width="50%"}
| rank | variable |
|---|---|
| 1 | under-5 mortality |
| 2 | electricity access |
| 3 | food production index |
| 4 | life expectancy |
| 5 | health expenditure per capita |
| **20** | **undernourishment prevalence** |
:::
::: {.column width="50%"}
Panels split along interpretable lines:

- **Vitamin A** → child-health composite
- **Folate** → the food system
- **Zinc** → infrastructure

**Calorie sufficiency is not micronutrient sufficiency** — the headline food-security indicator
ranks *last* of 20.
:::
::::

::: {.callout-note appearance="minimal"}
At Admin-2 the top variable was *soil zinc*, ranking first for vitamin A, B12 and folate, where it
has no mechanism. Much of that covariate signal is geography wearing a covariate label. The
national predictors are mechanistically far more defensible.
:::

---

## Negative results worth keeping

Four principled ideas tested that **did not** work, recorded so they are not retried:

| idea | result |
|---|---|
| binomial likelihood on survey counts | loses to Gaussian-on-logit at 30–170 districts |
| binomial GAM with a spatial field | rho 0.298 vs 0.354 for a plain spatial smoother |
| modelling the district biomarker **SD** | worse — that SD is estimated from ~13 people |
| hierarchical individual-biomarker model | MAE 15.5 pp vs 10.8 pp for the indicator route |

::: {.callout-note appearance="minimal"}
At this sample size **model machinery is not the binding constraint**. Predictor selection, getting
geography into the model, and being honest about the ceiling are what move the numbers.
:::

---

# Full Rebuild 2026-08 {background-color="#1f4d2e"}

## What this rebuild was

::: {.callout-note appearance="minimal"}
Full mode, 12-learner stack, GEE covariate hygiene on, 4 countries.
A fifth country (Tanzania, vitamin A only, TDHS 2010) was added and then
withdrawn after a project collaborator flagged its dried-blood-spot RBP
measurements as unreliable, consistent with data-quality anomalies found
independently in this pipeline. See `archive/ARCHIVE_MANIFEST.md`.
:::

Two changes since the last published numbers:

1. **Gaussian process dropped** from the full stack. It is $O(n^3)$ and made the
   pooled LOCO targets (n = 10,011 and 13,107) effectively non-terminating.
2. **Two baselines added** to the area-level comparison --- a national-mean null
   and a spatial-smoothing-only model --- so "better than the alternatives" can
   be separated from "better than nothing".

Point 2 is what changed the interpretation.

---

## Dropping the Gaussian process cost nothing

Measured directly rather than assumed: the `gp_sensitivity` target refits
**Ghana x child iron** with the GP restored and compares held-out performance.

| Model | without GP | with GP | difference |
|---|---|---|---|
| Binary (AUC) | 0.701550371878371 | 0.701550371878371 | **0.0000** |
| Continuous ($r$) | 0.2488 | 0.2540 | +0.0052 |

The binary result is identical to 16 decimal places --- the GP received **zero
weight** in the ensemble. A learner that contributes nothing while dominating
runtime is a pure cost.

::: {.callout-tip appearance="minimal"}
The check is retained as a pipeline target, so the decision stays testable rather
than becoming folklore.
:::

---

## The area-level models do not beat doing nothing

Median across 24 country x outcome cells, Admin-2:

| Approach | n | median $r$ | median MAE | $r$ / $r_{max}$ |
|---|---|---|---|---|
| Individual SL | 24 | **0.516** | 8.55 pp | **0.92** |
| Spatial only (no covariates) | 18 | 0.393 | 9.98 pp | 0.78 |
| Area SL (NLL) | 24 | 0.215 | 9.84 pp | 0.44 |
| Area SL (MSE) | 24 | 0.160 | 9.31 pp | 0.48 |
| **National mean (null)** | 24 | --- | **9.55 pp** | --- |

Two readings that only became visible once the baselines existed:

- **Neither area-level model beats the null.** MAE 9.31 and 9.84 pp against
  9.55 pp for predicting the national mean everywhere.
- **Spatial smoothing alone beats both.** At Admin-2, the covariates add
  nothing beyond "nearby districts are similar".

Only the individual-level model clearly earns its place.

---

## The ceiling, not the learner, is binding

Correcting for attenuation: with sampling variance $V$ and observed
between-district variance $\mathrm{Var}(p_{obs})$,

$$\lambda = 1 - \frac{E[V]}{\mathrm{Var}(p_{obs})}, \qquad r_{max} = \sqrt{\lambda}$$

- Median $r_{max}$ across cells: **0.098** (range 0--0.773)
- **4 of 24** cells have no detectable signal at all ($\lambda \le 0$)
- **19 of 114** approach-rows are flagged `signal = FALSE`

A perfect model would score $r \approx 0.10$ on the median cell. Individual SL's
median r_share (r divided by that cell's own r_max) is **0.92**. The gap between
achieved r and a theoretical 1.0 is not modelling headroom --- it is sampling
noise in the target.

::: {.callout-warning appearance="minimal"}
**12 of 24** cells show $r$ / $r_{max} \ge 1$ for the individual-level model ---
half of them. That is not superhuman performance; it means the ceiling itself
is estimated with substantial error at this sample size (as few as 14 areal
units per country) and should be read as an interval, not a point.
:::

---

## National prevalence: the models do work here

| Metric | Value |
|---|---|
| Cells | 24 |
| Median absolute error | **0.54 pp** |
| Median relative error | 8.6% |
| Predictions inside the survey's 95% CI | **24 of 24** |

Aggregating noisy district predictions to national level averages the noise away.
The same models that discriminate poorly *between* districts recover the national
total accurately.

**This is the deliverable that survives scrutiny.** Sub-national maps should be
presented as uncertainty-aware and provisional; national estimates are solid.

---

## Transportability remains the hard problem

Leave-one-country-out, median across held-out cells:

| Outcome | cells | Spearman | RMSE | national bias |
|---|---|---|---|---|
| child iron | 4 | 0.216 | 23.0 pp | -12.7 pp |
| child vitA | 4 | 0.167 | 11.7 pp | -4.8 pp |
| women iron | 4 | 0.273 | 21.9 pp | -0.3 pp |
| women vitA | 4 | 0.021 | 4.1 pp | -0.5 pp |

Across 21 benchmarked methods the best LOCO MAE is **10.6 pp** (penalized
regression) against **11.1 pp** for the null --- a modest gain with pattern
correlation near zero ($r$ = 0.09).

The methods with the best *pattern* correlation are the spatial ones
(`spatial_plus_soil` $r$ = 0.284) but they have *worse* level accuracy
(MAE 13.1 pp). **Level and pattern trade off, and neither is good.**

Report Spearman and national bias separately --- a single AUC hides which failed.

---

## Where the signal actually is

Median AUC drop when a predictor domain is removed (24 cells, vs the `none`
baseline):

| Domain removed | cells | median AUC drop |
|---|---|---|
| Infant & child mortality | 11 | 0.246 |
| HIV/AIDS | 10 | 0.125 |
| Survey indicators | 6 | 0.124 |
| **Environment** | 24 | **0.043** |
| **Climate** | 24 | **0.034** |
| Infectious disease | 23 | 0.020 |
| Everything else | --- | $<$ 0.01 |

::: {.callout-important appearance="minimal"}
The three largest drops are **survey-derived** domains, present in only 6--11 of
24 cells. They are not available for a proxy-only prediction model and cannot be
used for the intended use case. Among the proxy domains that exist everywhere,
Environment and Climate carry nearly all the signal --- and they are worth
0.043 and 0.034 AUC.
:::

**Calibration** is poor: median slope **0.44** (IQR 0.10--0.76), i.e.
predictions are systematically over-extreme and need shrinkage before use.

---

## What would actually move performance

Ranked by expected value, given the above:

1. **Raise the ceiling, not the model.** $r_{max}$ = 0.10 caps everything at
   Admin-2 (and is itself noisy enough that half of cells exceed it). Pooling
   survey rounds, modelling at Admin-1 where the design
   supports it, or borrowing strength across outcomes attacks the binding
   constraint. More learners do not.
2. **Calibrate before publishing** any district-level number (slope 0.44).
3. **Treat national estimates as the headline deliverable** and sub-national maps
   as exploratory, until the ceiling improves.
4. **Vet new countries before onboarding them.** Tanzania (vitamin A, TDHS 2010)
   was added then withdrawn after a data-quality flag on its dried-blood-spot RBP
   assay; `scripts/refresh_gee_coverage.R --run --rebuild` remains available for
   closing covariate gaps in whatever country replaces it.

---

## Summary

::: {.callout-note appearance="minimal"}
This pipeline estimates **Admin-2** and national prevalence of five micronutrient
deficiencies across **four** African countries (Gambia, Ghana, Sierra Leone,
Malawi) from proxy predictors in 13+ external sources, using a
12-learner SuperLearner ensemble. Rebuilt in full mode 2026-08: 831 targets,
0 errors. (A fifth country, Tanzania, was added and then withdrawn --- see
Full Rebuild slide.)

**What works.** National prevalence is recovered accurately: all **24 of 24**
country x outcome predictions fall inside the survey's own 95% CI, median
absolute error **0.54 pp** (5.4% relative).

**What does not.** Sub-national discrimination is weak. Median AUC across the 24
binary cells is **0.594**; AUC correlates *negatively* with the number of events
($r$ = **-0.521**, $p$ = **0.009**), so the high values sit in cells too small to
measure. In the well-powered tertile median AUC is **0.562**. At Admin-2 the
area-level models do **not** beat a national-mean null (MAE **9.3--9.8** vs **9.55** pp),
and a spatial-smoothing-only model beats both covariate models ($r$ **0.39** vs
**0.16--0.22**).

**Why.** The Admin-2 outcome is noise-dominated, and the ceiling itself is highly
uncertain at this sample size. The median reliability ceiling
is $r_{max}$ = **0.098** --- half of all cells exceed it, which is a sign of
estimation noise in the ceiling, not superhuman models. The individual-level model
already captures most of it (median r_share **92%**); the binding constraint is the data, not
the learner.
:::

---


## Discussion & Feedback {background-color="#1a3a5c"}

- Are the predictor domains comprehensive?
- Preferred visualization of uncertainty?
- Suggestions for stakeholder engagement?
- Priority micronutrients or countries for expansion?

# Covariate Rebuild & Honest Baselines 2026-08-30 {background-color="#0F6E66"}

## What this round actually changed

Three things, in descending order of how much they moved the numbers:

1. **Covariate harmonisation.** 294 predictors shared by all four countries,
   drawn from a registry of 23 providers, replacing two parallel vocabularies
   whose intersection was near zero.
2. **Nine join-key defects fixed.** GADM Admin-2 names are not unique within a
   country --- Malawi has 256 polygons under 243 names --- and nine consumers
   joined on the name alone.
3. **Two honest baselines added** to every comparison: a national-mean null and
   a covariate-free spatial smoother.

::: {.callout-note appearance="minimal"}
Point 2 changed row counts, not conclusions. Malawi's area approaches had been
scored on 90 rows against 87 surveyed districts, so they were not comparable
with the individual-level rows in the same table. Every arm now scores on 87.
:::

---

## Selecting covariates on all the data inflates accuracy by +0.18

Two otherwise identical pipelines. Both cross-validated by leaving out whole
regions. The **only** difference is whether the covariate screen saw the
held-out regions.

| | mean $r$ | positive in |
|---|---|---|
| Screen inside each fold (honest) | 0.187 | --- |
| Screen on all data, then cross-validate | 0.369 | --- |
| **Difference** | **+0.182** | **20 / 20 cells** |

The gap is largest exactly where true signal is weakest:

| Cell | in-fold | screened on all data |
|---|---|---|
| Sierra Leone, women vitamin A | 0.00 | **0.51** |
| Ghana, women B12 | −0.12 | **0.27** |
| Malawi, women zinc | −0.11 | **0.22** |

Three cells with zero or negative real skill become respectable-looking.

---

## A covariate-free spatial smoother matches 294 covariates

Mean rank correlation across 24 Admin-2 cells, all honestly cross-validated:

| Approach | mean $r$ |
|---|---|
| Spatial smoother, **no covariates** | **0.304** |
| Area wSL, prescreen in-fold | 0.187 |
| Area wSL, literature covariates (46) | 0.147 |
| Area wSL, precision-weighted (294) | 0.145 |

At Admin-2, geography carries the signal. The covariates add little on top of
knowing where a district is.

---

## But at Admin-1 the covariates clearly win

Same three arms, same leave-one-unit-out loop, same countries. Only the
geographic unit changes.

| | covariates | spatial only |
|---|---|---|
| **Admin-2** (24 cells) | $r$ = 0.156 | $r$ = 0.137 |
| **Admin-1** (14 cells) | $r$ = **0.437** | $r$ = 0.209 |

::: {.callout-warning appearance="minimal"}
The largest Admin-1 gains are on the **rarest** outcomes --- Malawi women's
vitamin A is 1.3% prevalence, 11 cases in 836. Across 87 districts that is
0--2 cases each. Much of the Admin-1 advantage is a sample-size effect, which
is a real finding but a different claim from "covariates work better there".
:::

---

## A hybrid, with the level chosen before looking at performance

Rule fixed in advance: use Admin-1 where a country has at least 8 regions,
otherwise Admin-2. It depends only on how a country is subdivided.

| Country | regions | level used |
|---|---|---|
| The Gambia | 6 | Admin-2 |
| Sierra Leone | 4 | Admin-2 |
| Ghana | 16 | **Admin-1** |
| Malawi | 27 | **Admin-1** |

Pooled over all 24 cells: **$r$ = 0.304, MAE 7.25 pp** --- double the
Admin-2-everywhere correlation, while still covering all four countries.

Choosing the level by *which scored better* would be selection on the test
data, which is the very bias measured two slides ago.

---

## Rank transports across borders. Level does not.

Leave-one-country-out, Admin-2:

- **7 of 15** cells reach significance on rank correlation against a
  permutation null, up to $r$ = 0.55
- Mean absolute national bias **10.6 pp**, worst case **42 pp**

```{r}
#| fig-height: 4.6
source(here::here("R", "plot_forest.R"))
forest_transport()
```

---

## Which is why anchoring to regional totals is the design that pays

The model supplies the ranking; the design-based survey supplies the level.

```{r}
knitr::kable(table_anchor_arms(), align = "lrrrr")
```

Anchoring more than doubles rank correlation and cuts bias from 3.60 to
1.49 pp --- better in 20 of 24 cells. National anchoring buys almost nothing
(+0.006).

---

## National prevalence is where the models genuinely work

```{r}
#| fig-height: 5
forest_national()
```

All 24 predictions fall inside the survey's own 95% CI; mean absolute error
0.96 pp.

---

## Data we hold and do not use

| Opportunity | Where | Size |
|---|---|---|
| **sTfR** --- body iron stores, less inflammation-sensitive | **all four countries** | full biomarker sample |
| School-age children (5--14 y), `vitA_def` + `iron_def` already derived | Malawi | 758 |
| Men | Malawi | 219 |
| Urinary iodine | Gambia, Sierra Leone | 1,383 / 1,422 |
| RBC folate (long-term status) | Malawi | 797 |

sTfR is the most promising: the cross-country failure lives in raw ferritin,
so an iron outcome that leans less on ferritin attacks the problem directly.

---

## One inconsistency worth fixing before publication

**Vitamin A is BRINDA-harmonised across all four countries. Iron is not.**

| Country | iron adjustment used |
|---|---|
| The Gambia | `gw_LogFerAdj` (log scale) |
| Ghana | `gw_cFerrAdjThurn` (Thurnham) |
| Sierra Leone | `gw_wFerAdjBR1` (BRINDA regression) |
| Malawi | `sf_reg` (regression-adjusted) |

The raw ingredients for a uniform BRINDA adjustment (marker, CRP, AGP) exist in
all four. Harmonising it does **not** fix transport --- that was tested, and the
offset is in raw ferritin --- but harmonising one outcome and not the other is
hard to defend in review.

---

## Where this leaves the programme

**Defensible now**

- National prevalence, within survey error, in every cell
- District *rankings*, anchored to design-based regional totals
- Targeting value: above 1.0x lift in 22 of 24 combinations

**Not defensible**

- Absolute district prevalence from covariates alone
- Any accuracy figure whose covariates were screened on all the data

**Most promising next**

- sTfR-based iron outcomes across all four countries
- Continuous biomarkers rather than binaries, especially for rare outcomes
- Malawi's school-age children as a fifth population

---

## Variable importance: what the ablations actually say

Two headline numbers from the domain ablation are **artifacts**, and finding
that is the result.

| Domain removed | mean AUC lost |
|---|---|
| Survey Indicators | 0.152 |
| Infant & Child Mortality | 0.121 |
| HIV/AIDS | 0.117 |
| Agriculture | 0.114 |

Removing **one** HIV/AIDS variable cannot cost 0.12 AUC. It doesn't.

---

## Two degenerate cells were driving every domain

In **Ghana women's vitamin A** (1.5% prevalence) and **Malawi women's B12**
(2.3%), removing *any* domain --- including a single variable --- collapses AUC
from 0.89 / 0.91 to **exactly 0.500**.

That is the model failing to fit at all, not information being lost. Those two
cells supply 35 rows at a loss of ~0.39--0.41 and drag every domain's mean
upward.

::: {.callout-important appearance="minimal"}
Both are rare outcomes. Removing even one predictor destabilises the fit
entirely --- the same rare-outcome fragility that shows up in the Admin-1
comparison.
:::

---

## Excluding those two cells, domain removal costs nothing

| Domain removed | variables | **median** AUC lost |
|---|---|---|
| Environment | 57 | **0.029** |
| Climate | 21 | 0.004 |
| Survey Indicators | 3 | 0.001 |
| Infectious Disease | 12 | 0.001 |
| Infrastructure | 2 | 0.001 |
| Water & Sanitation | 10 | 0.001 |

- **141 of 332** domain removals *improve* AUC
- **266 of 720** single-variable permutation drops are $\le 0$
- Median top-5 permutation drop: **0.008 AUC**

Only Environment --- 57 variables --- has a median cost distinguishable from
zero.

---

## No variable is consistently important

Across 24 country x outcome cells, the variable ranked **first** by permutation
importance is:

| Top-ranked variable | cells (of 24) |
|---|---|
| `dhs2014_AN_ANEM_W_ANY` | 3 |
| `dhs2015_CH_VACC_C_DP1_adm2` | 2 |
| every other variable | 1 |

Eighteen different variables take first place across 24 cells. With
`VARIMP_N_PERM = 2` the permutation estimates are themselves noisy, so this is
close to what a null would produce.

::: {.callout-note appearance="minimal"}
SHAP tells the same story: the top-ranked contributions are
`mean_abs_shap` ~ 0.004 on the probability scale --- a 0.4 percentage-point
influence per district.
:::

---

## So: is anything associated with the outcomes at all?

Bivariate association between every harmonised predictor and district
prevalence, ~296 predictors x 24 cells, with two nulls:

- **Naive** correlation test --- treats districts as independent
- **Restricted permutation** --- shuffles the outcome *within* Admin-1 regions,
  holding the regional pattern fixed, so a covariate must explain
  district-level deviation on top of geography

The gap between the two is the spatial-confounding correction. Malawi women's
B12: **48** predictors significant naively, **15** under the restricted null.

*(Full FDR results and the penalised-regression retained sets follow.)*

---

## Nothing survives FDR once space is accounted for

7,104 tests: ~296 harmonised predictors x 24 country x outcome cells.

| Test | significant | of 7,104 |
|---|---|---|
| Naive $p < 0.05$ | 1,209 | 17.0% |
| Naive, BH within cell | 476 | 6.7% |
| Naive, BH across all cells | 310 | 4.4% |
| **Restricted permutation $p < 0.05$** | **519** | **7.3%** |
| **Restricted, BH within cell** | **0** | **0.00%** |
| **Restricted, BH across all cells** | **0** | **0.00%** |

Expected by chance at $p<0.05$: **355**.

Naive testing gives 3.4x enrichment over chance. The spatially-restricted null
gives **1.5x** --- and after FDR correction, **not one predictor survives, in
any cell**.

---

## What that does and does not mean

::: {.callout-warning appearance="minimal"}
The restricted permutation holds the **regional** pattern fixed and asks what a
covariate adds *on top of geography*. It is deliberately conservative, and it is
not the same target as the LORO models, which predict held-out **regions**.
:::

Read together:

- **Between-region** association exists --- 310 predictors survive global BH on
  the naive test --- but is confounded with spatial smoothness
- **Within-region**, beyond what geography already gives you: **nothing**

Which is exactly why a covariate-free spatial smoother matches 294 covariates.
Two results, one explanation.

The strongest single associations are mechanistically sensible and still fail
multiplicity: Gambia women's iron vs women's secondary education
$r = -0.76$ ($p_{perm} = 0.002$, $q = 0.20$); Gambia women's vitamin A vs 2019
precipitation $r = -0.64$.

---

## Penalised regression agrees: mostly nothing is retained

Elastic net ($\alpha = 0.5$), CV-selected $\lambda_{1se}$, per cell:

| | |
|---|---|
| Median predictors retained | **0** |
| Cells retaining **nothing** | **13 of 18** |
| Mean deviance explained | 0.063 |

| Country | outcomes with $>0$ retained |
|---|---|
| Gambia | child iron (3), women iron (9) |
| Ghana | child vitA (3), child iron (16), women iron (15) |
| **Malawi** | **none --- 0 of 8, on 87 districts** |

At the looser $\lambda_{min}$ the median rises to 8.5 predictors and deviance
explained to 0.255 --- but that is the penalty CV rejects.

---

## The signal, where it exists, is iron

Every cell retaining predictors under $\lambda_{1se}$ is an **iron** outcome,
bar one. Vitamin A, folate, B12 and zinc retain nothing anywhere.

That is consistent across four independent lines of evidence:

1. Iron cells dominate the significant LOCO transport correlations (7 of 15)
2. Iron cells dominate the significant within-country correlations (8 of 24)
3. Iron is the only outcome family the penalised regression keeps
4. Malawi --- largest district sample, most outcomes --- retains **nothing**

::: {.callout-note appearance="minimal"}
Malawi retaining zero predictors across all eight outcomes on 87 districts is
the single most informative number here. It is not a power problem.
:::

---

## Model the biomarker, or the binary?

Every outcome here is a prevalence: the share below a cutoff. Dichotomising
discards information, most of it when the outcome is rare. We tested modelling
the district **mean of the log biomarker** and converting afterwards.

22 cells, leave-one-region-out, same covariates and stack:

| Arm | MAE (pp) | RMSE (pp) | rank $r$ |
|---|---|---|---|
| 1 binary (direct) | 9.40 | 12.36 | 0.168 |
| 2 continuous $\rightarrow$ prevalence | 10.80 | 13.08 | 0.196 |
| 3 continuous, oracle SD | 9.92 | 12.20 | 0.434 |
| **4 conversion ceiling (true mean)** | **7.49** | **9.11** | **0.672** |
| 5 null (train mean) | 8.06 | 10.14 | --- |

---

## The conversion is not the problem. Predicting the mean is.

Arm 4 feeds the **true** district mean through the same conversion --- no
prediction at all. It reaches MAE 7.49 and $r$ = 0.672, beating every other arm
**including the null**, and beats the binary arm in 14 of 22 cells.

::: {.callout-important appearance="minimal"}
So the mean-to-prevalence step is sound. A district mean predicted well would
convert to prevalence better than anything we currently produce. The entire
deficit is in predicting the mean --- which fails for the same reason
everything else does.
:::

District-mean prediction on the continuous scale, $r$ ranges **--0.68 to +0.69**.
In 9 of 22 cells the RMSE of the predicted mean *exceeds* the between-district
SD --- worse than predicting the overall mean.

---

## Rarity: the hypothesis holds, with one caveat

| Prevalence band | cells | binary MAE | continuous MAE |
|---|---|---|---|
| **rare (<5%)** | 6 | 4.61 | **2.99** |
| mid (5--20%) | 10 | 8.77 | 11.86 |
| common (>20%) | 6 | 15.25 | 16.84 |

Continuous cuts error by 35% where the outcome is rare, and is worse elsewhere.

::: {.callout-warning appearance="minimal"}
**One cell dominates that average.** Sierra Leone women's B12 (0.61%
prevalence): binary MAE **7.67 pp** --- a nonsensical error on a 0.6% outcome ---
against **0.53** for continuous. Excluding it, continuous wins 3 of the
remaining 5 rare cells by modest margins.
:::

Neither arm reliably beats the null: binary in 4 of 22 cells, continuous in 2.

---

## What the continuous route is and is not worth

**Worth doing**

- For **rare outcomes**, where the binary model can produce absurd errors
- Because the conversion ceiling shows real headroom ($r$ = 0.672)
- Because the oracle-SD arm ($r$ = 0.434 vs 0.196) shows the within-district
  spread carries information a single pooled SD discards

**Not a fix**

- It does not rescue ranking: 0.196 vs 0.168 overall
- It is *worse* on absolute error for mid and common outcomes
- It leaves the binding constraint untouched --- the district mean is no easier
  to predict from these covariates than the prevalence is

---

## WHO risk categories: the policy-legible version

Classifying each district into the WHO public-health categories
(None / Mild / Moderate / Severe) for child vitamin A:

| Level | units | covariates exact | **null exact** | covariates within-1 | null within-1 |
|---|---|---|---|---|---|
| **Admin-2** | 206 | 33% | **32%** | 69% | 63% |
| **Admin-1** | 43 | **56%** | 35% | 95% | 91% |

::: {.callout-important appearance="minimal"}
At Admin-2 the covariate model classifies risk categories **no better than
assigning every district the national prevalence** --- 33% against 32%. At
Admin-1 the gap is real: 56% against 35%.
:::

The 2026 Ghana deck reported 56% exact at Admin-1, which reproduces exactly.
Its Admin-2 figures (44% full model / 31% proxies-only) are consistent with the
31--33% seen here for a proxy-only model.

---

## Resolution result, confirmed on the rebuilt covariates

Re-run against the restored `area_covariates_*` targets --- identical to the
first pass, so the numbers are not an artifact of the fallback path:

| | covariates | spatial only | beats spatial |
|---|---|---|---|
| **Admin-2** (24 cells) | $r$ = 0.156 | $r$ = 0.137 | 6 / 24 |
| **Admin-1** (14 cells) | $r$ = **0.437** | $r$ = 0.209 | 7 / 14 |
| **Hybrid** (24 cells) | $r$ = **0.304**, MAE 7.25 pp | $r$ = 0.140, MAE 7.30 pp | 10 / 24 |

Gambia (6 regions) and Sierra Leone (4) are **not estimable** at Admin-1 and
fall back to Admin-2 under the a-priori rule.

---

## What should the model carry across a border?

District ranking transports; the absolute level does not. So stop asking the
model to carry the level. Two versions of that idea, opposite answers:

| target | Spearman | MAE (pp) | \|bias\| (pp) |
|---|---|---|---|
| level (baseline) | 0.146 | 13.8 | 10.2 |
| deviation from national mean | *worse* | --- | --- |
| **within-country z-score** | **0.199** | **10.0** | **5.9** |

Removing the country **mean** alone fails --- it leaves the between-country
**spread** in the target. Removing both is what works: +0.053 Spearman, better
in 3 of 4 outcomes.

::: {.callout-warning appearance="minimal"}
**Do not quote the two gains together.** A positive rescale plus an offset
cannot reorder predictions, so the rank gain is a real modelling improvement.
The larger error and bias gains come mostly from the **anchor** --- the
held-out country's national prevalence --- which is a deployment assumption,
not evidence the covariates improved.
:::

---

## Two changes that did not replicate

Reported because the negative result is the useful part.

| change | expected | measured | better in |
|---|---|---|---|
| raw prevalence scale (drop the logit clamp) | +0.048 | **--0.020** | 2 / 4 |
| district coordinates as predictors | +0.063 | **0.000** | 1 / 4 |

The raw-scale result had a strong prior: **8 of 24 cells have >40% of districts
observed at exactly 0%**, every one clamped to logit = --5.3. It still did not
help this recipe. The earlier gain was largest for a *spatial spline*, which is
a different model.

Coordinates fail for a mechanical reason: with a top-30 correlation prescreen
over 294 predictors, two coordinate columns rarely survive selection. The
earlier gain came from a curated 16-variable set where coordinates were a large
share of the model.

::: {.callout-note appearance="minimal"}
Both are kept in the code as options --- and the scale dispatcher fixed a real
defect, since `recipe$scale` previously did nothing at all.
:::
