---
title: "Micronutrient Deficiency Prediction: Multi-Country Results Report"
subtitle: "Proxy-Only Models --- Cross-Outcome and Cross-Country Analysis"
author: "Micronutrient Prediction Project"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
    self-contained: true
    smooth-scroll: true
execute:
  warning: false
  message: false
knitr:
  opts_chunk:
    comment: "#>"
    fig.align: "center"
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(readr)
library(scales)
library(sf)
library(geodata)
library(viridis)
library(patchwork)

theme_set(theme_minimal(base_size = 13))

# ── Configuration ──────────────────────────────────────────────────────────
countries <- c("gambia", "ghana", "sierraleone", "malawi")
outcomes  <- c("child_vitA", "women_vitA", "child_iron", "women_iron")
gadm_codes <- c(gambia = "GMB", ghana = "GHA", sierraleone = "SLE", malawi = "MWI")

country_labels <- c(
  gambia      = "Gambia",
  ghana       = "Ghana",
  sierraleone = "Sierra Leone",
  malawi      = "Malawi"
)

outcome_labels <- c(
  child_vitA = "Vitamin A (children)",
  women_vitA = "Vitamin A (women)",
  child_iron = "Iron deficiency (children)",
  women_iron = "Iron deficiency (women)"
)

# ── Helper: safe target loading ────────────────────────────────────────────
load_target <- function(name) {
  obj <- tryCatch(
    targets::tar_read_raw(name, store = here::here("_targets")),
    error = function(e) NULL
  )
  # Return NULL for non-data.frame results or empty data.frames
  if (is.data.frame(obj) && nrow(obj) == 0) return(NULL)
  obj
}

# ── Load ALL targets dynamically ───────────────────────────────────────────
# CV performance
cv_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("cv_perf_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      cv_list[[tname]] <- obj
    }
  }
}
cv_perf <- if (length(cv_list) > 0) bind_rows(cv_list) else NULL

# Admin-1 prevalence
a1_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin1_prev_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a1_list[[tname]] <- obj
    }
  }
}
a1_all_raw <- if (length(a1_list) > 0) bind_rows(a1_list) else NULL

# Admin-2 SL predictions
a2_sl <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin2_sl_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_sl[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Admin-2 survey prevalence
a2_svy <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("svy_admin2_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_svy[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Admin-2 error
a2_err_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin2_error_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_err_list[[tname]] <- obj
    }
  }
}
a2_error <- if (length(a2_err_list) > 0) bind_rows(a2_err_list) else NULL

# Area-level models
area_models <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("area_model_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      area_models[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Bootstrap CIs
boot_ci <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("bootstrap_ci_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      boot_ci[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Domain ablation
abl_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("ablation_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      abl_list[[tname]] <- obj
    }
  }
}
ablation <- if (length(abl_list) > 0) bind_rows(abl_list) else NULL

# Transportability (LOCO)
loco_results <- list()
for (oc in outcomes) {
  tname <- paste0("loco_", oc)
  obj <- load_target(tname)
  if (!is.null(obj)) loco_results[[oc]] <- obj
}
transportability_summary <- load_target("transportability_summary")

# Pooled datasets (metadata only)
pooled <- list()
for (oc in outcomes) {
  tname <- paste0("pooled_", oc)
  obj <- load_target(tname)
  if (!is.null(obj)) pooled[[oc]] <- obj
}

# ── GADM boundaries for all countries ──────────────────────────────────────
poly_a1 <- list()
poly_a2 <- list()
for (ctry in countries) {
  code <- gadm_codes[ctry]
  poly_a1[[ctry]] <- tryCatch({
    g <- geodata::gadm(country = code, level = 1, path = tempdir())
    sf::st_as_sf(g) |>
      dplyr::select(NAME_1) |>
      dplyr::rename(Admin1 = NAME_1) |>
      sf::st_transform(crs = 4326)
  }, error = function(e) NULL)

  poly_a2[[ctry]] <- tryCatch({
    g <- geodata::gadm(country = code, level = 2, path = tempdir())
    sf::st_as_sf(g) |>
      dplyr::select(NAME_1, NAME_2) |>
      dplyr::rename(Admin1 = NAME_1, Admin2 = NAME_2) |>
      sf::st_transform(crs = 4326)
  }, error = function(e) NULL)
}

# ── Shared helpers ─────────────────────────────────────────────────────────
# Standardise Admin-1 column names from targets output
standardize_a1 <- function(df) {
  if (is.null(df) || nrow(df) == 0) return(NULL)
  if ("sl_prev" %in% colnames(df)) df <- df |> rename(pred_prevalence = sl_prev)
  if ("obs_prev" %in% colnames(df)) df <- df |> rename(obs_prevalence = obs_prev)
  if (!"label" %in% colnames(df) && "outcome" %in% colnames(df)) {
    df <- df |> mutate(label = outcome_labels[outcome])
  }
  df
}

# Choropleth helper for Admin-1
map_admin1 <- function(poly, data, fill_col, title,
                       legend_lab = "Prevalence",
                       limits = NULL, option = "plasma") {
  if (is.null(poly) || is.null(data)) return(NULL)
  map_data <- poly |> dplyr::left_join(data, by = "Admin1")
  if (is.null(limits)) {
    vals <- map_data[[fill_col]]
    limits <- range(vals, na.rm = TRUE)
  }
  ggplot(map_data) +
    geom_sf(aes(fill = .data[[fill_col]]), colour = "white", linewidth = 0.3) +
    scale_fill_viridis_c(
      name   = legend_lab,
      labels = percent_format(accuracy = 1),
      limits = limits,
      na.value = "grey85",
      option = option
    ) +
    labs(title = title) +
    theme_void(base_size = 11) +
    theme(plot.title = element_text(size = 10, face = "bold"),
          legend.position = "bottom",
          legend.key.width = unit(1.2, "cm"))
}

# Choropleth helper for Admin-2
map_admin2 <- function(poly, data, fill_col, title,
                       legend_lab = "Prevalence",
                       limits = NULL, option = "plasma") {
  if (is.null(poly) || is.null(data)) return(NULL)
  map_data <- poly |> dplyr::left_join(data, by = "Admin2")
  if (is.null(limits)) {
    vals <- map_data[[fill_col]]
    limits <- range(vals, na.rm = TRUE)
  }
  ggplot(map_data) +
    geom_sf(aes(fill = .data[[fill_col]]), colour = "white", linewidth = 0.15) +
    scale_fill_viridis_c(
      name   = legend_lab,
      labels = percent_format(accuracy = 1),
      limits = limits,
      na.value = "grey85",
      option = option
    ) +
    labs(title = title) +
    theme_void(base_size = 10) +
    theme(plot.title = element_text(size = 9, face = "bold"),
          legend.position = "bottom",
          legend.key.width = unit(1, "cm"))
}

# Count loaded targets for diagnostic reporting
n_cv     <- length(cv_list)
n_a1     <- length(a1_list)
n_a2_err <- length(a2_err_list)
n_abl    <- length(abl_list)
n_boot   <- length(boot_ci)
n_loco   <- length(loco_results)
```


# Executive Summary

This report presents results from the micronutrient deficiency prediction pipeline applied across **four West and Southern African countries** --- the Gambia, Ghana, Sierra Leone, and Malawi --- covering four outcomes (vitamin A deficiency and iron deficiency in both children and women of reproductive age).

The pipeline uses **proxy-only predictors**: externally available data from DHS aggregate indicators, MICS, Google Earth Engine (GEE) satellite imagery, IHME disease burden estimates, MAP malaria surfaces, WFP food prices, and food security (FSEC) indicators. These represent data sources available nationally without requiring a concurrent biomarker survey. Leave-one-country-out (LOCO) cross-validation is used to assess transportability of predictions across countries.

::: {.callout-important}
## Proxy-only models

All primary results use **proxy predictors only** --- external data sources that are available nationally and do not require a concurrent biomarker survey. Survey-derived (`gw_`) variables are **excluded** because:

1. They cannot be used for prediction in unsurveyed areas or time periods
2. Including them produces artificially inflated AUC due to collinearity with the outcome
3. The policy-relevant question is: *how well can we predict deficiency from routinely available data?*
:::

```{r data-availability-note}
cat(sprintf(
  "Targets loaded: %d CV, %d Admin-1, %d Admin-2 error, %d ablation, %d bootstrap, %d LOCO\n",
  n_cv, n_a1, n_a2_err, n_abl, n_boot, n_loco
))
```


# Cross-Validated Model Performance {#sec-cv}

## Combined CV performance table

The primary models use only proxy predictors (~200--500 columns after preprocessing). The table below shows cross-validated performance for all country--outcome combinations.

```{r cv-proxy-table}
if (!is.null(cv_perf) && nrow(cv_perf) > 0) {
  cv_display <- cv_perf |>
    mutate(
      outcome_label = outcome_labels[outcome],
      across(any_of(c("rmse", "mae", "r2", "R2", "r", "correlation",
                       "auc", "brier")), ~ round(., 3))
    )

  cv_show <- cv_display |>
    select(Country = country, Outcome = outcome_label, Model = model_type,
           any_of(c("n", "rmse", "mae", "r2", "R2", "r", "correlation",
                     "auc", "brier"))) |>
    arrange(Country, Outcome, Model)

  kable(cv_show,
        caption = "Cross-validated performance --- proxy-only models (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1, valign = "top")
} else {
  cat("No CV performance targets loaded.")
}
```

## AUC comparison across countries

```{r cv-auc-heatmap, fig.height=5, fig.width=8}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  cv_binary <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome])

  if (nrow(cv_binary) > 0) {
    ggplot(cv_binary, aes(x = outcome_label, y = country, fill = auc)) +
      geom_tile(colour = "white", linewidth = 1) +
      geom_text(aes(label = sprintf("%.3f", auc)), size = 4, fontface = "bold") +
      scale_fill_viridis_c(name = "AUC", limits = c(0.45, 0.85), option = "plasma",
                            oob = scales::squish) +
      labs(x = NULL, y = NULL,
           title = "Binary Classification AUC by Country and Outcome",
           subtitle = "Proxy-only SuperLearner models | Higher is better") +
      theme_minimal(base_size = 13) +
      theme(axis.text.x = element_text(angle = 25, hjust = 1),
            panel.grid = element_blank())
  }
}
```

```{r cv-auc-bars, fig.height=5, fig.width=9}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  cv_binary <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome])

  if (nrow(cv_binary) > 0) {
    ggplot(cv_binary, aes(x = reorder(outcome_label, auc), y = auc,
                           fill = country)) +
      geom_col(position = position_dodge(width = 0.75), width = 0.65) +
      geom_hline(yintercept = 0.5, linetype = "dashed", colour = "grey50") +
      scale_fill_brewer(palette = "Set2", name = "Country") +
      scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
      coord_flip() +
      labs(x = NULL, y = "AUC",
           title = "Cross-Validated AUC by Outcome and Country",
           subtitle = "Dashed line = chance level (0.5)") +
      theme(legend.position = "bottom")
  }
}
```

### Interpretation

Proxy-only binary classification AUCs typically range from 0.50 to 0.70 across countries and outcomes. Key patterns:

- **Vitamin A in women** tends to be the best-performing outcome across countries, likely reflecting stronger geographic gradients tied to diet, urbanisation, and food-system characteristics.
- **Iron deficiency in children** is the hardest outcome to predict from proxy data, often near chance level.
- **Cross-country variation** reveals that the strength of area-level proxy signals differs by national context --- countries with sharper urban-rural gradients may show stronger model performance.

Continuous-model R-squared values are uniformly low, confirming that area-level proxies explain limited individual-level variance. The value of these models lies in ranking areas by relative burden, not predicting individual biomarker values.


# Admin-1 Prevalence Estimates {#sec-admin1}

## Regional prevalence table

```{r admin1-combined}
a1_all <- if (!is.null(a1_all_raw)) standardize_a1(a1_all_raw) else NULL

if (!is.null(a1_all) && nrow(a1_all) > 0 &&
    "obs_prevalence" %in% colnames(a1_all)) {
  a1_summary <- a1_all |>
    filter(!is.na(Admin1)) |>
    mutate(
      obs_pct  = round(obs_prevalence * 100, 1),
      pred_pct = round(pred_prevalence * 100, 1),
      error_pp = round((pred_prevalence - obs_prevalence) * 100, 1)
    ) |>
    select(Country = country, Region = Admin1, Outcome = label, n,
           `Observed (%)` = obs_pct,
           `Predicted (%)` = pred_pct,
           `Error (pp)` = error_pp)

  kable(a1_summary,
        caption = "Admin-1 prevalence: observed vs. predicted (all countries and outcomes)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10) |>
    collapse_rows(columns = 1:2, valign = "top")
} else {
  cat("No Admin-1 prevalence targets loaded.")
}
```

## Observed vs. predicted scatter plots

```{r admin1-scatter, fig.height=8, fig.width=10}
if (!is.null(a1_all) && nrow(a1_all) > 0 &&
    "obs_prevalence" %in% colnames(a1_all)) {
  a1_plot <- a1_all |> filter(!is.na(Admin1), !is.na(obs_prevalence))

  if (nrow(a1_plot) > 0) {
    ggplot(a1_plot, aes(x = obs_prevalence, y = pred_prevalence)) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
      geom_point(aes(size = n, colour = country), alpha = 0.7) +
      ggrepel::geom_text_repel(aes(label = Admin1), size = 2,
                                max.overlaps = 15, colour = "grey40") +
      scale_x_continuous(labels = percent_format(accuracy = 1)) +
      scale_y_continuous(labels = percent_format(accuracy = 1)) +
      scale_size_area(name = "n", max_size = 7) +
      scale_colour_brewer(name = "Country", palette = "Set2") +
      facet_wrap(~ label, scales = "free") +
      labs(
        x = "Observed prevalence", y = "Predicted prevalence",
        title = "Admin-1 Prevalence: Observed vs. Predicted",
        subtitle = "All countries | Dashed line = perfect agreement"
      ) +
      theme(legend.position = "bottom")
  }
}
```


## Admin-1 prevalence maps {#sec-admin1-maps}

```{r admin1-maps, results='asis'}
#| fig-height: 6
#| fig-width: 11

if (!is.null(a1_all) && nrow(a1_all) > 0) {
  for (ctry in countries) {
    ctry_label <- country_labels[ctry]
    ctry_poly <- poly_a1[[ctry]]
    if (is.null(ctry_poly)) next

    ctry_data <- a1_all |> filter(country == ctry_label)
    if (nrow(ctry_data) == 0) next

    cat(sprintf("\n\n### %s\n\n", ctry_label))

    map_panels <- list()
    for (oc in outcomes) {
      d <- ctry_data |> filter(outcome == oc, !is.na(Admin1))
      if (nrow(d) == 0) next

      all_vals <- c(d$obs_prevalence, d$pred_prevalence)
      lims <- range(all_vals, na.rm = TRUE)

      p_obs  <- map_admin1(ctry_poly, d, "obs_prevalence",
                            paste0(outcome_labels[oc], " - Observed"),
                            limits = lims)
      p_pred <- map_admin1(ctry_poly, d, "pred_prevalence",
                            paste0(outcome_labels[oc], " - Predicted"),
                            limits = lims)

      if (!is.null(p_obs) && !is.null(p_pred)) {
        map_panels[[oc]] <- p_obs + p_pred +
          plot_layout(ncol = 2, guides = "collect") &
          theme(legend.position = "bottom")
      }
    }

    if (length(map_panels) > 0) {
      p <- wrap_plots(map_panels, ncol = 1) +
        plot_annotation(
          title = paste0("Admin-1 Prevalence: ", ctry_label),
          subtitle = "Observed (left) vs. Predicted (right) | Proxy-only models",
          theme = theme(plot.title = element_text(size = 14, face = "bold"))
        )
      print(p)
    }
  }
}
```

Each row shows one outcome. Observed prevalence (left) is the survey-weighted mean within each Admin-1 region; predicted prevalence (right) is the mean binary SL prediction. Shared colour scales within each row enable direct comparison.


## Bootstrap confidence interval maps {#sec-ci-maps}

```{r admin1-ci-maps, results='asis'}
#| fig-height: 8
#| fig-width: 14

for (ctry in countries) {
  ctry_label <- country_labels[ctry]
  ctry_poly <- poly_a1[[ctry]]
  if (is.null(ctry_poly)) next

  ci_panels <- list()
  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    b <- boot_ci[[key]]
    if (is.null(b) || is.null(b$admin1_ci)) next
    ci <- b$admin1_ci
    if (nrow(ci) == 0) next

    all_vals <- c(ci$ci_lo, ci$boot_mean, ci$ci_hi)
    lims <- range(all_vals, na.rm = TRUE)

    p_lo   <- map_admin1(ctry_poly, ci, "ci_lo",
                          paste0(outcome_labels[oc], " - Lower"),
                          limits = lims, option = "mako")
    p_mean <- map_admin1(ctry_poly, ci, "boot_mean",
                          paste0(outcome_labels[oc], " - Mean"),
                          limits = lims, option = "mako")
    p_hi   <- map_admin1(ctry_poly, ci, "ci_hi",
                          paste0(outcome_labels[oc], " - Upper"),
                          limits = lims, option = "mako")

    if (!is.null(p_lo) && !is.null(p_mean) && !is.null(p_hi)) {
      ci_panels[[oc]] <- p_lo + p_mean + p_hi +
        plot_layout(ncol = 3, guides = "collect") &
        theme(legend.position = "bottom")
    }
  }

  if (length(ci_panels) > 0) {
    cat(sprintf("\n\n### %s\n\n", ctry_label))
    p <- wrap_plots(ci_panels, ncol = 1) +
      plot_annotation(
        title = paste0("Admin-1 Bootstrap 95% CI: ", ctry_label),
        subtitle = "Lower (left), Mean (centre), Upper (right)",
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
    print(p)
  }
}
```

Each row shows one outcome with three panels: lower bound (left), bootstrap mean (centre), and upper bound (right). Regions with wider gaps between lower and upper panels have greater prediction uncertainty and are priority areas for additional data collection.

::: {.callout-warning}
## Bootstrap CIs

If the pipeline was run in fast mode (B = 4), the CI bounds are illustrative only and too narrow for publication. Rerun with B >= 200 for stable intervals.
:::


# Admin-2 Prediction Results {#sec-admin2}

## Combined Admin-2 error table

```{r admin2-error-table}
if (!is.null(a2_error) && nrow(a2_error) > 0) {
  a2_err_show <- a2_error |>
    mutate(across(where(is.numeric), ~ round(., 2))) |>
    select(Country = country, Outcome = outcome,
           any_of(c("n_admin2", "mae_pp", "rmse_pp", "pearson_r", "mean_bias_pp")))

  # Rename columns for display
  col_rename <- c(
    n_admin2    = "N districts",
    mae_pp      = "MAE (pp)",
    rmse_pp     = "RMSE (pp)",
    pearson_r   = "Pearson r",
    mean_bias_pp = "Bias (pp)"
  )
  for (old in names(col_rename)) {
    if (old %in% colnames(a2_err_show)) {
      colnames(a2_err_show)[colnames(a2_err_show) == old] <- col_rename[old]
    }
  }

  kable(a2_err_show,
        caption = "Admin-2 prediction error summary (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1, valign = "top")
} else {
  cat("No Admin-2 error targets loaded.")
}
```

### Admin-2 interpretation

Admin-2 prediction quality varies across countries and outcomes. Key patterns to look for:

- **Vitamin A outcomes** typically show stronger district-level correlations, reflecting clearer geographic gradients in VAD risk.
- **Iron deficiency in children** is often weakest at the Admin-2 level, consistent with the near-chance individual-level AUC.
- **Cross-country variation** in MAE and correlation reveals how survey design, spatial heterogeneity, and data coverage affect model performance.
- Negative mean bias indicates systematic under-prediction; positive bias indicates over-prediction.


## Admin-2 prevalence maps

```{r admin2-maps, results='asis'}
#| fig-height: 8
#| fig-width: 11

for (ctry in countries) {
  ctry_label <- country_labels[ctry]
  ctry_poly <- poly_a2[[ctry]]
  if (is.null(ctry_poly)) next

  a2_map_panels <- list()

  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    sl_data  <- a2_sl[[key]]
    svy_data <- a2_svy[[key]]
    if (is.null(sl_data) || is.null(svy_data)) next

    svy_sub <- svy_data |> dplyr::select(Admin2, svy_prev)
    all_vals <- c(sl_data$sl_prev, svy_sub$svy_prev)
    lims <- range(all_vals, na.rm = TRUE)

    p_obs  <- map_admin2(ctry_poly, svy_sub, "svy_prev",
                          paste0(outcome_labels[oc], " - Survey"),
                          limits = lims)
    p_pred <- map_admin2(ctry_poly, sl_data, "sl_prev",
                          paste0(outcome_labels[oc], " - SL Predicted"),
                          limits = lims)

    if (!is.null(p_obs) && !is.null(p_pred)) {
      a2_map_panels[[oc]] <- p_obs + p_pred +
        plot_layout(ncol = 2, guides = "collect") &
        theme(legend.position = "bottom")
    }
  }

  if (length(a2_map_panels) > 0) {
    cat(sprintf("\n\n### %s\n\n", ctry_label))
    p <- wrap_plots(a2_map_panels, ncol = 1) +
      plot_annotation(
        title = paste0("Admin-2 Prevalence: ", ctry_label),
        subtitle = "Survey observed (left) vs. SL predicted (right) | Grey = missing data",
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
    print(p)
  }
}
```

Each row shows one outcome per country. Left panels display survey-weighted observed prevalence; right panels show mean SL-predicted prevalence aggregated to Admin-2 districts.

::: {.callout-note}
Admin-2 districts with very few survey respondents (< 15) may have unreliable observed prevalences. SL predictions smooth over sampling noise by aggregating individual-level model predictions.
:::


# Domain Importance {#sec-domains}

## Combined ablation table

```{r domain-ablation}
if (!is.null(ablation) && nrow(ablation) > 0) {
  abl_display <- ablation |>
    mutate(across(where(is.numeric), ~ round(., 3)))

  show_cols <- intersect(
    c("country", "outcome", "label", "domain", "domain_removed",
      "n_dom_cols", "n_removed",
      "auc_full", "auc_reduced", "auc",
      "delta_auc", "delta_brier", "brier"),
    colnames(abl_display)
  )

  kable(abl_display[, show_cols],
        caption = "Domain ablation --- proxy domains (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10) |>
    collapse_rows(columns = 1:2, valign = "top")
} else {
  cat("No domain ablation targets loaded.")
}
```

## Ablation heatmap

```{r ablation-heatmap, fig.height=7, fig.width=10}
if (!is.null(ablation) && nrow(ablation) > 0 && "delta_auc" %in% colnames(ablation)) {
  # Determine domain column name
  dom_col <- if ("domain_removed" %in% colnames(ablation)) "domain_removed" else "domain"

  abl_heat <- ablation |>
    mutate(
      outcome_label = outcome_labels[outcome],
      ctry_outcome = paste0(country, "\n", outcome_label),
      domain_name = .data[[dom_col]]
    ) |>
    filter(!is.na(delta_auc))

  if (nrow(abl_heat) > 0) {
    ggplot(abl_heat, aes(x = domain_name, y = ctry_outcome, fill = delta_auc)) +
      geom_tile(colour = "white", linewidth = 0.5) +
      geom_text(aes(label = sprintf("%.3f", delta_auc)), size = 2.8) +
      scale_fill_gradient2(
        name = "Delta AUC",
        low = "#c0392b", mid = "white", high = "#27ae60",
        midpoint = 0, limits = c(-0.05, 0.05), oob = scales::squish
      ) +
      labs(x = "Domain removed", y = NULL,
           title = "Domain Ablation: Change in AUC When Domain Removed",
           subtitle = "Negative = domain helpful (AUC drops) | Positive = domain noisy") +
      theme_minimal(base_size = 11) +
      theme(axis.text.x = element_text(angle = 35, hjust = 1),
            panel.grid = element_blank())
  }
}
```

### Interpretation

With survey (`gw_`) variables excluded, domain ablation across all four countries reveals:

1. **No single domain dominates prediction.** Delta-AUC values are generally small, indicating the predictive signal is distributed across data sources.
2. **Redundancy across domains.** DHS, MICS, IHME, MAP, and GEE capture overlapping geographic gradients, so removing one allows others to compensate.
3. **Food security (FSEC) and WFP** indicators may contribute non-redundant information in some country--outcome combinations, particularly for vitamin A outcomes.
4. **Country-specific patterns** emerge: MAP (malaria) indicators may be more important in high-endemicity countries, while GEE (satellite) features may add value where land cover correlates with dietary diversity.

**Policy implication:** Maintaining diverse data inputs is more important than investing heavily in any single source. The marginal value of each domain is small, reinforcing the complementary role of survey data.


# Transportability --- Leave-One-Country-Out {#sec-loco}

Leave-one-country-out (LOCO) cross-validation trains the model on three countries and predicts the held-out country. This assesses whether proxy-based predictions generalise across national boundaries.

## LOCO performance table

```{r loco-table}
if (!is.null(transportability_summary) && nrow(transportability_summary) > 0) {
  loco_show <- transportability_summary |>
    mutate(across(where(is.numeric), ~ round(., 3)))

  kable(loco_show,
        caption = "Leave-one-country-out transportability results") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11)
} else if (length(loco_results) > 0) {
  # Build summary from individual LOCO targets
  loco_rows <- list()
  for (oc in names(loco_results)) {
    obj <- loco_results[[oc]]
    if (is.data.frame(obj)) {
      obj$outcome <- oc
      loco_rows[[oc]] <- obj
    } else if (is.list(obj) && "results" %in% names(obj)) {
      r <- obj$results
      if (is.data.frame(r)) {
        r$outcome <- oc
        loco_rows[[oc]] <- r
      }
    }
  }

  if (length(loco_rows) > 0) {
    loco_combined <- bind_rows(loco_rows) |>
      mutate(across(where(is.numeric), ~ round(., 3)))

    kable(loco_combined,
          caption = "Leave-one-country-out cross-validation results") |>
      kable_styling(bootstrap_options = c("striped", "hover"),
                    full_width = FALSE, font_size = 11)
  } else {
    cat("LOCO targets loaded but format could not be parsed for display.")
  }
} else {
  cat("No LOCO transportability targets loaded.")
}
```

## LOCO AUC by held-out country

```{r loco-heatmap, fig.height=5, fig.width=8}
# Try to build a heatmap from LOCO results
loco_for_plot <- NULL

if (!is.null(transportability_summary) && nrow(transportability_summary) > 0) {
  # Attempt to find AUC and held-out-country columns
  auc_col <- intersect(c("auc", "AUC", "auc_heldout"), colnames(transportability_summary))
  heldout_col <- intersect(c("held_out", "heldout", "country", "held_out_country"),
                            colnames(transportability_summary))
  oc_col <- intersect(c("outcome", "Outcome"), colnames(transportability_summary))

  if (length(auc_col) > 0 && length(heldout_col) > 0 && length(oc_col) > 0) {
    loco_for_plot <- transportability_summary |>
      rename(auc_val = !!auc_col[1], held_out = !!heldout_col[1],
             outcome_name = !!oc_col[1]) |>
      mutate(outcome_label = ifelse(outcome_name %in% names(outcome_labels),
                                     outcome_labels[outcome_name], outcome_name))
  }
}

if (!is.null(loco_for_plot) && nrow(loco_for_plot) > 0) {
  ggplot(loco_for_plot, aes(x = outcome_label, y = held_out, fill = auc_val)) +
    geom_tile(colour = "white", linewidth = 1) +
    geom_text(aes(label = sprintf("%.3f", auc_val)), size = 4, fontface = "bold") +
    scale_fill_viridis_c(name = "AUC", limits = c(0.4, 0.8),
                          option = "plasma", oob = scales::squish) +
    labs(x = NULL, y = "Held-out country",
         title = "LOCO AUC: Performance on Held-Out Country",
         subtitle = "Lower AUC = harder to predict when excluded from training") +
    theme_minimal(base_size = 13) +
    theme(axis.text.x = element_text(angle = 25, hjust = 1),
          panel.grid = element_blank())
}
```

### Interpretation

Transportability analysis reveals how well proxy-based models generalise across national contexts:

- **Countries with unique covariate profiles** (e.g., distinct agro-ecological zones, dietary patterns) will show lower LOCO AUC when held out, indicating that country-specific signal is not easily borrowed from neighbours.
- **Outcomes with strong universal predictors** (shared across all countries) will show more consistent LOCO performance.
- Countries that are hardest to predict when held out may benefit most from country-specific survey data or targeted model calibration.


# National Prevalence with Uncertainty {#sec-national}

```{r national-ci}
nat_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    b <- boot_ci[[key]]
    if (is.null(b)) next
    nc <- if (is.list(b) && "national_ci" %in% names(b)) b$national_ci else NULL
    if (!is.null(nc) && nrow(nc) > 0) {
      if (!"country" %in% colnames(nc)) nc$country <- country_labels[ctry]
      nat_list[[key]] <- nc
    }
  }
}

if (length(nat_list) > 0) {
  nat_ci <- bind_rows(nat_list) |>
    mutate(
      mean_pct = round(boot_mean * 100, 1),
      lo_pct   = round(ci_lo * 100, 1),
      hi_pct   = round(ci_hi * 100, 1),
      ci_width = round((ci_hi - ci_lo) * 100, 1),
      outcome_label = outcome_labels[outcome]
    ) |>
    select(Country = country, Outcome = outcome_label,
           `Bootstrap mean (%)` = mean_pct,
           `95% CI lower (%)` = lo_pct, `95% CI upper (%)` = hi_pct,
           `CI width (pp)` = ci_width)

  kable(nat_ci,
        caption = "National prevalence estimates with bootstrap 95% CIs (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1, valign = "top")
} else {
  cat("No bootstrap CI targets loaded.")
}
```

```{r national-ci-forest, fig.height=8, fig.width=9}
if (length(nat_list) > 0) {
  nat_plot <- bind_rows(nat_list) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      ctry_label = country
    )

  if (nrow(nat_plot) > 0 && all(c("ci_lo", "ci_hi", "boot_mean") %in% colnames(nat_plot))) {
    ggplot(nat_plot, aes(y = ctry_label, x = boot_mean,
                          xmin = ci_lo, xmax = ci_hi, colour = ctry_label)) +
      geom_pointrange(size = 0.8, linewidth = 0.6) +
      scale_x_continuous(labels = percent_format(accuracy = 1)) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      facet_wrap(~ outcome_label, scales = "free_x", ncol = 2) +
      labs(x = "Prevalence", y = NULL,
           title = "National Prevalence Estimates with Bootstrap 95% CIs",
           subtitle = "All countries and outcomes") +
      theme(strip.text = element_text(face = "bold"))
  }
}
```

::: {.callout-warning}
## Bootstrap CIs

If the pipeline was run in fast mode (B = 4), the CIs are illustrative only. Rerun with B >= 200 for publication-quality intervals. Narrow CI widths (< 2 pp) in fast mode are artefacts of insufficient resampling.
:::


# Cross-Country Summary {#sec-summary}

## Predictability ranking

```{r cross-country-ranking, fig.height=6, fig.width=10}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf) &&
    !is.null(a2_error) && nrow(a2_error) > 0) {

  # Merge individual-level AUC with admin2-level correlation
  cv_auc <- cv_perf |>
    filter(!is.na(auc)) |>
    select(country, outcome, auc)

  r_col <- intersect(c("pearson_r", "r"), colnames(a2_error))
  if (length(r_col) > 0) {
    a2_r <- a2_error |>
      select(country, outcome, admin2_r = !!r_col[1])

    rank_df <- cv_auc |>
      left_join(a2_r, by = c("country", "outcome")) |>
      mutate(outcome_label = outcome_labels[outcome])

    if (nrow(rank_df) > 0) {
      ggplot(rank_df, aes(x = auc, y = admin2_r, colour = country, shape = outcome_label)) +
        geom_point(size = 4, alpha = 0.8) +
        geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
        geom_vline(xintercept = 0.5, linetype = "dashed", colour = "grey60") +
        scale_colour_brewer(palette = "Set2", name = "Country") +
        scale_shape_manual(values = c(16, 17, 15, 18), name = "Outcome") +
        labs(
          x = "Individual-level AUC (binary CV)",
          y = "Admin-2 Pearson r",
          title = "Predictability: Individual-Level vs. Area-Level Performance",
          subtitle = "Upper-right = best performing | Lower-left = hardest to predict"
        ) +
        theme(legend.position = "bottom",
              legend.box = "vertical")
    }
  }
}
```

```{r cross-country-table}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  rank_table <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome]) |>
    select(Country = country, Outcome = outcome_label, AUC = auc) |>
    arrange(desc(AUC)) |>
    mutate(
      Rank = row_number(),
      AUC = round(AUC, 3),
      Predictability = case_when(
        AUC >= 0.70 ~ "Good",
        AUC >= 0.60 ~ "Moderate",
        AUC >= 0.55 ~ "Weak",
        TRUE ~ "Near chance"
      )
    ) |>
    select(Rank, Country, Outcome, AUC, Predictability)

  kable(rank_table,
        caption = "All country-outcome combinations ranked by binary AUC") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

### Key findings across all countries

1. **Vitamin A deficiency is consistently more predictable than iron deficiency** from area-level proxy data. This pattern holds across all four countries and likely reflects stronger geographic structuring of VAD risk factors (dietary patterns, food systems, urbanisation).

2. **Women's outcomes tend to be more predictable than children's** for both micronutrients. Women's nutritional status may integrate longer-term area-level conditions more faithfully.

3. **Iron deficiency in children** is the hardest outcome to predict across all countries. Policy decisions for this outcome should rely on direct survey data rather than modelled estimates.

4. **Cross-country transportability** (LOCO results) indicates which countries share sufficient covariate structure for borrowing strength, and which require country-specific calibration.

5. **Food security data** (WFP prices, FSEC indicators) contributes to the predictor ensemble, though no single domain dominates ablation results.


# Limitations {#sec-limitations}

## Current limitations

1. **Modest proxy-only performance.** Without survey variables, models achieve AUC of 0.50--0.70 for binary classification. This reflects the genuine difficulty of predicting individual biomarker status from area-level indicators and is an honest assessment of what proxy data can deliver.

2. **Temporal snapshots.** Each country model is trained on a single survey wave with contemporaneous predictors. Temporal extrapolation has not been tested --- predictions for future years assume stable relationships between proxy indicators and micronutrient status.

3. **Admin-2 small-sample instability.** Districts with fewer than ~20 survey observations have uncertain ground truth. SL predictions smooth over this noise but validation is limited in thin-data areas.

4. **No explicit spatial random effects.** The current SuperLearner framework does not model spatial correlation directly. BYM2 or SPDE models via INLA could improve area-level predictions, particularly in countries with strong spatial structure.

5. **Bootstrap resampling.** If run in fast mode (B = 4), confidence intervals are illustrative only. Publication-quality CIs require B >= 200.

6. **Food security coverage.** FSEC and WFP data availability varies across countries and time periods. Missing food security indicators are imputed, which may attenuate their contribution.

7. **GADM boundary matching.** Admin-1 and Admin-2 names must match exactly between GADM boundaries and survey datasets. Minor mismatches (spelling, transliteration) may cause join failures that appear as grey (missing) areas on maps.

## Strengths

1. **Multi-country framework.** The same pipeline, predictor domains, and evaluation framework are applied consistently across four countries, enabling direct comparison.

2. **Proxy-only design.** By excluding survey variables, the models address the policy-relevant question: can we estimate micronutrient burden in areas without biomarker surveys?

3. **Transportability testing.** LOCO cross-validation provides the first systematic assessment of whether proxy-based micronutrient models generalise across national boundaries in sub-Saharan Africa.

4. **Diverse predictor ensemble.** Combining DHS, MICS, GEE, IHME, MAP, WFP, and FSEC data creates a robust, multi-domain predictor set that does not depend on any single data source.

5. **Reproducible pipeline.** All results are generated from a `{targets}` pipeline with cached, dependency-tracked execution. Any result can be reproduced or updated by rerunning the relevant targets.
