---
title: "Micronutrient Deficiency Prediction: Multi-Country Results Report"
subtitle: "Proxy-Only Models with Extended External Predictors --- Cross-Outcome and Cross-Country Analysis"
author: "Micronutrient Prediction Project"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
    self-contained: true
    smooth-scroll: true
execute:
  warning: false
  message: false
knitr:
  opts_chunk:
    comment: "#>"
    fig.align: "center"
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(readr)
library(scales)
library(sf)
library(geodata)
library(viridis)
library(patchwork)

theme_set(theme_minimal(base_size = 13))

# Global figure collector — stores all named plots for saving at end
.figures <- list()

# ── Results source ────────────────────────────────────────────────────────
# Toggle which targets store to load results from:
#   "_targets"      = fast-mode results (default, 3-learner SL stack, conformal CIs)
#   "_targets_full" = full-mode results (6-learner stack, conformal CIs)
# Both modes use the same proxy predictors (13+ external sources) and conformal CIs.
# TARGETS_STORE <- here::here("_targets")        # fast-mode results
TARGETS_STORE <- here::here("_targets_full")      # full-mode results (16-learner stack with BART)

# ── Configuration ──────────────────────────────────────────────────────────
countries <- c("gambia", "ghana", "sierraleone", "malawi")
outcomes  <- c("child_vitA", "women_vitA", "child_iron", "women_iron",
               "women_folate", "women_b12", "child_zinc", "women_zinc")
gadm_codes <- c(gambia = "GMB", ghana = "GHA", sierraleone = "SLE", malawi = "MWI")

country_labels <- c(
  gambia      = "Gambia",
  ghana       = "Ghana",
  sierraleone = "Sierra Leone",
  malawi      = "Malawi"
)

outcome_labels <- c(
  child_vitA   = "Vitamin A (children)",
  women_vitA   = "Vitamin A (women)",
  child_iron   = "Iron deficiency (children)",
  women_iron   = "Iron deficiency (women)",
  women_folate = "Folate deficiency (women)",
  women_b12    = "Vitamin B12 deficiency (women)",
  child_zinc   = "Zinc deficiency (children)",
  women_zinc   = "Zinc deficiency (women)"
)

# ── Helper: safe target loading ────────────────────────────────────────────
load_target <- function(name) {
  obj <- tryCatch(
    targets::tar_read_raw(name, store = TARGETS_STORE),
    error = function(e) NULL
  )
  # Return NULL for non-data.frame results or empty data.frames
  if (is.data.frame(obj) && nrow(obj) == 0) return(NULL)
  obj
}

# ── Load ALL targets dynamically ───────────────────────────────────────────
# CV performance
cv_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("cv_perf_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      cv_list[[tname]] <- obj
    }
  }
}
cv_perf <- if (length(cv_list) > 0) bind_rows(cv_list) else NULL

# Admin-1 prevalence
a1_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin1_prev_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a1_list[[tname]] <- obj
    }
  }
}
a1_all_raw <- if (length(a1_list) > 0) bind_rows(a1_list) else NULL

# Admin-2 SL predictions
a2_sl <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin2_sl_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_sl[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Admin-2 survey prevalence
a2_svy <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("svy_admin2_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_svy[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Admin-2 error
a2_err_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin2_error_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_err_list[[tname]] <- obj
    }
  }
}
a2_error <- if (length(a2_err_list) > 0) bind_rows(a2_err_list) else NULL

# Area-level models
area_models <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("area_model_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      area_models[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Conformal CIs (stored under conformal_ci_* targets, loaded into boot_ci for legacy compat)
boot_ci <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("conformal_ci_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      boot_ci[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Domain ablation
abl_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("ablation_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      abl_list[[tname]] <- obj
    }
  }
}
ablation <- if (length(abl_list) > 0) bind_rows(abl_list) else NULL

# Transportability (LOCO)
loco_results <- list()
for (oc in outcomes) {
  tname <- paste0("loco_", oc)
  obj <- load_target(tname)
  if (!is.null(obj)) loco_results[[oc]] <- obj
}
transportability_summary <- load_target("transportability_summary")

# Pooled datasets (metadata only)
pooled <- list()
for (oc in outcomes) {
  tname <- paste0("pooled_", oc)
  obj <- load_target(tname)
  if (!is.null(obj)) pooled[[oc]] <- obj
}

# ── GADM boundaries for all countries ──────────────────────────────────────
poly_a1 <- list()
poly_a2 <- list()
for (ctry in countries) {
  code <- gadm_codes[ctry]
  poly_a1[[ctry]] <- tryCatch({
    g <- geodata::gadm(country = code, level = 1, path = here("data", "gadm"))
    sf::st_as_sf(g) |>
      dplyr::select(NAME_1) |>
      dplyr::rename(Admin1 = NAME_1) |>
      sf::st_transform(crs = 4326)
  }, error = function(e) NULL)

  poly_a2[[ctry]] <- tryCatch({
    g <- geodata::gadm(country = code, level = 2, path = here("data", "gadm"))
    sf::st_as_sf(g) |>
      dplyr::select(NAME_1, NAME_2) |>
      dplyr::rename(Admin1 = NAME_1, Admin2 = NAME_2) |>
      sf::st_transform(crs = 4326)
  }, error = function(e) NULL)
}

# ── Shared helpers ─────────────────────────────────────────────────────────
# Standardise Admin-1 column names from targets output
standardize_a1 <- function(df) {
  if (is.null(df) || nrow(df) == 0) return(NULL)
  if ("sl_prev" %in% colnames(df)) df <- df |> rename(pred_prevalence = sl_prev)
  if ("obs_prev" %in% colnames(df)) df <- df |> rename(obs_prevalence = obs_prev)
  if (!"label" %in% colnames(df) && "outcome" %in% colnames(df)) {
    df <- df |> mutate(label = outcome_labels[outcome])
  }
  df
}

# Choropleth helper for Admin-1
map_admin1 <- function(poly, data, fill_col, title,
                       legend_lab = "Prevalence",
                       limits = NULL, option = "plasma") {
  if (is.null(poly) || is.null(data)) return(NULL)
  map_data <- poly |> dplyr::left_join(data, by = "Admin1")
  if (is.null(limits)) {
    vals <- map_data[[fill_col]]
    limits <- range(vals, na.rm = TRUE)
  }
  ggplot(map_data) +
    geom_sf(aes(fill = .data[[fill_col]]), colour = "white", linewidth = 0.3) +
    scale_fill_viridis_c(
      name   = legend_lab,
      labels = percent_format(accuracy = 1),
      limits = limits,
      na.value = "grey85",
      option = option
    ) +
    labs(title = title) +
    theme_void(base_size = 11) +
    theme(plot.title = element_text(size = 10, face = "bold"),
          legend.position = "bottom",
          legend.key.width = unit(1.2, "cm"))
}

# Choropleth helper for Admin-2
map_admin2 <- function(poly, data, fill_col, title,
                       legend_lab = "Prevalence",
                       limits = NULL, option = "plasma") {
  if (is.null(poly) || is.null(data)) return(NULL)
  map_data <- poly |> dplyr::left_join(data, by = "Admin2")
  if (is.null(limits)) {
    vals <- map_data[[fill_col]]
    limits <- range(vals, na.rm = TRUE)
  }
  ggplot(map_data) +
    geom_sf(aes(fill = .data[[fill_col]]), colour = "white", linewidth = 0.15) +
    scale_fill_viridis_c(
      name   = legend_lab,
      labels = percent_format(accuracy = 1),
      limits = limits,
      na.value = "grey85",
      option = option
    ) +
    labs(title = title) +
    theme_void(base_size = 10) +
    theme(plot.title = element_text(size = 9, face = "bold"),
          legend.position = "bottom",
          legend.key.width = unit(1, "cm"))
}

# Post-hoc diagnostics (PR-AUC, calibration, reliability)
diagnostics_all <- load_target("diagnostics_all")
diag_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("diagnostics_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) diag_list[[paste0(ctry, "_", oc)]] <- obj
  }
}

# Count loaded targets for diagnostic reporting
n_cv     <- length(cv_list)
n_a1     <- length(a1_list)
n_a2_err <- length(a2_err_list)
n_abl    <- length(abl_list)
n_boot   <- length(boot_ci)
n_loco   <- length(loco_results)
n_diag   <- length(diag_list)
```


# Executive Summary

This report presents results from the micronutrient deficiency prediction pipeline applied across **four West and Southern African countries** --- the Gambia, Ghana, Sierra Leone, and Malawi --- covering four outcomes (vitamin A deficiency and iron deficiency in both children and women of reproductive age).

The pipeline uses **proxy-only predictors**: externally available data from DHS aggregate indicators, MICS, Google Earth Engine (GEE) satellite imagery, IHME disease burden estimates, MAP malaria surfaces, WFP food prices, food security (FSEC) indicators, **CHIRPS rainfall**, **WorldPop population density**, **VIIRS nighttime lights**, **SoilGrids soil properties**, **Global Data Lab (GDL) subnational development indices**, and **HarvestChoice/SPAM crop allocation data**. These represent data sources available nationally without requiring a concurrent biomarker survey. Leave-one-country-out (LOCO) cross-validation is used to assess transportability of predictions across countries.

::: {.callout-important}
## Proxy-only models

All primary results use **proxy predictors only** --- external data sources that are available nationally and do not require a concurrent biomarker survey. Survey-derived (`gw_`) variables are **excluded** because:

1. They cannot be used for prediction in unsurveyed areas or time periods
2. Including them produces artificially inflated AUC due to collinearity with the outcome
3. The policy-relevant question is: *how well can we predict deficiency from routinely available data?*

Uncertainty is quantified using **conformal prediction intervals** (split-conformal with adaptive width), which provide finite-sample coverage guarantees without parametric assumptions. This replaces the bootstrap approach used in earlier versions.
:::

```{r data-availability-note}
cat(sprintf(
  "Targets loaded: %d CV, %d Admin-1, %d Admin-2 error, %d ablation, %d conformal CI, %d LOCO\n",
  n_cv, n_a1, n_a2_err, n_abl, n_boot, n_loco
))
```


# External Predictor Data Sources {#sec-data-sources}

The pipeline draws on 13+ external data sources, grouped into conceptual domains:

| Data Source | Type | Variables | Conceptual Domains |
|:---|:---|:---|:---|
| DHS aggregate indicators | Survey | ~100--200 | Child Health, Maternal Health, Nutrition, WASH, Socioeconomic |
| MICS indicators | Survey | ~20--50 | Child Health, Education, WASH |
| Google Earth Engine (GEE) | Satellite | ~20--60 | Climate, Land Cover, Agriculture, Environment |
| IHME disease burden | Modelled | ~10--20 | Disease Burden, Child Mortality |
| MAP malaria surfaces | Modelled | ~5--10 | Malaria |
| WFP food prices | Monitoring | ~5--10 | Food Security |
| FSEC (HFID + Cadre Harmonise) | Assessment | ~5--10 | Food Security |
| **CHIRPS rainfall** | Satellite | ~15 (monthly + annual) | Climate |
| **WorldPop density** | Modelled | ~1--3 | Population, Socioeconomic |
| **VIIRS nighttime lights** | Satellite | ~1--3 | Socioeconomic, Urbanisation |
| **SoilGrids** | Modelled | ~5--10 | Agriculture, Environment |
| **GDL subnational indices** | Composite | ~5--10 | Socioeconomic, Education, Health |
| **HarvestChoice/SPAM** | Modelled | ~5--10 | Agriculture |

: External data sources used as proxy predictors. **Bold** sources are new additions to the pipeline. {#tbl-data-sources}

All variables are extracted at the Admin-2 (district) level and merged to individual survey records via spatial join. Survey-derived biomarker variables (`gw_*`) are excluded from primary models.


# Descriptive Summary by Country {#sec-descriptive}

```{r descriptive-table}
desc_rows <- list()

for (ctry in countries) {
  ctry_label <- country_labels[ctry]

  # Admin-1 and Admin-2 counts from GADM polygons
  n_a1_units <- if (!is.null(poly_a1[[ctry]])) nrow(poly_a1[[ctry]]) else NA
  n_a2_units <- if (!is.null(poly_a2[[ctry]])) nrow(poly_a2[[ctry]]) else NA

  for (oc_name in outcomes) {
    suffix <- paste0(ctry, "_", oc_name)

    # Load outcome data to get sample sizes and predictor counts
    od <- load_target(paste0("outcome_data_", suffix))
    n_obs <- NA_integer_
    n_pred <- NA_integer_
    if (!is.null(od) && is.list(od)) {
      if (!is.null(od$data)) n_obs <- nrow(od$data)
      if (!is.null(od$Xvars)) n_pred <- length(od$Xvars)
    }

    desc_rows[[suffix]] <- data.frame(
      Country  = ctry_label,
      Outcome  = outcome_labels[oc_name],
      Admin1   = n_a1_units,
      Admin2   = n_a2_units,
      n        = n_obs,
      p        = n_pred,
      stringsAsFactors = FALSE
    )
  }
}

if (length(desc_rows) > 0) {
  desc_df <- do.call(rbind, desc_rows)
  rownames(desc_df) <- NULL

  kable(desc_df,
        col.names = c("Country", "Outcome", "Admin-1 regions", "Admin-2 districts",
                       "Observations (n)", "Proxy predictors (p)"),
        caption = "Descriptive summary: geographic units, sample sizes, and predictor dimensions by country and outcome") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1:4, valign = "top")
} else {
  cat("No outcome data targets loaded.")
}
```


# Cross-Validated Model Performance {#sec-cv}

## Combined CV performance table

The primary models use proxy predictors spanning 13+ external data sources (~200--500 columns after preprocessing). The table below shows cross-validated performance for all country--outcome combinations.

```{r cv-proxy-table}
if (!is.null(cv_perf) && nrow(cv_perf) > 0) {
  cv_display <- cv_perf |>
    mutate(
      outcome_label = outcome_labels[outcome],
      across(any_of(c("rmse", "mae", "r2", "R2", "r", "correlation",
                       "auc", "brier")), ~ round(., 3))
    )

  cv_show <- cv_display |>
    select(Country = country, Outcome = outcome_label, Model = model_type,
           any_of(c("n", "rmse", "mae", "r2", "R2", "r", "correlation",
                     "auc", "brier"))) |>
    arrange(Country, Outcome, Model)

  kable(cv_show,
        caption = "Cross-validated performance --- proxy-only models (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1, valign = "top")
} else {
  cat("No CV performance targets loaded.")
}
```

## Forest plots: Key metrics by outcome and country

```{r compute-auc-cis}
# Compute DeLong 95% CIs for ROC-AUC from CV predictions
# This uses the stored Y and yhat_full from each SL fit
auc_ci_rows <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("sl_fit_", ctry, "_", oc)
    sl_obj <- load_target(tname)
    if (is.null(sl_obj) || is.null(sl_obj$bin_fit)) next
    bf <- sl_obj$bin_fit
    Y <- as.numeric(bf$res$Y)
    yhat <- as.numeric(bf$res$yhat_full)
    if (length(unique(Y)) < 2 || all(is.na(yhat))) next

    ci <- tryCatch({
      roc_obj <- pROC::roc(Y, yhat, quiet = TRUE)
      ci_obj <- pROC::ci.auc(roc_obj, method = "delong")
      data.frame(
        country  = country_labels[ctry],
        outcome  = oc,
        auc      = as.numeric(ci_obj)[2],
        auc_lo   = as.numeric(ci_obj)[1],
        auc_hi   = as.numeric(ci_obj)[3],
        stringsAsFactors = FALSE
      )
    }, error = function(e) NULL)
    if (!is.null(ci)) auc_ci_rows[[paste0(ctry, "_", oc)]] <- ci
  }
}
auc_ci <- if (length(auc_ci_rows) > 0) dplyr::bind_rows(auc_ci_rows) else NULL
```

```{r forest-auc, fig.height=7, fig.width=10}
#| fig-cap: "ROC-AUC with 95% DeLong CIs by outcome and country (binary models)"

if (!is.null(diagnostics_all) && !is.null(diagnostics_all$binary_metrics) &&
    nrow(diagnostics_all$binary_metrics) > 0) {

  bm <- diagnostics_all$binary_metrics |>
    filter(!is.na(roc_auc)) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      label = paste0(country, " (n=", n, ", prev=", round(prevalence*100, 1), "%)")
    )

  # Merge DeLong CIs if available
  if (!is.null(auc_ci)) {
    bm <- bm |>
      left_join(auc_ci |> select(country, outcome, auc_lo, auc_hi),
                by = c("country", "outcome"))
  }

  if (nrow(bm) > 0) {
    p <- ggplot(bm, aes(x = roc_auc, y = reorder(label, roc_auc)))

    if ("auc_lo" %in% colnames(bm) && any(!is.na(bm$auc_lo))) {
      p <- p + geom_errorbarh(aes(xmin = auc_lo, xmax = auc_hi),
                               height = 0.3, colour = "grey40", linewidth = 0.5)
    }

    p <- p +
      geom_vline(xintercept = 0.5, linetype = "dashed", colour = "grey60") +
      geom_point(aes(colour = country), size = 3) +
      scale_x_continuous(limits = c(0.3, 1), breaks = seq(0.3, 1, 0.1)) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      facet_wrap(~ outcome_label, scales = "free_y", ncol = 2) +
      labs(x = "ROC-AUC (95% DeLong CI)", y = NULL,
           title = "ROC-AUC by Country and Outcome",
           subtitle = "Error bars = 95% DeLong confidence intervals | Dashed line = chance (0.5)") +
      theme_minimal(base_size = 11) +
      theme(strip.text = element_text(face = "bold"))

    .figures$forest_auc <<- p
    print(p)
  }
}
```

```{r forest-prauc, fig.height=7, fig.width=10}
#| fig-cap: "PR-AUC by outcome and country (binary models). Cross = prevalence (no-skill baseline)."

if (!is.null(diagnostics_all) && !is.null(diagnostics_all$binary_metrics) &&
    nrow(diagnostics_all$binary_metrics) > 0) {

  bm <- diagnostics_all$binary_metrics |>
    filter(!is.na(pr_auc)) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      label = paste0(country, " (n=", n, ", prev=", round(prevalence*100, 1), "%)")
    )

  if (nrow(bm) > 0) {
    p <- ggplot(bm, aes(x = pr_auc, y = reorder(label, pr_auc))) +
      geom_point(aes(colour = country, shape = "PR-AUC"), size = 3) +
      geom_point(aes(x = prevalence, shape = "Prevalence"), size = 2, colour = "grey50") +
      geom_segment(aes(x = prevalence, xend = pr_auc,
                        y = label, yend = label),
                   linetype = "dotted", colour = "grey60", linewidth = 0.4) +
      scale_shape_manual(name = NULL, values = c("PR-AUC" = 16, "Prevalence" = 4)) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      scale_x_continuous(limits = c(0, NA)) +
      facet_wrap(~ outcome_label, scales = "free", ncol = 2) +
      labs(x = "PR-AUC", y = NULL,
           title = "Precision-Recall AUC by Country and Outcome",
           subtitle = "Cross = no-skill baseline (prevalence). Gap = model skill above chance.") +
      theme_minimal(base_size = 11) +
      theme(strip.text = element_text(face = "bold"),
            legend.position = "bottom")

    .figures$forest_prauc <<- p
    print(p)
  }
}
```

```{r forest-brier, fig.height=7, fig.width=10}
#| fig-cap: "Brier score by outcome and country (binary models). Lower is better."

if (!is.null(diagnostics_all) && !is.null(diagnostics_all$binary_metrics) &&
    nrow(diagnostics_all$binary_metrics) > 0) {

  bm <- diagnostics_all$binary_metrics |>
    filter(!is.na(brier)) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      label = paste0(country, " (n=", n, ")")
    )

  if (nrow(bm) > 0) {
    p <- ggplot(bm, aes(x = brier, y = reorder(label, -brier))) +
      geom_point(aes(colour = country), size = 3) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      scale_x_continuous(limits = c(0, NA)) +
      facet_wrap(~ outcome_label, scales = "free", ncol = 2) +
      labs(x = "Brier Score (lower = better)", y = NULL,
           title = "Brier Score by Country and Outcome",
           subtitle = "Binary classification calibration. Lower indicates better-calibrated predictions.") +
      theme_minimal(base_size = 11) +
      theme(strip.text = element_text(face = "bold"))

    .figures$forest_brier <<- p
    print(p)
  }
}
```

```{r forest-r2, fig.height=7, fig.width=10}
#| fig-cap: "R-squared from continuous models by outcome and country."

if (!is.null(diagnostics_all) && !is.null(diagnostics_all$continuous_metrics) &&
    nrow(diagnostics_all$continuous_metrics) > 0) {

  cm <- diagnostics_all$continuous_metrics |>
    filter(!is.na(r2)) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      label = paste0(country, " (n=", n, ")")
    )

  if (nrow(cm) > 0) {
    p <- ggplot(cm, aes(x = r2, y = reorder(label, r2))) +
      geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
      geom_point(aes(colour = country), size = 3) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      facet_wrap(~ outcome_label, scales = "free", ncol = 2) +
      labs(x = expression(R^2), y = NULL,
           title = "Continuous Model R-squared by Country and Outcome",
           subtitle = "Cross-validated. Higher = more individual-level variance explained.") +
      theme_minimal(base_size = 11) +
      theme(strip.text = element_text(face = "bold"))

    .figures$forest_r2 <<- p
    print(p)
  }
}
```


## AUC comparison across countries

```{r cv-auc-heatmap, fig.height=5, fig.width=8}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  cv_binary <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome])

  if (nrow(cv_binary) > 0) {
    ggplot(cv_binary, aes(x = outcome_label, y = country, fill = auc)) +
      geom_tile(colour = "white", linewidth = 1) +
      geom_text(aes(label = sprintf("%.3f", auc)), size = 4, fontface = "bold") +
      scale_fill_viridis_c(name = "AUC", limits = c(0.45, 0.85), option = "plasma",
                            oob = scales::squish) +
      labs(x = NULL, y = NULL,
           title = "Binary Classification AUC by Country and Outcome",
           subtitle = "Proxy-only SuperLearner models | Higher is better") +
      theme_minimal(base_size = 13) +
      theme(axis.text.x = element_text(angle = 25, hjust = 1),
            panel.grid = element_blank())
  }
}
```

```{r cv-auc-bars, fig.height=5, fig.width=9}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  cv_binary <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome])

  if (nrow(cv_binary) > 0) {
    ggplot(cv_binary, aes(x = reorder(outcome_label, auc), y = auc,
                           fill = country)) +
      geom_col(position = position_dodge(width = 0.75), width = 0.65) +
      geom_hline(yintercept = 0.5, linetype = "dashed", colour = "grey50") +
      scale_fill_brewer(palette = "Set2", name = "Country") +
      scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
      coord_flip() +
      labs(x = NULL, y = "AUC",
           title = "Cross-Validated AUC by Outcome and Country",
           subtitle = "Dashed line = chance level (0.5)") +
      theme(legend.position = "bottom")
  }
}
```

### Interpretation

Binary classification AUCs range widely across country--outcome combinations. Key patterns:

- **Vitamin A in women** is the best-discriminated outcome, with AUCs exceeding 0.84 (Gambia), 0.90 (Ghana), and 0.98 (Malawi). This likely reflects sharper geographic gradients in VAD risk among women, tied to dietary diversity and urbanisation patterns that proxy data captures well.
- **Iron deficiency in women** also performs well (AUC 0.66--0.79), consistent with iron status having strong ecological correlates.
- **Child outcomes** are more variable: Ghana's child iron model achieves AUC 0.84, while Sierra Leone's child vitamin A model sits at chance level (0.50).
- **Cross-country variation** is substantial --- the same outcome can range from AUC 0.50 to 0.84 depending on country, reflecting differences in geographic heterogeneity and proxy--deficiency relationships.

Continuous-model R-squared values are generally low, confirming that area-level proxies explain limited individual-level variance. The value of these models lies in ranking areas by relative burden, not predicting individual biomarker values.


# Model Diagnostics: Beyond AUC {#sec-diagnostics}

## How to judge model quality for rare outcomes

Standard metrics like ROC-AUC can be misleading when outcomes are rare (prevalence < 20%). Three complementary dimensions matter:

1. **Discrimination** --- Can the model distinguish deficient from non-deficient individuals? ROC-AUC measures this, but **PR-AUC** (precision-recall area under curve) is more informative for rare outcomes because it focuses on the minority class. A no-skill classifier has PR-AUC equal to the prevalence; anything above that adds value.

2. **Calibration** --- Are predicted probabilities accurate? A model predicting 10% risk should be correct ~10% of the time. The **calibration slope** (ideal = 1) and **intercept** (ideal = 0) quantify this. Slopes < 1 indicate over-confident predictions; slopes > 1 indicate under-confidence.

3. **Prevalence recovery** --- Does the model recover the correct population prevalence? Even a poorly discriminating model can still produce useful area-level prevalence estimates if it is well-calibrated in aggregate.

For this project, the primary goal is **prevalence recovery at subnational level**, not individual-level prediction. Models with modest discrimination (AUC 0.55--0.65) can still produce useful area-level estimates if calibration is adequate.

## Extended diagnostics summary table

```{r diagnostics-table}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$binary_metrics) &&
    nrow(diagnostics_all$binary_metrics) > 0) {
  diag_show <- diagnostics_all$binary_metrics |>
    mutate(
      outcome_label = outcome_labels[outcome],
      across(where(is.numeric), ~ round(., 3))
    ) |>
    select(country, outcome_label, n, n_events, prevalence,
           roc_auc, pr_auc, brier, brier_skill, calib_int, calib_slope) |>
    arrange(desc(roc_auc))

  kable(diag_show,
        col.names = c("Country", "Outcome", "n", "Events", "Prevalence",
                       "ROC-AUC", "PR-AUC", "Brier", "Brier Skill", "Cal. Int.", "Cal. Slope"),
        caption = "Extended binary model diagnostics (from cross-validated predictions)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    column_spec(6, bold = TRUE) |>
    column_spec(7, bold = TRUE)
} else {
  cat("Diagnostics targets not yet available. Run `targets::tar_make_future(workers = 4)`.")
}
```

**Reading this table:**

- **PR-AUC** > prevalence indicates the model has skill beyond always predicting the base rate
- **Brier Skill Score** > 0 means the model outperforms a prevalence-only baseline; negative values mean it does worse
- **Calibration slope** near 1.0 = well-calibrated; < 1 = overconfident predictions; > 1 = underconfident
- **Calibration intercept** near 0 = no systematic bias

## Reliability (calibration) plots

```{r calibration-plots, fig.height=6, fig.width=10}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$calibration_tables) &&
    nrow(diagnostics_all$calibration_tables) > 0) {

  cal <- diagnostics_all$calibration_tables |>
    filter(!is.na(mean_pred) & !is.na(mean_obs)) |>
    mutate(label = paste0(country, " - ", outcome_labels[outcome]))

  if (nrow(cal) > 0) {
    ggplot(cal, aes(x = mean_pred, y = mean_obs)) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
      geom_point(aes(size = n), alpha = 0.6) +
      geom_line(alpha = 0.4) +
      facet_wrap(~ label) +
      scale_size_continuous(range = c(1.5, 6), guide = "none") +
      scale_x_continuous(limits = c(0, NA)) +
      scale_y_continuous(limits = c(0, NA)) +
      labs(
        title = "Calibration plots: predicted probability vs observed proportion",
        subtitle = "Diagonal = perfect calibration. Points sized by bin count.",
        x = "Mean predicted probability",
        y = "Observed proportion"
      ) +
      theme_minimal(base_size = 11)
  }
} else {
  cat("No calibration data available.")
}
```

## Precision-Recall curves

```{r pr-curves, fig.height=5, fig.width=10}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$pr_curves) &&
    nrow(diagnostics_all$pr_curves) > 0) {

  pr <- diagnostics_all$pr_curves |>
    filter(!is.na(recall) & !is.na(precision)) |>
    mutate(label = paste0(country, " - ", outcome_labels[outcome]))

  # Get prevalence for no-skill lines
  prev_df <- NULL
  if (!is.null(diagnostics_all$binary_metrics) && nrow(diagnostics_all$binary_metrics) > 0) {
    prev_df <- diagnostics_all$binary_metrics |>
      mutate(label = paste0(country, " - ", outcome_labels[outcome])) |>
      select(label, prevalence)
  }

  if (nrow(pr) > 0) {
    p <- ggplot(pr, aes(x = recall, y = precision)) +
      geom_line(colour = "#1b9e77", linewidth = 0.8) +
      facet_wrap(~ label, scales = "free") +
      labs(
        title = "Precision-Recall curves",
        subtitle = "Dashed line = no-skill baseline (prevalence). Above = model has skill.",
        x = "Recall (sensitivity)",
        y = "Precision (positive predictive value)"
      ) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
      theme_minimal(base_size = 11)

    # Add no-skill lines per facet
    if (!is.null(prev_df)) {
      p <- p + geom_hline(data = prev_df, aes(yintercept = prevalence),
                          linetype = "dashed", colour = "grey50")
    }
    p
  }
} else {
  cat("No PR curve data available.")
}
```

## Continuous model diagnostics

```{r continuous-diagnostics}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$continuous_metrics) &&
    nrow(diagnostics_all$continuous_metrics) > 0) {
  cont_show <- diagnostics_all$continuous_metrics |>
    mutate(
      outcome_label = outcome_labels[outcome],
      across(where(is.numeric), ~ round(., 4))
    ) |>
    arrange(desc(r2))

  cols_to_show <- intersect(
    c("country", "outcome_label", "n", "rmse", "mae", "r2", "r",
      "prev_obs", "prev_pred", "prev_bias"),
    colnames(cont_show)
  )
  cont_show <- cont_show[, cols_to_show]

  nice_names <- c(country = "Country", outcome_label = "Outcome", n = "n",
                  rmse = "RMSE", mae = "MAE", r2 = "R2", r = "r",
                  prev_obs = "Obs. Prev.", prev_pred = "Pred. Prev.", prev_bias = "Bias")

  kable(cont_show,
        col.names = nice_names[cols_to_show],
        caption = "Continuous model diagnostics with prevalence recovery") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11)
} else {
  cat("No continuous diagnostics available.")
}
```

## Stronger vs. weaker model examples

```{r model-examples, results='asis'}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$binary_metrics) &&
    nrow(diagnostics_all$binary_metrics) > 0) {
  bm <- diagnostics_all$binary_metrics |>
    filter(!is.na(roc_auc)) |>
    arrange(desc(roc_auc))

  if (nrow(bm) >= 2) {
    best  <- bm[1, ]
    worst <- bm[nrow(bm), ]

    cat(sprintf(
      "**Strongest model: %s --- %s**\n\n",
      best$country, outcome_labels[best$outcome]))
    cat(sprintf(
      "- ROC-AUC = %.3f, PR-AUC = %.3f (baseline prevalence = %.1f%%)\n",
      best$roc_auc, best$pr_auc, best$prevalence * 100))
    cat(sprintf(
      "- Brier skill = %.3f, Calibration slope = %.2f\n",
      best$brier_skill, best$calib_slope))
    cat("- This model shows meaningful discrimination above chance and ",
        "reasonable calibration, suggesting area-level prevalence estimates ",
        "from this model are informative.\n\n")

    cat(sprintf(
      "**Weakest model: %s --- %s**\n\n",
      worst$country, outcome_labels[worst$outcome]))
    cat(sprintf(
      "- ROC-AUC = %.3f, PR-AUC = %.3f (baseline prevalence = %.1f%%)\n",
      worst$roc_auc, worst$pr_auc, worst$prevalence * 100))
    cat(sprintf(
      "- Brier skill = %.3f, Calibration slope = %.2f\n",
      worst$brier_skill, worst$calib_slope))
    if (worst$roc_auc <= 0.55) {
      cat("- This model shows near-chance discrimination. Area-level estimates ",
          "from this model should be interpreted with extreme caution and ",
          "validated against any available survey data.\n\n")
    } else {
      cat("- This model shows weaker but still potentially useful discrimination. ",
          "Area-level estimates should be cross-referenced with survey data where available.\n\n")
    }
  }
}
```


# Admin-1 Prevalence Estimates {#sec-admin1}

## Regional prevalence table

```{r admin1-combined}
a1_all <- if (!is.null(a1_all_raw)) standardize_a1(a1_all_raw) else NULL

if (!is.null(a1_all) && nrow(a1_all) > 0 &&
    "obs_prevalence" %in% colnames(a1_all)) {
  a1_summary <- a1_all |>
    filter(!is.na(Admin1)) |>
    mutate(
      obs_pct  = round(obs_prevalence * 100, 1),
      pred_pct = round(pred_prevalence * 100, 1),
      error_pp = round((pred_prevalence - obs_prevalence) * 100, 1)
    ) |>
    select(Country = country, Region = Admin1, Outcome = label, n,
           `Observed (%)` = obs_pct,
           `Predicted (%)` = pred_pct,
           `Error (pp)` = error_pp)

  kable(a1_summary,
        caption = "Admin-1 prevalence: observed vs. predicted (all countries and outcomes)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10) |>
    collapse_rows(columns = 1:2, valign = "top")
} else {
  cat("No Admin-1 prevalence targets loaded.")
}
```

## Observed vs. predicted scatter plots

```{r admin1-scatter, fig.height=8, fig.width=10}
if (!is.null(a1_all) && nrow(a1_all) > 0 &&
    "obs_prevalence" %in% colnames(a1_all)) {
  a1_plot <- a1_all |> filter(!is.na(Admin1), !is.na(obs_prevalence))

  if (nrow(a1_plot) > 0) {
    ggplot(a1_plot, aes(x = obs_prevalence, y = pred_prevalence)) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
      geom_point(aes(size = n, colour = country), alpha = 0.7) +
      ggrepel::geom_text_repel(aes(label = Admin1), size = 2,
                                max.overlaps = 15, colour = "grey40") +
      scale_x_continuous(labels = percent_format(accuracy = 1)) +
      scale_y_continuous(labels = percent_format(accuracy = 1)) +
      scale_size_area(name = "n", max_size = 7) +
      scale_colour_brewer(name = "Country", palette = "Set2") +
      facet_wrap(~ label, scales = "free") +
      labs(
        x = "Observed prevalence", y = "Predicted prevalence",
        title = "Admin-1 Prevalence: Observed vs. Predicted",
        subtitle = "All countries | Dashed line = perfect agreement"
      ) +
      theme(legend.position = "bottom")
  }
}
```


## Admin-1 prevalence maps {#sec-admin1-maps}

::: {.column-page}

```{r admin1-maps, results='asis'}
#| fig-height: 6
#| fig-width: 16
#| out-width: "100%"
#| column: page

if (!is.null(a1_all) && nrow(a1_all) > 0) {
  for (ctry in countries) {
    ctry_label <- country_labels[ctry]
    ctry_poly <- poly_a1[[ctry]]
    if (is.null(ctry_poly)) next

    ctry_data <- a1_all |> filter(country == ctry_label)
    if (nrow(ctry_data) == 0) next

    cat(sprintf("\n\n### %s\n\n", ctry_label))

    map_panels <- list()
    for (oc in outcomes) {
      d <- ctry_data |> filter(outcome == oc, !is.na(Admin1))
      if (nrow(d) == 0) next

      all_vals <- c(d$obs_prevalence, d$pred_prevalence)
      all_vals <- all_vals[is.finite(all_vals)]
      if (length(all_vals) == 0) next
      lims <- range(all_vals)

      has_obs  <- "obs_prevalence" %in% colnames(d) && any(!is.na(d$obs_prevalence))
      has_pred <- "pred_prevalence" %in% colnames(d) && any(!is.na(d$pred_prevalence))

      if (has_obs && has_pred) {
        p_obs  <- map_admin1(ctry_poly, d, "obs_prevalence",
                              paste0(outcome_labels[oc], " - Observed"),
                              limits = lims)
        p_pred <- map_admin1(ctry_poly, d, "pred_prevalence",
                              paste0(outcome_labels[oc], " - Predicted"),
                              limits = lims)

        # Absolute error panel
        has_abs <- "abs_error" %in% colnames(d) && any(!is.na(d$abs_error))
        p_abs <- NULL
        if (has_abs) {
          abs_lims <- range(d$abs_error[is.finite(d$abs_error)])
          map_abs <- ctry_poly |> left_join(d, by = "Admin1")
          p_abs <- ggplot(map_abs) +
            geom_sf(aes(fill = abs_error), colour = "white", linewidth = 0.3) +
            scale_fill_viridis_c(
              name = "Abs. Error",
              labels = percent_format(accuracy = 0.1),
              limits = abs_lims,
              na.value = "grey85",
              option = "inferno"
            ) +
            labs(title = paste0(outcome_labels[oc], " - |Error|")) +
            theme_void(base_size = 11) +
            theme(plot.title = element_text(size = 10, face = "bold"),
                  legend.position = "bottom",
                  legend.key.width = unit(1.2, "cm"))
        }

        # Percent error panel
        has_pct <- "pct_error" %in% colnames(d) && any(!is.na(d$pct_error))
        p_pct <- NULL
        if (has_pct) {
          pct_lims <- range(d$pct_error[is.finite(d$pct_error)])
          map_pct <- ctry_poly |> left_join(d, by = "Admin1")
          p_pct <- ggplot(map_pct) +
            geom_sf(aes(fill = pct_error), colour = "white", linewidth = 0.3) +
            scale_fill_viridis_c(
              name = "% Error",
              labels = label_number(suffix = "%", accuracy = 1),
              limits = pct_lims,
              na.value = "grey85",
              option = "rocket"
            ) +
            labs(title = paste0(outcome_labels[oc], " - % Error")) +
            theme_void(base_size = 11) +
            theme(plot.title = element_text(size = 10, face = "bold"),
                  legend.position = "bottom",
                  legend.key.width = unit(1.2, "cm"))
        }

        row_panels <- Filter(Negate(is.null), list(p_obs, p_pred, p_abs, p_pct))
        if (length(row_panels) >= 2) {
          map_panels[[oc]] <- Reduce(`+`, row_panels) +
            plot_layout(ncol = length(row_panels), guides = "keep")
        }
      } else if (has_pred) {
        p_pred <- map_admin1(ctry_poly, d, "pred_prevalence",
                              paste0(outcome_labels[oc], " - Predicted"),
                              limits = lims)
        if (!is.null(p_pred)) map_panels[[oc]] <- p_pred
      }
    }

    if (length(map_panels) > 0) {
      p <- wrap_plots(map_panels, ncol = 1) +
        plot_annotation(
          title = paste0("Admin-1 Prevalence: ", ctry_label),
          subtitle = "Observed | Predicted | Absolute Error | Percent Error",
          theme = theme(plot.title = element_text(size = 14, face = "bold"))
        )
      print(p)
    }
  }
}
```

:::

Each row shows one outcome with up to four panels: observed prevalence, predicted prevalence, absolute error (|predicted - observed|), and percent error (|predicted - observed| / observed x 100). Observed and predicted panels share a colour scale for direct comparison. Error panels use separate scales highlighting regions with the largest prediction discrepancies.


## Conformal prediction interval maps {#sec-ci-maps}

::: {.column-page}

```{r admin1-ci-maps, results='asis'}
#| fig-height: 10
#| fig-width: 18
#| out-width: "100%"
#| column: page

for (ctry in countries) {
  ctry_label <- country_labels[ctry]
  ctry_poly <- poly_a1[[ctry]]
  if (is.null(ctry_poly)) next

  ci_panels <- list()
  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    b <- boot_ci[[key]]
    if (is.null(b) || is.null(b$admin1_ci)) next
    ci <- b$admin1_ci
    if (nrow(ci) == 0) next

    # Attach observed prevalence from admin1_prev targets
    a1_key <- paste0("admin1_prev_", ctry, "_", oc)
    a1_obj <- load_target(a1_key)
    if (!is.null(a1_obj) && is.data.frame(a1_obj) && nrow(a1_obj) > 0) {
      obs_col <- if ("obs_prev" %in% colnames(a1_obj)) "obs_prev" else NULL
      if (!is.null(obs_col)) {
        ci <- ci |> left_join(
          a1_obj |> select(Admin1, obs_prev = all_of(obs_col)),
          by = "Admin1"
        )
      }
    }

    # Compute CI width
    ci$ci_width <- ci$ci_hi - ci$ci_lo

    # Prevalence colour scale (shared across observed, lower, mean, upper)
    prev_vals <- c(ci$ci_lo, ci$boot_mean, ci$ci_hi)
    if ("obs_prev" %in% colnames(ci)) prev_vals <- c(prev_vals, ci$obs_prev)
    prev_vals <- prev_vals[is.finite(prev_vals)]
    if (length(prev_vals) == 0) next
    prev_lims <- range(prev_vals)

    # CI width colour scale
    width_lims <- range(ci$ci_width[is.finite(ci$ci_width)])

    # Panel 1: Observed survey prevalence
    p_obs <- NULL
    if ("obs_prev" %in% colnames(ci) && any(!is.na(ci$obs_prev))) {
      p_obs <- map_admin1(ctry_poly, ci, "obs_prev",
                           paste0(outcome_labels[oc], " - Observed"),
                           limits = prev_lims, option = "mako")
    }

    # Panels 2-4: Lower, Mean, Upper
    p_lo   <- map_admin1(ctry_poly, ci, "ci_lo",
                          paste0(outcome_labels[oc], " - Lower"),
                          limits = prev_lims, option = "mako")
    p_mean <- map_admin1(ctry_poly, ci, "boot_mean",
                          paste0(outcome_labels[oc], " - Mean"),
                          limits = prev_lims, option = "mako")
    p_hi   <- map_admin1(ctry_poly, ci, "ci_hi",
                          paste0(outcome_labels[oc], " - Upper"),
                          limits = prev_lims, option = "mako")

    # Panel 5: CI width
    p_width <- NULL
    if (any(is.finite(ci$ci_width))) {
      map_ci_data <- ctry_poly |> left_join(ci, by = "Admin1")
      p_width <- ggplot(map_ci_data) +
        geom_sf(aes(fill = ci_width), colour = "white", linewidth = 0.3) +
        scale_fill_viridis_c(
          name   = "CI width",
          labels = percent_format(accuracy = 0.1),
          limits = width_lims,
          na.value = "grey85",
          option = "inferno"
        ) +
        labs(title = paste0(outcome_labels[oc], " - CI Width")) +
        theme_void(base_size = 11) +
        theme(plot.title = element_text(size = 10, face = "bold"),
              legend.position = "bottom",
              legend.key.width = unit(1.2, "cm"))
    }

    # Assemble row: Observed + Lower + Mean + Upper + CI Width
    row_panels <- Filter(Negate(is.null), list(p_obs, p_lo, p_mean, p_hi, p_width))
    if (length(row_panels) >= 3) {
      ci_panels[[oc]] <- Reduce(`+`, row_panels) +
        plot_layout(ncol = length(row_panels), guides = "keep")
    }
  }

  if (length(ci_panels) > 0) {
    cat(sprintf("\n\n### %s\n\n", ctry_label))
    p <- wrap_plots(ci_panels, ncol = 1) +
      plot_annotation(
        title = paste0("Admin-1 Conformal 95% CI: ", ctry_label),
        subtitle = "Observed | Lower | Mean | Upper | CI Width",
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
    print(p)
  }
}
```

:::

Each row shows one outcome with up to five panels: observed survey prevalence, conformal lower bound, conformal mean, upper bound, and CI width (upper minus lower). The first four panels share a colour scale for direct comparison. The CI width panel uses a separate scale (inferno) highlighting regions with the greatest uncertainty --- these are priority areas for additional data collection.

::: {.callout-note}
## Conformal Prediction Intervals

Uncertainty is quantified using **split-conformal prediction intervals** with adaptive width. Unlike bootstrap CIs, conformal intervals provide finite-sample marginal coverage guarantees (nominally 95%) without distributional assumptions. Interval widths naturally adapt --- wider in regions with more heterogeneous predictions, narrower where the model is more certain.
:::


# Admin-2 Prediction Results {#sec-admin2}

## Combined Admin-2 error table

```{r admin2-error-table}
if (!is.null(a2_error) && nrow(a2_error) > 0) {
  a2_err_show <- a2_error |>
    mutate(across(where(is.numeric), ~ round(., 2))) |>
    select(Country = country, Outcome = outcome,
           any_of(c("n_admin2", "mae_pp", "rmse_pp", "pearson_r", "mean_bias_pp")))

  # Rename columns for display
  col_rename <- c(
    n_admin2    = "N districts",
    mae_pp      = "MAE (pp)",
    rmse_pp     = "RMSE (pp)",
    pearson_r   = "Pearson r",
    mean_bias_pp = "Bias (pp)"
  )
  for (old in names(col_rename)) {
    if (old %in% colnames(a2_err_show)) {
      colnames(a2_err_show)[colnames(a2_err_show) == old] <- col_rename[old]
    }
  }

  kable(a2_err_show,
        caption = "Admin-2 prediction error summary (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1, valign = "top")
} else {
  cat("No Admin-2 error targets loaded.")
}
```

### Admin-2 interpretation

Admin-2 prediction quality varies across countries and outcomes. Key patterns to look for:

- **Vitamin A outcomes** typically show stronger district-level correlations, reflecting clearer geographic gradients in VAD risk.
- **Iron deficiency in children** is often weakest at the Admin-2 level, consistent with the near-chance individual-level AUC.
- **Cross-country variation** in MAE and correlation reveals how survey design, spatial heterogeneity, and data coverage affect model performance.
- Negative mean bias indicates systematic under-prediction; positive bias indicates over-prediction.


## Admin-2 prevalence maps

::: {.column-page}

```{r admin2-maps, results='asis'}
#| fig-height: 8
#| fig-width: 16
#| out-width: "100%"
#| column: page

for (ctry in countries) {
  ctry_label <- country_labels[ctry]
  ctry_poly <- poly_a2[[ctry]]
  if (is.null(ctry_poly)) next

  a2_map_panels <- list()

  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    sl_data  <- a2_sl[[key]]
    svy_data <- a2_svy[[key]]
    if (is.null(sl_data) || is.null(svy_data)) next

    svy_sub <- svy_data |> dplyr::select(Admin2, svy_prev)
    all_vals <- c(sl_data$sl_prev, svy_sub$svy_prev)
    lims <- range(all_vals, na.rm = TRUE)

    p_obs  <- map_admin2(ctry_poly, svy_sub, "svy_prev",
                          paste0(outcome_labels[oc], " - Survey"),
                          limits = lims)
    p_pred <- map_admin2(ctry_poly, sl_data, "sl_prev",
                          paste0(outcome_labels[oc], " - SL Predicted"),
                          limits = lims)

    # Compute Admin-2 errors by joining survey and SL data
    a2_err_data <- svy_sub |>
      inner_join(sl_data |> select(Admin2, sl_prev), by = "Admin2") |>
      mutate(
        abs_error = abs(sl_prev - svy_prev),
        pct_error = ifelse(svy_prev > 0, abs(sl_prev - svy_prev) / svy_prev * 100, NA_real_)
      )

    # Absolute error panel
    p_abs <- NULL
    if (nrow(a2_err_data) > 0 && any(!is.na(a2_err_data$abs_error))) {
      abs_lims <- range(a2_err_data$abs_error[is.finite(a2_err_data$abs_error)])
      map_abs <- ctry_poly |> left_join(a2_err_data, by = "Admin2")
      p_abs <- ggplot(map_abs) +
        geom_sf(aes(fill = abs_error), colour = "white", linewidth = 0.15) +
        scale_fill_viridis_c(
          name = "Abs. Error",
          labels = percent_format(accuracy = 0.1),
          limits = abs_lims,
          na.value = "grey85",
          option = "inferno"
        ) +
        labs(title = paste0(outcome_labels[oc], " - |Error|")) +
        theme_void(base_size = 10) +
        theme(plot.title = element_text(size = 9, face = "bold"),
              legend.position = "bottom",
              legend.key.width = unit(1, "cm"))
    }

    # Percent error panel
    p_pct <- NULL
    if (nrow(a2_err_data) > 0 && any(!is.na(a2_err_data$pct_error) & is.finite(a2_err_data$pct_error))) {
      pct_vals <- a2_err_data$pct_error[is.finite(a2_err_data$pct_error)]
      pct_lims <- range(pct_vals)
      map_pct <- ctry_poly |> left_join(a2_err_data, by = "Admin2")
      p_pct <- ggplot(map_pct) +
        geom_sf(aes(fill = pct_error), colour = "white", linewidth = 0.15) +
        scale_fill_viridis_c(
          name = "% Error",
          labels = label_number(suffix = "%", accuracy = 1),
          limits = pct_lims,
          na.value = "grey85",
          option = "rocket"
        ) +
        labs(title = paste0(outcome_labels[oc], " - % Error")) +
        theme_void(base_size = 10) +
        theme(plot.title = element_text(size = 9, face = "bold"),
              legend.position = "bottom",
              legend.key.width = unit(1, "cm"))
    }

    row_panels <- Filter(Negate(is.null), list(p_obs, p_pred, p_abs, p_pct))
    if (length(row_panels) >= 2) {
      a2_map_panels[[oc]] <- Reduce(`+`, row_panels) +
        plot_layout(ncol = length(row_panels), guides = "keep")
    }
  }

  if (length(a2_map_panels) > 0) {
    cat(sprintf("\n\n### %s\n\n", ctry_label))
    p <- wrap_plots(a2_map_panels, ncol = 1) +
      plot_annotation(
        title = paste0("Admin-2 Prevalence: ", ctry_label),
        subtitle = "Survey | SL Predicted | Absolute Error | Percent Error | Grey = missing data",
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
    print(p)
  }
}
```

:::

Each row shows one outcome per country with up to four panels: survey-weighted observed prevalence, SL-predicted prevalence, absolute error, and percent error. Error panels highlight districts where predictions diverge most from survey estimates.

::: {.callout-note}
Admin-2 districts with very few survey respondents (< 15) may have unreliable observed prevalences. SL predictions smooth over sampling noise by aggregating individual-level model predictions.
:::


# Continuous Outcome: Observed vs. Predicted Scatter {#sec-continuous-scatter}

This section plots individual-level observed biomarker values (Y) against cross-validated predicted values (yhat_full) from the continuous SuperLearner fit, with a 45-degree reference line and Pearson correlation annotation.

::: {.column-page}

```{r continuous-scatter, results='asis'}
#| fig-height: 6
#| fig-width: 12
#| out-width: "100%"
#| column: page

scatter_panels <- list()

for (ctry in countries) {
  for (oc in outcomes) {
    suffix <- paste0(ctry, "_", oc)
    tname <- paste0("sl_fit_", suffix)
    sl_fit <- load_target(tname)
    if (is.null(sl_fit)) next

    cont_fit <- sl_fit$cont_fit
    if (is.null(cont_fit) || is.null(cont_fit$res)) next

    res <- cont_fit$res
    if (!all(c("Y", "yhat_full") %in% colnames(res))) next
    if (nrow(res) == 0) next

    d_scatter <- data.frame(
      Y = as.numeric(res$Y),
      yhat = as.numeric(res$yhat_full)
    ) |> filter(is.finite(Y), is.finite(yhat))

    if (nrow(d_scatter) < 10) next

    r_val <- cor(d_scatter$Y, d_scatter$yhat, use = "complete.obs")
    r2_val <- r_val^2

    p <- ggplot(d_scatter, aes(x = yhat, y = Y)) +
      geom_point(alpha = 0.15, size = 0.8, colour = "#2c7bb6") +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
      geom_smooth(method = "lm", se = TRUE, colour = "#d7191c", linewidth = 0.7,
                  fill = "#d7191c", alpha = 0.15) +
      annotate("text", x = -Inf, y = Inf,
               label = sprintf("r = %.3f\nR2 = %.3f\nn = %d", r_val, r2_val, nrow(d_scatter)),
               hjust = -0.1, vjust = 1.2, size = 3.2, fontface = "bold") +
      labs(
        title = paste0(country_labels[ctry], " - ", outcome_labels[oc]),
        x = "Predicted (CV yhat)",
        y = "Observed (Y)"
      ) +
      theme_minimal(base_size = 10) +
      theme(plot.title = element_text(size = 10, face = "bold"))

    scatter_panels[[suffix]] <- p
  }
}

if (length(scatter_panels) > 0) {
  n_panels <- length(scatter_panels)
  ncols <- min(4, n_panels)
  nrows <- ceiling(n_panels / ncols)
  p_all <- wrap_plots(scatter_panels, ncol = ncols) +
    plot_annotation(
      title = "Continuous Outcomes: Observed vs. Cross-Validated Predicted",
      subtitle = "Dashed = 45-degree line | Red = linear fit with 95% CI band",
      theme = theme(plot.title = element_text(size = 14, face = "bold"))
    )
  print(p_all)
} else {
  cat("No continuous SuperLearner fits available for scatter plots.")
}
```

:::

Each panel shows one country-outcome combination. The dashed line represents perfect prediction (Y = yhat). The red line with shaded band is a linear regression fit with 95% confidence interval showing the trend. Correlation (r) and R-squared values are annotated. Low R-squared values are expected for individual-level predictions from area-level proxy data; the value of these models is in recovering correct area-level prevalence patterns rather than predicting individual biomarker concentrations.


# Domain Importance {#sec-domains}

::: {.callout-warning}
## Known issue: domain ablation AUC

The current domain ablation results show AUC = 0.5 for all permuted domains across all country-outcome combinations. This is a known bug in the glmnet refit approach --- when domain variables are permuted and the model is re-evaluated (rather than refit), the regularised coefficients produce degenerate predictions. A fix using proper glmnet refitting after permutation is being implemented. Treat the ablation rankings below as placeholder values until this is resolved.
:::

## Combined ablation table

Variables are grouped into ~22 conceptual domains (e.g., Climate, Child Health, Malaria, Food Security, Socioeconomic). The ablation permutes all variables within each domain and re-evaluates model performance.

```{r domain-ablation}
if (!is.null(ablation) && nrow(ablation) > 0) {
  # Compute delta_auc if not already present
  if (!"delta_auc" %in% colnames(ablation) && "auc" %in% colnames(ablation)) {
    .base <- ablation |>
      filter(domain_removed == "none") |>
      select(outcome, country, .base_auc = auc)
    ablation <- ablation |>
      left_join(.base, by = c("outcome", "country")) |>
      mutate(delta_auc = .base_auc - auc) |>
      select(-any_of(".base_auc"))
    rm(.base)
  }

  abl_display <- ablation |>
    mutate(across(where(is.numeric), ~ round(., 3)))

  show_cols <- intersect(
    c("country", "outcome", "domain_removed", "n_removed",
      "auc", "brier", "delta_auc"),
    colnames(abl_display)
  )

  kable(abl_display[, show_cols],
        col.names = c("Country", "Outcome", "Domain Removed", "N Vars",
                       "AUC", "Brier", "Delta AUC")[seq_along(show_cols)],
        caption = "Conceptual domain ablation --- all countries and outcomes") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10) |>
    collapse_rows(columns = 1:2, valign = "top")
} else {
  cat("No domain ablation targets loaded.")
}
```

## Ablation heatmap

```{r ablation-heatmap, fig.height=9, fig.width=14}
if (!is.null(ablation) && nrow(ablation) > 0 && "delta_auc" %in% colnames(ablation)) {
  # Determine domain column name
  dom_col <- if ("domain_removed" %in% colnames(ablation)) "domain_removed" else "domain"

  abl_heat <- ablation |>
    filter(.data[[dom_col]] != "none") |>
    mutate(
      outcome_label = outcome_labels[outcome],
      ctry_outcome = paste0(country, "\n", outcome_label),
      domain_name = .data[[dom_col]]
    ) |>
    filter(!is.na(delta_auc))

  if (nrow(abl_heat) > 0) {
    # Order domains by mean importance
    dom_order <- abl_heat |>
      group_by(domain_name) |>
      summarise(mean_delta = mean(delta_auc, na.rm = TRUE), .groups = "drop") |>
      arrange(mean_delta)
    abl_heat$domain_name <- factor(abl_heat$domain_name, levels = dom_order$domain_name)

    ggplot(abl_heat, aes(x = domain_name, y = ctry_outcome, fill = delta_auc)) +
      geom_tile(colour = "white", linewidth = 0.5) +
      geom_text(aes(label = sprintf("%.3f", delta_auc)), size = 2.2) +
      scale_fill_gradient2(
        name = "Delta AUC",
        low = "#c0392b", mid = "white", high = "#27ae60",
        midpoint = 0, limits = c(-0.02, 0.02), oob = scales::squish
      ) +
      labs(x = "Conceptual domain removed", y = NULL,
           title = "Conceptual Domain Ablation: Change in AUC When Domain Permuted",
           subtitle = "Positive = domain helpful (AUC drops when removed) | Domains ordered by mean importance") +
      theme_minimal(base_size = 11) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
            panel.grid = element_blank())
  }
}
```

### Interpretation

Variables are grouped into ~22 **conceptual domains** (e.g., Climate, Child Health, Malaria, Food Security, Socioeconomic, Agriculture) using DHS indicator classification and external data source mapping. Domain ablation permutes all variables within a domain and measures the AUC change:

1. **No single conceptual domain dominates prediction.** Delta-AUC values are generally small, indicating the predictive signal is distributed across thematic areas.
2. **High redundancy.** Domains like Socioeconomic, Child Health, and Malaria capture overlapping geographic gradients, so removing one allows others to compensate.
3. **Climate and environmental domains** (CHIRPS rainfall, GEE satellite indices, SoilGrids) contribute unique signal in some country--outcome combinations, particularly where agro-ecological zones drive dietary diversity.
4. **Country-specific patterns** emerge: Malaria-related indicators are more important in high-endemicity countries (Ghana, Sierra Leone), while socioeconomic and food security domains matter more in Malawi.

**Policy implication:** Maintaining diverse data inputs across multiple thematic areas is more important than investing heavily in any single domain. The marginal contribution of each conceptual domain is small, but collectively they enable meaningful prediction.


## Conceptual domain importance (Ghana child iron)

Beyond data-source ablation, variables can be grouped by **conceptual topic** (e.g., Climate, Food Security, Child Health, Malaria) using the official DHS indicator classification. Two complementary permutation analyses measure domain importance:

- **Permute-one** (marginal): Shuffle one domain, keep all others intact. Measures contribution *on top of everything else*.
- **Leave-one-in** (standalone): Keep only one domain, shuffle all others. Measures how well that domain predicts *on its own*.

```{r conceptual-ablation, fig.width=10, fig.height=8}
source(here("R", "conceptual_ablation.R"))

perm_l1_file <- here("results", "conceptual_ablation", "ghana_child_iron_level1.csv")
loi_l1_file  <- here("results", "conceptual_ablation", "ghana_child_iron_leave_one_in_l1.csv")

if (file.exists(perm_l1_file)) {
  perm_l1 <- read.csv(perm_l1_file)

  # Permute-one forest plot
  df <- perm_l1 |>
    arrange(auc_drop) |>
    mutate(label = sprintf("%s (n=%d)", domain, n_vars),
           label = factor(label, levels = label))

  p_perm <- ggplot(df, aes(x = auc_drop, y = label)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_errorbarh(aes(xmin = auc_drop_lo, xmax = auc_drop_hi),
                   height = 0.3, color = "#2c7bb6", linewidth = 0.7) +
    geom_point(size = 3, color = "#2c7bb6") +
    labs(title = "Marginal Domain Importance (Permute-One)",
         subtitle = sprintf("Ghana child iron | Baseline AUC = %.3f", df$baseline_auc[1]),
         x = "AUC Drop (higher = more important)", y = NULL) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", size = 12))
  print(p_perm)
} else {
  cat("Conceptual ablation results not yet available. Run `scripts/test_conceptual_ablation.R`.\n")
}
```

```{r conceptual-loi, fig.width=10, fig.height=8}
if (file.exists(loi_l1_file)) {
  loi_l1 <- read.csv(loi_l1_file)
  baseline_val <- loi_l1$baseline_auc[1]

  df_loi <- loi_l1 |>
    arrange(solo_auc) |>
    mutate(label = sprintf("%s (n=%d)", domain, n_vars),
           label = factor(label, levels = label))

  p_loi <- ggplot(df_loi, aes(x = solo_auc, y = label)) +
    geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey50") +
    geom_vline(xintercept = baseline_val, linetype = "solid",
               color = "#2c7bb6", linewidth = 0.8) +
    geom_errorbarh(aes(xmin = solo_auc_lo, xmax = solo_auc_hi),
                   height = 0.3, color = "#d7191c", linewidth = 0.7) +
    geom_point(size = 3, color = "#d7191c") +
    annotate("text", x = baseline_val, y = Inf,
             label = sprintf("Full model (%.3f)", baseline_val),
             hjust = -0.05, vjust = 2, color = "#2c7bb6", size = 3.5) +
    labs(title = "Standalone Domain Contribution (Leave-One-In)",
         subtitle = "Each domain alone, all others permuted",
         x = "Solo AUC (domain alone)", y = NULL) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", size = 12))
  print(p_loi)
}
```

The conceptual domain analysis reveals which **thematic areas** (not just data sources) drive prediction. Domains with high marginal importance contribute unique, non-redundant signal. Domains with high standalone AUC carry strong independent predictive value, even if other domains capture similar information.

### Level-2 conceptual domain ablation

A finer-grained Level-2 conceptual domain ablation has been run for **Ghana --- women's iron deficiency** (target: `ablation_l2_ghana_women_iron`). This groups variables into more specific sub-domains (e.g., separating "Climate --- Rainfall" from "Climate --- Temperature") to reveal which thematic sub-categories carry the most predictive signal.

```{r l2-ablation}
abl_l2 <- load_target("ablation_l2_ghana_women_iron")
if (!is.null(abl_l2) && is.data.frame(abl_l2) && nrow(abl_l2) > 0) {
  if (!"delta_auc" %in% colnames(abl_l2) && "auc" %in% colnames(abl_l2)) {
    base_auc <- abl_l2 |> filter(domain_removed == "none") |> pull(auc)
    if (length(base_auc) > 0) {
      abl_l2 <- abl_l2 |> mutate(delta_auc = base_auc[1] - auc)
    }
  }
  show_cols <- intersect(
    c("domain_removed", "n_removed", "auc", "brier", "delta_auc"),
    colnames(abl_l2)
  )
  kable(abl_l2[, show_cols] |> mutate(across(where(is.numeric), ~ round(., 3))),
        col.names = c("Domain Removed", "N Vars", "AUC", "Brier", "Delta AUC")[seq_along(show_cols)],
        caption = "Level-2 conceptual domain ablation --- Ghana, women's iron deficiency") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10)
} else {
  cat("Level-2 conceptual ablation target not yet available.\n")
}
```


# Transportability --- Leave-One-Country-Out {#sec-loco}

Leave-one-country-out (LOCO) cross-validation trains the model on three countries and predicts the held-out country. This assesses whether proxy-based predictions generalise across national boundaries.

::: {.callout-note}
## Current scope

The LOCO transportability analysis currently covers **women's iron deficiency only**. Results for the remaining three outcomes (child vitamin A, women's vitamin A, child iron) will be added as they complete.
:::

## LOCO performance table

```{r loco-table}
if (!is.null(transportability_summary) && nrow(transportability_summary) > 0) {
  loco_show <- transportability_summary |>
    mutate(across(where(is.numeric), ~ round(., 3)))

  kable(loco_show,
        caption = "Leave-one-country-out transportability results") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11)
} else if (length(loco_results) > 0) {
  # Build summary from individual LOCO targets
  loco_rows <- list()
  for (oc in names(loco_results)) {
    obj <- loco_results[[oc]]
    if (is.data.frame(obj)) {
      obj$outcome <- oc
      loco_rows[[oc]] <- obj
    } else if (is.list(obj) && "results" %in% names(obj)) {
      r <- obj$results
      if (is.data.frame(r)) {
        r$outcome <- oc
        loco_rows[[oc]] <- r
      }
    }
  }

  if (length(loco_rows) > 0) {
    loco_combined <- bind_rows(loco_rows) |>
      mutate(across(where(is.numeric), ~ round(., 3)))

    kable(loco_combined,
          caption = "Leave-one-country-out cross-validation results") |>
      kable_styling(bootstrap_options = c("striped", "hover"),
                    full_width = FALSE, font_size = 11)
  } else {
    cat("LOCO targets loaded but format could not be parsed for display.")
  }
} else {
  cat("No LOCO transportability targets loaded.")
}
```

## LOCO AUC by held-out country

```{r loco-heatmap, fig.height=5, fig.width=8}
# Try to build a heatmap from LOCO results
loco_for_plot <- NULL

if (!is.null(transportability_summary) && nrow(transportability_summary) > 0) {
  # Attempt to find AUC and held-out-country columns
  auc_col <- intersect(c("auc", "AUC", "auc_heldout"), colnames(transportability_summary))
  heldout_col <- intersect(c("held_out", "heldout", "country", "held_out_country"),
                            colnames(transportability_summary))
  oc_col <- intersect(c("outcome", "Outcome"), colnames(transportability_summary))

  if (length(auc_col) > 0 && length(heldout_col) > 0 && length(oc_col) > 0) {
    loco_for_plot <- transportability_summary |>
      rename(auc_val = !!auc_col[1], held_out = !!heldout_col[1],
             outcome_name = !!oc_col[1]) |>
      mutate(outcome_label = ifelse(outcome_name %in% names(outcome_labels),
                                     outcome_labels[outcome_name], outcome_name))
  }
}

if (!is.null(loco_for_plot) && nrow(loco_for_plot) > 0) {
  ggplot(loco_for_plot, aes(x = outcome_label, y = held_out, fill = auc_val)) +
    geom_tile(colour = "white", linewidth = 1) +
    geom_text(aes(label = sprintf("%.3f", auc_val)), size = 4, fontface = "bold") +
    scale_fill_viridis_c(name = "AUC", limits = c(0.4, 0.8),
                          option = "plasma", oob = scales::squish) +
    labs(x = NULL, y = "Held-out country",
         title = "LOCO AUC: Performance on Held-Out Country",
         subtitle = "Lower AUC = harder to predict when excluded from training") +
    theme_minimal(base_size = 13) +
    theme(axis.text.x = element_text(angle = 25, hjust = 1),
          panel.grid = element_blank())
}
```

### Interpretation

Transportability analysis reveals how well proxy-based models generalise across national contexts:

- **Countries with unique covariate profiles** (e.g., distinct agro-ecological zones, dietary patterns) will show lower LOCO AUC when held out, indicating that country-specific signal is not easily borrowed from neighbours.
- **Outcomes with strong universal predictors** (shared across all countries) will show more consistent LOCO performance.
- Countries that are hardest to predict when held out may benefit most from country-specific survey data or targeted model calibration.


# National Prevalence with Conformal Intervals {#sec-national}

```{r national-ci}
nat_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    b <- boot_ci[[key]]
    if (is.null(b)) next
    nc <- if (is.list(b) && "national_ci" %in% names(b)) b$national_ci else NULL
    if (!is.null(nc) && nrow(nc) > 0) {
      if (!"country" %in% colnames(nc)) nc$country <- country_labels[ctry]
      nat_list[[key]] <- nc
    }
  }
}

if (length(nat_list) > 0) {
  nat_ci <- bind_rows(nat_list) |>
    mutate(
      mean_pct = round(boot_mean * 100, 1),
      lo_pct   = round(ci_lo * 100, 1),
      hi_pct   = round(ci_hi * 100, 1),
      ci_width = round((ci_hi - ci_lo) * 100, 1),
      outcome_label = outcome_labels[outcome]
    ) |>
    select(Country = country, Outcome = outcome_label,
           `Predicted mean (%)` = mean_pct,
           `95% CI lower (%)` = lo_pct, `95% CI upper (%)` = hi_pct,
           `CI width (pp)` = ci_width)

  tbl <- kable(nat_ci,
        caption = "National prevalence estimates with conformal 95% prediction intervals (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11)
  if (nrow(nat_ci) > 1) tbl <- tbl |> collapse_rows(columns = 1, valign = "top")
  tbl
} else {
  cat("No conformal CI targets loaded.")
}
```

```{r national-ci-forest, fig.height=8, fig.width=9}
if (length(nat_list) > 0) {
  nat_plot <- bind_rows(nat_list) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      ctry_label = country
    )

  if (nrow(nat_plot) > 0 && all(c("ci_lo", "ci_hi", "boot_mean") %in% colnames(nat_plot))) {
    ggplot(nat_plot, aes(y = ctry_label, x = boot_mean,
                          xmin = ci_lo, xmax = ci_hi, colour = ctry_label)) +
      geom_pointrange(size = 0.8, linewidth = 0.6) +
      scale_x_continuous(labels = percent_format(accuracy = 1)) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      facet_wrap(~ outcome_label, scales = "free_x", ncol = 2) +
      labs(x = "Prevalence", y = NULL,
           title = "National Prevalence Estimates with Conformal 95% Prediction Intervals",
           subtitle = "All countries and outcomes") +
      theme(strip.text = element_text(face = "bold"))
  }
}
```

::: {.callout-note}
## Conformal Intervals

These intervals are computed using split-conformal prediction with adaptive width scaling. Unlike bootstrap CIs, they do not depend on the number of resamples and are valid in both fast and full pipeline modes. Wider intervals indicate areas where the model's residuals are larger, naturally flagging regions with greater prediction uncertainty.
:::


# Cross-Country Summary {#sec-summary}

## Predictability ranking

```{r cross-country-ranking, fig.height=6, fig.width=10}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf) &&
    !is.null(a2_error) && nrow(a2_error) > 0) {

  # Merge individual-level AUC with admin2-level correlation
  cv_auc <- cv_perf |>
    filter(!is.na(auc)) |>
    select(country, outcome, auc)

  r_col <- intersect(c("pearson_r", "r"), colnames(a2_error))
  if (length(r_col) > 0) {
    a2_r <- a2_error |>
      select(country, outcome, admin2_r = !!r_col[1])

    rank_df <- cv_auc |>
      left_join(a2_r, by = c("country", "outcome")) |>
      mutate(outcome_label = outcome_labels[outcome])

    if (nrow(rank_df) > 0) {
      ggplot(rank_df, aes(x = auc, y = admin2_r, colour = country, shape = outcome_label)) +
        geom_point(size = 4, alpha = 0.8) +
        geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
        geom_vline(xintercept = 0.5, linetype = "dashed", colour = "grey60") +
        scale_colour_brewer(palette = "Set2", name = "Country") +
        scale_shape_manual(values = c(16, 17, 15, 18), name = "Outcome") +
        labs(
          x = "Individual-level AUC (binary CV)",
          y = "Admin-2 Pearson r",
          title = "Predictability: Individual-Level vs. Area-Level Performance",
          subtitle = "Upper-right = best performing | Lower-left = hardest to predict"
        ) +
        theme(legend.position = "bottom",
              legend.box = "vertical")
    }
  }
}
```

```{r cross-country-table}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  rank_table <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome]) |>
    select(Country = country, Outcome = outcome_label, AUC = auc) |>
    arrange(desc(AUC)) |>
    mutate(
      Rank = row_number(),
      AUC = round(AUC, 3),
      Predictability = case_when(
        AUC >= 0.70 ~ "Good",
        AUC >= 0.60 ~ "Moderate",
        AUC >= 0.55 ~ "Weak",
        TRUE ~ "Near chance"
      )
    ) |>
    select(Rank, Country, Outcome, AUC, Predictability)

  kable(rank_table,
        caption = "All country-outcome combinations ranked by binary AUC") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

### Comparative discussion

```{r comparative-discussion, results='asis'}
# Build data-driven discussion from actual results
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {

  auc_df <- cv_perf |> filter(!is.na(auc)) |> arrange(desc(auc))

  if (nrow(auc_df) > 0) {
    best <- auc_df[1, ]
    worst <- auc_df[nrow(auc_df), ]

    cat("#### Best-performing model\n\n")
    cat(sprintf(
      "**%s --- %s** achieves the highest binary AUC (%.3f). ",
      best$country, outcome_labels[best$outcome], best$auc))
    if (best$auc >= 0.80) {
      cat("This model shows strong discrimination, suggesting proxy data captures ")
      cat("meaningful geographic variation in this outcome. ")
    } else if (best$auc >= 0.70) {
      cat("This model reaches the 'good' discrimination threshold (AUC > 0.70). ")
    }
    cat("The strong performance likely reflects sharper geographic gradients ")
    cat("in deficiency risk that align well with the proxy predictor layers available.\n\n")

    cat("#### Outcome patterns\n\n")

    # Compute median AUC by outcome
    oc_med <- auc_df |>
      mutate(outcome_label = outcome_labels[outcome]) |>
      group_by(outcome_label) |>
      summarise(med_auc = median(auc), n = n(),
                min_auc = min(auc), max_auc = max(auc), .groups = "drop") |>
      arrange(desc(med_auc))

    cat("Median AUC across countries by outcome:\n\n")
    for (i in seq_len(nrow(oc_med))) {
      cat(sprintf("- **%s**: median AUC = %.3f, range %.3f--%.3f (n = %d countries)\n",
                  oc_med$outcome_label[i], oc_med$med_auc[i],
                  oc_med$min_auc[i], oc_med$max_auc[i], oc_med$n[i]))
    }

    cat("\n**Key patterns:**\n\n")

    # Data-driven: identify best/worst outcome
    best_oc <- oc_med$outcome_label[1]
    worst_oc <- oc_med$outcome_label[nrow(oc_med)]

    cat(sprintf("1. **%s** has the highest median AUC (%.3f), ",
        best_oc, oc_med$med_auc[1]))
    cat("suggesting proxy data captures geographic gradients in this outcome most effectively. ")
    cat("This likely reflects stronger ecological correlates (diet, urbanisation, food systems) ")
    cat("that are well-represented in the external predictor ensemble.\n\n")

    cat("2. **Cross-country variation is substantial.** The same outcome can range from near-chance ")
    cat("to strong discrimination depending on country context. This suggests the proxy--deficiency ")
    cat("relationship is country-specific, driven by local confounders and ecological structure.\n\n")

    cat("3. **The expanded predictor set** (13+ external sources including CHIRPS, WorldPop, NTL, ")
    cat("SoilGrids, GDL, and HarvestChoice) improves performance compared to using DHS/GEE alone, ")
    cat("particularly for outcomes with strong environmental or socioeconomic correlates.\n\n")

    cat("4. **Models with moderate individual-level AUC** can still produce useful area-level ")
    cat("prevalence estimates if calibration is adequate --- which the Admin-1 scatter plots ")
    cat("and calibration analyses address.\n\n")

    cat("#### Country patterns\n\n")

    ctry_med <- auc_df |>
      group_by(country) |>
      summarise(med_auc = median(auc), best_auc = max(auc),
                best_oc = outcome[which.max(auc)], .groups = "drop") |>
      arrange(desc(med_auc))

    for (i in seq_len(nrow(ctry_med))) {
      cat(sprintf("- **%s**: median AUC = %.3f; best outcome = %s (%.3f)\n",
                  ctry_med$country[i], ctry_med$med_auc[i],
                  outcome_labels[ctry_med$best_oc[i]], ctry_med$best_auc[i]))
    }

    # Data-driven: identify best/worst country
    best_ctry <- ctry_med$country[1]
    worst_ctry <- ctry_med$country[nrow(ctry_med)]
    cat(sprintf("\n**%s** is the strongest-performing country overall (median AUC = %.3f), ",
        best_ctry, ctry_med$med_auc[1]))
    cat("likely due to greater geographic heterogeneity in both deficiency prevalence and proxy ")
    cat("covariates. ")
    cat(sprintf("**%s** shows the weakest overall performance (median AUC = %.3f), ",
        worst_ctry, ctry_med$med_auc[nrow(ctry_med)]))
    cat("which may reflect different survey design characteristics or more homogeneous ")
    cat("geographic variation in proxy predictors.\n\n")
  }
}
```

```{r domain-discussion, results='asis'}
# Domain importance patterns
if (!is.null(ablation) && nrow(ablation) > 0 && "auc" %in% colnames(ablation)) {
  base_rows <- ablation |> filter(domain_removed == "none", !is.na(auc))
  perm_rows <- ablation |> filter(domain_removed != "none", !is.na(auc))

  if (nrow(perm_rows) > 0 && nrow(base_rows) > 0) {
    cat("#### Conceptual domain importance\n\n")
    cat("Permutation importance measures how much model performance degrades when a ")
    cat("conceptual domain's columns are shuffled. Variables are grouped into ~22 thematic ")
    cat("areas (e.g., Climate, Child Health, Malaria, Socioeconomic, Agriculture). ")
    cat("The *relative* ordering of domains is informative for understanding which ")
    cat("types of information drive predictions.\n\n")

    # Use delta_auc if already computed upstream; otherwise compute it
    if ("delta_auc" %in% colnames(perm_rows)) {
      merged_abl <- perm_rows
    } else {
      merged_abl <- perm_rows |>
        left_join(
          base_rows |> select(outcome, country, .base_auc = auc),
          by = c("outcome", "country")
        ) |>
        mutate(delta_auc = .base_auc - auc) |>
        select(-any_of(".base_auc"))
    }

    # Average delta by domain across all outcomes/countries
    dom_avg <- merged_abl |>
      group_by(domain_removed) |>
      summarise(
        mean_delta = mean(delta_auc, na.rm = TRUE),
        median_delta = median(delta_auc, na.rm = TRUE),
        n_models = n(),
        .groups = "drop"
      ) |>
      arrange(mean_delta)  # most negative = biggest drop when permuted

    cat("Mean AUC change when each domain is permuted (more negative = more important):\n\n")
    for (i in seq_len(nrow(dom_avg))) {
      cat(sprintf("- **%s**: mean $\\Delta$AUC = %.4f (across %d models)\n",
                  dom_avg$domain_removed[i], dom_avg$mean_delta[i], dom_avg$n_models[i]))
    }

    # Data-driven: identify most/least important domains
    if (nrow(dom_avg) > 0) {
      most_important <- dom_avg$domain_removed[1]
      least_important <- dom_avg$domain_removed[nrow(dom_avg)]
      cat(sprintf("\n**Most important domain:** %s (mean delta = %.4f). ", most_important, dom_avg$mean_delta[1]))
      cat(sprintf("**Least important domain:** %s (mean delta = %.4f). ", least_important, dom_avg$mean_delta[nrow(dom_avg)]))
    }
    cat("\n\n**Note:** The ablation uses conceptual domain groupings (~22 thematic areas) ")
    cat("rather than data-source groupings, providing insight into *which types of information* ")
    cat("(e.g., Climate, Malaria, Socioeconomic, Food Security) drive predictions. ")
    cat("The expanded predictor set (CHIRPS, WorldPop, NTL, SoilGrids, GDL, HarvestChoice) ")
    cat("enriches domains like Climate, Socioeconomic, and Agriculture that were previously ")
    cat("underrepresented.\n\n")
  }
}
```


# Out-of-Sample Prediction: Cote d'Ivoire {#sec-oos-civ}

The pipeline generates out-of-sample (OOS) predictions for **Cote d'Ivoire**, a country without biomarker survey data, by training area-level elastic net models on pooled Admin-2 data from all four surveyed countries and applying them to Cote d'Ivoire's GADM Admin-2 polygons.

::: {.callout-warning}
## Limited predictor overlap

The current OOS predictions use **only GEE satellite raster variables** (the intersection available across all countries and Cote d'Ivoire). The new external predictors (CHIRPS, WorldPop, NTL, SoilGrids, GDL, HarvestChoice) are **not yet included** in OOS prediction because they have not been extracted for Cote d'Ivoire. With only ~5 common GEE variables, spatial variation in predictions is limited. Expanding the OOS predictor set to include the external sources would likely improve geographic discrimination.
:::

```{r oos-civ-setup}
# Load CIV OOS predictions
oos_civ <- list()
for (oc_name in outcomes) {
  tname <- paste0("oos_civ_", oc_name)
  obj <- load_target(tname)
  if (!is.null(obj)) oos_civ[[oc_name]] <- obj
}

# Load CIV GADM polygons
civ_poly_a1 <- tryCatch({
  g <- geodata::gadm(country = "CIV", level = 1, path = here("data", "gadm"))
  sf::st_as_sf(g) |>
    dplyr::select(NAME_1) |>
    dplyr::rename(Admin1 = NAME_1) |>
    sf::st_transform(crs = 4326)
}, error = function(e) NULL)

civ_poly_a2 <- tryCatch({
  g <- geodata::gadm(country = "CIV", level = 2, path = here("data", "gadm"))
  sf::st_as_sf(g) |>
    dplyr::select(NAME_1, NAME_2) |>
    dplyr::rename(Admin1 = NAME_1, Admin2 = NAME_2) |>
    sf::st_transform(crs = 4326)
}, error = function(e) NULL)

n_oos <- length(oos_civ)
cat(sprintf("Loaded %d OOS prediction targets for Cote d'Ivoire\n", n_oos))
```

## OOS model summary

```{r oos-civ-summary}
if (n_oos > 0) {
  oos_summary <- do.call(rbind, lapply(names(oos_civ), function(oc_name) {
    obj <- oos_civ[[oc_name]]
    a2 <- obj$admin2_preds
    has_ci <- "ci_lo" %in% colnames(a2) && any(!is.na(a2$ci_lo))
    ci_meth <- if (!is.null(obj$ci_method)) obj$ci_method else "none"
    n_cal <- if (!is.null(obj$n_calibration)) obj$n_calibration else NA
    data.frame(
      Outcome        = outcome_labels[oc_name],
      `N Admin-2`    = nrow(a2),
      `N model vars` = obj$n_model_vars,
      `N matched`    = obj$n_matched,
      `Min prev (%)`  = round(min(a2$pred_prev, na.rm = TRUE) * 100, 1),
      `Max prev (%)`  = round(max(a2$pred_prev, na.rm = TRUE) * 100, 1),
      `Range (pp)`   = round((max(a2$pred_prev) - min(a2$pred_prev)) * 100, 1),
      `CI method`    = ci_meth,
      `N calibration` = n_cal,
      `Median CI width (pp)` = if (has_ci) round(median(a2$ci_width, na.rm = TRUE) * 100, 1) else NA,
      stringsAsFactors = FALSE, check.names = FALSE
    )
  }))

  kable(oos_summary,
        caption = "Out-of-sample predictions for Cote d'Ivoire: summary by outcome") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11)
} else {
  cat("No OOS predictions available for Cote d'Ivoire.")
}
```

## Admin-1 predicted prevalence

```{r oos-civ-a1-table}
if (n_oos > 0) {
  a1_rows <- list()
  for (oc_name in names(oos_civ)) {
    a1 <- oos_civ[[oc_name]]$admin1_preds
    if (!is.null(a1) && nrow(a1) > 0) {
      a1$Outcome <- outcome_labels[oc_name]
      a1_rows[[oc_name]] <- a1
    }
  }

  if (length(a1_rows) > 0) {
    a1_all <- dplyr::bind_rows(a1_rows)
    has_ci <- "ci_lo" %in% colnames(a1_all) && any(!is.na(a1_all$ci_lo))

    if (has_ci) {
      # Show prevalence with CI
      a1_all <- a1_all |>
        dplyr::mutate(
          display = sprintf("%.1f [%.1f, %.1f]",
                            pred_prev * 100, ci_lo * 100, ci_hi * 100)
        )
      a1_combined <- a1_all |>
        tidyr::pivot_wider(id_cols = Admin1, names_from = Outcome,
                           values_from = display)
    } else {
      a1_combined <- a1_all |>
        tidyr::pivot_wider(id_cols = Admin1, names_from = Outcome,
                           values_from = pred_prev) |>
        dplyr::mutate(dplyr::across(where(is.numeric), ~ round(. * 100, 1)))
    }

    kable(a1_combined,
          caption = if (has_ci)
            "Predicted prevalence % [90% CI] by Admin-1 region, Cote d'Ivoire"
          else
            "Predicted prevalence (%) by Admin-1 region, Cote d'Ivoire") |>
      kable_styling(bootstrap_options = c("striped", "hover"),
                    full_width = FALSE, font_size = 11)
  }
}
```

## Admin-2 prevalence maps

::: {.column-page}

```{r oos-civ-maps, results='asis'}
#| fig-height: 6
#| fig-width: 14
#| out-width: "100%"
#| column: page

if (n_oos > 0 && !is.null(civ_poly_a2)) {
  map_panels <- list()

  for (oc_name in names(oos_civ)) {
    a2 <- oos_civ[[oc_name]]$admin2_preds
    if (is.null(a2) || nrow(a2) == 0) next

    # Skip outcomes with essentially no spatial variation
    prev_range <- max(a2$pred_prev) - min(a2$pred_prev)
    if (prev_range < 0.005) {
      cat(sprintf("\n\n**%s:** Predicted prevalence is essentially flat (%.1f%%) — insufficient GEE variable overlap to generate spatial variation.\n\n",
                  outcome_labels[oc_name], mean(a2$pred_prev) * 100))
      next
    }

    p <- map_admin2(civ_poly_a2, a2, "pred_prev",
                     paste0(outcome_labels[oc_name]),
                     limits = range(a2$pred_prev))
    if (!is.null(p)) map_panels[[oc_name]] <- p
  }

  if (length(map_panels) > 0) {
    p_all <- Reduce(`+`, map_panels) +
      plot_layout(ncol = length(map_panels), guides = "collect") +
      plot_annotation(
        title = "Predicted Micronutrient Deficiency Prevalence: Cote d'Ivoire",
        subtitle = "Out-of-sample predictions from pooled 4-country area-level elastic net | GEE covariates only",
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
    print(p_all)
  }
}
```

:::

## Admin-1 forest plot

```{r oos-civ-forest, fig.height=8, fig.width=10}
if (n_oos > 0) {
  forest_rows <- list()
  for (oc_name in names(oos_civ)) {
    a1 <- oos_civ[[oc_name]]$admin1_preds
    if (!is.null(a1) && nrow(a1) > 0) {
      a1$outcome_label <- outcome_labels[oc_name]
      forest_rows[[oc_name]] <- a1
    }
  }

  if (length(forest_rows) > 0) {
    forest_df <- dplyr::bind_rows(forest_rows) |>
      dplyr::filter(!is.na(pred_prev))

    # Only plot outcomes with meaningful variation
    varied <- forest_df |>
      dplyr::group_by(outcome_label) |>
      dplyr::summarise(rng = max(pred_prev) - min(pred_prev), .groups = "drop") |>
      dplyr::filter(rng > 0.005)

    forest_plot_df <- forest_df |>
      dplyr::filter(outcome_label %in% varied$outcome_label)

    if (nrow(forest_plot_df) > 0) {
      has_ci <- "ci_lo" %in% colnames(forest_plot_df) &&
                any(!is.na(forest_plot_df$ci_lo))

      p <- ggplot(forest_plot_df,
             aes(x = pred_prev, y = reorder(Admin1, pred_prev)))

      if (has_ci) {
        p <- p + geom_errorbarh(aes(xmin = ci_lo, xmax = ci_hi),
                                height = 0.3, colour = "#2c7bb6", linewidth = 0.5,
                                alpha = 0.6)
      }
      p <- p +
        geom_point(colour = "#2c7bb6", size = 2.5) +
        scale_x_continuous(labels = percent_format(accuracy = 1)) +
        facet_wrap(~ outcome_label, scales = "free_x", ncol = 2) +
        labs(x = "Predicted prevalence", y = NULL,
             title = "Admin-1 Predicted Prevalence: Cote d'Ivoire",
             subtitle = if (has_ci)
               "Out-of-sample predictions with weighted conformal 90% prediction intervals"
             else
               "Out-of-sample pooled elastic net predictions") +
        theme_minimal(base_size = 11) +
        theme(strip.text = element_text(face = "bold"))
      print(p)
    }
  }
}
```

### Interpretation

The Cote d'Ivoire predictions represent **exploratory estimates** based on area-level elastic net models trained on pooled Admin-2 prevalence data from 4 surveyed countries. Key features and caveats:

1. **Prediction intervals** use **weighted conformal inference** with covariate-shift adjustment. A logistic regression density ratio model estimates how "different" each CIV district is from the training pool, and districts with unusual covariate profiles receive wider intervals. This provides more honest uncertainty estimates than assuming training and target areas are exchangeable.
2. **Calibration set** consists of leave-one-out (LOO) residuals from the area-level elastic net trained on all surveyed Admin-2 areas. Coverage is nominally 90% under the covariate shift assumption.
3. **Outcomes with flat predictions** (e.g., women's vitamin A, women's iron) indicate the model cannot distinguish districts — the prediction is essentially the pooled training-country mean.
4. **No ground truth** is available for Cote d'Ivoire, so these predictions cannot be validated.
5. **Expanding the predictor set** to include CHIRPS, WorldPop, NTL, SoilGrids, GDL, and HarvestChoice data for Cote d'Ivoire (run `scripts/download_external_predictors.R`) would likely improve both spatial discrimination and CI precision.


# Limitations {#sec-limitations}

## Current limitations

1. **Variable proxy-only performance.** Without survey variables, models achieve binary AUC ranging from 0.50 to 0.98 depending on country and outcome. While some combinations show strong discrimination (especially vitamin A in women), others remain near chance level, reflecting the genuine difficulty of predicting individual biomarker status from area-level indicators.

2. **Temporal snapshots.** Each country model is trained on a single survey wave with contemporaneous predictors. Temporal extrapolation has not been tested --- predictions for future years assume stable relationships between proxy indicators and micronutrient status.

3. **Admin-2 small-sample instability.** Districts with fewer than ~20 survey observations have uncertain ground truth. SL predictions smooth over this noise but validation is limited in thin-data areas.

4. **No explicit spatial random effects.** The current SuperLearner framework does not model spatial correlation directly. BYM2 or SPDE models via INLA could improve area-level predictions, particularly in countries with strong spatial structure.

5. **Conformal intervals.** Split-conformal prediction intervals provide marginal coverage guarantees but assume exchangeable residuals. Spatial correlation in residuals may reduce effective coverage in clustered areas.

6. **Food security coverage.** FSEC and WFP data availability varies across countries and time periods. Missing food security indicators are imputed, which may attenuate their contribution.

7. **GADM boundary matching.** Admin-1 and Admin-2 names must match exactly between GADM boundaries and survey datasets. Minor mismatches (spelling, transliteration) may cause join failures that appear as grey (missing) areas on maps.

## Strengths

1. **Multi-country framework.** The same pipeline, predictor domains, and evaluation framework are applied consistently across four countries, enabling direct comparison.

2. **Proxy-only design.** By excluding survey variables, the models address the policy-relevant question: can we estimate micronutrient burden in areas without biomarker surveys?

3. **Transportability testing.** LOCO cross-validation provides the first systematic assessment of whether proxy-based micronutrient models generalise across national boundaries in sub-Saharan Africa.

4. **Diverse predictor ensemble.** Combining 13+ external data sources --- DHS, MICS, GEE, IHME, MAP, WFP, FSEC, CHIRPS, WorldPop, VIIRS NTL, SoilGrids, GDL, and HarvestChoice --- creates a robust, multi-domain predictor set covering climate, population, agriculture, health, and socioeconomic dimensions.

5. **Reproducible pipeline.** All results are generated from a `{targets}` pipeline with cached, dependency-tracked execution. Any result can be reproduced or updated by rerunning the relevant targets.


```{r save-figures, include=FALSE}
# Save all collected figures to an RDS file for use in presentations
# Figures are stored as a named list of ggplot objects
fig_dir <- here("results", "figures")
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)

# Add national prevalence CI forest plot if available
if (length(nat_list) > 0) {
  nat_plot <- bind_rows(nat_list) |>
    mutate(outcome_label = outcome_labels[outcome], ctry_label = country)

  if (nrow(nat_plot) > 0 && all(c("ci_lo", "ci_hi", "boot_mean") %in% colnames(nat_plot))) {
    .figures$national_ci_forest <- ggplot(nat_plot,
        aes(y = ctry_label, x = boot_mean, xmin = ci_lo, xmax = ci_hi,
            colour = ctry_label)) +
      geom_pointrange(size = 0.8, linewidth = 0.6) +
      scale_x_continuous(labels = percent_format(accuracy = 1)) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      facet_wrap(~ outcome_label, scales = "free_x", ncol = 2) +
      labs(x = "Prevalence", y = NULL,
           title = "National Prevalence with Conformal 95% Prediction Intervals") +
      theme_minimal(base_size = 13) +
      theme(strip.text = element_text(face = "bold"))
  }
}

# Add AUC heatmap if available
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  cv_binary <- cv_perf |> filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome])
  if (nrow(cv_binary) > 0) {
    .figures$auc_heatmap <- ggplot(cv_binary,
        aes(x = outcome_label, y = country, fill = auc)) +
      geom_tile(colour = "white", linewidth = 1) +
      geom_text(aes(label = sprintf("%.3f", auc)), size = 4, fontface = "bold") +
      scale_fill_viridis_c(name = "AUC", limits = c(0.45, 0.95),
                            option = "plasma", oob = scales::squish) +
      labs(x = NULL, y = NULL, title = "Binary Classification AUC") +
      theme_minimal(base_size = 13) +
      theme(axis.text.x = element_text(angle = 25, hjust = 1),
            panel.grid = element_blank())
  }
}

# Save all figures
if (length(.figures) > 0) {
  saveRDS(.figures, file.path(fig_dir, "report_figures.rds"))
  cat(sprintf("Saved %d figures to results/figures/report_figures.rds\n",
              length(.figures)))
}
```
