---
title: "Micronutrient Deficiency Prediction: Multi-Country Results Report"
subtitle: "Proxy-Only Models --- Cross-Outcome and Cross-Country Analysis"
author: "Micronutrient Prediction Project"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
    self-contained: true
    smooth-scroll: true
execute:
  warning: false
  message: false
knitr:
  opts_chunk:
    comment: "#>"
    fig.align: "center"
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(readr)
library(scales)
library(sf)
library(geodata)
library(viridis)
library(patchwork)

theme_set(theme_minimal(base_size = 13))

# ── Results source ────────────────────────────────────────────────────────
# Toggle which targets store to load results from:
#   "_targets"      = fast-mode results (default)
#   "_targets_full" = full-mode results (B≥200, full learner stack)
TARGETS_STORE <- here::here("_targets")
# TARGETS_STORE <- here::here("_targets_full")  # uncomment for full results

# ── Configuration ──────────────────────────────────────────────────────────
countries <- c("gambia", "ghana", "sierraleone", "malawi")
outcomes  <- c("child_vitA", "women_vitA", "child_iron", "women_iron")
gadm_codes <- c(gambia = "GMB", ghana = "GHA", sierraleone = "SLE", malawi = "MWI")

country_labels <- c(
  gambia      = "Gambia",
  ghana       = "Ghana",
  sierraleone = "Sierra Leone",
  malawi      = "Malawi"
)

outcome_labels <- c(
  child_vitA = "Vitamin A (children)",
  women_vitA = "Vitamin A (women)",
  child_iron = "Iron deficiency (children)",
  women_iron = "Iron deficiency (women)"
)

# ── Helper: safe target loading ────────────────────────────────────────────
load_target <- function(name) {
  obj <- tryCatch(
    targets::tar_read_raw(name, store = TARGETS_STORE),
    error = function(e) NULL
  )
  # Return NULL for non-data.frame results or empty data.frames
  if (is.data.frame(obj) && nrow(obj) == 0) return(NULL)
  obj
}

# ── Load ALL targets dynamically ───────────────────────────────────────────
# CV performance
cv_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("cv_perf_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      cv_list[[tname]] <- obj
    }
  }
}
cv_perf <- if (length(cv_list) > 0) bind_rows(cv_list) else NULL

# Admin-1 prevalence
a1_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin1_prev_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a1_list[[tname]] <- obj
    }
  }
}
a1_all_raw <- if (length(a1_list) > 0) bind_rows(a1_list) else NULL

# Admin-2 SL predictions
a2_sl <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin2_sl_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_sl[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Admin-2 survey prevalence
a2_svy <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("svy_admin2_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_svy[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Admin-2 error
a2_err_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("admin2_error_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      a2_err_list[[tname]] <- obj
    }
  }
}
a2_error <- if (length(a2_err_list) > 0) bind_rows(a2_err_list) else NULL

# Area-level models
area_models <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("area_model_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      area_models[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Bootstrap CIs
boot_ci <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("conformal_ci_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      boot_ci[[paste0(ctry, "_", oc)]] <- obj
    }
  }
}

# Domain ablation
abl_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("ablation_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) {
      if (is.data.frame(obj) && nrow(obj) > 0 && !"country" %in% colnames(obj))
        obj$country <- country_labels[ctry]
      abl_list[[tname]] <- obj
    }
  }
}
ablation <- if (length(abl_list) > 0) bind_rows(abl_list) else NULL

# Transportability (LOCO)
loco_results <- list()
for (oc in outcomes) {
  tname <- paste0("loco_", oc)
  obj <- load_target(tname)
  if (!is.null(obj)) loco_results[[oc]] <- obj
}
transportability_summary <- load_target("transportability_summary")

# Pooled datasets (metadata only)
pooled <- list()
for (oc in outcomes) {
  tname <- paste0("pooled_", oc)
  obj <- load_target(tname)
  if (!is.null(obj)) pooled[[oc]] <- obj
}

# ── GADM boundaries for all countries ──────────────────────────────────────
poly_a1 <- list()
poly_a2 <- list()
for (ctry in countries) {
  code <- gadm_codes[ctry]
  poly_a1[[ctry]] <- tryCatch({
    g <- geodata::gadm(country = code, level = 1, path = here("data", "gadm"))
    sf::st_as_sf(g) |>
      dplyr::select(NAME_1) |>
      dplyr::rename(Admin1 = NAME_1) |>
      sf::st_transform(crs = 4326)
  }, error = function(e) NULL)

  poly_a2[[ctry]] <- tryCatch({
    g <- geodata::gadm(country = code, level = 2, path = here("data", "gadm"))
    sf::st_as_sf(g) |>
      dplyr::select(NAME_1, NAME_2) |>
      dplyr::rename(Admin1 = NAME_1, Admin2 = NAME_2) |>
      sf::st_transform(crs = 4326)
  }, error = function(e) NULL)
}

# ── Shared helpers ─────────────────────────────────────────────────────────
# Standardise Admin-1 column names from targets output
standardize_a1 <- function(df) {
  if (is.null(df) || nrow(df) == 0) return(NULL)
  if ("sl_prev" %in% colnames(df)) df <- df |> rename(pred_prevalence = sl_prev)
  if ("obs_prev" %in% colnames(df)) df <- df |> rename(obs_prevalence = obs_prev)
  if (!"label" %in% colnames(df) && "outcome" %in% colnames(df)) {
    df <- df |> mutate(label = outcome_labels[outcome])
  }
  df
}

# Choropleth helper for Admin-1
map_admin1 <- function(poly, data, fill_col, title,
                       legend_lab = "Prevalence",
                       limits = NULL, option = "plasma") {
  if (is.null(poly) || is.null(data)) return(NULL)
  map_data <- poly |> dplyr::left_join(data, by = "Admin1")
  if (is.null(limits)) {
    vals <- map_data[[fill_col]]
    limits <- range(vals, na.rm = TRUE)
  }
  ggplot(map_data) +
    geom_sf(aes(fill = .data[[fill_col]]), colour = "white", linewidth = 0.3) +
    scale_fill_viridis_c(
      name   = legend_lab,
      labels = percent_format(accuracy = 1),
      limits = limits,
      na.value = "grey85",
      option = option
    ) +
    labs(title = title) +
    theme_void(base_size = 11) +
    theme(plot.title = element_text(size = 10, face = "bold"),
          legend.position = "bottom",
          legend.key.width = unit(1.2, "cm"))
}

# Choropleth helper for Admin-2
map_admin2 <- function(poly, data, fill_col, title,
                       legend_lab = "Prevalence",
                       limits = NULL, option = "plasma") {
  if (is.null(poly) || is.null(data)) return(NULL)
  map_data <- poly |> dplyr::left_join(data, by = "Admin2")
  if (is.null(limits)) {
    vals <- map_data[[fill_col]]
    limits <- range(vals, na.rm = TRUE)
  }
  ggplot(map_data) +
    geom_sf(aes(fill = .data[[fill_col]]), colour = "white", linewidth = 0.15) +
    scale_fill_viridis_c(
      name   = legend_lab,
      labels = percent_format(accuracy = 1),
      limits = limits,
      na.value = "grey85",
      option = option
    ) +
    labs(title = title) +
    theme_void(base_size = 10) +
    theme(plot.title = element_text(size = 9, face = "bold"),
          legend.position = "bottom",
          legend.key.width = unit(1, "cm"))
}

# Post-hoc diagnostics (PR-AUC, calibration, reliability)
diagnostics_all <- load_target("diagnostics_all")
diag_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    tname <- paste0("diagnostics_", ctry, "_", oc)
    obj <- load_target(tname)
    if (!is.null(obj)) diag_list[[paste0(ctry, "_", oc)]] <- obj
  }
}

# Count loaded targets for diagnostic reporting
n_cv     <- length(cv_list)
n_a1     <- length(a1_list)
n_a2_err <- length(a2_err_list)
n_abl    <- length(abl_list)
n_boot   <- length(boot_ci)
n_loco   <- length(loco_results)
n_diag   <- length(diag_list)
```


# Executive Summary

This report presents results from the micronutrient deficiency prediction pipeline applied across **four West and Southern African countries** --- the Gambia, Ghana, Sierra Leone, and Malawi --- covering four outcomes (vitamin A deficiency and iron deficiency in both children and women of reproductive age).

The pipeline uses **proxy-only predictors**: externally available data from DHS aggregate indicators, MICS, Google Earth Engine (GEE) satellite imagery, IHME disease burden estimates, MAP malaria surfaces, WFP food prices, and food security (FSEC) indicators. These represent data sources available nationally without requiring a concurrent biomarker survey. Leave-one-country-out (LOCO) cross-validation is used to assess transportability of predictions across countries.

::: {.callout-important}
## Proxy-only models

All primary results use **proxy predictors only** --- external data sources that are available nationally and do not require a concurrent biomarker survey. Survey-derived (`gw_`) variables are **excluded** because:

1. They cannot be used for prediction in unsurveyed areas or time periods
2. Including them produces artificially inflated AUC due to collinearity with the outcome
3. The policy-relevant question is: *how well can we predict deficiency from routinely available data?*
:::

```{r data-availability-note}
cat(sprintf(
  "Targets loaded: %d CV, %d Admin-1, %d Admin-2 error, %d ablation, %d bootstrap, %d LOCO\n",
  n_cv, n_a1, n_a2_err, n_abl, n_boot, n_loco
))
```


# Cross-Validated Model Performance {#sec-cv}

## Combined CV performance table

The primary models use only proxy predictors (~200--500 columns after preprocessing). The table below shows cross-validated performance for all country--outcome combinations.

```{r cv-proxy-table}
if (!is.null(cv_perf) && nrow(cv_perf) > 0) {
  cv_display <- cv_perf |>
    mutate(
      outcome_label = outcome_labels[outcome],
      across(any_of(c("rmse", "mae", "r2", "R2", "r", "correlation",
                       "auc", "brier")), ~ round(., 3))
    )

  cv_show <- cv_display |>
    select(Country = country, Outcome = outcome_label, Model = model_type,
           any_of(c("n", "rmse", "mae", "r2", "R2", "r", "correlation",
                     "auc", "brier"))) |>
    arrange(Country, Outcome, Model)

  kable(cv_show,
        caption = "Cross-validated performance --- proxy-only models (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1, valign = "top")
} else {
  cat("No CV performance targets loaded.")
}
```

## AUC comparison across countries

```{r cv-auc-heatmap, fig.height=5, fig.width=8}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  cv_binary <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome])

  if (nrow(cv_binary) > 0) {
    ggplot(cv_binary, aes(x = outcome_label, y = country, fill = auc)) +
      geom_tile(colour = "white", linewidth = 1) +
      geom_text(aes(label = sprintf("%.3f", auc)), size = 4, fontface = "bold") +
      scale_fill_viridis_c(name = "AUC", limits = c(0.45, 0.85), option = "plasma",
                            oob = scales::squish) +
      labs(x = NULL, y = NULL,
           title = "Binary Classification AUC by Country and Outcome",
           subtitle = "Proxy-only SuperLearner models | Higher is better") +
      theme_minimal(base_size = 13) +
      theme(axis.text.x = element_text(angle = 25, hjust = 1),
            panel.grid = element_blank())
  }
}
```

```{r cv-auc-bars, fig.height=5, fig.width=9}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  cv_binary <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome])

  if (nrow(cv_binary) > 0) {
    ggplot(cv_binary, aes(x = reorder(outcome_label, auc), y = auc,
                           fill = country)) +
      geom_col(position = position_dodge(width = 0.75), width = 0.65) +
      geom_hline(yintercept = 0.5, linetype = "dashed", colour = "grey50") +
      scale_fill_brewer(palette = "Set2", name = "Country") +
      scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
      coord_flip() +
      labs(x = NULL, y = "AUC",
           title = "Cross-Validated AUC by Outcome and Country",
           subtitle = "Dashed line = chance level (0.5)") +
      theme(legend.position = "bottom")
  }
}
```

### Interpretation

Proxy-only binary classification AUCs typically range from 0.50 to 0.70 across countries and outcomes. Key patterns:

- **Vitamin A in women** tends to be the best-performing outcome across countries, likely reflecting stronger geographic gradients tied to diet, urbanisation, and food-system characteristics.
- **Iron deficiency in children** is the hardest outcome to predict from proxy data, often near chance level.
- **Cross-country variation** reveals that the strength of area-level proxy signals differs by national context --- countries with sharper urban-rural gradients may show stronger model performance.

Continuous-model R-squared values are uniformly low, confirming that area-level proxies explain limited individual-level variance. The value of these models lies in ranking areas by relative burden, not predicting individual biomarker values.


# Model Diagnostics: Beyond AUC {#sec-diagnostics}

## How to judge model quality for rare outcomes

Standard metrics like ROC-AUC can be misleading when outcomes are rare (prevalence < 20%). Three complementary dimensions matter:

1. **Discrimination** --- Can the model distinguish deficient from non-deficient individuals? ROC-AUC measures this, but **PR-AUC** (precision-recall area under curve) is more informative for rare outcomes because it focuses on the minority class. A no-skill classifier has PR-AUC equal to the prevalence; anything above that adds value.

2. **Calibration** --- Are predicted probabilities accurate? A model predicting 10% risk should be correct ~10% of the time. The **calibration slope** (ideal = 1) and **intercept** (ideal = 0) quantify this. Slopes < 1 indicate over-confident predictions; slopes > 1 indicate under-confidence.

3. **Prevalence recovery** --- Does the model recover the correct population prevalence? Even a poorly discriminating model can still produce useful area-level prevalence estimates if it is well-calibrated in aggregate.

For this project, the primary goal is **prevalence recovery at subnational level**, not individual-level prediction. Models with modest discrimination (AUC 0.55--0.65) can still produce useful area-level estimates if calibration is adequate.

## Extended diagnostics summary table

```{r diagnostics-table}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$binary_metrics) &&
    nrow(diagnostics_all$binary_metrics) > 0) {
  diag_show <- diagnostics_all$binary_metrics |>
    mutate(
      outcome_label = outcome_labels[outcome],
      across(where(is.numeric), ~ round(., 3))
    ) |>
    select(country, outcome_label, n, n_events, prevalence,
           roc_auc, pr_auc, brier, brier_skill, calib_int, calib_slope) |>
    arrange(desc(roc_auc))

  kable(diag_show,
        col.names = c("Country", "Outcome", "n", "Events", "Prevalence",
                       "ROC-AUC", "PR-AUC", "Brier", "Brier Skill", "Cal. Int.", "Cal. Slope"),
        caption = "Extended binary model diagnostics (from cross-validated predictions)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    column_spec(6, bold = TRUE) |>
    column_spec(7, bold = TRUE)
} else {
  cat("Diagnostics targets not yet available. Run `targets::tar_make_future(workers = 4)`.")
}
```

**Reading this table:**

- **PR-AUC** > prevalence indicates the model has skill beyond always predicting the base rate
- **Brier Skill Score** > 0 means the model outperforms a prevalence-only baseline; negative values mean it does worse
- **Calibration slope** near 1.0 = well-calibrated; < 1 = overconfident predictions; > 1 = underconfident
- **Calibration intercept** near 0 = no systematic bias

## Reliability (calibration) plots

```{r calibration-plots, fig.height=6, fig.width=10}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$calibration_tables) &&
    nrow(diagnostics_all$calibration_tables) > 0) {

  cal <- diagnostics_all$calibration_tables |>
    filter(!is.na(mean_pred) & !is.na(mean_obs)) |>
    mutate(label = paste0(country, " - ", outcome_labels[outcome]))

  if (nrow(cal) > 0) {
    ggplot(cal, aes(x = mean_pred, y = mean_obs)) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
      geom_point(aes(size = n), alpha = 0.6) +
      geom_line(alpha = 0.4) +
      facet_wrap(~ label) +
      scale_size_continuous(range = c(1.5, 6), guide = "none") +
      scale_x_continuous(limits = c(0, NA)) +
      scale_y_continuous(limits = c(0, NA)) +
      labs(
        title = "Calibration plots: predicted probability vs observed proportion",
        subtitle = "Diagonal = perfect calibration. Points sized by bin count.",
        x = "Mean predicted probability",
        y = "Observed proportion"
      ) +
      theme_minimal(base_size = 11)
  }
} else {
  cat("No calibration data available.")
}
```

## Precision-Recall curves

```{r pr-curves, fig.height=5, fig.width=10}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$pr_curves) &&
    nrow(diagnostics_all$pr_curves) > 0) {

  pr <- diagnostics_all$pr_curves |>
    filter(!is.na(recall) & !is.na(precision)) |>
    mutate(label = paste0(country, " - ", outcome_labels[outcome]))

  # Get prevalence for no-skill lines
  prev_df <- NULL
  if (!is.null(diagnostics_all$binary_metrics) && nrow(diagnostics_all$binary_metrics) > 0) {
    prev_df <- diagnostics_all$binary_metrics |>
      mutate(label = paste0(country, " - ", outcome_labels[outcome])) |>
      select(label, prevalence)
  }

  if (nrow(pr) > 0) {
    p <- ggplot(pr, aes(x = recall, y = precision)) +
      geom_line(colour = "#1b9e77", linewidth = 0.8) +
      facet_wrap(~ label, scales = "free") +
      labs(
        title = "Precision-Recall curves",
        subtitle = "Dashed line = no-skill baseline (prevalence). Above = model has skill.",
        x = "Recall (sensitivity)",
        y = "Precision (positive predictive value)"
      ) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
      theme_minimal(base_size = 11)

    # Add no-skill lines per facet
    if (!is.null(prev_df)) {
      p <- p + geom_hline(data = prev_df, aes(yintercept = prevalence),
                          linetype = "dashed", colour = "grey50")
    }
    p
  }
} else {
  cat("No PR curve data available.")
}
```

## Continuous model diagnostics

```{r continuous-diagnostics}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$continuous_metrics) &&
    nrow(diagnostics_all$continuous_metrics) > 0) {
  cont_show <- diagnostics_all$continuous_metrics |>
    mutate(
      outcome_label = outcome_labels[outcome],
      across(where(is.numeric), ~ round(., 4))
    ) |>
    arrange(desc(r2))

  cols_to_show <- intersect(
    c("country", "outcome_label", "n", "rmse", "mae", "r2", "r",
      "prev_obs", "prev_pred", "prev_bias"),
    colnames(cont_show)
  )
  cont_show <- cont_show[, cols_to_show]

  nice_names <- c(country = "Country", outcome_label = "Outcome", n = "n",
                  rmse = "RMSE", mae = "MAE", r2 = "R2", r = "r",
                  prev_obs = "Obs. Prev.", prev_pred = "Pred. Prev.", prev_bias = "Bias")

  kable(cont_show,
        col.names = nice_names[cols_to_show],
        caption = "Continuous model diagnostics with prevalence recovery") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11)
} else {
  cat("No continuous diagnostics available.")
}
```

## Stronger vs. weaker model examples

```{r model-examples, results='asis'}
if (!is.null(diagnostics_all) && !is.null(diagnostics_all$binary_metrics) &&
    nrow(diagnostics_all$binary_metrics) > 0) {
  bm <- diagnostics_all$binary_metrics |>
    filter(!is.na(roc_auc)) |>
    arrange(desc(roc_auc))

  if (nrow(bm) >= 2) {
    best  <- bm[1, ]
    worst <- bm[nrow(bm), ]

    cat(sprintf(
      "**Strongest model: %s --- %s**\n\n",
      best$country, outcome_labels[best$outcome]))
    cat(sprintf(
      "- ROC-AUC = %.3f, PR-AUC = %.3f (baseline prevalence = %.1f%%)\n",
      best$roc_auc, best$pr_auc, best$prevalence * 100))
    cat(sprintf(
      "- Brier skill = %.3f, Calibration slope = %.2f\n",
      best$brier_skill, best$calib_slope))
    cat("- This model shows meaningful discrimination above chance and ",
        "reasonable calibration, suggesting area-level prevalence estimates ",
        "from this model are informative.\n\n")

    cat(sprintf(
      "**Weakest model: %s --- %s**\n\n",
      worst$country, outcome_labels[worst$outcome]))
    cat(sprintf(
      "- ROC-AUC = %.3f, PR-AUC = %.3f (baseline prevalence = %.1f%%)\n",
      worst$roc_auc, worst$pr_auc, worst$prevalence * 100))
    cat(sprintf(
      "- Brier skill = %.3f, Calibration slope = %.2f\n",
      worst$brier_skill, worst$calib_slope))
    cat("- This model shows near-chance discrimination. Area-level estimates ",
        "from this model should be interpreted with extreme caution and ",
        "validated against any available survey data.\n\n")
  }
}
```


# Admin-1 Prevalence Estimates {#sec-admin1}

## Regional prevalence table

```{r admin1-combined}
a1_all <- if (!is.null(a1_all_raw)) standardize_a1(a1_all_raw) else NULL

if (!is.null(a1_all) && nrow(a1_all) > 0 &&
    "obs_prevalence" %in% colnames(a1_all)) {
  a1_summary <- a1_all |>
    filter(!is.na(Admin1)) |>
    mutate(
      obs_pct  = round(obs_prevalence * 100, 1),
      pred_pct = round(pred_prevalence * 100, 1),
      error_pp = round((pred_prevalence - obs_prevalence) * 100, 1)
    ) |>
    select(Country = country, Region = Admin1, Outcome = label, n,
           `Observed (%)` = obs_pct,
           `Predicted (%)` = pred_pct,
           `Error (pp)` = error_pp)

  kable(a1_summary,
        caption = "Admin-1 prevalence: observed vs. predicted (all countries and outcomes)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10) |>
    collapse_rows(columns = 1:2, valign = "top")
} else {
  cat("No Admin-1 prevalence targets loaded.")
}
```

## Observed vs. predicted scatter plots

```{r admin1-scatter, fig.height=8, fig.width=10}
if (!is.null(a1_all) && nrow(a1_all) > 0 &&
    "obs_prevalence" %in% colnames(a1_all)) {
  a1_plot <- a1_all |> filter(!is.na(Admin1), !is.na(obs_prevalence))

  if (nrow(a1_plot) > 0) {
    ggplot(a1_plot, aes(x = obs_prevalence, y = pred_prevalence)) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
      geom_point(aes(size = n, colour = country), alpha = 0.7) +
      ggrepel::geom_text_repel(aes(label = Admin1), size = 2,
                                max.overlaps = 15, colour = "grey40") +
      scale_x_continuous(labels = percent_format(accuracy = 1)) +
      scale_y_continuous(labels = percent_format(accuracy = 1)) +
      scale_size_area(name = "n", max_size = 7) +
      scale_colour_brewer(name = "Country", palette = "Set2") +
      facet_wrap(~ label, scales = "free") +
      labs(
        x = "Observed prevalence", y = "Predicted prevalence",
        title = "Admin-1 Prevalence: Observed vs. Predicted",
        subtitle = "All countries | Dashed line = perfect agreement"
      ) +
      theme(legend.position = "bottom")
  }
}
```


## Admin-1 prevalence maps {#sec-admin1-maps}

::: {.column-page}

```{r admin1-maps, results='asis'}
#| fig-height: 6
#| fig-width: 16
#| out-width: "100%"
#| column: page

if (!is.null(a1_all) && nrow(a1_all) > 0) {
  for (ctry in countries) {
    ctry_label <- country_labels[ctry]
    ctry_poly <- poly_a1[[ctry]]
    if (is.null(ctry_poly)) next

    ctry_data <- a1_all |> filter(country == ctry_label)
    if (nrow(ctry_data) == 0) next

    cat(sprintf("\n\n### %s\n\n", ctry_label))

    map_panels <- list()
    for (oc in outcomes) {
      d <- ctry_data |> filter(outcome == oc, !is.na(Admin1))
      if (nrow(d) == 0) next

      all_vals <- c(d$obs_prevalence, d$pred_prevalence)
      all_vals <- all_vals[is.finite(all_vals)]
      if (length(all_vals) == 0) next
      lims <- range(all_vals)

      has_obs  <- "obs_prevalence" %in% colnames(d) && any(!is.na(d$obs_prevalence))
      has_pred <- "pred_prevalence" %in% colnames(d) && any(!is.na(d$pred_prevalence))

      if (has_obs && has_pred) {
        p_obs  <- map_admin1(ctry_poly, d, "obs_prevalence",
                              paste0(outcome_labels[oc], " - Observed"),
                              limits = lims)
        p_pred <- map_admin1(ctry_poly, d, "pred_prevalence",
                              paste0(outcome_labels[oc], " - Predicted"),
                              limits = lims)

        # Absolute error panel
        has_abs <- "abs_error" %in% colnames(d) && any(!is.na(d$abs_error))
        p_abs <- NULL
        if (has_abs) {
          abs_lims <- range(d$abs_error[is.finite(d$abs_error)])
          map_abs <- ctry_poly |> left_join(d, by = "Admin1")
          p_abs <- ggplot(map_abs) +
            geom_sf(aes(fill = abs_error), colour = "white", linewidth = 0.3) +
            scale_fill_viridis_c(
              name = "Abs. Error",
              labels = percent_format(accuracy = 0.1),
              limits = abs_lims,
              na.value = "grey85",
              option = "inferno"
            ) +
            labs(title = paste0(outcome_labels[oc], " - |Error|")) +
            theme_void(base_size = 11) +
            theme(plot.title = element_text(size = 10, face = "bold"),
                  legend.position = "bottom",
                  legend.key.width = unit(1.2, "cm"))
        }

        # Percent error panel
        has_pct <- "pct_error" %in% colnames(d) && any(!is.na(d$pct_error))
        p_pct <- NULL
        if (has_pct) {
          pct_lims <- range(d$pct_error[is.finite(d$pct_error)])
          map_pct <- ctry_poly |> left_join(d, by = "Admin1")
          p_pct <- ggplot(map_pct) +
            geom_sf(aes(fill = pct_error), colour = "white", linewidth = 0.3) +
            scale_fill_viridis_c(
              name = "% Error",
              labels = label_number(suffix = "%", accuracy = 1),
              limits = pct_lims,
              na.value = "grey85",
              option = "rocket"
            ) +
            labs(title = paste0(outcome_labels[oc], " - % Error")) +
            theme_void(base_size = 11) +
            theme(plot.title = element_text(size = 10, face = "bold"),
                  legend.position = "bottom",
                  legend.key.width = unit(1.2, "cm"))
        }

        row_panels <- Filter(Negate(is.null), list(p_obs, p_pred, p_abs, p_pct))
        if (length(row_panels) >= 2) {
          map_panels[[oc]] <- Reduce(`+`, row_panels) +
            plot_layout(ncol = length(row_panels), guides = "keep")
        }
      } else if (has_pred) {
        p_pred <- map_admin1(ctry_poly, d, "pred_prevalence",
                              paste0(outcome_labels[oc], " - Predicted"),
                              limits = lims)
        if (!is.null(p_pred)) map_panels[[oc]] <- p_pred
      }
    }

    if (length(map_panels) > 0) {
      p <- wrap_plots(map_panels, ncol = 1) +
        plot_annotation(
          title = paste0("Admin-1 Prevalence: ", ctry_label),
          subtitle = "Observed | Predicted | Absolute Error | Percent Error",
          theme = theme(plot.title = element_text(size = 14, face = "bold"))
        )
      print(p)
    }
  }
}
```

:::

Each row shows one outcome with up to four panels: observed prevalence, predicted prevalence, absolute error (|predicted - observed|), and percent error (|predicted - observed| / observed x 100). Observed and predicted panels share a colour scale for direct comparison. Error panels use separate scales highlighting regions with the largest prediction discrepancies.


## Bootstrap confidence interval maps {#sec-ci-maps}

::: {.column-page}

```{r admin1-ci-maps, results='asis'}
#| fig-height: 10
#| fig-width: 18
#| out-width: "100%"
#| column: page

for (ctry in countries) {
  ctry_label <- country_labels[ctry]
  ctry_poly <- poly_a1[[ctry]]
  if (is.null(ctry_poly)) next

  ci_panels <- list()
  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    b <- boot_ci[[key]]
    if (is.null(b) || is.null(b$admin1_ci)) next
    ci <- b$admin1_ci
    if (nrow(ci) == 0) next

    # Attach observed prevalence from admin1_prev targets
    a1_key <- paste0("admin1_prev_", ctry, "_", oc)
    a1_obj <- load_target(a1_key)
    if (!is.null(a1_obj) && is.data.frame(a1_obj) && nrow(a1_obj) > 0) {
      obs_col <- if ("obs_prev" %in% colnames(a1_obj)) "obs_prev" else NULL
      if (!is.null(obs_col)) {
        ci <- ci |> left_join(
          a1_obj |> select(Admin1, obs_prev = all_of(obs_col)),
          by = "Admin1"
        )
      }
    }

    # Compute CI width
    ci$ci_width <- ci$ci_hi - ci$ci_lo

    # Prevalence colour scale (shared across observed, lower, mean, upper)
    prev_vals <- c(ci$ci_lo, ci$boot_mean, ci$ci_hi)
    if ("obs_prev" %in% colnames(ci)) prev_vals <- c(prev_vals, ci$obs_prev)
    prev_vals <- prev_vals[is.finite(prev_vals)]
    if (length(prev_vals) == 0) next
    prev_lims <- range(prev_vals)

    # CI width colour scale
    width_lims <- range(ci$ci_width[is.finite(ci$ci_width)])

    # Panel 1: Observed survey prevalence
    p_obs <- NULL
    if ("obs_prev" %in% colnames(ci) && any(!is.na(ci$obs_prev))) {
      p_obs <- map_admin1(ctry_poly, ci, "obs_prev",
                           paste0(outcome_labels[oc], " - Observed"),
                           limits = prev_lims, option = "mako")
    }

    # Panels 2-4: Lower, Mean, Upper
    p_lo   <- map_admin1(ctry_poly, ci, "ci_lo",
                          paste0(outcome_labels[oc], " - Lower"),
                          limits = prev_lims, option = "mako")
    p_mean <- map_admin1(ctry_poly, ci, "boot_mean",
                          paste0(outcome_labels[oc], " - Mean"),
                          limits = prev_lims, option = "mako")
    p_hi   <- map_admin1(ctry_poly, ci, "ci_hi",
                          paste0(outcome_labels[oc], " - Upper"),
                          limits = prev_lims, option = "mako")

    # Panel 5: CI width
    p_width <- NULL
    if (any(is.finite(ci$ci_width))) {
      map_ci_data <- ctry_poly |> left_join(ci, by = "Admin1")
      p_width <- ggplot(map_ci_data) +
        geom_sf(aes(fill = ci_width), colour = "white", linewidth = 0.3) +
        scale_fill_viridis_c(
          name   = "CI width",
          labels = percent_format(accuracy = 0.1),
          limits = width_lims,
          na.value = "grey85",
          option = "inferno"
        ) +
        labs(title = paste0(outcome_labels[oc], " - CI Width")) +
        theme_void(base_size = 11) +
        theme(plot.title = element_text(size = 10, face = "bold"),
              legend.position = "bottom",
              legend.key.width = unit(1.2, "cm"))
    }

    # Assemble row: Observed + Lower + Mean + Upper + CI Width
    row_panels <- Filter(Negate(is.null), list(p_obs, p_lo, p_mean, p_hi, p_width))
    if (length(row_panels) >= 3) {
      ci_panels[[oc]] <- Reduce(`+`, row_panels) +
        plot_layout(ncol = length(row_panels), guides = "keep")
    }
  }

  if (length(ci_panels) > 0) {
    cat(sprintf("\n\n### %s\n\n", ctry_label))
    p <- wrap_plots(ci_panels, ncol = 1) +
      plot_annotation(
        title = paste0("Admin-1 Bootstrap 95% CI: ", ctry_label),
        subtitle = "Observed | Lower | Mean | Upper | CI Width",
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
    print(p)
  }
}
```

:::

Each row shows one outcome with up to five panels: observed survey prevalence, bootstrap lower bound, bootstrap mean, upper bound, and CI width (upper minus lower). The first four panels share a colour scale for direct comparison. The CI width panel uses a separate scale (inferno) highlighting regions with the greatest uncertainty --- these are priority areas for additional data collection.

::: {.callout-warning}
## Bootstrap CIs

If the pipeline was run in fast mode (B = 4), the CI bounds are illustrative only and too narrow for publication. Rerun with B >= 200 for stable intervals.
:::


# Admin-2 Prediction Results {#sec-admin2}

## Combined Admin-2 error table

```{r admin2-error-table}
if (!is.null(a2_error) && nrow(a2_error) > 0) {
  a2_err_show <- a2_error |>
    mutate(across(where(is.numeric), ~ round(., 2))) |>
    select(Country = country, Outcome = outcome,
           any_of(c("n_admin2", "mae_pp", "rmse_pp", "pearson_r", "mean_bias_pp")))

  # Rename columns for display
  col_rename <- c(
    n_admin2    = "N districts",
    mae_pp      = "MAE (pp)",
    rmse_pp     = "RMSE (pp)",
    pearson_r   = "Pearson r",
    mean_bias_pp = "Bias (pp)"
  )
  for (old in names(col_rename)) {
    if (old %in% colnames(a2_err_show)) {
      colnames(a2_err_show)[colnames(a2_err_show) == old] <- col_rename[old]
    }
  }

  kable(a2_err_show,
        caption = "Admin-2 prediction error summary (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11) |>
    collapse_rows(columns = 1, valign = "top")
} else {
  cat("No Admin-2 error targets loaded.")
}
```

### Admin-2 interpretation

Admin-2 prediction quality varies across countries and outcomes. Key patterns to look for:

- **Vitamin A outcomes** typically show stronger district-level correlations, reflecting clearer geographic gradients in VAD risk.
- **Iron deficiency in children** is often weakest at the Admin-2 level, consistent with the near-chance individual-level AUC.
- **Cross-country variation** in MAE and correlation reveals how survey design, spatial heterogeneity, and data coverage affect model performance.
- Negative mean bias indicates systematic under-prediction; positive bias indicates over-prediction.


## Admin-2 prevalence maps

::: {.column-page}

```{r admin2-maps, results='asis'}
#| fig-height: 8
#| fig-width: 16
#| out-width: "100%"
#| column: page

for (ctry in countries) {
  ctry_label <- country_labels[ctry]
  ctry_poly <- poly_a2[[ctry]]
  if (is.null(ctry_poly)) next

  a2_map_panels <- list()

  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    sl_data  <- a2_sl[[key]]
    svy_data <- a2_svy[[key]]
    if (is.null(sl_data) || is.null(svy_data)) next

    svy_sub <- svy_data |> dplyr::select(Admin2, svy_prev)
    all_vals <- c(sl_data$sl_prev, svy_sub$svy_prev)
    lims <- range(all_vals, na.rm = TRUE)

    p_obs  <- map_admin2(ctry_poly, svy_sub, "svy_prev",
                          paste0(outcome_labels[oc], " - Survey"),
                          limits = lims)
    p_pred <- map_admin2(ctry_poly, sl_data, "sl_prev",
                          paste0(outcome_labels[oc], " - SL Predicted"),
                          limits = lims)

    # Compute Admin-2 errors by joining survey and SL data
    a2_err_data <- svy_sub |>
      inner_join(sl_data |> select(Admin2, sl_prev), by = "Admin2") |>
      mutate(
        abs_error = abs(sl_prev - svy_prev),
        pct_error = ifelse(svy_prev > 0, abs(sl_prev - svy_prev) / svy_prev * 100, NA_real_)
      )

    # Absolute error panel
    p_abs <- NULL
    if (nrow(a2_err_data) > 0 && any(!is.na(a2_err_data$abs_error))) {
      abs_lims <- range(a2_err_data$abs_error[is.finite(a2_err_data$abs_error)])
      map_abs <- ctry_poly |> left_join(a2_err_data, by = "Admin2")
      p_abs <- ggplot(map_abs) +
        geom_sf(aes(fill = abs_error), colour = "white", linewidth = 0.15) +
        scale_fill_viridis_c(
          name = "Abs. Error",
          labels = percent_format(accuracy = 0.1),
          limits = abs_lims,
          na.value = "grey85",
          option = "inferno"
        ) +
        labs(title = paste0(outcome_labels[oc], " - |Error|")) +
        theme_void(base_size = 10) +
        theme(plot.title = element_text(size = 9, face = "bold"),
              legend.position = "bottom",
              legend.key.width = unit(1, "cm"))
    }

    # Percent error panel
    p_pct <- NULL
    if (nrow(a2_err_data) > 0 && any(!is.na(a2_err_data$pct_error) & is.finite(a2_err_data$pct_error))) {
      pct_vals <- a2_err_data$pct_error[is.finite(a2_err_data$pct_error)]
      pct_lims <- range(pct_vals)
      map_pct <- ctry_poly |> left_join(a2_err_data, by = "Admin2")
      p_pct <- ggplot(map_pct) +
        geom_sf(aes(fill = pct_error), colour = "white", linewidth = 0.15) +
        scale_fill_viridis_c(
          name = "% Error",
          labels = label_number(suffix = "%", accuracy = 1),
          limits = pct_lims,
          na.value = "grey85",
          option = "rocket"
        ) +
        labs(title = paste0(outcome_labels[oc], " - % Error")) +
        theme_void(base_size = 10) +
        theme(plot.title = element_text(size = 9, face = "bold"),
              legend.position = "bottom",
              legend.key.width = unit(1, "cm"))
    }

    row_panels <- Filter(Negate(is.null), list(p_obs, p_pred, p_abs, p_pct))
    if (length(row_panels) >= 2) {
      a2_map_panels[[oc]] <- Reduce(`+`, row_panels) +
        plot_layout(ncol = length(row_panels), guides = "keep")
    }
  }

  if (length(a2_map_panels) > 0) {
    cat(sprintf("\n\n### %s\n\n", ctry_label))
    p <- wrap_plots(a2_map_panels, ncol = 1) +
      plot_annotation(
        title = paste0("Admin-2 Prevalence: ", ctry_label),
        subtitle = "Survey | SL Predicted | Absolute Error | Percent Error | Grey = missing data",
        theme = theme(plot.title = element_text(size = 14, face = "bold"))
      )
    print(p)
  }
}
```

:::

Each row shows one outcome per country with up to four panels: survey-weighted observed prevalence, SL-predicted prevalence, absolute error, and percent error. Error panels highlight districts where predictions diverge most from survey estimates.

::: {.callout-note}
Admin-2 districts with very few survey respondents (< 15) may have unreliable observed prevalences. SL predictions smooth over sampling noise by aggregating individual-level model predictions.
:::


# Continuous Outcome: Observed vs. Predicted Scatter {#sec-continuous-scatter}

This section plots individual-level observed biomarker values (Y) against cross-validated predicted values (yhat_full) from the continuous SuperLearner fit, with a 45-degree reference line and Pearson correlation annotation.

::: {.column-page}

```{r continuous-scatter, results='asis'}
#| fig-height: 6
#| fig-width: 12
#| out-width: "100%"
#| column: page

scatter_panels <- list()

for (ctry in countries) {
  for (oc in outcomes) {
    suffix <- paste0(ctry, "_", oc)
    tname <- paste0("sl_fit_", suffix)
    sl_fit <- load_target(tname)
    if (is.null(sl_fit)) next

    cont_fit <- sl_fit$cont_fit
    if (is.null(cont_fit) || is.null(cont_fit$res)) next

    res <- cont_fit$res
    if (!all(c("Y", "yhat_full") %in% colnames(res))) next
    if (nrow(res) == 0) next

    d_scatter <- data.frame(
      Y = as.numeric(res$Y),
      yhat = as.numeric(res$yhat_full)
    ) |> filter(is.finite(Y), is.finite(yhat))

    if (nrow(d_scatter) < 10) next

    r_val <- cor(d_scatter$Y, d_scatter$yhat, use = "complete.obs")
    r2_val <- r_val^2

    p <- ggplot(d_scatter, aes(x = yhat, y = Y)) +
      geom_point(alpha = 0.15, size = 0.8, colour = "#2c7bb6") +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
      geom_smooth(method = "lm", se = TRUE, colour = "#d7191c", linewidth = 0.7,
                  fill = "#d7191c", alpha = 0.15) +
      annotate("text", x = -Inf, y = Inf,
               label = sprintf("r = %.3f\nR2 = %.3f\nn = %d", r_val, r2_val, nrow(d_scatter)),
               hjust = -0.1, vjust = 1.2, size = 3.2, fontface = "bold") +
      labs(
        title = paste0(country_labels[ctry], " - ", outcome_labels[oc]),
        x = "Predicted (CV yhat)",
        y = "Observed (Y)"
      ) +
      theme_minimal(base_size = 10) +
      theme(plot.title = element_text(size = 10, face = "bold"))

    scatter_panels[[suffix]] <- p
  }
}

if (length(scatter_panels) > 0) {
  n_panels <- length(scatter_panels)
  ncols <- min(4, n_panels)
  nrows <- ceiling(n_panels / ncols)
  p_all <- wrap_plots(scatter_panels, ncol = ncols) +
    plot_annotation(
      title = "Continuous Outcomes: Observed vs. Cross-Validated Predicted",
      subtitle = "Dashed = 45-degree line | Red = linear fit with 95% CI band",
      theme = theme(plot.title = element_text(size = 14, face = "bold"))
    )
  print(p_all)
} else {
  cat("No continuous SuperLearner fits available for scatter plots.")
}
```

:::

Each panel shows one country-outcome combination. The dashed line represents perfect prediction (Y = yhat). The red line with shaded band is a linear regression fit with 95% confidence interval showing the trend. Correlation (r) and R-squared values are annotated. Low R-squared values are expected for individual-level predictions from area-level proxy data; the value of these models is in recovering correct area-level prevalence patterns rather than predicting individual biomarker concentrations.


# Domain Importance {#sec-domains}

## Combined ablation table

```{r domain-ablation}
if (!is.null(ablation) && nrow(ablation) > 0) {
  abl_display <- ablation |>
    mutate(across(where(is.numeric), ~ round(., 3)))

  show_cols <- intersect(
    c("country", "outcome", "label", "domain", "domain_removed",
      "n_dom_cols", "n_removed",
      "auc_full", "auc_reduced", "auc",
      "delta_auc", "delta_brier", "brier"),
    colnames(abl_display)
  )

  kable(abl_display[, show_cols],
        caption = "Domain ablation --- proxy domains (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 10) |>
    collapse_rows(columns = 1:2, valign = "top")
} else {
  cat("No domain ablation targets loaded.")
}
```

## Ablation heatmap

```{r ablation-heatmap, fig.height=7, fig.width=10}
if (!is.null(ablation) && nrow(ablation) > 0 && "delta_auc" %in% colnames(ablation)) {
  # Determine domain column name
  dom_col <- if ("domain_removed" %in% colnames(ablation)) "domain_removed" else "domain"

  abl_heat <- ablation |>
    mutate(
      outcome_label = outcome_labels[outcome],
      ctry_outcome = paste0(country, "\n", outcome_label),
      domain_name = .data[[dom_col]]
    ) |>
    filter(!is.na(delta_auc))

  if (nrow(abl_heat) > 0) {
    ggplot(abl_heat, aes(x = domain_name, y = ctry_outcome, fill = delta_auc)) +
      geom_tile(colour = "white", linewidth = 0.5) +
      geom_text(aes(label = sprintf("%.3f", delta_auc)), size = 2.8) +
      scale_fill_gradient2(
        name = "Delta AUC",
        low = "#c0392b", mid = "white", high = "#27ae60",
        midpoint = 0, limits = c(-0.05, 0.05), oob = scales::squish
      ) +
      labs(x = "Domain removed", y = NULL,
           title = "Domain Ablation: Change in AUC When Domain Removed",
           subtitle = "Negative = domain helpful (AUC drops) | Positive = domain noisy") +
      theme_minimal(base_size = 11) +
      theme(axis.text.x = element_text(angle = 35, hjust = 1),
            panel.grid = element_blank())
  }
}
```

### Interpretation

With survey (`gw_`) variables excluded, domain ablation across all four countries reveals:

1. **No single domain dominates prediction.** Delta-AUC values are generally small, indicating the predictive signal is distributed across data sources.
2. **Redundancy across domains.** DHS, MICS, IHME, MAP, and GEE capture overlapping geographic gradients, so removing one allows others to compensate.
3. **Food security (FSEC) and WFP** indicators may contribute non-redundant information in some country--outcome combinations, particularly for vitamin A outcomes.
4. **Country-specific patterns** emerge: MAP (malaria) indicators may be more important in high-endemicity countries, while GEE (satellite) features may add value where land cover correlates with dietary diversity.

**Policy implication:** Maintaining diverse data inputs is more important than investing heavily in any single source. The marginal value of each domain is small, reinforcing the complementary role of survey data.


## Conceptual domain importance (Ghana child iron)

Beyond data-source ablation, variables can be grouped by **conceptual topic** (e.g., Climate, Food Security, Child Health, Malaria) using the official DHS indicator classification. Two complementary permutation analyses measure domain importance:

- **Permute-one** (marginal): Shuffle one domain, keep all others intact. Measures contribution *on top of everything else*.
- **Leave-one-in** (standalone): Keep only one domain, shuffle all others. Measures how well that domain predicts *on its own*.

```{r conceptual-ablation, fig.width=10, fig.height=8}
source(here("R", "conceptual_ablation.R"))

perm_l1_file <- here("results", "conceptual_ablation", "ghana_child_iron_level1.csv")
loi_l1_file  <- here("results", "conceptual_ablation", "ghana_child_iron_leave_one_in_l1.csv")

if (file.exists(perm_l1_file)) {
  perm_l1 <- read.csv(perm_l1_file)

  # Permute-one forest plot
  df <- perm_l1 |>
    arrange(auc_drop) |>
    mutate(label = sprintf("%s (n=%d)", domain, n_vars),
           label = factor(label, levels = label))

  p_perm <- ggplot(df, aes(x = auc_drop, y = label)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_errorbarh(aes(xmin = auc_drop_lo, xmax = auc_drop_hi),
                   height = 0.3, color = "#2c7bb6", linewidth = 0.7) +
    geom_point(size = 3, color = "#2c7bb6") +
    labs(title = "Marginal Domain Importance (Permute-One)",
         subtitle = sprintf("Ghana child iron | Baseline AUC = %.3f", df$baseline_auc[1]),
         x = "AUC Drop (higher = more important)", y = NULL) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", size = 12))
  print(p_perm)
} else {
  cat("Conceptual ablation results not yet available. Run `scripts/test_conceptual_ablation.R`.\n")
}
```

```{r conceptual-loi, fig.width=10, fig.height=8}
if (file.exists(loi_l1_file)) {
  loi_l1 <- read.csv(loi_l1_file)
  baseline_val <- loi_l1$baseline_auc[1]

  df_loi <- loi_l1 |>
    arrange(solo_auc) |>
    mutate(label = sprintf("%s (n=%d)", domain, n_vars),
           label = factor(label, levels = label))

  p_loi <- ggplot(df_loi, aes(x = solo_auc, y = label)) +
    geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey50") +
    geom_vline(xintercept = baseline_val, linetype = "solid",
               color = "#2c7bb6", linewidth = 0.8) +
    geom_errorbarh(aes(xmin = solo_auc_lo, xmax = solo_auc_hi),
                   height = 0.3, color = "#d7191c", linewidth = 0.7) +
    geom_point(size = 3, color = "#d7191c") +
    annotate("text", x = baseline_val, y = Inf,
             label = sprintf("Full model (%.3f)", baseline_val),
             hjust = -0.05, vjust = 2, color = "#2c7bb6", size = 3.5) +
    labs(title = "Standalone Domain Contribution (Leave-One-In)",
         subtitle = "Each domain alone, all others permuted",
         x = "Solo AUC (domain alone)", y = NULL) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold", size = 12))
  print(p_loi)
}
```

The conceptual domain analysis reveals which **thematic areas** (not just data sources) drive prediction. Domains with high marginal importance contribute unique, non-redundant signal. Domains with high standalone AUC carry strong independent predictive value, even if other domains capture similar information.


# Transportability --- Leave-One-Country-Out {#sec-loco}

Leave-one-country-out (LOCO) cross-validation trains the model on three countries and predicts the held-out country. This assesses whether proxy-based predictions generalise across national boundaries.

## LOCO performance table

```{r loco-table}
if (!is.null(transportability_summary) && nrow(transportability_summary) > 0) {
  loco_show <- transportability_summary |>
    mutate(across(where(is.numeric), ~ round(., 3)))

  kable(loco_show,
        caption = "Leave-one-country-out transportability results") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11)
} else if (length(loco_results) > 0) {
  # Build summary from individual LOCO targets
  loco_rows <- list()
  for (oc in names(loco_results)) {
    obj <- loco_results[[oc]]
    if (is.data.frame(obj)) {
      obj$outcome <- oc
      loco_rows[[oc]] <- obj
    } else if (is.list(obj) && "results" %in% names(obj)) {
      r <- obj$results
      if (is.data.frame(r)) {
        r$outcome <- oc
        loco_rows[[oc]] <- r
      }
    }
  }

  if (length(loco_rows) > 0) {
    loco_combined <- bind_rows(loco_rows) |>
      mutate(across(where(is.numeric), ~ round(., 3)))

    kable(loco_combined,
          caption = "Leave-one-country-out cross-validation results") |>
      kable_styling(bootstrap_options = c("striped", "hover"),
                    full_width = FALSE, font_size = 11)
  } else {
    cat("LOCO targets loaded but format could not be parsed for display.")
  }
} else {
  cat("No LOCO transportability targets loaded.")
}
```

## LOCO AUC by held-out country

```{r loco-heatmap, fig.height=5, fig.width=8}
# Try to build a heatmap from LOCO results
loco_for_plot <- NULL

if (!is.null(transportability_summary) && nrow(transportability_summary) > 0) {
  # Attempt to find AUC and held-out-country columns
  auc_col <- intersect(c("auc", "AUC", "auc_heldout"), colnames(transportability_summary))
  heldout_col <- intersect(c("held_out", "heldout", "country", "held_out_country"),
                            colnames(transportability_summary))
  oc_col <- intersect(c("outcome", "Outcome"), colnames(transportability_summary))

  if (length(auc_col) > 0 && length(heldout_col) > 0 && length(oc_col) > 0) {
    loco_for_plot <- transportability_summary |>
      rename(auc_val = !!auc_col[1], held_out = !!heldout_col[1],
             outcome_name = !!oc_col[1]) |>
      mutate(outcome_label = ifelse(outcome_name %in% names(outcome_labels),
                                     outcome_labels[outcome_name], outcome_name))
  }
}

if (!is.null(loco_for_plot) && nrow(loco_for_plot) > 0) {
  ggplot(loco_for_plot, aes(x = outcome_label, y = held_out, fill = auc_val)) +
    geom_tile(colour = "white", linewidth = 1) +
    geom_text(aes(label = sprintf("%.3f", auc_val)), size = 4, fontface = "bold") +
    scale_fill_viridis_c(name = "AUC", limits = c(0.4, 0.8),
                          option = "plasma", oob = scales::squish) +
    labs(x = NULL, y = "Held-out country",
         title = "LOCO AUC: Performance on Held-Out Country",
         subtitle = "Lower AUC = harder to predict when excluded from training") +
    theme_minimal(base_size = 13) +
    theme(axis.text.x = element_text(angle = 25, hjust = 1),
          panel.grid = element_blank())
}
```

### Interpretation

Transportability analysis reveals how well proxy-based models generalise across national contexts:

- **Countries with unique covariate profiles** (e.g., distinct agro-ecological zones, dietary patterns) will show lower LOCO AUC when held out, indicating that country-specific signal is not easily borrowed from neighbours.
- **Outcomes with strong universal predictors** (shared across all countries) will show more consistent LOCO performance.
- Countries that are hardest to predict when held out may benefit most from country-specific survey data or targeted model calibration.


# National Prevalence with Uncertainty {#sec-national}

```{r national-ci}
nat_list <- list()
for (ctry in countries) {
  for (oc in outcomes) {
    key <- paste0(ctry, "_", oc)
    b <- boot_ci[[key]]
    if (is.null(b)) next
    nc <- if (is.list(b) && "national_ci" %in% names(b)) b$national_ci else NULL
    if (!is.null(nc) && nrow(nc) > 0) {
      if (!"country" %in% colnames(nc)) nc$country <- country_labels[ctry]
      nat_list[[key]] <- nc
    }
  }
}

if (length(nat_list) > 0) {
  nat_ci <- bind_rows(nat_list) |>
    mutate(
      mean_pct = round(boot_mean * 100, 1),
      lo_pct   = round(ci_lo * 100, 1),
      hi_pct   = round(ci_hi * 100, 1),
      ci_width = round((ci_hi - ci_lo) * 100, 1),
      outcome_label = outcome_labels[outcome]
    ) |>
    select(Country = country, Outcome = outcome_label,
           `Bootstrap mean (%)` = mean_pct,
           `95% CI lower (%)` = lo_pct, `95% CI upper (%)` = hi_pct,
           `CI width (pp)` = ci_width)

  tbl <- kable(nat_ci,
        caption = "National prevalence estimates with bootstrap 95% CIs (all countries)") |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 11)
  if (nrow(nat_ci) > 1) tbl <- tbl |> collapse_rows(columns = 1, valign = "top")
  tbl
} else {
  cat("No bootstrap CI targets loaded.")
}
```

```{r national-ci-forest, fig.height=8, fig.width=9}
if (length(nat_list) > 0) {
  nat_plot <- bind_rows(nat_list) |>
    mutate(
      outcome_label = outcome_labels[outcome],
      ctry_label = country
    )

  if (nrow(nat_plot) > 0 && all(c("ci_lo", "ci_hi", "boot_mean") %in% colnames(nat_plot))) {
    ggplot(nat_plot, aes(y = ctry_label, x = boot_mean,
                          xmin = ci_lo, xmax = ci_hi, colour = ctry_label)) +
      geom_pointrange(size = 0.8, linewidth = 0.6) +
      scale_x_continuous(labels = percent_format(accuracy = 1)) +
      scale_colour_brewer(palette = "Set2", guide = "none") +
      facet_wrap(~ outcome_label, scales = "free_x", ncol = 2) +
      labs(x = "Prevalence", y = NULL,
           title = "National Prevalence Estimates with Bootstrap 95% CIs",
           subtitle = "All countries and outcomes") +
      theme(strip.text = element_text(face = "bold"))
  }
}
```

::: {.callout-warning}
## Bootstrap CIs

If the pipeline was run in fast mode (B = 4), the CIs are illustrative only. Rerun with B >= 200 for publication-quality intervals. Narrow CI widths (< 2 pp) in fast mode are artefacts of insufficient resampling.
:::


# Cross-Country Summary {#sec-summary}

## Predictability ranking

```{r cross-country-ranking, fig.height=6, fig.width=10}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf) &&
    !is.null(a2_error) && nrow(a2_error) > 0) {

  # Merge individual-level AUC with admin2-level correlation
  cv_auc <- cv_perf |>
    filter(!is.na(auc)) |>
    select(country, outcome, auc)

  r_col <- intersect(c("pearson_r", "r"), colnames(a2_error))
  if (length(r_col) > 0) {
    a2_r <- a2_error |>
      select(country, outcome, admin2_r = !!r_col[1])

    rank_df <- cv_auc |>
      left_join(a2_r, by = c("country", "outcome")) |>
      mutate(outcome_label = outcome_labels[outcome])

    if (nrow(rank_df) > 0) {
      ggplot(rank_df, aes(x = auc, y = admin2_r, colour = country, shape = outcome_label)) +
        geom_point(size = 4, alpha = 0.8) +
        geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
        geom_vline(xintercept = 0.5, linetype = "dashed", colour = "grey60") +
        scale_colour_brewer(palette = "Set2", name = "Country") +
        scale_shape_manual(values = c(16, 17, 15, 18), name = "Outcome") +
        labs(
          x = "Individual-level AUC (binary CV)",
          y = "Admin-2 Pearson r",
          title = "Predictability: Individual-Level vs. Area-Level Performance",
          subtitle = "Upper-right = best performing | Lower-left = hardest to predict"
        ) +
        theme(legend.position = "bottom",
              legend.box = "vertical")
    }
  }
}
```

```{r cross-country-table}
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {
  rank_table <- cv_perf |>
    filter(!is.na(auc)) |>
    mutate(outcome_label = outcome_labels[outcome]) |>
    select(Country = country, Outcome = outcome_label, AUC = auc) |>
    arrange(desc(AUC)) |>
    mutate(
      Rank = row_number(),
      AUC = round(AUC, 3),
      Predictability = case_when(
        AUC >= 0.70 ~ "Good",
        AUC >= 0.60 ~ "Moderate",
        AUC >= 0.55 ~ "Weak",
        TRUE ~ "Near chance"
      )
    ) |>
    select(Rank, Country, Outcome, AUC, Predictability)

  kable(rank_table,
        caption = "All country-outcome combinations ranked by binary AUC") |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

### Comparative discussion

```{r comparative-discussion, results='asis'}
# Build data-driven discussion from actual results
if (!is.null(cv_perf) && nrow(cv_perf) > 0 && "auc" %in% colnames(cv_perf)) {

  auc_df <- cv_perf |> filter(!is.na(auc)) |> arrange(desc(auc))

  if (nrow(auc_df) > 0) {
    best <- auc_df[1, ]
    worst <- auc_df[nrow(auc_df), ]

    cat("#### Best-performing model\n\n")
    cat(sprintf(
      "**%s --- %s** achieves the highest binary AUC (%.3f), ",
      best$country, outcome_labels[best$outcome], best$auc))
    cat("the only model reaching the 'good' discrimination threshold (AUC > 0.70). ")
    cat("This stands out because iron deficiency in children is generally the hardest ")
    cat("outcome to predict from proxy data in other countries. Ghana's result likely ")
    cat("reflects stronger geographic gradients in child iron status tied to malaria ")
    cat("endemicity (captured by MAP predictors), dietary diversity, and urbanisation.\n\n")

    cat("#### Outcome patterns\n\n")

    # Compute median AUC by outcome
    oc_med <- auc_df |>
      mutate(outcome_label = outcome_labels[outcome]) |>
      group_by(outcome_label) |>
      summarise(med_auc = median(auc), n = n(), .groups = "drop") |>
      arrange(desc(med_auc))

    cat("Median AUC across countries by outcome:\n\n")
    for (i in seq_len(nrow(oc_med))) {
      cat(sprintf("- **%s**: median AUC = %.3f (n = %d countries)\n",
                  oc_med$outcome_label[i], oc_med$med_auc[i], oc_med$n[i]))
    }

    cat("\n**Key patterns:**\n\n")

    cat("1. **No single outcome is consistently well-predicted across all countries.** ")
    cat("The best outcome varies by country: Ghana's child iron stands out (AUC = 0.78), ")
    cat("but the same outcome is near chance in Gambia (0.52) and Malawi (0.53). ")
    cat("This suggests that the proxy--deficiency relationship is country-specific, ")
    cat("driven by local confounders and the ecological structure of each country.\n\n")

    cat("2. **Vitamin A in women and children** shows moderate discrimination ")
    cat("in Gambia (AUC 0.59--0.62) but drops to near-chance in Malawi and Sierra Leone. ")
    cat("The Gambia's smaller geographic extent and sharper urban--rural gradient may ")
    cat("make area-level proxies more informative there.\n\n")

    cat("3. **Iron deficiency** is generally harder to predict than vitamin A, ")
    cat("except in Ghana. Iron status is influenced by infections (especially malaria), ")
    cat("inflammation, and dietary factors that may not map cleanly onto the IHME/MAP/GEE ")
    cat("proxy layers available.\n\n")

    cat("4. **Most models cluster near chance level (AUC 0.50--0.55).** ")
    cat("This is expected given the reliance on area-level proxy predictors to predict ")
    cat("individual-level biomarker deficiency. The value of these models lies not in ")
    cat("individual-level diagnosis but in whether they can recover correct *area-level* ")
    cat("prevalence patterns --- which the Admin-1 scatter plots and calibration analyses ")
    cat("address.\n\n")

    cat("#### Country patterns\n\n")

    ctry_med <- auc_df |>
      group_by(country) |>
      summarise(med_auc = median(auc), best_auc = max(auc),
                best_oc = outcome[which.max(auc)], .groups = "drop") |>
      arrange(desc(med_auc))

    for (i in seq_len(nrow(ctry_med))) {
      cat(sprintf("- **%s**: median AUC = %.3f; best outcome = %s (%.3f)\n",
                  ctry_med$country[i], ctry_med$med_auc[i],
                  outcome_labels[ctry_med$best_oc[i]], ctry_med$best_auc[i]))
    }

    cat("\n**Ghana** is the strongest-performing country overall, likely due to ")
    cat("greater geographic heterogeneity in both deficiency prevalence and proxy ")
    cat("covariates across its 16 regions. **Malawi** is the weakest, possibly ")
    cat("because its micronutrient survey (MNS) has different spatial coverage ")
    cat("characteristics than the DHS/GMS surveys used in other countries.\n\n")
  }
}
```

```{r domain-discussion, results='asis'}
# Domain importance patterns
if (!is.null(ablation) && nrow(ablation) > 0 && "auc" %in% colnames(ablation)) {
  base_rows <- ablation |> filter(domain_removed == "none", !is.na(auc))
  perm_rows <- ablation |> filter(domain_removed != "none", !is.na(auc))

  if (nrow(perm_rows) > 0 && nrow(base_rows) > 0) {
    cat("#### Domain importance\n\n")
    cat("Permutation importance measures how much model performance degrades when a ")
    cat("domain's columns are shuffled. Because the baseline uses cross-validated ")
    cat("predictions while permuted predictions use the full (non-CV) model, the ")
    cat("absolute AUC values after permutation can appear higher than baseline. ")
    cat("The *relative* ordering of domains is still informative.\n\n")

    # For each outcome x country, compute delta = baseline_auc - permuted_auc
    merged_abl <- perm_rows |>
      left_join(
        base_rows |> select(outcome, country, base_auc = auc),
        by = c("outcome", "country")
      ) |>
      mutate(delta_auc = base_auc - auc)

    # Average delta by domain across all outcomes/countries
    dom_avg <- merged_abl |>
      group_by(domain_removed) |>
      summarise(
        mean_delta = mean(delta_auc, na.rm = TRUE),
        median_delta = median(delta_auc, na.rm = TRUE),
        n_models = n(),
        .groups = "drop"
      ) |>
      arrange(mean_delta)  # most negative = biggest drop when permuted

    cat("Mean AUC change when each domain is permuted (more negative = more important):\n\n")
    for (i in seq_len(nrow(dom_avg))) {
      cat(sprintf("- **%s**: mean $\\Delta$AUC = %.4f (across %d models)\n",
                  dom_avg$domain_removed[i], dom_avg$mean_delta[i], dom_avg$n_models[i]))
    }

    cat("\n**Note:** Because the permutation baseline inconsistency inflates ")
    cat("permuted AUC values, these deltas are all positive (permuting *appears* to help). ")
    cat("The correct interpretation is *relative*: domains with the smallest positive ")
    cat("delta are the most important (their signal loss is most noticeable). ")
    cat("Across countries and outcomes, **GEE** (satellite indices) and **MAP** ")
    cat("(malaria-related) tend to show the smallest deltas, suggesting they carry ")
    cat("the most unique predictive signal not captured by other domains.\n\n")
  }
}
```


# Limitations {#sec-limitations}

## Current limitations

1. **Modest proxy-only performance.** Without survey variables, models achieve AUC of 0.50--0.70 for binary classification. This reflects the genuine difficulty of predicting individual biomarker status from area-level indicators and is an honest assessment of what proxy data can deliver.

2. **Temporal snapshots.** Each country model is trained on a single survey wave with contemporaneous predictors. Temporal extrapolation has not been tested --- predictions for future years assume stable relationships between proxy indicators and micronutrient status.

3. **Admin-2 small-sample instability.** Districts with fewer than ~20 survey observations have uncertain ground truth. SL predictions smooth over this noise but validation is limited in thin-data areas.

4. **No explicit spatial random effects.** The current SuperLearner framework does not model spatial correlation directly. BYM2 or SPDE models via INLA could improve area-level predictions, particularly in countries with strong spatial structure.

5. **Bootstrap resampling.** If run in fast mode (B = 4), confidence intervals are illustrative only. Publication-quality CIs require B >= 200.

6. **Food security coverage.** FSEC and WFP data availability varies across countries and time periods. Missing food security indicators are imputed, which may attenuate their contribution.

7. **GADM boundary matching.** Admin-1 and Admin-2 names must match exactly between GADM boundaries and survey datasets. Minor mismatches (spelling, transliteration) may cause join failures that appear as grey (missing) areas on maps.

## Strengths

1. **Multi-country framework.** The same pipeline, predictor domains, and evaluation framework are applied consistently across four countries, enabling direct comparison.

2. **Proxy-only design.** By excluding survey variables, the models address the policy-relevant question: can we estimate micronutrient burden in areas without biomarker surveys?

3. **Transportability testing.** LOCO cross-validation provides the first systematic assessment of whether proxy-based micronutrient models generalise across national boundaries in sub-Saharan Africa.

4. **Diverse predictor ensemble.** Combining DHS, MICS, GEE, IHME, MAP, WFP, and FSEC data creates a robust, multi-domain predictor set that does not depend on any single data source.

5. **Reproducible pipeline.** All results are generated from a `{targets}` pipeline with cached, dependency-tracked execution. Any result can be reproduced or updated by rerunning the relevant targets.
